---
title: "Section 5 Analysis"
author: "Anonymous"
date: today
date-format: "iso"
format:
  html:
    reference-location: margin
    embed-resources: true
---

```{r}
#| code-fold: true
#| label: set up
#| echo: false

library(tidyverse)
library(jsonlite)
library(ggplot2)
library(tidybayes)
library(brms)
options(brms.backend = "cmdstanr")
options(mc.cores = parallel::detectCores())
library(posterior)
library(bayesplot)
library(ggdist)
library(purrr)
library(here)
theme_set(theme_tidybayes())
bayesplot::bayesplot_theme_update(text = element_text(family = "sans"))
# source the helper functions
source(here("R", "helpers.R"))
```

Read in response data: 

```{r}
#| code-fold: true

raw_df <- read_csv(here::here("data", "part2", "exp2-pilot.csv")) %>%
  as_tibble(.) %>%
  filter(status != "rejected")

pilot_df <- raw_df %>% filter(responseId != "windowEvents")

# participant-relevant df
# i.e., the part that is relevant for pixelsPerMM and distToScreen
participant_df <- pilot_df %>%
  filter(str_detect(trialId, "chinrest")) %>%
  select(participantId, responseId, answer) %>%
  filter(
    str_detect(responseId, "pixelsPerMM") |
      str_detect(responseId, "dist-calibration-MM")
  ) %>%
  mutate(answer = as.numeric(answer)) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  dplyr::rename(distToScreen = `dist-calibration-MM`)

# part 1 relevant df
part1_df <- pilot_df %>%
  filter(str_detect(trialId, "block5_test")) %>%
  select(participantId, trialOrder, responseId, answer) %>%
  mutate(answer = as.numeric(answer)) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  dplyr::rename(
    num.ans.x = `location-x`,
    num.ans.y = `location-y`,
    num.resp.x = `slider-x`,
    num.resp.y = `slider-y`,
    px.resp.x = `px-slider-x`,
    px.resp.y = `px-slider-y`,
    px.ans.x = `px-location-x`,
    px.ans.y = `px-location-y`
  )

# part 2 relevant df
part2_df <- pilot_df %>%
  filter(str_detect(trialId, "task")) %>%
  select(
    participantId,
    trialOrder,
    trialId,
    responseId,
    condition,
    answer,
    parameters_taskType,
    parameters_taskIndex
  ) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  dplyr::rename(
    type = parameters_taskType,
    ogIndex = parameters_taskIndex,
    px.resp.y = pixelResponse,
    num.resp.y = numericResponse
  ) %>%
  mutate(
    px.resp.y = as.numeric(px.resp.y),
    num.resp.y = as.numeric(num.resp.y),
    type = ifelse(type == "pointArc", "point_arc", "point")
  )

# join to make the participant_df also contain condition
participant_df <- participant_df %>%
  left_join(
    part2_df %>%
      select(-condition) %>%
      separate_wider_delim(
        trialId,
        delim = "_",
        names = c("condition", "A", "B")
      ) %>%
      select(participantId, condition) %>%
      unique(.)
  )
```

Read in stimuli data for experiment 2 from Moritz et al.: 

```{r}
#| code-fold: true

exp_json_data <- jsonlite::fromJSON(here::here(
  "data",
  "moritz",
  "stimuli",
  "exp2-stimuli.json"
)) %>%
  as_tibble(.)

stimuli_df <- exp_json_data %>%
  select(index, seed, noise, flip, type, data) %>%
  mutate(noise = as.character(noise))

# get the mapping between og index and the index used in exp
stimuli_df %>%
  filter(type == "point") %>%
  mutate(expIndex = 1:nrow(.)) %>%
  select(index, type, expIndex) -> point_map
stimuli_df %>%
  filter(type == "point_arc") %>%
  mutate(expIndex = 1:nrow(.)) %>%
  select(index, type, expIndex) -> pointArc_map

# get rid of the type that we don't use
stimuli_df <- stimuli_df %>% filter(type != "line")

# get the answer
part2.ans_df <- stimuli_df %>%
  unnest(col = data) %>%
  dplyr::rename(num.x = x, num.y = y, ogIndex = index) %>%
  dplyr::summarise(
    num.mean.y = mean(num.y),
    num.median.y = median(num.y),
    .by = c(ogIndex, type, flip, noise, seed)
  ) %>%
  mutate(
    # this is what Moritz did, and what we're not doing
    # num.ans.y = num.mean.y[type == "point"][[1]],
    num.ans.y = num.mean.y,
    .by = c(flip, noise, seed)
  )
```

Read in helper functions for transforming between spaces: 

```{r}
#| code-fold: true
#| label: helper-functions

# First Part -- Block 5
x_block5 <- make_va_scale(
  data_min = -5,
  data_max = 5,
  px_min = 50,
  px_max = 585
)
y_block5 <- make_va_scale(
  data_min = 0,
  data_max = 10,
  px_min = 410,
  px_max = 15
)

# Second part -- Mortiz et al. related
x_moritz <- make_va_scale(
  data_min = 0,
  data_max = 120,
  px_min = 0,
  px_max = 500
)
y_moritz <- make_va_scale(data_min = 0, data_max = 1, px_min = 200, px_max = 0)
```

:::{.callout-note collapse="true"}

# See whether data is sampling correctly 

```{r}
# get a sense of the distribution of part1_df
part1_df %>%
  ggplot(aes(y = num.resp.y)) +
  stat_halfeye() +
  coord_cartesian(ylim = c(0, 10))
part1_df %>%
  ggplot(aes(x = num.resp.x)) +
  stat_halfeye() +
  coord_cartesian(xlim = c(-5, 5))
part1_df %>% ggplot(aes(x = num.resp.x, y = num.resp.y)) + geom_bin2d()
```

The above tells us that data is being sampled correctly. 

:::

# Fit model for part 1 

## Get the data ready

... by adding the visual angle transformed data. 

```{r}
#| code-fold: true

part1_df <- part1_df %>%
  left_join(participant_df) %>%
  mutate(
    va.ans.x = x_block5$data_to_va(num.ans.x, distToScreen, pixelsPerMM),
    va.ans.y = y_block5$data_to_va(num.ans.y, distToScreen, pixelsPerMM),
    va.resp.x = x_block5$data_to_va(num.resp.x, distToScreen, pixelsPerMM),
    va.resp.y = y_block5$data_to_va(num.resp.y, distToScreen, pixelsPerMM)
  ) %>%
  mutate(va.error.x = va.resp.x - va.ans.x, va.error.y = va.resp.y - va.ans.y)
```

Get a sense of error distribution 

```{r}
#| layout-ncol: 2
#| code-fold: true

part1_df %>%
  ggplot(aes(y = va.error.x, x = va.ans.y)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  labs(title = "Project to X-Axis Error (Vis Angle)")
part1_df %>%
  ggplot(aes(y = va.error.y, x = va.ans.x)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  labs(title = "Project to Y-Axis Error (Vis Angle)")
```

Note that for matching with part 2, we are *only interested in the second graph* for model fitting purposes. 

## Model Specification 

Recall the we defined the following model for the "project from dot to axis" operator: 

$$
\begin{align*}
\texttt{error}_i &\sim \mathcal{N}(\beta_0 + u_{\text{PID[i]}}, \sigma_{i, \text{PID}[i]}) \\ 
\log(\sigma_{i, j}) &= \alpha_j + \log(\texttt{va.ans.x}_i) \\ 
\text{or:} \log(\sigma_{i, j}) &\sim \mathcal{N}(\mu_\alpha + \log(\texttt{va.ans.x}_i), \sigma_\alpha)\\ 
\alpha_j &\sim N(\mu_\alpha, \sigma_\alpha) \\ 
\mu_\alpha &\sim N(0, 1) \\ 
\sigma_\alpha &\sim \mathcal{N}^+(0, 1)
\end{align*}
$$

Here is the specification we used for fitting: 

```{r}
#| label: brms specs

# Recall that we defined error = resp - ans
f <- bf(
  va.error.y ~ (1 | participantId),
  sigma ~ (1 | participantId) + offset(log(va.ans.x))
)

p <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 1), class = sd, group = participantId, lb = 0),
  prior(normal(0, 1), class = Intercept, dpar = "sigma"),
  prior(normal(0, 1), class = sd, group = participantId, dpar = sigma, lb = 0)
)
```

Fit the model: 

```{r}
#| label: fit model for part 1
#| code-fold: true
#| code-line-numbers: true

m1 <- part1_df %>%
  brm(
    formula = f,
    data = .,
    family = gaussian(),
    prior = p,
    chains = 4,
    cores = 4,
    control = list(adapt_delta = 0.99),
    iter = 6000,
    file = here::here("analysis", "fitted_models", "part2", "m1"),
    file_refit = "on_change",
    save_pars = save_pars(all = TRUE)
  )
```

Get a sense of posterior fit: 

```{r}
summary(m1)
```

Note that the Bulk_ESS and Tail_ESS all indicate good fit. 

```{r}
plot(m1)
```

The four chains mix well. Now we move on to perform posterior predictive checks: 

```{r}
#| layout-ncol: 2
#| code-fold: true
pp_check(m1, ndraws = 100) +
  theme(legend.position = c(0.9, 0.5)) +
  theme(text = element_text(size = 12)) -> p1
p1

pp_check(m1, ndraws = 100, type = "pit_ecdf") +
  theme(text = element_text(size = 12)) -> p2
p2

pp_check(m1, ndraws = 100, type = "pit_ecdf", plot_diff = TRUE) +
  theme(text = element_text(size = 10)) -> p3
p3
```

```{r}
#| eval: false
#| echo: false

ggsave(
  here::here("figs", "part2-task1-pp-check.pdf"),
  p1,
  width = 12,
  height = 6,
  units = "cm",
  dpi = 300
)
ggsave(
  here::here("figs", "part2-task1-pit-ecdf.pdf"),
  p2,
  height = 6,
  width = 6,
  units = "cm",
  dpi = 300
)
```

Plot existing data on top of prediction, for an average participant: 

```{r}
#| code-fold: true
#| label: part 1 posterior fit check

prediction_grid <- part1_df %>%
  modelr::data_grid(
    va.ans.x = modelr::seq_range(va.ans.x, n = 100),
    participantId = unique(participantId)
  ) %>%
  add_predicted_draws(m1, ndraws = 100)

ggplot(prediction_grid, aes(x = va.ans.x, y = .prediction)) +
  stat_lineribbon(
    .width = c(.95, .80, .50),
    # alpha = 0.5,
    linewidth = 0.5
  ) +
  scale_fill_brewer(palette = "Blues") +
  geom_point(
    data = part1_df,
    aes(y = va.error.y, x = va.ans.x),
    alpha = 0.5,
    size = 0.5
  ) +
  labs(
    title = "Part 1: Project to Y-axis",
    y = "Error",
    x = "Distance to Y-Axis"
  ) +
  # theme_minimal(base_size = 8) +
  # coord_cartesian(ylim = c(-0.25, 0.4)) +
  theme(legend.position = "none") -> p1

p1
```

```{r}
#| eval: false
#| echo: false

ggsave(
  here::here("figs", "part2-task1-project-y.pdf"),
  p1,
  height = 8,
  width = 8,
  units = "cm",
  dpi = 300
)
```

## Prior sensitivity analysis

:::{.callout-caution}
The following is quite time-consuming. Only run this if your computer has good power. 
:::

To test whether our model is sensitive to prior specifications, we fit them on differnt priors: 

```{r}
# Wider — less informative
p_wide <- c(
  prior(normal(0, 5), class = Intercept),
  prior(normal(0, 5), class = sd, group = participantId, lb = 0),
  prior(normal(0, 5), class = Intercept, dpar = "sigma"),
  prior(normal(0, 5), class = sd, group = participantId, dpar = sigma, lb = 0)
)

# Tighter — more informative
p_tight <- c(
  prior(normal(0, 0.1), class = Intercept),
  prior(normal(0, 0.1), class = sd, group = participantId, lb = 0),
  prior(normal(0, 0.1), class = Intercept, dpar = "sigma"),
  prior(normal(0, 0.1), class = sd, group = participantId, dpar = sigma, lb = 0)
)

# Different family — heavier tails
p_student <- c(
  prior(student_t(3, 0, 1), class = Intercept),
  prior(student_t(3, 0, 1), class = sd, group = participantId, lb = 0),
  prior(student_t(3, 0, 1), class = Intercept, dpar = "sigma"),
  prior(
    student_t(3, 0, 1),
    class = sd,
    group = participantId,
    dpar = sigma,
    lb = 0
  )
)
```

```{r}
m1_wide <- part1_df %>%
  brm(
    formula = f,
    data = .,
    family = gaussian(),
    prior = p_wide,
    chains = 4,
    cores = 4,
    control = list(adapt_delta = 0.99),
    iter = 6000,
    file = here::here("analysis", "fitted_models", "part2", "m1_wide"),
    file_refit = "on_change",
    save_pars = save_pars(all = TRUE)
  )

m1_tight <- part1_df %>%
  brm(
    formula = f,
    data = .,
    family = gaussian(),
    prior = p_tight,
    chains = 4,
    cores = 4,
    control = list(adapt_delta = 0.99),
    iter = 6000,
    file = here::here("analysis", "fitted_models", "part2", "m1_tight"),
    file_refit = "on_change",
    save_pars = save_pars(all = TRUE)
  )

m1_student <- part1_df %>%
  brm(
    formula = f,
    data = .,
    family = gaussian(),
    prior = p_student,
    chains = 4,
    cores = 4,
    control = list(adapt_delta = 0.99),
    iter = 6000,
    file = here::here("analysis", "fitted_models", "part2", "m1_student"),
    file_refit = "on_change",
    save_pars = save_pars(all = TRUE)
  )
```

All models fit without convergence issues. Now we combine posteriors into a list and plot them: 

```{r}
#| code-fold: true

# Combine posteriors from all models
models <- list(
  "default (0,1)" = m1,
  "wide (0,5)" = m1_wide,
  "tight (0,0.1)" = m1_tight,
  "student_t(3,0,1)" = m1_student
)

combined <- models %>%
  imap_dfr(
    ~ gather_draws(.x, b_Intercept, b_sigma_Intercept) %>%
      mutate(prior = .y)
  )
```

Interval plot: 

```{r}
#| code-fold: true
combined %>%
  # uncomment to see what variance instead of log(variance) is.
  # mutate(
  #   .value = if_else(.variable == "b_sigma_Intercept", exp(.value), .value)
  # ) %>%
  ggplot(aes(x = .value, y = prior)) +
  stat_pointinterval(.width = c(.50, .95)) +
  facet_wrap(~.variable, scales = "free_x") +
  labs(x = "Estimate", y = "Prior specification")
```

Densities overlaid: 

```{r}
#| code-fold: true
combined %>%
  ggplot(aes(x = .value, color = prior, fill = prior)) +
  stat_slab(alpha = 0.3) +
  facet_wrap(~.variable, scales = "free") +
  labs(x = "Estimate", y = NULL)
```

To see the actual table: 

```{r}
#| code-fold: true
combined %>%
  group_by(prior, .variable) %>%
  median_qi(.value, .width = .95) %>%
  select(prior, .variable, .value, .lower, .upper) %>%
  arrange(.variable, prior)
```

We can also check random effects: 

```{r}
#| code-fold: true
combined_sd <- models %>%
  imap_dfr(
    ~ gather_draws(
      .x,
      sd_participantId__Intercept,
      sd_participantId__sigma_Intercept
    ) %>%
      mutate(prior = .y)
  )

combined_sd %>%
  ggplot(aes(x = .value, y = prior)) +
  stat_pointinterval(.width = c(.50, .95)) +
  facet_wrap(~.variable, scales = "free_x") +
  labs(x = "Estimate", y = "Prior specification")
```

**Takeaway**: posteriors were robust to prior specification across default, wide, and heavy-tailed priors. The tight prior $\sigma = 0.1$ constrained the residual variance parameter, confirming it is unreasonably informative for this parameter; the remaining three specifications produced essentially identical posteriors. 

# Infer model for part 2 

**Step 1:** Extract participant-level draws (20 participants $\times$ 3000 per chain $\times$ 4 chains)

```{r}
m1_draws_df <- m1 %>%
  spread_draws(
    b_Intercept,
    r_participantId[participantId, ],
    b_sigma_Intercept,
    r_participantId__sigma[participantId, ]
  ) %>%
  mutate(
    mu = b_Intercept + r_participantId,
    var_scale = exp(b_sigma_Intercept + r_participantId__sigma)
  ) %>%
  select(.draw, participantId, mu, var_scale)
```

**Step 2**: Then, we get a cross between participant and the stimuli they saw, in order to generate **predictions**: 

```{r}
predict_df <- participant_df %>%
  rename(type = condition) %>%
  mutate(type = ifelse(type == "pointArc", "point_arc", "point")) %>%
  inner_join(
    stimuli_df %>% filter(type != "line") %>% rename(ogIndex = index),
    by = "type",
    relationship = "many-to-many"
  ) %>%
  unnest(col = data) %>%
  rename(num.ans.x = x, num.ans.y = y) %>%
  mutate(
    va.60 = x_moritz$data_to_va(60, distToScreen, pixelsPerMM),
    va.ans.x = x_moritz$data_to_va(num.ans.x, distToScreen, pixelsPerMM),
    va.ans.y = y_moritz$data_to_va(num.ans.y, distToScreen, pixelsPerMM),
  )

predict_df %>% head(5)
```

## Project Once 

**Step 3: Draw the posterior prediction for project once to the y-axis:**

Note that the following is going to take a while to run. On a mac mini it takes about 5 mins to fit. 

```{r}
#| eval: false
posterior_predict_df <- map_dfr(
  .progress = TRUE,
  unique(predict_df$participantId),
  function(pid) {
    # 12,000 draws for this participant
    d <- m1_draws_df %>%
      filter(participantId == pid) %>%
      arrange(.draw)

    # 2,880 observations (48 trials × 60 dots)
    obs <- predict_df %>%
      filter(participantId == pid) %>%
      select(ogIndex, va.60, va.ans.x, va.ans.y, pixelsPerMM, distToScreen)

    cross_join(d, obs) %>%
      mutate(
        sigma = var_scale * va.ans.x, # dist to y-axis is va.ans.x
        .prediction = rnorm(n(), mean = mu, sd = sigma),
        pred_y = va.ans.y + .prediction,
        w = 1 / (mu^2 + sigma^2) # inverse-mse
      ) %>%
      group_by(
        .draw,
        mu,
        var_scale,
        pixelsPerMM,
        distToScreen,
        va.60,
        ogIndex
      ) %>%
      summarise(
        va.1proj.mean = mean(pred_y),
        va.1proj.median = median(pred_y),
        va.1proj.weighted = sum(w * pred_y) / sum(w),
        .groups = "drop_last"
      ) %>%
      mutate(
        participantId = pid
      )
  }
)
```

```{r}
#| eval: false
posterior_predict_df <- posterior_predict_df %>%
  mutate(
    num.1proj.mean = y_moritz$va_to_data(
      va.1proj.mean,
      distToScreen,
      pixelsPerMM
    ),
    num.1proj.median = y_moritz$va_to_data(
      va.1proj.median,
      distToScreen,
      pixelsPerMM
    ),
    num.1proj.weighted = y_moritz$va_to_data(
      va.1proj.weighted,
      distToScreen,
      pixelsPerMM
    )
  )
```

```{r}
#| eval: false
#| echo: false
saveRDS(
  posterior_predict_df,
  file = here::here("analysis/fitted_models/part2/part2_posterior_1proj.rds")
)
```

```{r}
#| echo: false
#| label: short cut to read in data
posterior_predict_df <- readRDS(
  file = here::here("analysis/fitted_models/part2/part2_posterior_1proj.rds")
)
```

### Posterior checks with `bayesplot` 

First we get the data in the right format to plot with `bayesplot`: 

```{r}
#| code-fold: true

# get the relevant ones for each strategy
response_ordered <- part2_df %>% arrange(participantId, ogIndex)
y <- response_ordered$num.resp.y

# group can be any condition-level variable, or a combination
group <- response_ordered %>%
  left_join(part2.ans_df %>% select(ogIndex, type, flip, noise)) %>%
  # we don't care about type, only care about noise and flip
  mutate(
    noise_label = ifelse(noise == "c(0, 0)", "Var 0", "Var 0.4"),
    flip_label = ifelse(flip == FALSE, "Var Lower", "Var Upper")
  ) %>%
  mutate(g = paste(noise_label, flip_label, sep = " | ")) %>%
  pull(g)

yrep_mean <- posterior_predict_df %>%
  ungroup() %>%
  select(participantId, ogIndex, .draw, num.1proj.mean) %>%
  arrange(participantId, ogIndex, .draw) %>%
  unite("obs_id", participantId, ogIndex) %>%
  pivot_wider(
    id_cols = .draw,
    names_from = obs_id,
    values_from = num.1proj.mean
  ) %>%
  select(-.draw) %>%
  as.matrix()

yrep_median <- posterior_predict_df %>%
  ungroup() %>%
  select(participantId, ogIndex, .draw, num.1proj.median) %>%
  arrange(participantId, ogIndex, .draw) %>%
  unite("obs_id", participantId, ogIndex) %>%
  pivot_wider(
    id_cols = .draw,
    names_from = obs_id,
    values_from = num.1proj.median
  ) %>%
  select(-.draw) %>%
  as.matrix()

yrep_weighted <- posterior_predict_df %>%
  ungroup() %>%
  select(participantId, ogIndex, .draw, num.1proj.weighted) %>%
  arrange(participantId, ogIndex, .draw) %>%
  unite("obs_id", participantId, ogIndex) %>%
  pivot_wider(
    id_cols = .draw,
    names_from = obs_id,
    values_from = num.1proj.weighted
  ) %>%
  select(-.draw) %>%
  as.matrix()
```

**KDE-Overlay**: 

```{r}
#| layout-ncol: 3
#| code-fold: true

ppc_dens_overlay(y, yrep_mean[sample(nrow(yrep_mean), 100), ]) +
  labs(title = "Mean Strategy") +
  coord_cartesian(xlim = c(0, 1))
ppc_dens_overlay(y, yrep_median[sample(nrow(yrep_median), 100), ]) +
  labs(title = "Median Strategy") +
  coord_cartesian(xlim = c(0, 1))
ppc_dens_overlay(y, yrep_weighted[sample(nrow(yrep_weighted), 100), ]) +
  labs(title = "Weighted Strategy") +
  coord_cartesian(xlim = c(0, 1))
```

**KDE-Overlay faceted by factor**: 

```{r}
#| layout-ncol: 3
#| code-fold: true

p1 <- ppc_dens_overlay_grouped(
  y,
  yrep_mean[sample(nrow(yrep_mean), 100), ],
  group
) +
  labs(title = "Mean Strategy") +
  coord_cartesian(xlim = c(0, 1))
p2 <- ppc_dens_overlay_grouped(
  y,
  yrep_median[sample(nrow(yrep_median), 100), ],
  group
) +
  labs(title = "Median Strategy") +
  coord_cartesian(xlim = c(0, 1))
p3 <- ppc_dens_overlay_grouped(
  y,
  yrep_weighted[sample(nrow(yrep_weighted), 100), ],
  group
) +
  labs(title = "Weighted Strategy") +
  coord_cartesian(xlim = c(0, 1))
p1
p2
p3
```

```{r}
#| echo: false
#| eval: false

ggsave(
  here::here("figs/1proj_mean.pdf"),
  p1 + scale_y_continuous(breaks = 0) + theme_tidybayes(base_size = 8),
  width = 3.5,
  height = 3.5 / 4 * 3,
  unit = "in",
  dpi = 300
)
```

---

**PIT-ECDFs**: 

```{r}
#| layout-ncol: 3
#| code-fold: true
ppc_pit_ecdf(y, yrep_mean) + labs(title = "Mean Strategy")
ppc_pit_ecdf(y, yrep_median) + labs(title = "Median Strategy")
ppc_pit_ecdf(y, yrep_weighted) + labs(title = "Weighted Strategy")
```

Faceted version: 

```{r}
#| layout-ncol: 3
#| code-fold: true
ppc_pit_ecdf_grouped(y, yrep_mean, group) + labs(title = "Mean Strategy")
ppc_pit_ecdf_grouped(y, yrep_median, group) + labs(title = "Median Strategy")
ppc_pit_ecdf_grouped(y, yrep_weighted, group) +
  labs(title = "Weighted Strategy")
```

This demonstrates that the fit does not capture outliers. 

## Project Twice 

Same step 1 and step 2 as before, so we don't generate anything new, but start from step 3 with generating `posterior_predict_df`: 

**Step 3: Draw the posterior prediction for project once to the middle AND project twice to the y-axis:**

Again, this is going to take a long time to run. 

```{r}
#| code-fold: true
#| eval: false

posterior_predict_df <- map_dfr(
  .progress = TRUE,
  unique(predict_df$participantId),
  function(pid) {
    # 12,000 draws for this participant
    d <- m1_draws_df %>%
      filter(participantId == pid) %>%
      arrange(.draw)

    # 2,880 observations (48 trials × 60 dots)
    obs <- predict_df %>%
      filter(participantId == pid) %>%
      select(ogIndex, va.60, va.ans.x, va.ans.y, pixelsPerMM, distToScreen)

    cross_join(d, obs) %>%
      mutate(
        sigma = var_scale * abs(va.60 - va.ans.x), # dist to center
        .prediction = rnorm(n(), mean = mu, sd = sigma),
        pred_y = va.ans.y + .prediction,
        w = 1 / (mu^2 + sigma^2) # inverse-mse
      ) %>%
      group_by(
        .draw,
        mu,
        var_scale,
        pixelsPerMM,
        distToScreen,
        va.60,
        ogIndex
      ) %>%
      summarise(
        va.1proj.mean = mean(pred_y),
        va.1proj.median = median(pred_y),
        va.1proj.weighted = sum(w * pred_y) / sum(w),
        .groups = "drop_last"
      ) %>%
      mutate(
        proj_noise = rnorm(n(), mean = mu, sd = var_scale * va.60),
        va.2proj.mean = va.1proj.mean + proj_noise,
        va.2proj.median = va.1proj.median + proj_noise,
        va.2proj.weighted = va.1proj.weighted + proj_noise,
        participantId = pid
      )
  }
)
```

We then transform things from visual angle prediction to actual numerical value, for comparison against the actual response we got: 

```{r}
#| code-fold: true
#| eval: false

posterior_predict_df <- posterior_predict_df %>%
  mutate(
    num.2proj.mean = y_moritz$va_to_data(
      va.2proj.mean,
      distToScreen,
      pixelsPerMM
    ),
    num.2proj.median = y_moritz$va_to_data(
      va.2proj.median,
      distToScreen,
      pixelsPerMM
    ),
    num.2proj.weighted = y_moritz$va_to_data(
      va.2proj.weighted,
      distToScreen,
      pixelsPerMM
    ),
  )
```

```{r}
#| eval: false
#| echo: false
saveRDS(
  posterior_predict_df,
  file = here::here("analysis/fitted_models/part2/part2_posterior_2proj.rds")
)
```

```{r}
#| echo: false
posterior_predict_df <- readRDS(here::here(
  "analysis/fitted_models/part2/part2_posterior_2proj.rds"
))
```

### Posterior Checks with `bayesplot` 

First we get the data format ready to be plotted using `bayesplot`: 

```{r}
#| code-fold: true
response_ordered <- part2_df %>% arrange(participantId, ogIndex)
y <- response_ordered$num.resp.y

# group can be any condition-level variable, or a combination
group <- response_ordered %>%
  left_join(
    part2.ans_df %>%
      select(ogIndex, type, flip, noise)
  ) %>%
  mutate(
    noise_label = ifelse(noise == "c(0, 0)", "Var 0", "Var 0.4"),
    flip_label = ifelse(flip == FALSE, "Var Lower", "Var Upper")
  ) %>%
  mutate(g = paste(noise_label, flip_label, sep = " | ")) %>%
  pull(g)

yrep_mean <- posterior_predict_df %>%
  ungroup() %>%
  select(participantId, ogIndex, .draw, num.2proj.mean) %>%
  arrange(participantId, ogIndex, .draw) %>%
  unite("obs_id", participantId, ogIndex) %>%
  pivot_wider(
    id_cols = .draw,
    names_from = obs_id,
    values_from = num.2proj.mean
  ) %>%
  select(-.draw) %>%
  as.matrix()

yrep_median <- posterior_predict_df %>%
  ungroup() %>%
  select(participantId, ogIndex, .draw, num.2proj.median) %>%
  arrange(participantId, ogIndex, .draw) %>%
  unite("obs_id", participantId, ogIndex) %>%
  pivot_wider(
    id_cols = .draw,
    names_from = obs_id,
    values_from = num.2proj.median
  ) %>%
  select(-.draw) %>%
  as.matrix()

yrep_weighted <- posterior_predict_df %>%
  ungroup() %>%
  select(participantId, ogIndex, .draw, num.2proj.weighted) %>%
  arrange(participantId, ogIndex, .draw) %>%
  unite("obs_id", participantId, ogIndex) %>%
  pivot_wider(
    id_cols = .draw,
    names_from = obs_id,
    values_from = num.2proj.weighted
  ) %>%
  select(-.draw) %>%
  as.matrix()
```

**KDE-plots**: 

```{r}
#| layout-ncol: 3
#| code-fold: true
ppc_dens_overlay(y, yrep_mean[sample(nrow(yrep_mean), 100), ]) +
  labs(title = "Mean Strategy")
ppc_dens_overlay(y, yrep_median[sample(nrow(yrep_median), 100), ]) +
  labs(title = "Median Strategy")
ppc_dens_overlay(y, yrep_weighted[sample(nrow(yrep_weighted), 100), ]) +
  labs(title = "Weighted Mean Strategy")
```

**KDE faceted**: 

```{r}
#| code-fold: true
#| layout-ncol: 3
p1 <- ppc_dens_overlay_grouped(
  y,
  yrep_mean[sample(nrow(yrep_mean), 100), ],
  group = group
) +
  labs(title = "Mean Strategy") +
  coord_cartesian(xlim = c(0, 1))
p2 <- ppc_dens_overlay_grouped(
  y,
  yrep_median[sample(nrow(yrep_median), 100), ],
  group = group
) +
  labs(title = "Median Strategy")
p3 <- ppc_dens_overlay_grouped(
  y,
  yrep_weighted[sample(nrow(yrep_weighted), 100), ],
  group = group
) +
  labs(title = "Weighted Mean Strategy")
p1
p2
p3
```

```{r}
#| echo: false
#| eval: false

ggsave(
  here::here("figs/2proj_weighted.pdf"),
  p3 + scale_y_continuous(breaks = 0) + theme_tidybayes(base_size = 8),
  width = 3.5,
  height = 3.5 / 4 * 3,
  unit = "in",
  dpi = 300
)
```

---

**PIT-ECDF**: 

```{r}
#| code-fold: true
#| layout-ncol: 3
ppc_pit_ecdf(y, yrep_mean) + labs(title = "Mean Strategy")
ppc_pit_ecdf(y, yrep_median) + labs(title = "Median Strategy")
ppc_pit_ecdf(y, yrep_weighted) + labs(title = "Weighted Mean Strategy")
# ppc_pit_ecdf(y, yrep_mean, plot_diff = TRUE) +
#   labs(title = "Mean Strategy")
# ppc_pit_ecdf(y, yrep_median, plot_diff = TRUE) + labs(title = "Median Strategy")
# ppc_pit_ecdf(y, yrep_weighted, plot_diff = TRUE) +
#   labs(title = "Weighted Mean Strategy")
```

**PIT-ECDF Faceted**: 

```{r}
#| layout-ncol: 3
#| code-fold: true
ppc_pit_ecdf_grouped(y, yrep_mean, group) + labs(title = "Mean Strategy")
ppc_pit_ecdf_grouped(y, yrep_median, group) + labs(title = "Median Strategy")
ppc_pit_ecdf_grouped(y, yrep_weighted, group) +
  labs(title = "Weighted Mean Strategy")
```

### Posterior Checks with `ggdist` 

For better customization and easier to manipulate dataframes, we will use `stat_slab` in `ggdist` instead of `bayesplot::ppc_dens_overlay`. 

We extract the shared layers across all three strategies first: 

```{r}
#| code-fold: true
# get the plotting relevant data ready

part2_joined <- part2_df %>%
  left_join(
    part2.ans_df %>% select(ogIndex, type, flip, noise),
    by = join_by(type, ogIndex)
  )

# the density of the response that participants did
participant_resp <- stat_slab(
  data = part2_joined,
  aes(x = num.resp.y, group = NULL, thickness = after_stat(pdf)),
  fill = NA,
  color = "darkblue",
  linewidth = 1,
  density = density_bounded(bounds = c(0, 1))
)

# the shared "annotation-like" details
shared_layers <- list(
  geom_vline(
    data = part2.ans_df %>%
      group_by(flip, noise) %>%
      summarise(mean.y = mean(num.mean.y)),
    aes(xintercept = mean.y),
    linetype = "dashed",
    color = "darkgray"
  ),
  facet_grid(
    noise ~ flip,
    labeller = labeller(
      noise = c("c(0, 0)" = "Variability 0", "c(0, 0.4)" = "Variability .4"),
      flip = c("TRUE" = "Variability Upper", "FALSE" = "Variability Lower")
    )
  ),
  coord_cartesian(xlim = c(0, 1), ylim = c(-0.05, 1)),
  scale_x_continuous(breaks = c(0, 0.5, 1)),
  scale_y_continuous(breaks = 0)
)
```

```{r}
#| code-fold: true
draw_subset <- sample(unique(posterior_predict_df$.draw), 100)

# bias + var for facets
biasVar <- posterior_predict_df %>%
  ungroup(.) %>%
  select(.draw, ogIndex, num.2proj.mean) %>%
  left_join(
    part2.ans_df %>%
      select(ogIndex, type, flip, noise)
  ) %>%
  group_by(flip, noise) %>%
  mean_qi(num.2proj.mean, .width = 0.68)


p1 <- posterior_predict_df %>%
  ungroup(.) %>%
  filter(.draw %in% draw_subset) %>%
  select(.draw, ogIndex, num.2proj.mean) %>%
  left_join(
    part2.ans_df %>%
      select(ogIndex, type, flip, noise)
  ) %>%
  ggplot(aes(x = num.2proj.mean, group = .draw, thickness = after_stat(pdf))) +
  stat_slab(
    fill = NA,
    color = "#7CCBA2",
    alpha = 0.7,
    size = 0.25,
    density = density_bounded(bounds = c(0, 1))
  ) +
  participant_resp +
  scale_thickness_shared() +
  # bias and variannce from response
  stat_pointinterval(
    data = part2_joined,
    aes(x = num.resp.y, group = NULL),
    color = "darkblue",
    point_interval = mean_qi,
    .width = .68, # about +/- 1SD
    inherit.aes = FALSE,
    position = position_nudge(y = 0.05),
    size = 1
  ) +
  geom_pointinterval(
    data = biasVar,
    aes(x = num.2proj.mean, xmin = .lower, xmax = .upper),
    color = "#7CCBA2",
    position = position_nudge(y = -0.05),
    inherit.aes = FALSE,
    size = 1
  ) +
  shared_layers +
  labs(y = "", x = "numerical value", title = "Mean Strategy") +
  theme_light(base_size = 8)

p1
```

Median stategy with ggdist: 

```{r}
#| code-fold: true
draw_subset <- sample(unique(posterior_predict_df$.draw), 100)

# bias + var for facets
biasVar <- posterior_predict_df %>%
  ungroup(.) %>%
  select(.draw, ogIndex, num.2proj.median) %>%
  left_join(
    part2.ans_df %>%
      select(ogIndex, type, flip, noise)
  ) %>%
  group_by(flip, noise) %>%
  mean_qi(num.2proj.median, .width = 0.68)

p2 <- posterior_predict_df %>%
  ungroup(.) %>%
  filter(.draw %in% draw_subset) %>%
  select(.draw, ogIndex, num.2proj.median) %>%
  left_join(
    part2.ans_df %>%
      select(ogIndex, type, flip, noise)
  ) %>%
  ggplot(aes(
    x = num.2proj.median,
    group = .draw,
    thickness = after_stat(pdf)
  )) +
  stat_slab(
    fill = NA,
    color = "#7CCBA2",
    alpha = 0.7,
    size = 0.25,
    density = density_bounded(bounds = c(0, 1))
  ) +
  participant_resp +
  scale_thickness_shared() +
  stat_pointinterval(
    data = part2_joined,
    aes(x = num.resp.y, group = NULL),
    color = "darkblue",
    point_interval = mean_qi,
    .width = .68, # about +/- 1SD
    inherit.aes = FALSE,
    position = position_nudge(y = 0.05),
    size = 1
  ) +
  geom_pointinterval(
    data = biasVar,
    aes(x = num.2proj.median, xmin = .lower, xmax = .upper),
    color = "#7CCBA2",
    position = position_nudge(y = -0.05),
    inherit.aes = FALSE,
    size = 1
  ) +
  shared_layers +
  labs(y = "", x = "numerical value", title = "Median Strategy") +
  theme_light(base_size = 8)

p2
```

Weighted mean strategy: 

```{r}
#| code-fold: true
draw_subset <- sample(unique(posterior_predict_df$.draw), 100)

# bias + var for facets
biasVar <- posterior_predict_df %>%
  ungroup(.) %>%
  select(.draw, ogIndex, num.2proj.weighted) %>%
  left_join(
    part2.ans_df %>%
      select(ogIndex, type, flip, noise)
  ) %>%
  group_by(flip, noise) %>%
  mean_qi(num.2proj.weighted, .width = 0.68)

p3 <- posterior_predict_df %>%
  ungroup(.) %>%
  filter(.draw %in% draw_subset) %>%
  select(.draw, ogIndex, num.2proj.weighted) %>%
  left_join(
    part2.ans_df %>%
      select(ogIndex, type, flip, noise)
  ) %>%
  ggplot(aes(
    x = num.2proj.weighted,
    group = .draw,
    thickness = after_stat(pdf)
  )) +
  stat_slab(
    fill = NA,
    color = "#7CCBA2",
    alpha = 0.7,
    size = 0.25,
    # density = density_bounded(bounds = c(0, 1))
    # Note: the density goes beyond [0, 1] which is why we can't used density_bounded
    density = "unbounded"
  ) +
  participant_resp +
  stat_pointinterval(
    data = part2_joined,
    aes(x = num.resp.y, group = NULL),
    color = "darkblue",
    point_interval = mean_qi,
    .width = .68, # about +/- 1SD
    inherit.aes = FALSE,
    position = position_nudge(y = 0.05),
    size = 1
  ) +
  geom_pointinterval(
    data = biasVar,
    aes(x = num.2proj.weighted, xmin = .lower, xmax = .upper),
    color = "#7CCBA2",
    position = position_nudge(y = -0.05),
    inherit.aes = FALSE,
    size = 1
  ) +
  scale_thickness_shared() +
  shared_layers +
  labs(y = "", x = "numerical value", title = "Weighted Mean Strategy") +
  theme_light(base_size = 8)

p3
```

```{r}
#| echo: false
#| eval: false
ggsave(
  here::here(
    "figs",
    "part2_task2_mean.svg"
  ),
  p1,
  width = 3,
  height = 2.5,
  units = c("in"),
  dpi = 300
)

ggsave(
  here::here(
    "figs",
    "part2_task2_median.svg"
  ),
  p2,
  width = 3,
  height = 2.5,
  units = c("in"),
  dpi = 300
)

ggsave(
  here::here(
    "figs",
    "part2_task2_weighted.svg"
  ),
  p3,
  width = 3,
  height = 2.5,
  units = c("in"),
  dpi = 300
)
```
