---
title: "Part 2: Data Cleaning and Descriptive Checks"
author: "Anonymous"
date: today
date-format: "iso"
format:
  html:
    callout-appearance: simple
    reference-location: margin
    embed-resources: true
---

```{r}
#| echo: false
#| output: false
#| label: setup

library(tidyverse)
library(ggplot2)
library(sgt)
library(numDeriv)
library(ggdist)
# setting the theme
theme_set(theme_minimal())
# model
library(brms)
# plot results
library(tidybayes)
library(here)
```

# Set up

Read in response data

```{r}
#| code-fold: true
raw_df <- read_csv(here::here("data", "part2", "exp2-pilot.csv")) %>%
  as_tibble(.) %>%
  filter(status != "rejected")

pilot_df <- raw_df %>% filter(responseId != "windowEvents")
```

We follow the same strategy as the first study, by checking 1) whether participant's screen to distance is above 200 mm; and 2) whether participant's response to the true answer is above 50%. 

# Eye distance to screen 

```{r}
#| code-fold: true

p <- pilot_df %>%
  filter(status == "completed") %>%
  select(participantId, trialId, responseId, answer) %>%
  filter(grepl("virtual", trialId))

ball_positions <- p %>%
  select(participantId, responseId, answer) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  mutate(`ball-positions` = gsub("\\[|\\]", "", `ball-positions`)) %>%
  mutate(split_answer = str_split(`ball-positions`, ",")) %>%
  mutate(numeric_vectors = map(split_answer, ~ as.numeric(.x))) %>%
  unnest(split_answer) %>%
  mutate(split_answer = as.numeric(split_answer)) %>%
  group_by(participantId) %>%
  summarise(
    avg.ball.pos = sum(split_answer) / 5,
    min.ball.pos = min(split_answer),
    max.ball.pos = max(split_answer),
    true_ball_pos = (avg.ball.pos * 5 - min.ball.pos - max.ball.pos) / 3
  )

calibrations <- p %>%
  filter(grepl("pixelsPerMM|dist-calibration-MM", responseId)) %>%
  select(-trialId) %>%
  mutate(answer = as.numeric(answer)) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  rename(pixelToMM = pixelsPerMM, dist_to_screen = `dist-calibration-MM`)

participants <- calibrations %>%
  right_join(ball_positions, by = join_by(participantId)) %>%
  mutate(
    phy = dist_to_screen * tan(13.5 * pi / 180),
    px = phy * pixelToMM,
    square_pos = round(px + avg.ball.pos)
  ) %>%
  mutate(
    ball.square.dist = (square_pos - true_ball_pos) / pixelToMM,
    d = ball.square.dist / tan(13.5 * pi / 180)
  ) %>%
  select(participantId, pixelToMM, d) %>%
  rename(dist_to_screen = d)

# plot distance to screen distribution
participants %>%
  ggplot(aes(x = dist_to_screen)) +
  geom_dotplot() +
  labs(x = "distance to screen (mm)")
```

Everyone has `dist_to_screen` greater than 200 mm here, so we are not filtering any participant out here. 

# Correlation for the `ProjToAxis` task

First we get the part 1 relevant data: 

```{r}
#| code-fold: true
#| label: get part 1 relevant df
part1_df <- pilot_df %>%
  filter(str_detect(trialId, "block5_test")) %>%
  select(participantId, trialOrder, responseId, answer) %>%
  mutate(answer = as.numeric(answer)) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  dplyr::rename(
    num.ans.x = `location-x`,
    num.ans.y = `location-y`,
    num.resp.x = `slider-x`,
    num.resp.y = `slider-y`,
    px.resp.x = `px-slider-x`,
    px.resp.y = `px-slider-y`,
    px.ans.x = `px-location-x`,
    px.ans.y = `px-location-y`
  )
```

Then we check correlation: 

```{r}
part1_df %>%
  group_by(participantId) %>%
  summarise(
    correlation.x = cor(num.resp.x, num.ans.x, method = "spearman"),
    correlation.y = cor(num.resp.y, num.ans.y, method = "spearman"),
  )
```

Everyone has correlation above 0.50, so we keep all the data here and give everyone full bonus. 

# Timing

On average, how long did people take? 

```{r}
#| code-fold: true

pilot_df %>%
  filter(trialOrder >= 3, trialOrder <= 72) %>%
  group_by(participantId, trialOrder, cleanedDuration) %>%
  summarise(.) %>%
  group_by(participantId) %>%
  summarise(sum_cleanedDuration = sum(cleanedDuration)) %>%
  mutate(totalDuration = sum_cleanedDuration / (1000 * 60))
```

So the average across participants is ... 

```{r}
#| code-fold: true
pilot_df %>%
  # no need for introduction
  filter(trialOrder >= 3, trialOrder <= 72) %>%
  group_by(participantId, trialOrder, cleanedDuration) %>%
  summarise(.) %>%
  group_by(participantId) %>%
  summarise(sum_cleanedDuration = sum(cleanedDuration)) %>%
  mutate(totalDuration = sum_cleanedDuration / (1000 * 60)) %>%
  pull(totalDuration) %>%
  mean(.)
```

What about demographics? 

```{r}
#| code-fold: true
pilot_df %>%
  filter(trialId == "post_study") %>%
  select(participantId, responseId, answer) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  select(-`gender-other`, -`feedback-text`) %>%
  ggplot(aes(x = age, y = gender)) +
  geom_dotplot()
```

Gender is balanced: 

```{r}
#| code-fold: true
pilot_df %>%
  filter(trialId == "post_study") %>%
  select(participantId, responseId, answer) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  select(-`gender-other`, -`feedback-text`) %>%
  count(gender)
```

Mean age is 38. 

```{r}
#| code-fold: true
pilot_df %>%
  filter(trialId == "post_study") %>%
  select(participantId, responseId, answer) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  select(-`gender-other`, -`feedback-text`) %>%
  mutate(age = as.numeric(age)) %>%
  pull(age) %>%
  mean(.)
```

4 our of 20 people provided textual feedback; similar to the ones that Moritz et al. obtained. 