---
title: "Part 2"
author: "Anonymous"
date: today
date-format: "iso"
format:
  html:
    reference-location: margin
---

```{r}
#| code-fold: true

library(tidyverse)
library(jsonlite)
library(ggplot2)
library(tidybayes)
library(brms)
library(posterior)
library(bayesplot)
library(purrr)
theme_set(theme_tidybayes())
bayesplot::bayesplot_theme_update(text = element_text(family = "sans"))
```

Read in response data: 

```{r}
raw_df <- read_csv(here::here("data", "part_2", "exp2-pilot.csv")) %>%
  as_tibble(.) %>%
  filter(status != "rejected")

pilot_df <- raw_df %>% filter(responseId != "windowEvents")

# participant-relevant df
# i.e., the part that is relevant for pixelsPerMM and distToScreen
participant_df <- pilot_df %>%
  filter(str_detect(trialId, "chinrest")) %>%
  select(participantId, responseId, answer) %>%
  filter(
    str_detect(responseId, "pixelsPerMM") |
      str_detect(responseId, "dist-calibration-MM")
  ) %>%
  mutate(answer = as.numeric(answer)) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  dplyr::rename(distToScreen = `dist-calibration-MM`)

# part 1 relevant df
part1_df <- pilot_df %>%
  filter(str_detect(trialId, "block5_test")) %>%
  select(participantId, trialOrder, responseId, answer) %>%
  mutate(answer = as.numeric(answer)) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  dplyr::rename(
    num.ans.x = `location-x`,
    num.ans.y = `location-y`,
    num.resp.x = `slider-x`,
    num.resp.y = `slider-y`,
    px.resp.x = `px-slider-x`,
    px.resp.y = `px-slider-y`,
    px.ans.x = `px-location-x`,
    px.ans.y = `px-location-y`
  )

# part 2 relevant df
part2_df <- pilot_df %>%
  filter(str_detect(trialId, "task")) %>%
  select(
    participantId,
    trialOrder,
    trialId,
    responseId,
    condition,
    answer,
    parameters_taskType,
    parameters_taskIndex
  ) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  dplyr::rename(
    type = parameters_taskType,
    ogIndex = parameters_taskIndex,
    px.resp.y = pixelResponse,
    num.resp.y = numericResponse
  ) %>%
  mutate(
    px.resp.y = as.numeric(px.resp.y),
    num.resp.y = as.numeric(num.resp.y),
    type = ifelse(type == "pointArc", "point_arc", "point")
  )

# join to make the participant_df also contain condition
participant_df <- participant_df %>%
  left_join(
    part2_df %>%
      select(-condition) %>%
      separate_wider_delim(
        trialId,
        delim = "_",
        names = c("condition", "A", "B")
      ) %>%
      select(participantId, condition) %>%
      unique(.)
  )
```

Read in stimuli data for experiment 2 from Moritz et al.: 

```{r}
exp_json_data <- jsonlite::fromJSON(here::here(
  "data",
  "moritz",
  "stimuli",
  "exp2-stimuli.json"
)) %>%
  as_tibble(.)

stimuli_df <- exp_json_data %>%
  select(index, seed, noise, flip, type, data) %>%
  mutate(noise = as.character(noise))

# get the mapping between og index and the index used in exp
stimuli_df %>%
  filter(type == "point") %>%
  mutate(expIndex = 1:nrow(.)) %>%
  select(index, type, expIndex) -> point_map
stimuli_df %>%
  filter(type == "point_arc") %>%
  mutate(expIndex = 1:nrow(.)) %>%
  select(index, type, expIndex) -> pointArc_map

# get rid of the type that we don't use
stimuli_df <- stimuli_df %>% filter(type != "line")

# get the answer
part2.ans_df <- stimuli_df %>%
  unnest(col = data) %>%
  dplyr::rename(num.x = x, num.y = y, ogIndex = index) %>%
  dplyr::summarise(
    num.mean.y = mean(num.y),
    num.median.y = median(num.y),
    .by = c(ogIndex, type, flip, noise, seed)
  ) %>%
  mutate(
    # this is what Moritz did, and what we're not doing
    # num.ans.y = num.mean.y[type == "point"][[1]],
    num.ans.y = num.mean.y,
    .by = c(flip, noise, seed)
  )
```

Read in helper functions for transforming between spaces: 

```{r}
#| code-fold: true
#| label: helper-functions

# --- share helper functions
# return visual angle in degrees and not radian
vis_angle <- function(size, distance) {
  return(2 * atan(size / (2 * distance)) * 180 / pi)
}

# define function for turning vis angle to phy size
phys_size <- function(va, distance) {
  return(2 * distance * tan(va * pi / 360))
}

px_to_phys = \(px, px_per_mm) px / px_per_mm
phys_to_px = \(phys, px_per_mm) phys * px_per_mm

# --- Factory function
make_va_scale <- function(data_min, data_max, px_min, px_max) {
  px_extent <- abs(px_max - px_min)
  data_range <- data_max - data_min

  data_to_phys <- function(data, px_per_mm) {
    (data - data_min) / data_range * px_extent / px_per_mm
  }
  phys_to_data <- function(phys, px_per_mm) {
    phys * px_per_mm / px_extent * data_range + data_min
  }

  data_to_va <- function(data, distance, px_per_mm) {
    data |> data_to_phys(px_per_mm) |> vis_angle(distance = distance)
  }
  va_to_data <- function(va, distance, px_per_mm) {
    va |> phys_size(distance = distance) |> phys_to_data(px_per_mm)
  }

  data_to_px <- function(data) {
    (data - data_min) / data_range * (px_max - px_min) + px_min
  }
  px_to_data <- function(px) {
    (px - px_min) / (px_max - px_min) * data_range + data_min
  }

  list(
    data_to_va = data_to_va,
    va_to_data = va_to_data,
    data_to_phys = data_to_phys,
    phys_to_data = phys_to_data,
    data_to_px = data_to_px,
    px_to_data = px_to_data
  )
}

# First Part -- Block 5
x_block5 <- make_va_scale(
  data_min = -5,
  data_max = 5,
  px_min = 50,
  px_max = 585
)
y_block5 <- make_va_scale(
  data_min = 0,
  data_max = 10,
  px_min = 410,
  px_max = 15
)

# Second part -- Mortiz
x_moritz <- make_va_scale(
  data_min = 0,
  data_max = 120,
  px_min = 0,
  px_max = 500
)
y_moritz <- make_va_scale(data_min = 0, data_max = 1, px_min = 0, px_max = 200)

# # Code copied from Matt; this prevents us from having mistakes
# x_num_to_px = \(x, x_range = 120, px_range = 500) x / x_range * px_range
# y_num_to_px = \(y, y_range = 1, px_range = 200) y / y_range * px_range

# x_px_to_num = \(x, x_range = 120, px_range = 500) x * x_range / px_range
# y_px_to_num = \(y, y_range = 1, px_range = 200) y * y_range / px_range

# x_va_to_num = \(va, distance, px_per_mm, x_range = 120, px_range = 500) {
#   va |>
#     phys_size(distance) |>
#     phys_to_px(px_per_mm) |>
#     x_px_to_num(x_range, px_range)
# }

# y_va_to_num = \(va, distance, px_per_mm, y_range = 1, px_range = 200) {
#   va |>
#     phys_size(distance) |>
#     phys_to_px(px_per_mm) |>
#     y_px_to_num(y_range, px_range)
# }

# data_to_pixel_y <- function(data_y) {
#   return(-395 * data_y + 410)
# }
# data_to_pixel_x <- function(data_x) {
#   return(53.5 * data_x + 317.5)
# }

# # origin is bottom left
# pixel_to_phy_x <- function(pixel, pxMM) {
#   (pixel - 50) / pxMM
# }
# pixel_to_phy_y <- function(pixel, pxMM) {
#   (410 - pixel) / pxMM
# }
```

# See whether data is sampling correctly 

```{r}
# get a sense of the distribution of part1_df
part1_df %>%
  ggplot(aes(y = num.resp.y)) +
  stat_halfeye() +
  coord_cartesian(ylim = c(0, 10))
part1_df %>%
  ggplot(aes(x = num.resp.x)) +
  stat_halfeye() +
  coord_cartesian(xlim = c(-5, 5))
part1_df %>% ggplot(aes(x = num.resp.x, y = num.resp.y)) + geom_bin2d()
```

The above tells us that data is being sampled correctly. 

--- 

Get a sense of whether participants answered the calibration task correctly: 

```{r}
participant_df %>% ggplot(aes(x = pixelsPerMM)) + geom_dotplot()
participant_df %>% ggplot(aes(x = distToScreen)) + geom_dotplot()
```

These numbers look normal. 

---

Calculate spearman correlation to determine whether we want to filter out participant who fail our check: 

```{r}
part1_df %>%
  summarise(
    rho = cor(num.ans.x, num.resp.x, method = "spearman"),
    .by = participantId
  )

part1_df %>%
  summarise(
    rho = cor(num.ans.y, num.resp.y, method = "spearman"),
    .by = participantId
  )
```

So far everyone seems to fit. So we are not filtering out anyone from this. 

# Fit model for part 1 

## Get the data ready: 

```{r}
# adding the visual angle related stuff

part1_df <- part1_df %>%
  left_join(participant_df) %>%
  mutate(
    va.ans.x = x_block5$data_to_va(num.ans.x, distToScreen, pixelsPerMM),
    va.ans.y = y_block5$data_to_va(num.ans.y, distToScreen, pixelsPerMM),
    va.resp.x = x_block5$data_to_va(num.resp.x, distToScreen, pixelsPerMM),
    va.resp.y = y_block5$data_to_va(num.resp.y, distToScreen, pixelsPerMM)
  ) %>%
  mutate(va.error.x = va.resp.x - va.ans.x, va.error.y = va.resp.y - va.ans.y)
```

Get a sense of error distribution 

```{r}
#| layout-ncol: 2
#| code-fold: true

part1_df %>%
  ggplot(aes(y = va.error.x, x = va.ans.y)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  labs(title = "Project to X-Axis Error (Vis Angle)")
part1_df %>%
  ggplot(aes(y = va.error.y, x = va.ans.x)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  labs(title = "Project to Y-Axis Error (Vis Angle)")
```

Note that for matching with part 2, we are interested in the second graph. 

## Model Specification 

Recall the we defined the following model for the "project from dot to axis" operator: 

$$
\begin{align*}
\texttt{error}_i &\sim \mathcal{N}(\beta_0 + u_{\text{PID[i]}}, \sigma_{i, \text{PID}[i]}) \\ 
\log(\sigma_{i, j}) &= \alpha_j + \log(\texttt{va.ans.x}_i) \\ 
\text{or:} \log(\sigma_{i, j}) &\sim \mathcal{N}(\mu_\alpha + \log(\texttt{va.ans.x}_i), \sigma_\alpha)\\ 
\alpha_j &\sim N(\mu_\alpha, \sigma_\alpha) \\ 
\mu_\alpha &\sim N(0, 1) \\ 
\sigma_\alpha &\sim \mathcal{N}^+(0, 1)
\end{align*}
$$

Here is the specification we used for fitting: 

```{r}
#| label: brms specs

# Recall that we defined error = resp - ans
f <- bf(
  va.error.y ~ (1 | participantId),
  sigma ~ (1 | participantId) + offset(log(va.ans.x))
)

p <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 1), class = sd, group = participantId, lb = 0),
  prior(normal(0, 1), class = Intercept, dpar = "sigma"),
  prior(normal(0, 1), class = sd, group = participantId, dpar = sigma, lb = 0)
)
```

Fit the model: 

```{r}
#| label: fit model for part 1

m1 <- part1_df %>%
  brm(
    formula = f,
    data = .,
    family = gaussian(),
    prior = p,
    chains = 4,
    cores = 4,
    control = list(adapt_delta = 0.99),
    iter = 6000,
    file = here::here("analysis", "fitted_models", "exp2", "m1"),
    file_refit = "on_change",
    save_pars = save_pars(all = TRUE)
  )
```

Get a sense of posterior fit: 

```{r}
summary(m1)
plot(m1)
```

Note that the Bulk_ESS and Tail_ESS all indicate good fit. 

Posterior predictive checks: 

```{r}
#| layout-ncol: 3
#| code-fold: true
pp_check(m1, ndraws = 100) +
  theme(legend.position = c(0.9, 0.5)) +
  theme(text = element_text(size = 12)) -> p1
p1

pp_check(m1, ndraws = 100, type = "pit_ecdf") +
  theme(text = element_text(size = 12)) -> p2
p2

pp_check(m1, ndraws = 100, type = "pit_ecdf", plot_diff = TRUE) +
  theme(text = element_text(size = 10)) -> p3
p3
```

```{r}
#| eval: false

ggsave(
  here::here("figs", "part2-task1-pp-check.pdf"),
  p1,
  width = 12,
  height = 6,
  units = "cm",
  dpi = 300
)
ggsave(
  here::here("figs", "part2-task1-pit-ecdf.pdf"),
  p2,
  height = 6,
  width = 6,
  units = "cm",
  dpi = 300
)
```

Plot existing data on top of prediction, for an average participant: 

```{r}
#| code-fold: true
#| label: part 1 posterior fit check

prediction_grid <- part1_df %>%
  modelr::data_grid(
    va.ans.x = modelr::seq_range(va.ans.x, n = 100),
    participantId = unique(participantId)
  ) %>%
  add_predicted_draws(m1, ndraws = 100)

ggplot(prediction_grid, aes(x = va.ans.x, y = .prediction)) +
  stat_lineribbon(
    .width = c(.95, .80, .50),
    # alpha = 0.5,
    linewidth = 0.5
  ) +
  scale_fill_brewer(palette = "Blues") +
  geom_point(
    data = part1_df,
    aes(y = va.error.y, x = va.ans.x),
    alpha = 0.5,
    size = 0.5
  ) +
  labs(
    title = "Part 1: Project to Y-axis",
    y = "Error",
    x = "Distance to Y-Axis"
  ) +
  # theme_minimal(base_size = 8) +
  # coord_cartesian(ylim = c(-0.25, 0.4)) +
  theme(legend.position = "none") -> p1

p1
```

```{r}
#| eval: false

ggsave(
  here::here("figs", "part2-task1-project-y.pdf"),
  p1,
  height = 8,
  width = 8,
  units = "cm",
  dpi = 300
)
```


## Prior sensitivity analysis -- Only Run if time permitting 

To test whether our model is sensitive to prior specifications, we fit them on differnt priors: 

```{r}
#| eval: false
# Wider — less informative
p_wide <- c(
  prior(normal(0, 5), class = Intercept),
  prior(normal(0, 5), class = sd, group = participantId, lb = 0),
  prior(normal(0, 5), class = Intercept, dpar = "sigma"),
  prior(normal(0, 5), class = sd, group = participantId, dpar = sigma, lb = 0)
)

# Tighter — more informative
p_tight <- c(
  prior(normal(0, 0.1), class = Intercept),
  prior(normal(0, 0.1), class = sd, group = participantId, lb = 0),
  prior(normal(0, 0.1), class = Intercept, dpar = "sigma"),
  prior(normal(0, 0.1), class = sd, group = participantId, dpar = sigma, lb = 0)
)

# Different family — heavier tails
p_student <- c(
  prior(student_t(3, 0, 1), class = Intercept),
  prior(student_t(3, 0, 1), class = sd, group = participantId, lb = 0),
  prior(student_t(3, 0, 1), class = Intercept, dpar = "sigma"),
  prior(
    student_t(3, 0, 1),
    class = sd,
    group = participantId,
    dpar = sigma,
    lb = 0
  )
)
```

```{r}
#| eval: false
m1_wide <- part1_df %>%
  brm(
    formula = f,
    data = .,
    family = gaussian(),
    prior = p_wide,
    chains = 4,
    cores = 4,
    control = list(adapt_delta = 0.99),
    iter = 6000,
    file = here::here("analysis", "fitted_models", "exp2", "m1_wide"),
    file_refit = "on_change",
    save_pars = save_pars(all = TRUE)
  )

m1_tight <- part1_df %>%
  brm(
    formula = f,
    data = .,
    family = gaussian(),
    prior = p_tight,
    chains = 4,
    cores = 4,
    control = list(adapt_delta = 0.99),
    iter = 6000,
    file = here::here("analysis", "fitted_models", "exp2", "m1_tight"),
    file_refit = "on_change",
    save_pars = save_pars(all = TRUE)
  )

m1_student <- part1_df %>%
  brm(
    formula = f,
    data = .,
    family = gaussian(),
    prior = p_student,
    chains = 4,
    cores = 4,
    control = list(adapt_delta = 0.99),
    iter = 6000,
    file = here::here("analysis", "fitted_models", "exp2", "m1_student"),
    file_refit = "on_change",
    save_pars = save_pars(all = TRUE)
  )
```

All models fit without convergence issues. Now we combine posterior and plot them: 

```{r}
# Combine posteriors from all models
models <- list(
  "default (0,1)" = m1,
  "wide (0,5)" = m1_wide,
  "tight (0,0.1)" = m1_tight,
  "student_t(3,0,1)" = m1_student
)

combined <- models %>%
  imap_dfr(
    ~ gather_draws(.x, b_Intercept, b_sigma_Intercept) %>%
      mutate(prior = .y)
  )
```

Interval plot: 

```{r}
combined %>%
  # uncomment to see what variance instead of log(variance) is.
  # mutate(
  #   .value = if_else(.variable == "b_sigma_Intercept", exp(.value), .value)
  # ) %>%
  ggplot(aes(x = .value, y = prior)) +
  stat_pointinterval(.width = c(.50, .95)) +
  facet_wrap(~.variable, scales = "free_x") +
  labs(x = "Estimate", y = "Prior specification")
```

Densities overlaid: 

```{r}
combined %>%
  ggplot(aes(x = .value, color = prior, fill = prior)) +
  stat_slab(alpha = 0.3) +
  facet_wrap(~.variable, scales = "free") +
  labs(x = "Estimate", y = NULL)
```

To see the actual table: 

```{r}
combined %>%
  group_by(prior, .variable) %>%
  median_qi(.value, .width = .95) %>%
  select(prior, .variable, .value, .lower, .upper) %>%
  arrange(.variable, prior)
```

We can also check random effects: 

```{r}
combined_sd <- models %>%
  imap_dfr(
    ~ gather_draws(
      .x,
      sd_participantId__Intercept,
      sd_participantId__sigma_Intercept
    ) %>%
      mutate(prior = .y)
  )
```

```{r}
combined_sd %>%
  ggplot(aes(x = .value, y = prior)) +
  stat_pointinterval(.width = c(.50, .95)) +
  facet_wrap(~.variable, scales = "free_x") +
  labs(x = "Estimate", y = "Prior specification")
```

**Takeaway**: posteriors were robust to prior specification across default, wide, and heavy-tailed priors. The tight prior $\sigma = 0.1$ constrained the residual variance parameter, confirming it is unreasonably informative for this parameter; the remaining three specifications produced essentially identical posteriors. 

# Infer model for part 2 

**Step 1:** Extract participant-level draws (20 participants $\times$ 3000 per chain $\times$ 4 chains)

```{r}
m1_draws_df <- m1 %>%
  spread_draws(
    b_Intercept,
    r_participantId[participantId, ],
    b_sigma_Intercept,
    r_participantId__sigma[participantId, ]
  ) %>%
  mutate(
    mu = b_Intercept + r_participantId,
    var_scale = exp(b_sigma_Intercept + r_participantId__sigma)
  ) %>%
  select(.draw, participantId, mu, var_scale)
```


**Step 2**: Then, we get a cross between participant and the stimuli they saw, in order to generate **predictions**: 

```{r}
predict_df <- participant_df %>%
  rename(type = condition) %>%
  mutate(type = ifelse(type == "pointArc", "point_arc", "point")) %>%
  inner_join(
    stimuli_df %>% filter(type != "line") %>% rename(ogIndex = index),
    by = "type",
    relationship = "many-to-many"
  ) %>%
  unnest(col = data) %>%
  rename(num.ans.x = x, num.ans.y = y) %>%
  mutate(
    va.60 = x_moritz$data_to_va(60, distToScreen, pixelsPerMM),
    va.ans.x = x_moritz$data_to_va(num.ans.x, distToScreen, pixelsPerMM),
    va.ans.y = y_moritz$data_to_va(num.ans.y, distToScreen, pixelsPerMM),
  )

predict_df %>% head(5)
```

## Project Once 

**Step 3: Draw the posterior prediction for project once to the y-axis:**

```{r}
posterior_predict_df <- map_dfr(
  .progress = TRUE,
  unique(predict_df$participantId),
  function(pid) {
    # 12,000 draws for this participant
    d <- m1_draws_df %>%
      filter(participantId == pid) %>%
      arrange(.draw)

    # 2,880 observations (48 trials × 60 dots)
    obs <- predict_df %>%
      filter(participantId == pid) %>%
      select(ogIndex, va.60, va.ans.x, va.ans.y, pixelsPerMM, distToScreen)

    cross_join(d, obs) %>%
      mutate(
        sigma = var_scale * va.ans.x, # dist to y-axis is va.ans.x
        .prediction = rnorm(n(), mean = mu, sd = sigma),
        pred_y = va.ans.y + .prediction,
        w = 1 / (mu^2 + sigma^2) # inverse-mse
      ) %>%
      group_by(
        .draw,
        mu,
        var_scale,
        pixelsPerMM,
        distToScreen,
        va.60,
        ogIndex
      ) %>%
      summarise(
        va.1proj.mean = mean(pred_y),
        va.1proj.median = median(pred_y),
        va.1proj.weighted = sum(w * pred_y) / sum(w),
        .groups = "drop_last"
      ) %>%
      mutate(
        participantId = pid
      )
  }
)
```

```{r}
posterior_predict_df <- posterior_predict_df %>%
  mutate(
    num.1proj.mean = y_moritz$va_to_data(
      va.1proj.mean,
      distToScreen,
      pixelsPerMM
    ),
    num.1proj.median = y_moritz$va_to_data(
      va.1proj.median,
      distToScreen,
      pixelsPerMM
    ),
    num.1proj.weighted = y_moritz$va_to_data(
      va.1proj.weighted,
      distToScreen,
      pixelsPerMM
    )
  )
```

```{r}
#| eval: false
#| echo: false

saveRDS(
  posterior_predict_df,
  file = here::here("analysis/fitted_models/exp2/part2_posterior_1proj.rds")
)
```

### Posterior checks 

```{r}
#| code-fold: true

# get the relevant ones for each strategy
response_ordered <- part2_df %>% arrange(participantId, ogIndex)
y <- response_ordered$num.resp.y

# group can be any condition-level variable, or a combination
group <- response_ordered %>%
  left_join(part2_ans_df %>% select(ogIndex, type, flip, noise)) %>%
  mutate(g = paste(type, noise, flip, sep = " | ")) %>%
  pull(g)

yrep_mean <- posterior_predict_df %>%
  ungroup() %>%
  select(participantId, ogIndex, .draw, num.1proj.mean) %>%
  arrange(participantId, ogIndex, .draw) %>%
  unite("obs_id", participantId, ogIndex) %>%
  pivot_wider(
    id_cols = .draw,
    names_from = obs_id,
    values_from = num.1proj.mean
  ) %>%
  select(-.draw) %>%
  as.matrix()

yrep_median <- posterior_predict_df %>%
  ungroup() %>%
  select(participantId, ogIndex, .draw, num.1proj.median) %>%
  arrange(participantId, ogIndex, .draw) %>%
  unite("obs_id", participantId, ogIndex) %>%
  pivot_wider(
    id_cols = .draw,
    names_from = obs_id,
    values_from = num.1proj.median
  ) %>%
  select(-.draw) %>%
  as.matrix()

yrep_weighted <- posterior_predict_df %>%
  ungroup() %>%
  select(participantId, ogIndex, .draw, num.1proj.weighted) %>%
  arrange(participantId, ogIndex, .draw) %>%
  unite("obs_id", participantId, ogIndex) %>%
  pivot_wider(
    id_cols = .draw,
    names_from = obs_id,
    values_from = num.1proj.weighted
  ) %>%
  select(-.draw) %>%
  as.matrix()
```

**PIT-ECDFs**: 

```{r}
#| layout-ncol: 3
ppc_pit_ecdf(y, yrep_mean) + labs(title = "Mean Strategy")
ppc_pit_ecdf(y, yrep_median) + labs(title = "Median Strategy")
ppc_pit_ecdf(y, yrep_weighted) + labs(title = "Weighted Strategy")
```

Grouped version: 

```{r}
#| layout-ncol: 3
ppc_pit_ecdf_grouped(y, yrep_mean, group) + labs(title = "Mean Strategy")
ppc_pit_ecdf_grouped(y, yrep_median, group) + labs(title = "Median Strategy")
ppc_pit_ecdf_grouped(y, yrep_weighted, group) +
  labs(title = "Weighted Strategy")
```

**KDE-Overlay**: 

```{r}
ppc_dens_overlay(y, yrep_mean[sample(nrow(yrep_mean), 100), ]) +
  labs(title = "Mean Strategy")
ppc_dens_overlay(y, yrep_median[sample(nrow(yrep_median), 100), ]) +
  labs(title = "Median Strategy")
ppc_dens_overlay(y, yrep_weighted[sample(nrow(yrep_weighted), 100), ]) +
  labs(title = "Weighted Strategy")
```

## Project Twice 

Same step 1 and step 2 as before, so we reuse them. 

**Step 3: Draw the posterior prediction for project once to the middle AND project twice to the y-axis:**

```{r}
posterior_predict_df <- map_dfr(
  .progress = TRUE,
  unique(predict_df$participantId),
  function(pid) {
    # 12,000 draws for this participant
    d <- m1_draws_df %>%
      filter(participantId == pid) %>%
      arrange(.draw)

    # 2,880 observations (48 trials × 60 dots)
    obs <- predict_df %>%
      filter(participantId == pid) %>%
      select(ogIndex, va.60, va.ans.x, va.ans.y, pixelsPerMM, distToScreen)

    cross_join(d, obs) %>%
      mutate(
        sigma = var_scale * abs(va.60 - va.ans.x), # dist to center
        .prediction = rnorm(n(), mean = mu, sd = sigma),
        pred_y = va.ans.y + .prediction,
        w = 1 / (mu^2 + sigma^2) # inverse-mse
      ) %>%
      group_by(
        .draw,
        mu,
        var_scale,
        pixelsPerMM,
        distToScreen,
        va.60,
        ogIndex
      ) %>%
      summarise(
        va.1proj.mean = mean(pred_y),
        va.1proj.median = median(pred_y),
        va.1proj.weighted = sum(w * pred_y) / sum(w), # not sure if this is the right level ...? This feels more like across the things ...?
        .groups = "drop_last"
      ) %>%
      mutate(
        proj_noise = rnorm(n(), mean = mu, sd = var_scale * va.60),
        va.2proj.mean = va.1proj.mean + proj_noise,
        va.2proj.median = va.1proj.median + proj_noise,
        va.2proj.weighted = va.1proj.weighted + proj_noise,
        participantId = pid
      )
  }
)
```

We then transform things from visual angle prediction to actual numerical value, for comparison against the actual response we got: 

```{r}
posterior_predict_df <- posterior_predict_df %>%
  mutate(
    num.2proj.mean = y_moritz$va_to_data(
      va.2proj.mean,
      distToScreen,
      pixelsPerMM
    ),
    num.2proj.median = y_moritz$va_to_data(
      va.2proj.median,
      distToScreen,
      pixelsPerMM
    ),
    num.2proj.weighted = y_moritz$va_to_data(
      va.2proj.weighted,
      distToScreen,
      pixelsPerMM
    ),
  )
```

```{r}
#| eval: false
#| echo: false

saveRDS(
  posterior_predict_df,
  file = here::here("analysis/fitted_models/exp2/part2_posterior_2proj.rds")
)

posterior_predict_df <- readRDS(here::here(
  "analysis/fitted_models/exp2/part2_posterior_2proj.rds"
))
```

### Posterior Checks 

PIT-ECDF Plots 

```{r}
response_ordered <- part2_df %>% arrange(participantId, ogIndex)
y <- response_ordered$num.resp.y

# group can be any condition-level variable, or a combination
group <- response_ordered %>%
  left_join(
    part2_ans_df %>%
      select(ogIndex, type, flip, noise)
  ) %>%
  # mutate(g = paste(type, noise, flip, sep = " | ")) %>%
  mutate(g = paste(noise, flip, sep = " | ")) %>%
  pull(g)

yrep_mean <- posterior_predict_df %>%
  ungroup() %>%
  select(participantId, ogIndex, .draw, num.2proj.mean) %>%
  arrange(participantId, ogIndex, .draw) %>%
  unite("obs_id", participantId, ogIndex) %>%
  pivot_wider(
    id_cols = .draw,
    names_from = obs_id,
    values_from = num.2proj.mean
  ) %>%
  select(-.draw) %>%
  as.matrix()

yrep_median <- posterior_predict_df %>%
  ungroup() %>%
  select(participantId, ogIndex, .draw, num.2proj.median) %>%
  arrange(participantId, ogIndex, .draw) %>%
  unite("obs_id", participantId, ogIndex) %>%
  pivot_wider(
    id_cols = .draw,
    names_from = obs_id,
    values_from = num.2proj.median
  ) %>%
  select(-.draw) %>%
  as.matrix()

yrep_weighted <- posterior_predict_df %>%
  ungroup() %>%
  select(participantId, ogIndex, .draw, num.2proj.weighted) %>%
  arrange(participantId, ogIndex, .draw) %>%
  unite("obs_id", participantId, ogIndex) %>%
  pivot_wider(
    id_cols = .draw,
    names_from = obs_id,
    values_from = num.2proj.weighted
  ) %>%
  select(-.draw) %>%
  as.matrix()
```

```{r}
ppc_pit_ecdf(y, yrep_mean) + labs(title = "Mean Strategy")
ppc_pit_ecdf(y, yrep_mean, plot_diff = TRUE) +
  labs(title = "Mean Strategy")
ppc_pit_ecdf_grouped(y, yrep_mean, group) + labs(title = "Mean Strategy")
```

```{r}
ppc_pit_ecdf(y, yrep_median) + labs(title = "Median Strategy")
ppc_pit_ecdf(y, yrep_median, plot_diff = TRUE) + labs(title = "Median Strategy")
ppc_pit_ecdf_grouped(y, yrep_median, group) + labs(title = "Median Strategy")
```

```{r}
ppc_pit_ecdf(y, yrep_weighted) + labs(title = "Weighted Mean Strategy")
ppc_pit_ecdf(y, yrep_weighted, plot_diff = TRUE) +
  labs(title = "Weighted Mean Strategy")
ppc_pit_ecdf_grouped(y, yrep_weighted, group) +
  labs(title = "Weighted Mean Strategy")
```

### KDE plots 

```{r}
#| layout-ncol: 3
ppc_dens_overlay(y, yrep_mean[sample(nrow(yrep_mean), 100), ]) +
  labs(title = "Mean Strategy")
ppc_dens_overlay(y, yrep_median[sample(nrow(yrep_median), 100), ]) +
  labs(title = "Median Strategy")
ppc_dens_overlay(y, yrep_weighted[sample(nrow(yrep_weighted), 100), ]) +
  labs(title = "Weighted Mean Strategy")

ppc_dens_overlay_grouped(
  y,
  yrep_mean[sample(nrow(yrep_mean), 100), ],
  group = group
) +
  labs(title = "Mean Strategy") +
  geom_vline(xintercept = 0.5, linetype = "dashed") +
  coord_cartesian(xlim = c(0, 1))
ppc_dens_overlay_grouped(
  y,
  yrep_median[sample(nrow(yrep_median), 100), ],
  group = group
) +
  labs(title = "Median Strategy")
ppc_dens_overlay_grouped(
  y,
  yrep_weighted[sample(nrow(yrep_weighted), 100), ],
  group = group
) +
  labs(title = "Weighted Mean Strategy")
```

## Custom KDE plots with `ggdist` and `stat_slab` instead of `bayesplot::ppc_dens_overlay`

```{r}
library(ggdist)

draw_subset <- sample(unique(posterior_predict_df$.draw), 100)

p1 <- posterior_predict_df %>%
  ungroup(.) %>%
  filter(.draw %in% draw_subset) %>%
  select(.draw, ogIndex, num.2proj.mean) %>%
  left_join(
    part2.ans_df %>%
      select(ogIndex, type, flip, noise)
  ) %>%
  ggplot(aes(x = num.2proj.mean, group = .draw, thickness = after_stat(pdf))) +
  stat_slab(
    fill = NA,
    color = "#7CCBA2",
    alpha = 0.7,
    # normalize = "all",
    size = 0.25,
    density = density_bounded(bounds = c(0, 1))
  ) +
  stat_slab(
    data = part2_df %>%
      left_join(
        part2.ans_df %>% select(ogIndex, type, flip, noise),
        by = join_by(type, ogIndex)
      ),
    aes(x = num.resp.y, group = NULL, thickness = after_stat(pdf)),
    fill = NA,
    color = "navy",
    linewidth = 1,
    # normalize = "all",
    density = density_bounded(bounds = c(0, 1))
  ) +
  scale_thickness_shared() +
  geom_vline(
    data = part2.ans_df %>%
      group_by(flip, noise) %>%
      summarise(mean.y = mean(num.mean.y)),
    aes(xintercept = mean.y),
    linetype = "dashed",
    color = "darkgray"
  ) +
  facet_grid(
    noise ~ flip,
    labeller = labeller(
      noise = c("c(0, 0)" = "Variability 0", "c(0, 0.4)" = "Variability .4"),
      flip = c("TRUE" = "Variability Upper", "FALSE" = "Variability Lower")
    )
  ) +
  coord_cartesian(xlim = c(0, 1)) +
  labs(y = "", x = "numerical value", title = "Mean Strategy") +
  theme_light(base_size = 8)

p1
```

Median stategy with ggdist: 

```{r}
draw_subset <- sample(unique(posterior_predict_df$.draw), 100)

p2 <- posterior_predict_df %>%
  ungroup(.) %>%
  filter(.draw %in% draw_subset) %>%
  select(.draw, ogIndex, num.2proj.median) %>%
  left_join(
    part2.ans_df %>%
      select(ogIndex, type, flip, noise)
  ) %>%
  ggplot(aes(
    x = num.2proj.median,
    group = .draw,
    thickness = after_stat(pdf)
  )) +
  stat_slab(
    fill = NA,
    color = "#7CCBA2",
    alpha = 0.7,
    # normalize = "all",
    size = 0.25,
    density = density_bounded(bounds = c(0, 1))
  ) +
  stat_slab(
    data = part2_df %>%
      left_join(
        part2.ans_df %>% select(ogIndex, type, flip, noise),
        by = join_by(type, ogIndex)
      ),
    aes(x = num.resp.y, group = NULL, thickness = after_stat(pdf)),
    fill = NA,
    color = "navy",
    linewidth = 1,
    # normalize = "all",
    density = density_bounded(bounds = c(0, 1))
  ) +
  scale_thickness_shared() +
  geom_vline(
    data = part2.ans_df %>%
      group_by(flip, noise) %>%
      summarise(mean.y = mean(num.mean.y)),
    aes(xintercept = mean.y),
    linetype = "dashed",
    color = "darkgray"
  ) +
  facet_grid(
    noise ~ flip,
    labeller = labeller(
      noise = c("c(0, 0)" = "Variability 0", "c(0, 0.4)" = "Variability .4"),
      flip = c("TRUE" = "Variability Upper", "FALSE" = "Variability Lower")
    )
  ) +
  coord_cartesian(xlim = c(0, 1)) +
  labs(y = "", x = "numerical value", title = "Median Strategy") +
  theme_light(base_size = 8)

p2
```

Sensor fusion strategy: 

```{r}
draw_subset <- sample(unique(posterior_predict_df$.draw), 100)

p3 <- posterior_predict_df %>%
  ungroup(.) %>%
  filter(.draw %in% draw_subset) %>%
  select(.draw, ogIndex, num.2proj.weighted) %>%
  left_join(
    part2.ans_df %>%
      select(ogIndex, type, flip, noise)
  ) %>%
  ggplot(aes(
    x = num.2proj.weighted,
    group = .draw,
    thickness = after_stat(pdf)
  )) +
  stat_slab(
    fill = NA,
    color = "#7CCBA2",
    alpha = 0.7,
    # normalize = "all",
    size = 0.25,
    # density = density_bounded(bounds = c(0, 1))
    density = "unbounded"
  ) +
  stat_slab(
    data = part2_df %>%
      left_join(
        part2.ans_df %>% select(ogIndex, type, flip, noise),
        by = join_by(type, ogIndex)
      ),
    aes(x = num.resp.y, group = NULL, thickness = after_stat(pdf)),
    fill = NA,
    color = "navy",
    linewidth = 1,
    # normalize = "all",
    density = density_bounded(bounds = c(0, 1))
  ) +
  scale_thickness_shared() +
  geom_vline(
    data = part2.ans_df %>%
      group_by(flip, noise) %>%
      summarise(mean.y = mean(num.mean.y)),
    aes(xintercept = mean.y),
    linetype = "dashed",
    color = "darkgray"
  ) +
  facet_grid(
    noise ~ flip,
    labeller = labeller(
      noise = c("c(0, 0)" = "Variability 0", "c(0, 0.4)" = "Variability .4"),
      flip = c("TRUE" = "Variability Upper", "FALSE" = "Variability Lower")
    )
  ) +
  coord_cartesian(xlim = c(0, 1)) +
  labs(y = "", x = "numerical value", title = "Sensor Fusion Strategy") +
  theme_light(base_size = 8)

p3
```

```{r}
#| echo: false
#| eval: false
ggsave(
  here::here(
    "figs",
    "part2_task2_mean.svg"
  ),
  p1,
  width = 3,
  height = 2.5,
  units = c("in"),
  dpi = 300
)

ggsave(
  here::here(
    "figs",
    "part2_task2_median.svg"
  ),
  p2,
  width = 3,
  height = 2.5,
  units = c("in"),
  dpi = 300
)

ggsave(
  here::here(
    "figs",
    "part2_task2_weighted.svg"
  ),
  p3,
  width = 3,
  height = 2.5,
  units = c("in"),
  dpi = 300
)
```

### SKIP --- Group-level checks 

```{r}
# average observed response per stimulus
y_avg <- part2_df %>%
  group_by(ogIndex) %>%
  summarise(response = mean(num.resp.y), .groups = "drop") %>%
  arrange(ogIndex)

# average predicted response per stimulus per draw
yrep_avg <- posterior_predict_df %>%
  group_by(.draw, ogIndex) %>%
  summarise(pred = mean(num.2proj.mean), .groups = "drop") %>%
  arrange(ogIndex, .draw) %>%
  pivot_wider(id_cols = .draw, names_from = ogIndex, values_from = pred) %>%
  select(-.draw) %>%
  as.matrix()

ppc_pit_ecdf(y_avg$response, yrep_avg, prob = 0.99, plot_diff = TRUE)
ppc_dens_overlay(y_avg$response, yrep_avg[sample(nrow(yrep_avg), 100), ])
```

Hmm nope doesn't really capture this two hump thing going on here ... 

<!-- 
### Checking Results across participants 

We can also plot them by the diagonal to see whether they predict things correcty: 

```{r}
#| code-fold: true
combined <- posterior_predict_df %>%
  group_by(ogIndex, participantId) %>%
  median_qi(num.pred.avg, .width = .95) %>%
  left_join(part2_df %>% select(ogIndex, participantId, num.resp.y))

ggplot(combined, aes(x = num.pred.avg, y = num.resp.y)) +
  geom_point(alpha = 0.3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  labs(x = "Predicted", y = "Observed", title = "Mean Strategy") +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))
```

```{r}
#| code-fold: true
combined <- posterior_predict_df %>%
  group_by(ogIndex, participantId) %>%
  median_qi(num.pred.median, .width = .95) %>%
  left_join(part2_df %>% select(ogIndex, participantId, num.resp.y))

ggplot(combined, aes(x = num.pred.median, y = num.resp.y)) +
  geom_point(alpha = 0.3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  labs(x = "Predicted", y = "Observed", title = "Median Strategy") +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))
```

Let's also do **coverage check for mean**: 

```{r}
#| code-fold: true
combined <- posterior_predict_df %>%
  group_by(ogIndex, participantId) %>%
  summarise(
    lo95 = quantile(num.pred.avg, 0.025),
    hi95 = quantile(num.pred.avg, 0.975),
    .groups = "drop"
  ) %>%
  left_join(part2_df %>% select(ogIndex, participantId, num.resp.y)) %>%
  mutate(covered = num.resp.y >= lo95 & num.resp.y <= hi95)

mean(combined$covered)
```

**coverage check for median**:

```{r}
#| code-fold: true
combined <- posterior_predict_df %>%
  group_by(ogIndex, participantId) %>%
  summarise(
    lo95 = quantile(num.pred.median, 0.025),
    hi95 = quantile(num.pred.median, 0.975),
    .groups = "drop"
  ) %>%
  left_join(part2_df %>% select(ogIndex, participantId, num.resp.y)) %>%
  mutate(covered = num.resp.y >= lo95 & num.resp.y <= hi95)

mean(combined$covered)
```

**create PIT-ECDF plot**: 



```{r}
ppc_dens_overlay(y, yrep[sample(nrow(yrep), 50), ]) +
  labs(title = "Mean Strategy")
```

```{r}
response_ordered <- part2_df %>% arrange(participantId, ogIndex)
y <- response_ordered$num.resp.y

yrep <- posterior_predict_df %>%
  ungroup() %>%
  select(participantId, ogIndex, .draw, num.pred.median) %>%
  arrange(participantId, ogIndex, .draw) %>%
  unite("obs_id", participantId, ogIndex) %>%
  pivot_wider(
    id_cols = .draw,
    names_from = obs_id,
    values_from = num.pred.median
  ) %>%
  select(-.draw) %>%
  as.matrix()

bayesplot::ppc_pit_ecdf(y, yrep) + labs(title = "Median Strategy")
```

We can also try KDE overlay: 

```{r}
ppc_dens_overlay(y, yrep[sample(nrow(yrep), 50), ]) +
  labs(title = "Median Strategy")
```

```{r}
# boxplots: observed vs replicated summary stats
ppc_stat(y, yrep, stat = "mean")
ppc_stat(y, yrep, stat = "sd")

# histogram of y with density lines from yrep
ppc_hist(y, yrep[sample(nrow(yrep), 8), ])

# ECDF overlay (non-PIT — just raw values)
ppc_ecdf_overlay(y, yrep[sample(nrow(yrep), 50), ])
```



---

Note that increasing the number of draws does not significantly increase the coverage we have, suggesting that these likely do not capture important structure that humans might actually be using. 

### Checking results for one participant -- skipping for now 

```{r}
# Pick one participant
#| eval: false
pid <- unique(part1_df$participantId)[1]

pred_one <- preds %>%
  filter(participantId == pid)

obs_one <- part2_df %>%
  filter(participantId == pid)
```

Let's first plot mean strategy: 

```{r}
#| code-fold: true
#| eval: false

ggplot() +
  stat_pointinterval(
    data = pred_one,
    aes(y = ogIndex, x = num.pred.mean.y),
    .width = c(.50, .95)
  ) +
  geom_point(
    data = obs_one,
    aes(y = ogIndex, x = num.resp.y),
    color = "red",
    size = 2
  ) +
  labs(y = "Trial", x = "Response (numeric)", title = "Mean Strategy") +
  facet_wrap(~flip)
```

Then let's plot median strategy: 

```{r}
#| code-fold: true
#| eval: false
ggplot() +
  stat_pointinterval(
    data = pred_one,
    aes(y = ogIndex, x = num.pred.median.y),
    .width = c(.50, .95)
  ) +
  geom_point(
    data = obs_one,
    aes(y = ogIndex, x = num.resp.y),
    color = "red",
    size = 2
  ) +
  labs(y = "Trial", x = "Response (numeric)", title = "Median Strategy") +
  facet_wrap(~flip)
```

## Sensor Fusion Strategy 

$$
\hat{y} = \sum w_i \times y_i, \text{ where } w_i = \frac{MSE^{-1}(y_i)}{\sum_j MSE(y_j)}
$$

```{r}
# remove values that are no longer useful
if (exists("preds")) {
  rm(preds)
}
if (exists("combined")) {
  rm(combined)
}
if (exists("predict_df")) {
  rm(predict_df)
}
```

**Step 0:** Redefine `predict_df`, since we're now projecting to the vertical line of `xintercept == 60`: 

```{r}
predict_df <- participant_df %>%
  rename(type = condition) %>%
  mutate(type = ifelse(type == "pointArc", "point_arc", "point")) %>%
  inner_join(
    stimuli_df %>% filter(type != "line") %>% rename(ogIndex = index),
    by = "type",
    relationship = "many-to-many"
  ) %>%
  unnest(col = data) %>%
  rename(num.ans.x = x, num.ans.y = y) %>%
  mutate(
    va_60 = x_moritz$data_to_va(60, distToScreen, pixelsPerMM),
    va.ans.x = abs(
      va_60 - x_moritz$data_to_va(num.ans.x, distToScreen, pixelsPerMM)
    ),
    va.ans.y = y_moritz$data_to_va(num.ans.y, distToScreen, pixelsPerMM),
  )
```

**Step 1:** Extract participant-level draws (20 participants $\times$ 3000 per chain $\times$ 4 chains)

```{r}
m1_draws_df <- m1 %>%
  spread_draws(
    b_Intercept,
    r_participantId[participantId, ],
    b_sigma_Intercept,
    r_participantId__sigma[participantId, ]
  ) %>%
  mutate(
    mu = b_Intercept + r_participantId,
    log_sigma_base = b_sigma_Intercept + r_participantId__sigma
  ) %>%
  select(participantId, .draw, mu, log_sigma_base)
```

Step 2: Do it for a single participant and then extend it to all participants using `purrr::map_dfr`: 

```{r}
# Step 2: Loop over participants, use matrix ops, summarize per trial
posterior_predict_df <- map_dfr(
  .progress = TRUE,
  unique(predict_df$participantId),
  function(pid) {
    # 12,000 draws for this participant
    d <- m1_draws_df %>%
      filter(participantId == pid) %>%
      arrange(.draw)
    n_draws <- nrow(d)

    # 2,880 observations (48 trials × 60 dots)
    obs <- predict_df %>% filter(participantId == pid)
    n_obs <- nrow(obs)

    mu_vec <- d$mu # length 12,000
    log_sb <- d$log_sigma_base # length 12,000
    log_vax <- log(obs$va.ans.x) # length 2,880
    log_va_proj <- log(obs$va_60)

    # sigma matrix: 12,000 × 2,880 (~276 MB)
    sigma_mat <- exp(outer(log_sb, log_vax, "+"))

    # weights: 1 / (mu^2 + sigma^2)
    w_mat <- 1 / (outer(mu_vec^2, rep(1, n_obs)) + sigma_mat^2)

    # predicted y: va.ans.y + rnorm(mu, sigma)
    Z <- matrix(rnorm(n_draws * n_obs), nrow = n_draws, ncol = n_obs)
    pred_y_mat <- matrix(obs$va.ans.y, n_draws, n_obs, byrow = TRUE) +
      mu_vec +
      sigma_mat * Z
    rm(Z, sigma_mat)
    gc()

    # weighted average per trial per draw
    trial_ids <- obs$ogIndex
    trial_results <- map_dfr(unique(trial_ids), function(tr) {
      cols <- which(trial_ids == tr)
      w_sub <- w_mat[, cols, drop = FALSE]
      pred_sub <- pred_y_mat[, cols, drop = FALSE]

      tibble(
        participantId = pid,
        ogIndex = tr,
        .draw = seq_len(n_draws),
        va.weighted_pred = rowSums(w_sub * pred_sub) / rowSums(w_sub)
      )
    })
    # second projection
    n_trials <- length(unique(trial_ids))
    trial_results %>%
      mutate(
        sigma_proj = exp(rep(log_sb, times = n_trials) + log_va_proj[1]),
        va.final_pred = va.weighted_pred +
          rnorm(n(), mean = rep(mu_vec, times = n_trials), sd = sigma_proj)
      )
  }
)
```

This will give us 12000 draws per participant-trial, and we have `20 * 48` trials, which is overall speaking, `11520000` rows. Which is the same as `nrow(posterior_predict_df)`. 

We will translate from visual angle space back to numeric space: 

```{r}
full_pp_df <- posterior_predict_df %>%
  left_join(participant_df, by = join_by(participantId)) %>%
  mutate(
    num.weighted_pred = y_moritz$va_to_data(
      va.final_pred,
      distToScreen,
      pixelsPerMM
    )
  )
```

We can then compare this against the actual response, starting with the **coverage check**:

```{r}
#| code-fold: true
combined <- full_pp_df %>%
  group_by(ogIndex, participantId) %>%
  median_qi(num.weighted_pred, .width = .95) %>%
  left_join(part2_df %>% select(ogIndex, participantId, num.resp.y)) %>%
  mutate(covered = num.resp.y >= .lower & num.resp.y <= .upper)

mean(combined$covered) # should be close to 0.95 if well-calibrated
```

Note this is twice the coverage compared to the median strategy. 

```{r}
combined <- full_pp_df %>%
  group_by(ogIndex, participantId) %>%
  median_qi(num.weighted_pred, .width = .95) %>%
  left_join(part2_df %>% select(ogIndex, participantId, num.resp.y))

ggplot(combined, aes(x = num.weighted_pred, y = num.resp.y)) +
  geom_point(alpha = 0.3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  labs(x = "Predicted", y = "Observed", title = "Weighted Mean Strategy") +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))
```

And then there's also the PIT histogram: 

```{r}
pit_df <- posterior_predict_df %>%
  left_join(part2_df %>% select(ogIndex, participantId, num.resp.y)) %>%
  group_by(ogIndex, participantId) %>%
  summarize(
    pit = mean(num.pred.median <= num.resp.y),
    .groups = "drop"
  )

ggplot(pit_df, aes(x = pit)) +
  geom_histogram(bins = 20, boundary = 0) +
  geom_hline(
    yintercept = nrow(pit_df) / 20,
    linetype = "dashed",
    color = "gray"
  ) +
  labs(
    x = "PIT value",
    y = "Count",
    title = "PIT Histogram (uniform = well-calibrated)"
  )
```

This seems like a good source of inference ...? 
https://topmodels.r-forge.r-project.org/articles/topmodels.html

  
## pre-process data 
  
Look at the empirical distribution: 
  
```{r}
part2_df %>%
  left_join(part2_ans_df) %>%
  mutate(diff = num.resp.y - num.ans.y) %>%
  ggplot(aes(x = diff, y = noise, fill = type)) +
  stat_halfeye(alpha = 0.5) +
  facet_wrap(~flip, ncol = 1) +
  geom_vline(xintercept = 0, linetype = "dashed")
```

```{r}
# plotting function for Moritz result
plot_results <- function(data, xvar) {
  data %>%
    ggplot(aes(
      y = noise,
      x = {{ xvar }},
      dist = "norm",
      arg1 = {{ xvar }},
      arg2 = se,
      fill = type
    )) + #uncertainty
    stat_dist_halfeye(
      show_interval = TRUE,
      position = position_dodge(width = 0.2),
      .width = 0.95,
      size = .5,
      alpha = .8
    ) +
    theme(legend.position = "top") +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray") +
    scale_fill_manual(values = c("#76B7B2", "#E15759")) + # take out "#59A14F"
    ylab("Variability") # + #add x label
  # xlab("Error") #add y label
}
```

```{r}
part2_df %>%
  left_join(part2_ans_df) %>%
  mutate(
    error = num.resp.y - num.ans.y,
    noise = factor(noise, levels = c("c(0, 0)", "c(0, 0.4)"))
  ) %>%
  summarySEwithin(
    data = .,
    measurevar = "error",
    withinvars = c("noise", "flip"),
    betweenvars = "type",
    idvar = "participantId"
  ) %>%
  mutate(
    flip_lab = ifelse(
      flip == FALSE,
      "Variability in top",
      "Variability in bottom"
    ),
    flip_lab = factor(
      flip_lab,
      levels = c("Variability in top", "Variability in bottom")
    )
  ) %>%
  # change error to the direction of "variability"
  # mutate(error = ifelse(flip == FALSE, error, -error)) %>%
  plot_results(., error) +
  # xlab("Error (in the direction of variability)") +
  facet_wrap(~flip_lab, ncol = 1) +
  labs(title = "N = 20")
```

Now this is not something that you would see in Moritz ... 

Posterior distribution of answer for a particular stimuli, for all participants who saw the stimuli: 

```{r}
#| code-fold: true
#| label: plot_stimuli_forAll

# define a helper function for this plot
plot_stimuli_forAll <- function(i, s) {
  target_col <- switch(
    s,
    "mean" = "num.pred.avg",
    "median" = "num.pred.median",
    stop("Error: 's' must be either 'mean' or 'median'.")
  )
  posterior_predict_df %>%
    ungroup(.) %>%
    filter(ogIndex == i) %>%
    ggplot(aes(y = .data[[target_col]])) +
    stat_halfeye() +
    geom_hline(
      data = part2_df %>% filter(ogIndex == i),
      aes(yintercept = num.resp.y),
      color = "red",
      linetype = "dashed"
    ) +
    facet_wrap(~participantId) +
    labs(
      title = paste0(
        "Stimuli Index ",
        i,
        " across 10 participants, ",
        s,
        " strategy"
      )
    )
}
```

```{r}
#| layout-ncol: 2
plot_stimuli_forAll(4, "mean") # point
plot_stimuli_forAll(5, "mean")
```

```{r}
#| layout-ncol: 2
plot_stimuli_forAll(8, "median") # point_arc
plot_stimuli_forAll(9, "median")
``` -->