---
title: "Moritz exp1"
---

```{r}
library(ggplot2)
library(tidyverse)
library(tidybayes)
library(here)
```

Read in data: 

- 142 participants 
- each participant did 48 trials 
- `computed_error` is calculated as `average - reportedaverage`

```{r}
df <- read.csv(here("data", "moritz", "exp", "exp1.csv")) %>% as.data.frame(.)

p_df <- df %>% select(assignmentid, trial, stimuliindex, variability, flip, average, error, errorFrom50, reportedaverage, computed_error)
p_df %>% group_by(trial) %>% count(.)
```


```{r}
p_df %>% ggplot(aes(x = computed_error, y = variability, fill = flip)) + 
    stat_halfeye(alpha = 0.5) + 
    theme_minimal()
```

Note that `flip = TRUE` entails that more variability in lower values of y, while `flip = FALSE` entails higher variability in higher values of y. 

I guess it does look kind of Gaussian ... ?

What about absolute error ? 


```{r}
p_df %>% ggplot(aes(x = abs(computed_error), y = variability)) + 
    stat_halfeye() + 
    theme_minimal()
```

What about the distribution of the true average values ...? 


```{r}
p_df %>% select(average) %>% unique(.) %>% 
    ggplot(aes(x = average)) + stat_halfeye() + 
    coord_cartesian(xlim=c(0, 1)) + 
    theme_minimal()
```