---
title: "Moritz exp1 and exp 2"
---

```{r}
library(ggplot2)
library(tidyverse)
library(tidybayes)
library(here)
library(jsonlite)
```

# Exp 1 

Read in data for experiment 1: 

- 142 participants 
- each participant did 48 trials 
- `computed_error` is calculated as `average - reportedaverage`

```{r}
df <- read.csv(here("data", "moritz", "exp", "exp1.csv")) %>% as.data.frame(.)

p_df <- df %>% select(assignmentid, trial, stimuliindex, variability, flip, average, error, errorFrom50, reportedaverage, computed_error)
p_df %>% group_by(trial) %>% count(.)
```


```{r}
p_df %>% ggplot(aes(x = computed_error, y = variability, fill = flip)) + 
    stat_halfeye(alpha = 0.5) + 
    theme_minimal()
```

Note that `flip = TRUE` entails that more variability in lower values of y, while `flip = FALSE` entails higher variability in higher values of y. 

I guess it does look kind of Gaussian ... ?

What about absolute error ? 


```{r}
p_df %>% ggplot(aes(x = abs(computed_error), y = variability)) + 
    stat_halfeye() + 
    theme_minimal()
```

What about the distribution of the true average values ...? 

```{r}
p_df %>% group_by(stimuliindex, average) %>% summarize(.) %>% 
    ggplot(aes(x = average)) + 
    stat_halfeye() + 
    coord_cartesian(xlim=c(0, 1)) + 
    theme_minimal() + 
    labs(title = "signed error")

p_df %>% group_by(stimuliindex, average) %>% summarize(.)
```

If we answer 0.5 all the time, then the distribution of error is going to look like: 


```{r}
p_df %>% group_by(stimuliindex, average) %>% summarize(.) %>% 
    ggplot(aes(x = average - 0.5)) + 
    stat_halfeye() + 
    # coord_cartesian(xlim=c(0, 1)) + 
    theme_minimal() + 
    labs(title = "signed error")

p_df %>% group_by(stimuliindex, average) %>% summarize(.) %>% 
    ggplot(aes(x = abs(average - 0.5))) + 
    stat_halfeye() + 
    # coord_cartesian(xlim=c(0, 1)) + 
    theme_minimal() + 
    labs(title = "abs error")
```

# Experiment 2 

Experiment design: used a 2 (variability: 0 and .4) `noise` x 2 (variability: upper vs. lower) `flip` x 3 (mark type: line graph, Cartesian spaced points, arc spaced points) `type` mix-design. Using 12 seeds. 

The takeaway is that for datasets that share the same ssed, same noise, same type --- the data is exactly the same except it's flipped. 
same seed, same flip, same type, different noise --> ??? What does this entail ...? 

I just want to understand --- how many of the underlying "line is the same", but they just differ in the spacing ...?? 


```{r}
json_data <- jsonlite::fromJSON(here("data", "moritz", "stimuli", "exp2-stimuli.json" )) %>% as_tibble(.)
# json_data

df <- json_data %>% select(index, seed, noise, flip, type, data) %>% 
    mutate(noise = as.character(noise))

line_df <- json_data %>% select(index, seed, noise, flip, type, lineData) %>% 
    mutate(noise = as.character(noise))

highNoise_df <- json_data %>% select(index, seed, noise, flip, type, highNoiseData) %>% 
    mutate(noise = as.character(noise))

# let's check what the mean value for each group is ...  
```

```{r}
df %>% filter(type != 'line') %>% 
    unnest(cols=c(data)) %>% 
    group_by(index, seed, noise, flip, type) %>% 
    summarise(x_nrow = mean(x), mean_y = mean(y))

line_df %>% filter(type != 'line') %>% 
    unnest(lineData) %>% 
    group_by(index, seed, noise, flip, type) %>% 
    summarise(x_nrow = mean(x), mean_y = mean(y))

highNoise_df %>% filter(type != 'line') %>% 
    unnest(highNoiseData) %>% 
    group_by(index, seed, noise, flip, type) %>% 
    summarise(x_nrow = n(), mean_y = mean(y))
```


```{r}
df %>% filter(type != "line") %>% 
    filter(index %in% c(4, 6)) %>% 
    unnest(data) %>% 
    ggplot(aes(x = x, y = y, color=noise)) + 
    geom_point() + 
    theme_minimal() + 
    labs(title = "Data share seed, flip and type; differ in noise") + 
    coord_cartesian(ylim = c(0, 1))
```

```{r}
#| layout-ncol: 2

df %>% filter(type != "line") %>% 
    filter(index %in% c(4, 8)) %>% 
    unnest(data) %>% 
    ggplot(aes(x = x, y = y, color=type)) + 
    geom_point() + 
    coord_cartesian(ylim = c(0, 1)) + 
    theme_minimal() + 
    labs(title = "Data share seed, flip and noise; differ in type; no noise")

df %>% filter(type != "line") %>% 
    filter(index %in% c(6, 10)) %>% 
    unnest(data) %>% 
    ggplot(aes(x = x, y = y, color=type)) + 
    geom_point() + 
    coord_cartesian(ylim = c(0, 1)) +
    theme_minimal() + 
    labs(title = "Data share seed, flip and noise; differ in type; small noise")
```

```{r}
df %>% filter(type != "line") %>% 
    filter(index %in% c(4, 5)) %>% 
    unnest(data) %>% 
    ggplot(aes(x = x, y = y, color=flip)) + 
    geom_point() + 
    coord_cartesian(ylim = c(0, 1)) +
    theme_minimal() + 
    labs(title = "Data share seed, type and noise; differ in flip")
```

To fill this up, we also consider 4 and 7: 

```{r}
df %>% filter(type != "line") %>% 
    filter(index %in% c(4, 7)) %>% 
    unnest(data) %>% 
    ggplot(aes(x = x, y = y, color=flip)) + 
    geom_point() + 
    theme_minimal() + 
    coord_cartesian(ylim = c(0, 1)) +
    labs(title = "Data share seed, type; differ in flip and noise")
```

Let us check the mean for the first group that share the same seed: 

```{r}
df %>% filter(type != "line") %>% 
    filter(seed == 1) %>% 
    unnest(cols = c(data)) %>% 
    group_by(index, seed, noise, flip, type) %>% 
    summarise(x_nrow = n(), mean_y = mean(y))
```

```{r}
# df %>% unnest(cols=c(data)) %>% 
#     group_by(index, seed, noise, flip, type) %>% 
#     summarise(mean_x = mean(x), mean_y = mean(y)) %>% 
#     ungroup(.) %>% 
#     select(seed, noise, flip, type, mean_y) %>% 
#     filter(flip == FALSE) %>% 
#     ggplot(aes(x = seed, y = mean_y, color = type)) + 
#     geom_point() + 
#     theme_minimal()
```
