---
title: "Moritz exp1 and exp 2"
---

```{r}
library(ggplot2)
library(tidyverse)
library(tidybayes)
library(here)
library(jsonlite)
```

# Exp 1 

Read in data for experiment 1: 

- 142 participants 
- each participant did 48 trials 
- `computed_error` is calculated as `average - reportedaverage`

```{r}
df <- read.csv(here("data", "moritz", "exp", "exp1.csv")) %>% as.data.frame(.)

p_df <- df %>% select(assignmentid, trial, stimuliindex, variability, flip, average, error, errorFrom50, reportedaverage, computed_error)
p_df %>% group_by(trial) %>% count(.)
```


```{r}
p_df %>% ggplot(aes(x = computed_error, y = variability, fill = flip)) + 
    stat_halfeye(alpha = 0.5) + 
    theme_minimal()
```

Note that `flip = TRUE` entails that more variability in lower values of y, while `flip = FALSE` entails higher variability in higher values of y. 

I guess it does look kind of Gaussian ... ?

What about absolute error ? 


```{r}
p_df %>% ggplot(aes(x = abs(computed_error), y = variability)) + 
    stat_halfeye() + 
    theme_minimal()
```

What about the distribution of the true average values ...? 

```{r}
p_df %>% group_by(stimuliindex, average) %>% summarize(.) %>% 
    ggplot(aes(x = average)) + 
    stat_halfeye() + 
    coord_cartesian(xlim=c(0, 1)) + 
    theme_minimal() + 
    labs(title = "signed error")

p_df %>% group_by(stimuliindex, average) %>% summarize(.)
```

If we answer 0.5 all the time, then the distribution of error is going to look like: 


```{r}
p_df %>% group_by(stimuliindex, average) %>% summarize(.) %>% 
    ggplot(aes(x = average - 0.5)) + 
    stat_halfeye() + 
    # coord_cartesian(xlim=c(0, 1)) + 
    theme_minimal() + 
    labs(title = "signed error")

p_df %>% group_by(stimuliindex, average) %>% summarize(.) %>% 
    ggplot(aes(x = abs(average - 0.5))) + 
    stat_halfeye() + 
    # coord_cartesian(xlim=c(0, 1)) + 
    theme_minimal() + 
    labs(title = "abs error")
```

# Read in data for experiment 2 


```{r}
json_data <- jsonlite::fromJSON(here("data", "moritz", "stimuli", "exp2-stimuli.json" ))
json_data

df <- json_data %>% as_tibble(.) %>% 
    select(index, seed, noise, flip, type, data)

# we only want data that has type equal to `point` or `point_arc` 

point_df <- df %>% filter(type != "line")

point_df %>% unnest(cols=c(data)) %>% 
    group_by(index, seed, noise, flip, type) %>% 
    summarise(mean_x = mean(x), mean_y = mean(y))
```