---
title: "Extend from current data to Moritz"
---

```{r}
library(ggplot2)
library(tidyverse)
library(tidybayes)
library(here)
library(jsonlite)
```

Assume that we have the following model: start from the dots, project to the y-axis. This corresponds to the model xxx that we've fit. 

We will first read in the stimuli data from Mortiz experiment 2: 

```{r}
exp_json_data <- jsonlite::fromJSON(here("data", "moritz", "stimuli", "exp2-stimuli.json" )) %>% as_tibble(.)

stimuli_df <- exp_json_data %>% select(index, seed, noise, flip, type, data) %>% 
    mutate(noise = as.character(noise))
```

Let us from the simple case where index == 0: 

```{r}
stimuli_df %>% filter(index == 0) %>% 
    unnest(data) %>% 
    rename(data_x = x, data_y = y)
```

# TODO we need to translate from the data space to visual angle space 

The following code was obtained from the original analysis. 

```{r}
#| label: define-function
#| code-fold: true

# origin is top left  
data_to_pixel_y <- function(data_y) {
  return(-395 * data_y + 410)
}
data_to_pixel_x <- function(data_x) {
  return (53.5 * data_x + 317.5)
}

# origin is bottom left 
pixel_to_phy_x <- function(pixel, pxMM){
  (pixel - 50) / pxMM
}
pixel_to_phy_y <- function(pixel, pxMM){
  (410 - pixel) / pxMM
}

# return visual angle in degrees and not radian
vis_angle <- function(size, distance){
  return(2 * atan(size / (2 * distance)) * 180 / pi)
}

# tolerance for numerical precision 
tolerance <- 1e-10
```

In terms of the models that projected to y-axis, we want to look at the model we fitted for task 5, where we projected to y-axis. This is the math model: 

$$
\begin{align*}
\texttt{error}_i &\sim \mathcal{N}(0, \sigma_{i, \text{PID}[i]}) \\ 
\log(\sigma_{i, j}) &= \alpha_j + \log(\texttt{va.ans.x}_i) \\ 
\alpha_j &\sim N(\mu_\alpha, \sigma_\alpha) \\ 
\mu_\alpha &\sim N(0, 1) \\ 
\sigma_\alpha &\sim \text{Half-Cauchy}(0, 1)
\end{align*}
$$

(Again, copied and pasted from stuff...)

```{r}
f <- bf(error_va ~ (1 | participantId), 
        sigma ~ (1|participantId) + offset(log(va.ans.x)))

p <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 1), class = sd, group = participantId, lb = 0), 
  prior(normal(0, 1), class = Intercept, dpar = "sigma"), 
  prior(normal(0, 1), class = sd, group = participantId, dpar = sigma, lb = 0)
)
```

Since we have a model fitted before, now we just need to read in the fitted rds file. 


```{r}
# TODO 
```