---
title: "Extend from current data to Moritz"
---

```{r}
#| code-fold: true
library(ggplot2)
library(tidyverse)
library(tidybayes)
library(here)
library(jsonlite)
library(brms)
```

# Replicating Mortiz Exp 2

Here are the details of the original experiment: 

- **Hypothesis:** “the effect of variability will be smaller for graphs with points spaced along the x-axis than graphs with points spaced along the arc” ([Moritz et al., 2023, p. 5](zotero://select/library/items/ZPTBTRIF)) ([pdf](zotero://open-pdf/library/items/AZRSTSWZ?page=5&annotation=4IPHWX8I))
- **Experiment Design**:
  - 2 variability: `0` and `0.4`
  - 2 flipped: `upper` and `lower` (i.e., whether the plot was flipped or not)
  - 3 mark type: `line`, `cartesian spaced points`, `arc spaced points` -> between-subjects
    - In our experiment we will only use cartesian spaced points and arc spaced points. 
  - 12 different seeds -> `2 * 2 * 12 = 48` trials per participant.


## Read in Stimuli Data

First, read in the stimuli data from Mortiz experiment 2: 

```{r}
exp_json_data <- jsonlite::fromJSON(here("data", "moritz", "stimuli", "exp2-stimuli.json" )) %>% as_tibble(.)

stimuli_df <- exp_json_data %>% select(index, seed, noise, flip, type, data) %>% 
    mutate(noise = as.character(noise))

stimuli_df %>% head(5)
```

<!-- We will then save the part of it where the `type` is not line, to serve as the `json` input for our experiment:  -->

```{r}
#| eval: false
#| echo: false
# list of index to pull things from 
stimuli_df %>% filter(type!= "line") %>% 
  pull(index)
```

We will start from a toy example, when `index == 4`: 

```{r}
stimuli_df <- stimuli_df %>% filter(type != "line")

ex_df <- stimuli_df %>% filter(index == 4) %>% 
    unnest(data) %>% 
    rename(data_x = x, data_y = y)
ex_df %>% head(5)
```

Let's also read in the data collected from pilot: 

```{r}
raw_df <- read_csv(here("data", "part_2", "part2_pilot.csv")) %>% as_tibble(.) 

pilot_df <- raw_df %>% select(participantId, trialId, trialOrder, responseId, answer, duration, parameters_taskid, parameters_params)

# responses
pilot_df %>% filter(responseId %in% c("numericResponse", "pixelResponse")) %>%
  select(participantId, trialId, responseId, answer) %>%
  separate_wider_delim(col = trialId, delim = "_", names = c("taskType", NA, "index")) %>% 
  pivot_wider(names_from = "responseId", values_from = "answer") %>% 
  mutate(numericResponse = as.numeric(numericResponse), 
    pixelResponse = as.numeric(pixelResponse), 
    index = as.integer(index)
  ) %>% 
  mutate(pixelResponse = 200 - pixelResponse + 5)
  # TODO: 1) px to physical; 2) find out the correct numeric ans

# visual angle related 
pilot_df %>% filter(!responseId %in% c("numericResponse", "pixelResponse")) %>% 
  group_by(participantId) %>% 
  filter(responseId %in% c("prolificId", "pixelsPerMM", "dist-calibration-MM", "dist-calibration-CM", "ball-positions", "square-position")) %>% View(.)
```

## Translate btw spaces

The goal is to translate from the pixel space to the physical space, and then to the visual angle space. 
In the previous study, since we built the x-axis and y-axis using custom d3 scale, we had to hard code the information. 
This time, we used `VegaEmbed` so the translation would be a bit more natural.

```{r}
# y-axis: numerical [0, 1] -> pixel [0, 200] -> phys [...?] 
# for some reason, the pixel response seems to be on the 210 scale and not the 200 scale ...? 
```

```{r}
#| label: define-function
#| code-fold: true

# TODO 

# origin is top left  
data_to_pixel_y <- function(data_y) {
  return(-395 * data_y + 410)
}
data_to_pixel_x <- function(data_x) {
  return (53.5 * data_x + 317.5)
}

# origin is bottom left 
pixel_to_phy_x <- function(pixel, pxMM){
  (pixel - 50) / pxMM
}
pixel_to_phy_y <- function(pixel, pxMM){
  (410 - pixel) / pxMM
}

# return visual angle in degrees and not radian
vis_angle <- function(size, distance){
  return(2 * atan(size / (2 * distance)) * 180 / pi)
}

# tolerance for numerical precision 
tolerance <- 1e-10
```


## Read in fitted model

We assume that the baseline model is this: participants project from each of the dots to the y-axis, and then perform an averaging. We will use the **Project to Y-Axis Model**. 

$$
\begin{align*}
\texttt{error}_i &\sim \mathcal{N}(0, \sigma_{i, \text{PID}[i]}) \\ 
\log(\sigma_{i, j}) &= \alpha_j + \log(\texttt{va.ans.x}_i) \\ 
\alpha_j &\sim N(\mu_\alpha, \sigma_\alpha) \\ 
\mu_\alpha &\sim N(0, 1) \\ 
\sigma_\alpha &\sim \mathcal{N}^+(0, 1)
\end{align*}
$$

Read in the fitted model by first reading in its formula and prior specification: 

```{r}
f <- bf(error_va ~ (1 | participantId), 
        sigma ~ (1|participantId) + offset(log(va.ans.x)))

p <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 1), class = sd, group = participantId, lb = 0), 
  prior(normal(0, 1), class = Intercept, dpar = "sigma"), 
  prior(normal(0, 1), class = sd, group = participantId, dpar = sigma, lb = 0)
)
```

```{r}
# model <- brm(
#     formula = f,
#     data = .,
#     family = gaussian(),
#     prior = p,
#     chains = 4, 
#     cores = 4,
#     control = list(adapt_delta = 0.99),
#     iter = 6000, 
#     file = here("analysis", "fitted_models", "task5.y"), 
#     file_refit = "on_change", 
#     save_pars = save_pars(all = TRUE)
# )
model <- readRDS(here("analysis", "fitted_models", "task5.y.rds"))
model
```

Get the posterior draw: 


```{r}
draws <- as_draws_df(model)
```

# What does an average participant see? 

Assume the pixels to MM ratio is `3.25`. 

```{r}
# know which one it's making back to 
pixelToMM <- 3.25
screenDist <- 612

stimuli_df <- stimuli_df %>% mutate(expIndex = 1:nrow(.))
ex_df <- stimuli_df %>% filter(index == 4) %>% unnest(col = data) %>% 
  rename(numX = x, numY = y) %>%
  # calculate pixel values from numerical values
  mutate(pxX = 500 / 120 * numX,
         pxY = 200 * numY) %>% 
  # calcuate physical values (in MM) from pixel values
  mutate(phyX = pxX / pixelToMM, 
         phyY = pxY / pixelToMM
  ) %>%
  # calculate visual angle 
  mutate(va.ans.x = vis_angle(phyX, screenDist),
         va.ans.y = vis_angle(phyY, screenDist)
  )

ex_df
```

What would the posterior prediction look like? 

```{r}
# define function for turning vis angle to phy size 
phys_size <- function(va, distance) {
  return(2 * distance * tan(va * pi / 360))
}
```

```{r}
ex_df %>% select(index, va.ans.x, va.ans.y) %>% 
  mutate(participantId = "extend") %>% 
  tidybayes::add_predicted_draws(newdata = ., model, re_formula = NA) %>% 
  mutate(va.pred.y = va.ans.y + .prediction) %>% 
  group_by(.draw) %>%
  summarise(va.pred.y.avg = mean(va.pred.y)) %>%
  mutate(phy.pred.y.avg = phys_size(va.pred.y.avg, screenDist), 
    px.pred.y.avg = phy.pred.y.avg * pixelToMM, 
    num.pred.y.avg = px.pred.y.avg / 200
  ) -> pred_df

pred_df
```

```{r}
pred_df %>% median_qi(num.pred.y.avg) 
```

Let's also try to plot this uncertainty: 

```{r}
library(ggplot2)
library(tidybayes)
library(ggdist)

pred_df %>% ggplot(data = ., aes(y = num.pred.y.avg)) +
  stat_halfeye() + 
  theme_minimal() + 
  coord_cartesian(ylim = c(0, 1))
```