---
title: "Extend from current data to Moritz"
---

```{r}
#| code-fold: true
library(ggplot2)
library(tidyverse)
library(tidybayes)
library(here)
library(jsonlite)
```

# Replicating Mortiz Exp 2

Here are the details of the original experiment: 

- **Hypothesis:** “the effect of variability will be smaller for graphs with points spaced along the x-axis than graphs with points spaced along the arc” ([Moritz et al., 2023, p. 5](zotero://select/library/items/ZPTBTRIF)) ([pdf](zotero://open-pdf/library/items/AZRSTSWZ?page=5&annotation=4IPHWX8I))
- **Experiment Design**:
  - 2 variability: `0` and `0.4`
  - 2 flipped: `upper` and `lower` (i.e., whether the plot was flipped or not)
  - 3 mark type: `line`, `cartesian spaced points`, `arc spaced points` -> between-subjects
    - In our experiment we will only use cartesian spaced points and arc spaced points. 
  - 12 different seeds -> `2 * 2 * 12 = 48` trials per participant.


## Read in Stimuli Data

First, read in the stimuli data from Mortiz experiment 2: 

```{r}
exp_json_data <- jsonlite::fromJSON(here("data", "moritz", "stimuli", "exp2-stimuli.json" )) %>% as_tibble(.)

stimuli_df <- exp_json_data %>% select(index, seed, noise, flip, type, data) %>% 
    mutate(noise = as.character(noise))

stimuli_df %>% head(5)
```

<!-- We will then save the part of it where the `type` is not line, to serve as the `json` input for our experiment:  -->

```{r}
#| eval: false
#| echo: false
# list of index to pull things from 
stimuli_df %>% filter(type!= "line") %>% 
  pull(index)
```

We will start from a toy example, when `index == 4`: 

```{r}
ex_df <- stimuli_df %>% filter(index == 4) %>% 
    unnest(data) %>% 
    rename(data_x = x, data_y = y)
ex_df %>% head(5)
```

## Translate btw spaces



```{r}
#| label: define-function
#| code-fold: true

# origin is top left  
data_to_pixel_y <- function(data_y) {
  return(-395 * data_y + 410)
}
data_to_pixel_x <- function(data_x) {
  return (53.5 * data_x + 317.5)
}

# origin is bottom left 
pixel_to_phy_x <- function(pixel, pxMM){
  (pixel - 50) / pxMM
}
pixel_to_phy_y <- function(pixel, pxMM){
  (410 - pixel) / pxMM
}

# return visual angle in degrees and not radian
vis_angle <- function(size, distance){
  return(2 * atan(size / (2 * distance)) * 180 / pi)
}

# tolerance for numerical precision 
tolerance <- 1e-10
```


## Read in fitted model 

Assume that we have the following model: start from the dots, project to the y-axis. This corresponds to the model xxx that we've fit. 

# TODO we need to translate from the data space to visual angle space 

The following code was obtained from the original analysis and therefore needs to be changed. 



In terms of the models that projected to y-axis, we want to look at the model we fitted for task 5, where we projected to y-axis. This is the math model: 

$$
\begin{align*}
\texttt{error}_i &\sim \mathcal{N}(0, \sigma_{i, \text{PID}[i]}) \\ 
\log(\sigma_{i, j}) &= \alpha_j + \log(\texttt{va.ans.x}_i) \\ 
\alpha_j &\sim N(\mu_\alpha, \sigma_\alpha) \\ 
\mu_\alpha &\sim N(0, 1) \\ 
\sigma_\alpha &\sim \text{Half-Cauchy}(0, 1)
\end{align*}
$$

(Again, copied and pasted from stuff...)

```{r}
# f <- bf(error_va ~ (1 | participantId), 
#         sigma ~ (1|participantId) + offset(log(va.ans.x)))

# p <- c(
#   prior(normal(0, 1), class = Intercept),
#   prior(normal(0, 1), class = sd, group = participantId, lb = 0), 
#   prior(normal(0, 1), class = Intercept, dpar = "sigma"), 
#   prior(normal(0, 1), class = sd, group = participantId, dpar = sigma, lb = 0)
# )
```

Since we have a model fitted before, now we just need to read in the fitted rds file. 


```{r}
# TODO 
```

---

Example vega-lite config: 

```
{
  "$schema": "https://vega.github.io/schema/vega-lite/v6.json",
  "description": "A simple bar chart with embedded data.",
  "data": {
    "values": [
    {
      "x": 0,
      "y": 0.8303027032037189
    },
    {
      "x": 2,
      "y": 0.7933579863898285
    },
    {
      "x": 4,
      "y": 0.7323336215851782
    },
    {
      "x": 6,
      "y": 0.6792478913612636
    },
    {
      "x": 8,
      "y": 0.6266749399986509
    },
    {
      "x": 10,
      "y": 0.5874960907002436
    },
    {
      "x": 12,
      "y": 0.5131307369088348
    },
    {
      "x": 14,
      "y": 0.4530739601959595
    },
    {
      "x": 16,
      "y": 0.3949372090732871
    },
    {
      "x": 18,
      "y": 0.33654483050205275
    },
    {
      "x": 20,
      "y": 0.2807749889916027
    },
    {
      "x": 22,
      "y": 0.2457998243565798
    },
    {
      "x": 24,
      "y": 0.23461030068322902
    },
    {
      "x": 26,
      "y": 0.237901653192568
    },
    {
      "x": 28,
      "y": 0.25535590539690317
    },
    {
      "x": 30,
      "y": 0.27752864575624153
    },
    {
      "x": 32,
      "y": 0.28725630720455464
    },
    {
      "x": 34,
      "y": 0.29940586367576155
    },
    {
      "x": 36,
      "y": 0.3121498733033438
    },
    {
      "x": 38,
      "y": 0.32504969820670915
    },
    {
      "x": 40,
      "y": 0.3192286583323132
    },
    {
      "x": 42,
      "y": 0.3363265756201552
    },
    {
      "x": 44,
      "y": 0.3625566706335878
    },
    {
      "x": 46,
      "y": 0.40486026425783006
    },
    {
      "x": 48,
      "y": 0.44377845412846856
    },
    {
      "x": 50,
      "y": 0.49445047261015856
    },
    {
      "x": 52,
      "y": 0.5467532210258335
    },
    {
      "x": 54,
      "y": 0.6146711781236365
    },
    {
      "x": 56,
      "y": 0.6640100223872704
    },
    {
      "x": 58,
      "y": 0.6961895797548191
    },
    {
      "x": 60,
      "y": 0.7360024431605858
    },
    {
      "x": 62,
      "y": 0.766210868669597
    },
    {
      "x": 64,
      "y": 0.7773871415383563
    },
    {
      "x": 66,
      "y": 0.7908152399999402
    },
    {
      "x": 68,
      "y": 0.7916199390452369
    },
    {
      "x": 70,
      "y": 0.814623481209869
    },
    {
      "x": 72,
      "y": 0.8257722434051891
    },
    {
      "x": 74,
      "y": 0.8155876631354766
    },
    {
      "x": 76,
      "y": 0.7950448716026716
    },
    {
      "x": 78,
      "y": 0.7797881633092574
    },
    {
      "x": 80,
      "y": 0.7892118660712901
    },
    {
      "x": 82,
      "y": 0.7872880517478441
    },
    {
      "x": 84,
      "y": 0.7506101327948177
    },
    {
      "x": 86,
      "y": 0.7262035585897042
    },
    {
      "x": 88,
      "y": 0.6878298851943453
    },
    {
      "x": 90,
      "y": 0.6424837175176794
    },
    {
      "x": 92,
      "y": 0.6157889552795547
    },
    {
      "x": 94,
      "y": 0.5997331854414678
    },
    {
      "x": 96,
      "y": 0.5720235188234739
    },
    {
      "x": 98,
      "y": 0.5522607637076893
    },
    {
      "x": 100,
      "y": 0.5082117520782833
    },
    {
      "x": 102,
      "y": 0.4592010851079365
    },
    {
      "x": 104,
      "y": 0.41532372517972366
    },
    {
      "x": 106,
      "y": 0.36720108171038296
    },
    {
      "x": 108,
      "y": 0.306623132752718
    },
    {
      "x": 110,
      "y": 0.21383714339601448
    },
    {
      "x": 112,
      "y": 0.13789091135114828
    },
    {
      "x": 114,
      "y": 0.09298387485943124
    },
    {
      "x": 116,
      "y": 0.044019550695727
    },
    {
      "x": 118,
      "y": 0.01469670080080308
    }
  ]
  },
  "mark": "point",
  "encoding": {
    "x": {"field": "x", "axis": {"ticks": false, "labels": false}},
    "y": {"field": "y", "axis": {"ticks": false, "labels": false}}
  },
  "width": 500,
  "height": 200,
  "autosize": "none",
  "padding": 50
}

```