---
title: "Extend from current data to Moritz"
author: "Anonymous"
date: today
date-format: "iso"
format:
  html:
    reference-location: margin
---

```{r}
#| code-fold: true
#| label: read in libraries

library(plyr)
source(here::here("analysis", "moritz", "Functions.R"))
library(ggplot2)
library(tidyverse)
library(tidybayes)

library(jsonlite)
library(brms)
library(posterior)
library(patchwork)
theme_set(theme_tidybayes())
```

# TLDR (and TODO)

- **High-level goal**: Our previous experiment gave us some decoding operators. We want to showcase its capabilities. For that purpose, we've picked the Moritz et al. paper. 
- **How did Mortiz et al. perform analysis**? 
  - the `lmer` code is for calculating $p$-values for hypothesis testing. 
  - the graphs are actually **descriptive summary statistics in distributional form**.  
- What is Mortiz's plan of analysis here (as in, are they trying to model an average participant? Something else??)
- What is the "apples-to-apples" comparison we need to show here 
- What does this reveal about their modeling choice, and how could we potentially improve on top of it 

For details related to Mortiz's original experiment, and different ways of analyzing it, see `reanalyze.qmd`. 

# Generate posterior predictions for an "average" participant 

Here are the details of the original experiment: 

- **Hypothesis:** “the effect of variability will be smaller for graphs with points spaced along the x-axis than graphs with points spaced along the arc” ([Moritz et al., 2023, p. 5](zotero://select/library/items/ZPTBTRIF)) ([pdf](zotero://open-pdf/library/items/AZRSTSWZ?page=5&annotation=4IPHWX8I))
- **Experiment Design**:
  - *Within Subject*: 
    - 2 variability: `0` and `0.4`
    - 2 flipped^[i.e., whether the plot was flipped or not, where `flip == FALSE` implies that upper-y region has more variability, and `flip == TRUE` implies that the lower-y region has more variability]: `upper` and `lower` 
  - *Between Subject*: 
    - 3 mark type: ~~`line`~~, `cartesian spaced points`, `arc spaced points`
      - Note that we will only use cartesian spaced points and arc spaced points for replication.
  - 12 different seeds for randomization -> `2 * 2 * 12 = 48` trials per participant.

## Read in Stimuli Data

First, read in the stimuli data from Mortiz experiment 2: 

```{r}
#| code-fold: true
#| label: read in moritz stimuli data 

exp_json_data <- jsonlite::fromJSON(here::here(
  "data",
  "moritz",
  "stimuli",
  "exp2-stimuli.json"
)) %>%
  as_tibble(.)

stimuli_df <- exp_json_data %>%
  select(index, seed, noise, flip, type, data) %>%
  mutate(noise = as.character(noise))

# get the mapping between og index and the index used in exp
stimuli_df %>%
  filter(type == "point") %>%
  mutate(expIndex = 1:nrow(.)) %>%
  select(index, type, expIndex) -> point_map
stimuli_df %>%
  filter(type == "point_arc") %>%
  mutate(expIndex = 1:nrow(.)) %>%
  select(index, type, expIndex) -> pointArc_map

# calculate the answer
stimuli_df <- stimuli_df %>%
  filter(type != "line") %>%
  unnest(data) %>%
  left_join(point_map, by = join_by(index, type)) %>%
  left_join(pointArc_map, by = join_by(index, type)) %>%
  mutate(expIndex = coalesce(expIndex.x, expIndex.y)) %>%
  select(index, expIndex, type, flip, noise, seed, x, y) %>%
  ungroup(.) %>%
  rename(ogIndex = index, num.x = x, num.y = y)
```

```{r}
#| label: show stimuli dataframe
stimuli_df
```

Define helper functions: 

```{r}
#| code-fold: true
#| label: define functions

get_og_stimuli <- function(i) {
  mean <- stimuli_df %>%
    filter(ogIndex == i) %>%
    pull(num.y) %>%
    mean(.)

  stimuli_df %>%
    filter(ogIndex == i) %>%
    ggplot(aes(x = num.x, y = num.y)) +
    geom_point() +
    coord_cartesian(ylim = c(0, 1)) +
    geom_hline(yintercept = mean, color = "gray", linetype = "dashed") +
    annotate("text", x = 120, y = mean, label = mean) +
    labs(title = paste0("ogIndex", " ", i))
}

# return visual angle in degrees and not radian
vis_angle <- function(size, distance) {
  return(2 * atan(size / (2 * distance)) * 180 / pi)
}

# define function for turning vis angle to phy size
phys_size <- function(va, distance) {
  return(2 * distance * tan(va * pi / 360))
}

# Code copied from Matt; this prevents us from having mistakes 
x_num_to_px = \(x, x_range = 120, px_range = 500) x / x_range * px_range
y_num_to_px = \(y, y_range = 1, px_range = 200) y / y_range * px_range

x_px_to_num = \(x, x_range = 120, px_range = 500) x * x_range / px_range
y_px_to_num = \(y, y_range = 1, px_range = 200) y * y_range / px_range

px_to_phys = \(px, px_per_mm) px / px_per_mm
phys_to_px = \(phys, px_per_mm) phys * px_per_mm

x_va_to_num = \(va, distance, px_per_mm, x_range = 120, px_range = 500) va |>
  phys_size(distance) |>
  phys_to_px(px_per_mm) |>
  x_px_to_num(x_range, px_range)

y_va_to_num = \(va, distance, px_per_mm, y_range = 1, px_range = 200) va |>
  phys_size(distance) |>
  phys_to_px(px_per_mm) |>
  y_px_to_num(y_range, px_range)
```

Let's also read in the data collected from pilot: 

:::{.callout-note}
## A note on translating between spaces 
The goal is to translate from the pixel space to the physical space, and then to the visual angle space. 

1. Data to Pixel: This should be given by the custom `VegaEmbed` component we built (directly fetching)
  - In this case, data is from $x \in [0, 120]$ and $y \in [0, 1]$ with the x-axis mapping to `500 px` and y-axis to `200 px`.
2. Pixel to Physical: need to know the pixel to MM ratio, which is *different per Participant*.
3. Physical to Vis Angle: need to know the distance to screen, again, which is *different per Participant*. 
:::

```{r}
#| code-fold: true
#| label: read in pilot df

raw_df <- read_csv(here::here("data", "part_2", "part2_pilot.csv")) %>%
  as_tibble(.)

pilot_df <- raw_df %>%
  select(
    participantId,
    trialId,
    trialOrder,
    responseId,
    answer,
    duration,
    parameters_taskid,
    parameters_params
  )

# visual angle related participant-level data
participant_df <- pilot_df %>%
  filter(!responseId %in% c("numericResponse", "pixelResponse")) %>%
  group_by(participantId) %>%
  filter(
    responseId %in%
      c(
        "prolificId",
        "pixelsPerMM",
        "dist-calibration-MM",
        "dist-calibration-CM",
        "ball-positions",
        "square-position"
      )
  ) %>%
  select(participantId, responseId, answer) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  # we will not use participants who have distance calibration < 250 mm
  mutate(distToScreen = as.numeric(`dist-calibration-MM`)) %>%
  filter(distToScreen >= 250) %>%
  mutate(
    pixelsPerMM = as.numeric(pixelsPerMM)
  )

# only keep people whose ans are somewhat useful
keep_id <- participant_df %>% pull(participantId)

# responses
resp_df <- pilot_df %>%
  filter(responseId %in% c("numericResponse", "pixelResponse")) %>%
  select(participantId, trialId, responseId, answer) %>%
  separate_wider_delim(
    col = trialId,
    delim = "_",
    names = c("type", NA, "index")
  ) %>%
  pivot_wider(names_from = "responseId", values_from = "answer") %>%
  mutate(
    numericResponse = as.numeric(numericResponse),
    pixelResponse = as.numeric(pixelResponse),
    index = as.integer(index)
  ) %>%
  rename(expIndex = index) %>%
  mutate(pixelResponse = 200 - pixelResponse + 5) %>%
  # only keep the ones that matter
  filter(participantId %in% keep_id) %>%
  left_join(participant_df) %>%
  rename(
    num.y = numericResponse,
    px.y = pixelResponse,
  ) %>%
  mutate(
    pixelsPerMM = as.numeric(pixelsPerMM),
    phy.y = px.y / pixelsPerMM,
    va.y = vis_angle(phy.y, distToScreen)
  )

participant_df <- resp_df %>%
  count(participantId, type) %>%
  select(-n) %>%
  left_join(participant_df) %>%
  mutate(type = ifelse(type == "point", type, "point_arc"))
```

And then we also calculate the answer dataframe `ans_df`. **Note that we are NOT using the same strategy that Mortiz et al.** used to calculate **error** and **normalized error**^[... which is, instead of calculating the actual mean across 60 points for each trial, for trials that share the same value for `seed` and `flip`, they share the same "default mean", i.e., the mean of the `point` type, instead of its actual mean.]. 

```{r}
#| code-fold: true
#| label: get ans df

ans_df <- stimuli_df |>
  summarise(
    num.mean.y = mean(num.y),
    num.median.y = median(num.y),
    .by = c(ogIndex, expIndex, type, flip, noise, seed)
  ) |>
  mutate(
    # num.ans.y = num.mean.y[type == "point"][[1]], # this is what Moritz did, and what we're not doing 
    num.ans.y = num.mean.y,
    .by = c(flip, noise, seed)
  )

```

Here are the processed dataframes: 

```{r}
participant_df
ans_df
resp_df %>% count(participantId, type)
```

## Look at participants' error

```{r}
#| code-fold: true

resp_df %>%
  select(participantId, type, expIndex, num.y) %>%
  mutate(type = ifelse(type == "pointArc", "point_arc", "point")) %>%
  left_join(ans_df, by = c("type", "expIndex")) %>%
  mutate(error = num.y - num.ans.y) %>%
  summarySEwithin(
    data = .,
    measurevar = "error",
    betweenvars = "type",
    withinvars = c("noise", "flip"),
    idvar = "participantId"
  ) %>%
  mutate(
    noise = factor(noise, levels = c("c(0, 0)", "c(0, 0.4)")),
    flip_lab = ifelse(flip == FALSE, "Noise in Upper", "Noise in Lower"),
    flip_lab = factor(flip_lab, levels = c("Noise in Upper", "Noise in Lower"))
  ) %>%
  ggplot(aes(
    y = noise,
    x = error,
    dist = "norm",
    arg1 = error,
    arg2 = se,
    fill = type
  )) + #uncertainty
  stat_dist_halfeye(
    show_interval = TRUE,
    position = position_dodge(width = 0.2),
    .width = 0.95,
    size = .5,
    alpha = .8
  ) +
  theme(legend.position = "top") +
  scale_fill_manual(values = c("#76B7B2", "#E15759")) + # take out "#59A14F"
  ylab("Variability") + #add x label
  xlab("Error") + #add y label
  facet_wrap(flip_lab ~ ., ncol = 1) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(y = "Variability", x = "True Error", title = "Pilot study (N = 6)")
```

## Read in fitted model for `project_to_y_axis`

Recall the we defined the following model for the "project from dot to axis" operator: 

$$
\begin{align*}
\texttt{error}_i &\sim \mathcal{N}(\beta_0 + u_{\text{PID[i]}}, \sigma_{i, \text{PID}[i]}) \\ 
\log(\sigma_{i, j}) &= \alpha_j + \log(\texttt{va.ans.x}_i) \\ 
\text{or:} \log(\sigma_{i, j}) &\sim \mathcal{N}(\mu_\alpha + \log(\texttt{va.ans.x}_i), \sigma_\alpha)\\ 
\alpha_j &\sim N(\mu_\alpha, \sigma_\alpha) \\ 
\mu_\alpha &\sim N(0, 1) \\ 
\sigma_\alpha &\sim \mathcal{N}^+(0, 1)
\end{align*}
$$

Here is the specification we used for fitting: 

```{r}
#| eval: false
#| label: brms specs

# Recall that we defined error = resp - ans
f <- bf(
  error_va ~ (1 | participantId),
  sigma ~ (1 | participantId) + offset(log(va.ans.x))
)

p <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 1), class = sd, group = participantId, lb = 0),
  prior(normal(0, 1), class = Intercept, dpar = "sigma"),
  prior(normal(0, 1), class = sd, group = participantId, dpar = sigma, lb = 0)
)
```

And here we load the actual fitted model: 

```{r}
model <- readRDS(here::here("analysis", "fitted_models", "task5.y.rds"))
model
```

:::{.callout-note}
### How do things map to each other?

```{r}
model %>% get_variables(.) %>% head(4)
```

- `b_Intercept` corresponds to $\beta_0$ 
- `b_sigma_Intercept` corresponds to $\alpha_0$
- `sd_participantId__Intercept` and `sd_participantId__sigma_Intercept` corresponds to $\tau_\mu$ and $\tau_\alpha$
- `r_participantId[<id>, Intercept]`, `r_participantId__sigma[<id>, Intercept]` correspond to $u_j$ and $v_j$ respectively. 
:::

## ... with specific fixed `pixelToMM` and `distToScreen` 

We will make the predictions for an average participant, conditioning on specific values of `pixelToMM` and `distToScreen`. 

**Note here that we are also implicitly making the assumption that the same participant saw both the `point` type condition and the `point_arc` type condition, which is NOT the case for the actual study.** 

### Strategy 1 --- Project to vertical line `x==0`

For a baseline strategy, we assume that the participant projects from the dot to the y-axis, and then performs some mental averaging. 

First, we calculate the visual angle values for all the stimuli. 
This is the route from `num -> px -> MM -> VA`. This will allow us to make predictions of error in the visual angle space. 

```{r}
# 5760 rows = 2 (plot type) * 2 (noise type) * 2 (flipped or not) *
#  12 repeated measures * 60 points per graph
new_data <- stimuli_df %>% 
  # assume that the value of pixelToMM and distToScreen is fixed
  mutate(pixelToMM = 5.12, distToScreen = 473) %>% 
  # calculate pixel values from numerical values
  mutate(
    px.x = x_num_to_px(num.x),
    px.y = y_num_to_px(num.y)
  ) %>%
  # calcuate physical values (in MM) from pixel values
  mutate(
    phy.x = px_to_phys(px.x, pixelToMM), 
    phy.y = px_to_phys(px.y, pixelToMM)
  ) %>%
  # calculate visual angle
  mutate(
    va.ans.x = vis_angle(phy.x, distToScreen),
    va.ans.y = vis_angle(phy.y, distToScreen)
  )
```

Get (thinned) posterior draws: 

```{r}
# 3000 iterations per chain * 4 chains = 12,000 rows
# if we thin by k, then we get 12,000 / k rows
draws <- model %>%
  spread_draws(b_Intercept, b_sigma_Intercept) %>%
  as_draws_df(.) %>%
  # if too slow we can also do more aggressive thinning 
  # by making k large
  thin_draws(thin = 10)
```

Then we go the prediction route: `VA -> mm -> px -> num`:
  
```{r}
pred <- draws %>%
  # every draw * every dot;
  # will result in 5760 rows * 3000 (thinned version) dots / row
  expand_grid(new_data) %>%
  mutate(
    # mu = b_Intercept,
    sigma = exp(b_sigma_Intercept + log(va.ans.x)),
    # va.pred.error = rnorm(n(), mean = mu, sd = sigma),
    va.pred.error = rnorm(n(), mean = b_Intercept, sd = sigma),
    va.pred.y = va.pred.error + va.ans.y
  ) %>%
  # turn vis angle back to numerical values
  mutate(
    num.pred.y = y_va_to_num(va.pred.y, distToScreen, pixelToMM)
  )
```

:::{.callout-note collapse="true"}
Let's say we're only interested in `ogIndex == 4`: 

```{r}
left <- pred %>%
  filter(ogIndex == 4) %>%
  group_by(.draw) %>%
  summarise(mean.y = mean(num.pred.y)) %>%
  ggplot(aes(y = mean.y)) +
  stat_halfeye() +
  coord_cartesian(ylim = c(0, 1))

right <- get_og_stimuli(4)

left + right
```

One could tell from this that the mean strategy is too precise. 
:::

Now we "average"^[**Averaging Strategy vs. Median Strategy**: “Equal number of points or line below and above. A strategy we did not anticipate was ensuring an equal number of points or line lengths above and below the judgment indicator. To indicate their responses, participants drug a line on the graph. By allowing participants to place a line directly on the graphs, participants could then easily count the number of points above and below the line. One participant simply wrote, “I tried to have the number of points above and below the line be approximately equal.” Unsurprisingly, this strategy was most common for participants who viewed graphs with points equally spaced along the x-axis (24%).” ([Moritz et al., 2023, p. 7](zotero://select/library/items/ZPTBTRIF)) ([pdf](zotero://open-pdf/library/items/AZRSTSWZ?page=7&annotation=4AENLNDJ))
] across the 60 dots for each trial: 

```{r}
avg_pred <- pred %>%
  group_by(ogIndex, expIndex, type, flip, noise, .draw) %>%
  summarise(
    # suppose participants can calculate mean
    num.pred.avg.y = mean(num.pred.y),
    # suppose participants use the median instead
    num.pred.median.y = median(num.pred.y),
  )
```

:::{.callout-note}
## How many rows? 
To calculate how many rows that `pred` has we need to look at these two other dataframes: 

- The `draws` dataframe has 3000 iterations / chain $\times$ 4 chains $\div k = 2$ = 6000 rows 
- The `new_data` dataframe has 96 trials $\times$ 60 dots per trial = 5760 rows
- Therefore `pred` is going to be 6000 $\times$ 5760 = 34560000 rows 

If we don't consider _rejection sampling_, and just calculate the average across all 60 dots for each trial, we will have 34560000/60 = 576000 rows for `avg_pred`. 
:::

And now we get the dataframe for plotting: 

```{r}
plot_df <- avg_pred %>%
  ungroup(.) %>%
  left_join(ans_df) %>%
  mutate(
    error.median = num.pred.median.y - num.ans.y,
    error.avg = num.pred.avg.y - num.ans.y,
    noise = factor(noise, levels = c("c(0, 0)", "c(0, 0.4)"))
  )
```

And now we plot the results: 

```{r}
#| code-fold: true
#| layout-ncol: 2
#| label: plot-check-strategy1

plot_df %>%
  group_by(type, noise, flip, .draw) %>%
  summarise(error = mean(error.avg)) %>%
  ggplot(aes(x = error, y = noise, fill = type)) +
  stat_halfeye(alpha = 0.8) +
  scale_fill_manual(values = c("#76B7B2", "#E15759")) +
  theme(legend.position = "top") +
  labs(
    title = "Posterior error (mean strategy)"
  ) +
  facet_wrap(~flip, ncol = 1) +
  geom_vline(xintercept = 0, linetype = "dashed")

plot_df %>%
  group_by(type, noise, flip, .draw) %>%
  summarise(error = mean(error.median)) %>%
  ggplot(aes(x = error, y = noise, fill = type)) +
  stat_halfeye(alpha = 0.8) +
  scale_fill_manual(values = c("#76B7B2", "#E15759")) +
  theme(legend.position = "top") +
  labs(
    title = "Posterior error (median strategy)"
  ) +
  facet_wrap(~flip, ncol = 1) +
  geom_vline(xintercept = 0, linetype = "dashed")
```

As one can see, the above posterior error has very small variance. Median has slightly larger variance, but the "normalized error" goes negative. 

---

What if the projected responses use some kind of sensor-fusion based weighting? 

This implies `draws` or `new_data`, but we would change generate `pred` differently: 

```{r}
avg_pred <- draws %>%
  # every draw * every dot;
  expand_grid(new_data) %>%
  mutate(
    sigma = exp(b_sigma_Intercept + log(va.ans.x)),
    va.pred.error = rnorm(n(), mean = b_Intercept, sd = sigma),
    va.pred.y = va.pred.error + va.ans.y,
    mse = b_Intercept^2 + sigma^2,
    w = 1 / mse
  ) %>%
  # this is typically the part done with avg_pred above
  group_by(ogIndex, expIndex, pixelToMM, distToScreen, type, flip, noise, .draw) %>%
  summarise(va.pred.y = weighted.mean(va.pred.y, w), .groups = "drop") %>%
  # turn vis angle back to numerical values
  mutate(
    num.pred.y = y_va_to_num(va.pred.y, pixelToMM, distToScreen)
  )
```

Now we get the data used for plotting: 

```{r}
plot_df <- avg_pred %>%
  ungroup(.) %>%
  left_join(ans_df) %>%
  mutate(
    error = num.pred.y - num.ans.y,
    noise = factor(noise, levels = c("c(0, 0)", "c(0, 0.4)"))
  )
```

```{r}
#| code-fold: true

plot_df %>%
  group_by(type, noise, flip, .draw) %>%
  summarise(error = mean(error)) %>%
  ggplot(aes(x = error, y = noise, fill = type)) +
  stat_halfeye(alpha = 0.8) +
  scale_fill_manual(values = c("#76B7B2", "#E15759")) +
  theme(legend.position = "top") +
  labs(
    title = "Posterior error (weighted mean)"
  ) +
  facet_wrap(~flip, ncol = 1) +
  geom_vline(xintercept = 0, linetype = "dashed")

```

:::{.callout-note collapse="true"}

We can also do the per-graph comparison: 

```{r}
left <- pred %>%
  filter(ogIndex == 4) %>%
  mutate(mse = b_Intercept^2 + sigma^2, w = 1 / mse) %>%
  group_by(.draw) %>%
  summarise(va.pred.y = weighted.mean(va.pred.y, w), .groups = "drop") %>%
  mutate(
    num.pred.y = y_va_to_num(va.pred.y, pixelToMM, distToScreen)
  ) %>%
  ggplot(aes(y = num.pred.y)) +
  stat_halfeye() +
  coord_cartesian(ylim = c(0, 1))

left + right
```

From this I think it's quite clear that the dots closer to the y-axis seems to have significant weighting power here ... 

:::


### Strategy 2 --- Project to vertical line where `xintercept == 60` and then `xintercept == 0`

Instead of trying to sum over the results projected to the y-axis, we could have a result projecting to the middle point, i.e., where everyone projects to $x = 60$, and then use that as an input to then project to the y-axis in order to potentially introduce more variance into the results. 

Again, assume that we have fixed values for `pixelToMM` and `distToScreen`: 

```{r}
# 5760 rows = 2 (plot type) * 2 (noise type) * 2 (flipped or not) *
# 12 repeated measures * 60 points per graph
new_data <- stimuli_df %>%
  mutate(
    pixelToMM = 5.12, 
    distToScreen = 473, 
    # physical distance between vertical line with 
    # xintercept == 60 and xintercept == 0
    phy.x_ref = x_num_to_px(60) %>% px_to_phys(pixelToMM)
  ) %>% 
  # calculate pixel values from numerical values
  # NEW: instead of using num.x we're using the following:
  mutate(
    px.x = x_num_to_px(abs(60 - num.x)),
    px.y = y_num_to_px(num.y)
  ) %>%
  # calcuate physical values (in MM) from pixel values
  mutate(
    phy.x = px_to_phys(px.x, pixelToMM),
    phy.y = px_to_phys(px.y, pixelToMM)
  ) %>%
  # calculate visual angle
  mutate(
    va.ans.x = vis_angle(phy.x, distToScreen),
    va.ans.y = vis_angle(phy.y, distToScreen)
  )
```

Create the posterior draws: 

```{r}
#| label: strategy 2 thin dr aws

rm(draws)

draws <- model %>%
  spread_draws(b_Intercept, b_sigma_Intercept) %>%
  as_draws_df(.) %>%
  thin_draws(thin = 10)
```


Combine the posterior with new data: 

```{r}
#| label: strategy2-project-twice
rm(pred)

# project to the middle axis
pred <- draws %>%
  # every draw * every dot;
  # will result in 5760 * 3000 (thinned version) row
  expand_grid(new_data) %>%
  mutate(
    # mu = b_Intercept,
    sigma = exp(b_sigma_Intercept + log(va.ans.x)),
    # va.pred.error = rnorm(n(), mean = mu, sd = sigma),
    va.pred.error = rnorm(n(), mean = b_Intercept, sd = sigma),
    va.pred.y = va.pred.error + va.ans.y
  ) %>%
  group_by(ogIndex, expIndex, type, flip, noise, .draw) %>%
  summarise(
    # assume participants use the mean for second projection
    va.mid_pred.y.avg = mean(va.pred.y),
    # assume participants use the median for second projection
    va.mid_pred.y.median = median(va.pred.y),
    .groups = "drop_last"
  )

# project from middle to the left axis
pred2 <- pred %>%
  left_join(draws) %>%
  mutate(
    # distance to y-axis is vis angle distance from x == 60
    sigma = exp(b_sigma_Intercept + log(vis_angle(phy.x_ref, distToScreen))),
    va.pred.error = rnorm(n(), mean = b_Intercept, sd = sigma),
    va.pred.y.avg = va.pred.error + va.mid_pred.y.avg,
    va.pred.y.median = va.pred.error + va.mid_pred.y.median
  ) %>%
  # turn vis angle back to numerical values
  mutate(
    num.pred.y.avg = y_va_to_num(va.pred.y.avg, distToScreen, pixelToMM), 
    num.pred.y.median = y_va_to_num(va.pred.y.median, distToScreen, pixelToMM),
  )
```

:::{.callout-note collapse="true"}
Check to see the posterior prediction for one graph (`ogIndex == 4`): 

```{r}
#| code-fold: true
#| layout-ncol: 2
#| label: strategy2-check-singlegraph

pred2 %>%
  filter(ogIndex == 4) %>%
  ggplot(aes(x = va.mid_pred.y.avg, y = "x == 60")) +
  stat_halfeye() +
  stat_halfeye(aes(x = va.pred.y.avg, y = "x == 0")) +
  coord_cartesian(xlim = c(2.2, 2.8)) +
  labs(
    title = "Mean strategy; project to x == 60, then to x == 0",
    x = "Visual Angle of Predicted Y value"
  )

pred2 %>%
  filter(ogIndex == 4) %>%
  ggplot(aes(x = va.mid_pred.y.median, y = "x == 60")) +
  stat_halfeye() +
  stat_halfeye(aes(x = va.pred.y.median, y = "x == 0")) +
  coord_cartesian(xlim = c(2.2, 2.8)) +
  labs(
    title = "Median strategy; project to x == 60, then to x == 0",
    x = "Visual Angle of Predicted Y value"
  )
```

:::

So we are indeed seeing more variations. The median at `x == 60` is just a bit wider than the mean at `x == 60` but when projected to `x == 0` they result in a similarly shaped distribution. 

:::{.column-margin}
What would this variation result in changes in the numerical space? 

```{r}
phys_size(0.25, distToScreen) * pixelToMM / 200
```

So 5% on the [0, 1] scale. 
:::

Let's also see what the average looks like across trials: 

```{r}
rm(avg_pred)

avg_pred <- pred2 %>%
  group_by(ogIndex, expIndex, type, flip, noise, .draw) %>%
  summarise(
    # suppose participants can calculate mean
    num.pred.avg.avg.y = mean(num.pred.y.avg),
    num.pred.avg.median.y = median(num.pred.y.median),
  )
```

Get the dataframe for plotting: 

```{r}
plot_df <- avg_pred %>%
  ungroup(.) %>%
  left_join(ans_df) %>%
  mutate(
    error.median = num.pred.avg.median.y - num.ans.y,
    error.avg = num.pred.avg.avg.y - num.ans.y,
  ) %>%
  group_by(type, noise, flip, .draw) %>%
  summarise(
    error.avg = mean(error.avg),
    error.median = mean(error.median),
  ) %>%
  mutate(noise = factor(noise, levels = c("c(0, 0)", "c(0, 0.4)")))
```

Plotting: 

```{r}
#| code-fold: true
#| layout-ncol: 2
#| label: strategy 2 plot check

plot_df %>%
  ggplot(aes(x = error.avg, y = noise, fill = type)) +
  stat_halfeye(alpha = 0.8) +
  scale_fill_manual(values = c("#76B7B2", "#E15759")) +
  labs(
    title = "Mean Strategy on twice projected value on y-axis",
    x = "Error"
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  facet_wrap(~flip, ncol = 1)

plot_df %>%
  ggplot(aes(x = error.median, y = noise, fill = type)) +
  stat_halfeye(alpha = 0.8) +
  scale_fill_manual(values = c("#76B7B2", "#E15759")) +
  labs(
    title = "Median Strategy on twice projected value on y-axis",
    x = "Error"
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  facet_wrap(~flip, ncol = 1)
```

---

Now let's also try the sensor-fusion approach here: 

```{r}
rm(pred)

# project to the middle axis
pred <- draws %>%
  # every draw * every dot;
  # will result in 5760 * 3000 (thinned version) row
  expand_grid(new_data) %>%
  mutate(
    # mu = b_Intercept,
    sigma = exp(b_sigma_Intercept + log(va.ans.x)),
    # va.pred.error = rnorm(n(), mean = mu, sd = sigma),
    va.pred.error = rnorm(n(), mean = b_Intercept, sd = sigma),
    va.pred.y = va.pred.error + va.ans.y,
    mse = b_Intercept^2 + sigma^2,
    w = 1 / mse
  ) %>%
  group_by(ogIndex, expIndex, type, flip, noise, .draw) %>%
  summarise(va.pred.y = weighted.mean(va.pred.y, w), .groups = "drop")

# project from middle to the left axis
pred2 <- pred %>%
  left_join(draws) %>%
  mutate(
    # distance to y-axis is vis angle distance from x == 60
    sigma = exp(b_sigma_Intercept + log(vis_angle(phy.x_ref, distToScreen))),
    va.pred.error = rnorm(n(), mean = b_Intercept, sd = sigma),
    va.pred.y = va.pred.error + va.pred.y,
    # no need to reweight because it's just projecting from one thing 
  ) %>%
  # turn vis angle back to numerical values
  mutate(
    num.pred.y = y_va_to_num(va.pred.y, distToScreen, pixelToMM)
  )
```

```{r}
plot_df <- pred2 %>%
  ungroup(.) %>%
  left_join(ans_df) %>%
  mutate(
    error = num.pred.y - num.ans.y,
  ) %>%
  group_by(type, noise, flip, .draw) %>%
  summarise(
    error.avg = mean(error),
  ) %>%
  mutate(noise = factor(noise, levels = c("c(0, 0)", "c(0, 0.4)")))
```

```{r}
#| code-fold: true 

plot_df %>%
  ggplot(aes(x = error.avg, y = noise, fill = type)) +
  stat_halfeye(alpha = 0.8) +
  scale_fill_manual(values = c("#76B7B2", "#E15759")) +
  labs(
    title = "Weighted Mean on twice projected value on y-axis",
    x = "Error"
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black")  +
  facet_wrap(~flip, ncol = 1)
```

## NEW -- What if we do truncated normal sampling? 

```{r}
rm(draws)

draws <- model %>%
  spread_draws(b_Intercept, b_sigma_Intercept) %>%
  as_draws_df(.) %>%
  thin_draws(thin = 10)
```

```{r}
# library("devtools")
# install_github("olafmersmann/truncnorm")
library(truncnorm)

pred <- draws %>%
  expand_grid(new_data) %>%
  mutate(
    sigma = exp(b_sigma_Intercept + log(va.ans.x)),

    # 1. Calculate analytical bounds for the error
    va.min = vis_angle(0, distToScreen),
    # Upper bound: when physical size is maxed out (200 / pixelToMM)
    va.max = vis_angle(200 / pixelToMM, distToScreen),

    error_lb = va.min - va.ans.y,
    error_ub = va.max - va.ans.y,

    # 2. Draw directly from the truncated distribution
    va.pred.error = rtruncnorm(
      n(),
      a = error_lb,
      b = error_ub,
      mean = b_Intercept,
      sd = sigma
    ),

    # 3. Proceed forward (guaranteed to fall in [0,1])
    va.pred.y = va.pred.error + va.ans.y,
    phy.pred.y = phys_size(va.pred.y, distToScreen),
    px.pred.y = phy.pred.y * pixelToMM,
    num.pred.y = px.pred.y / 200
  )
```

```{r}
avg_pred <- pred %>%
  group_by(ogIndex, expIndex, type, flip, noise, .draw) %>%
  summarise(
    # suppose participants can calculate mean
    num.pred.avg.y = mean(num.pred.y),
    # suppose participants use the median instead
    num.pred.median.y = median(num.pred.y),
  )
```

Get the plotting dataframe and plots: 

```{r}
plot_df <- avg_pred %>%
  ungroup(.) %>%
  left_join(ans_df) %>%
  mutate(
    error.median = num.pred.median.y - num.ans.y,
    error.avg = num.pred.avg.y - num.ans.y,
    noise = factor(noise, levels = c("c(0, 0)", "c(0, 0.4)"))
  )
```

```{r}
#| layout-ncol: 2
#| code-fold: true

plot_df %>%
  group_by(type, noise, flip, .draw) %>%
  summarise(error = mean(error.avg)) %>%
  ggplot(aes(x = error, y = noise, fill = type)) +
  stat_halfeye(alpha = 0.8) +
  scale_fill_manual(values = c("#76B7B2", "#E15759")) +
  theme(legend.position = "top") +
  labs(
    title = "Posterior normalized error (mean strategy)"
  ) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_wrap(~flip, ncol = 1)

plot_df %>%
  group_by(type, noise, flip, .draw) %>%
  summarise(error = mean(error.median)) %>%
  ggplot(aes(x = error, y = noise, fill = type)) +
  stat_halfeye(alpha = 0.8) +
  scale_fill_manual(values = c("#76B7B2", "#E15759")) +
  theme(legend.position = "top") +
  labs(
    title = "Posterior normalized error (median strategy)"
  ) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_wrap(~flip, ncol = 1)
```

The conclusion from this is that truncated normal does not save the day ... 

## NEW -- Add more randomness 

Doing it with random effect: 

```{r}
draws <- model %>%
  spread_draws(
    b_Intercept,
    b_sigma_Intercept,
    sd_participantId__Intercept,
    sd_participantId__sigma_Intercept
  ) %>%
  as_draws_df(.) %>%
  thin_draws(thin = 10) %>%
  mutate(
    mu = b_Intercept + rnorm(n(), mean = 0, sd = sd_participantId__Intercept),
    sigma = rnorm(
      n(),
      mean = b_sigma_Intercept,
      sd_participantId__sigma_Intercept
    )
  )
```

```{r}
draws <- model %>%
  spread_draws(
    b_Intercept,
    b_sigma_Intercept,
    sd_participantId__Intercept, # <- NEW
    sd_participantId__sigma_Intercept # <- NEW
  ) %>%
  as_draws_df(.) %>%
  thin_draws(thin = 4)

pred <- draws %>%
  expand_grid(new_data) %>%
  mutate(
    # simulate a new participant's offsets from the population distribution
    re_mu = rnorm(n(), mean = 0, sd = sd_participantId__Intercept),
    re_sigma = rnorm(n(), mean = 0, sd = sd_participantId__sigma_Intercept),
    mu = b_Intercept + re_mu,
    sigma = exp(b_sigma_Intercept + re_sigma + log(va.ans.x)),
    va.pred.error = rnorm(n(), mean = mu, sd = sigma),
    va.pred.y = va.pred.error + va.ans.y
  ) %>%
  # turn vis angle back to numerical values
  mutate(
    phy.pred.y = phys_size(va.pred.y, distToScreen),
    px.pred.y = phy.pred.y * pixelToMM,
    num.pred.y = px.pred.y / 200
  )

avg_pred <- pred %>%
  group_by(ogIndex, expIndex, type, flip, noise, .draw) %>%
  summarise(
    # suppose participants can calculate mean
    num.pred.avg.y = mean(num.pred.y),
    # suppose participants use the median instead
    num.pred.median.y = median(num.pred.y),
  )
```

Create the plotting dataframe: 

```{r}
plot_df <- avg_pred %>%
  ungroup(.) %>%
  left_join(ans_df) %>%
  mutate(
    error.median = num.pred.median.y - num.ans.y,
    norm.error.median = ifelse(
      flip == FALSE,
      num.pred.median.y - num.ans.y,
      num.ans.y - num.pred.median.y
    ),
    error.avg = num.pred.avg.y - num.ans.y,
    norm.error.avg = ifelse(
      flip == FALSE,
      num.pred.avg.y - num.ans.y,
      num.ans.y - num.pred.avg.y
    )
  ) %>%
  mutate(noise = factor(noise, levels = c("c(0, 0)", "c(0, 0.4)")))

plot_df %>%
  group_by(type, noise, .draw) %>%
  summarise(error = mean(norm.error.avg)) %>%
  ggplot(aes(x = error, y = noise, fill = type)) +
  stat_halfeye(alpha = 0.8) +
  scale_fill_manual(values = c("#76B7B2", "#E15759")) +
  theme(legend.position = "top") +
  labs(
    title = "Posterior normalized error (mean strategy)"
  )

plot_df %>%
  mutate(noise = factor(noise, levels = c("c(0, 0)", "c(0, 0.4)"))) %>%
  group_by(type, noise, .draw) %>%
  summarise(error = mean(norm.error.median)) %>%
  ggplot(aes(x = error, y = noise, fill = type)) +
  stat_halfeye(alpha = 0.8) +
  scale_fill_manual(values = c("#76B7B2", "#E15759")) +
  theme(legend.position = "top") +
  labs(
    title = "Posterior normalized error (median strategy)"
  )
```

Hmm maybe I did this wrong but why does it look so similar ...??? 

---

Recall that all the above is for a single participant with a particular value for `pixelToMM` and `distToScreen`. We should also be able to calculate it for all participants in the pilot. 

## Generate posterior predictions and average across participants  

Recall this is the list of participants we have from pilot: 

```{r}
participant_df
```

Let us first do more aggressive thinning: 

```{r}
#| label: thin posterior draws for all participants
#| code-fold: true

draws <- model %>%
  spread_draws(b_Intercept, b_sigma_Intercept) %>%
  as_draws_df(.) %>%
  thin_draws(thin = 20)
```
 

### Strategy 1 -- Project once to y-axis 

```{r}

# 6 participant * 48 trials per participant * 60 dots per trial
# 17280 rows
new_data <- participant_df %>%
  left_join(stimuli_df, relationship = "many-to-many") %>%
  # calculate pixel values from numerical values
  mutate(px.x = 500 / 120 * num.x, px.y = 200 * num.y) %>%
  # calcuate physical values (in MM) from pixel values
  mutate(phy.x = px.x / pixelsPerMM, phy.y = px.y / pixelsPerMM) %>%
  # calculate visual angle
  mutate(
    va.ans.x = vis_angle(phy.x, distToScreen),
    va.ans.y = vis_angle(phy.y, distToScreen)
  )

new_data
```

Build the posterior prediction: 
  
```{r}
rm(pred)

# 17280 rows * 600 draws per row
# 1036800
pred <- new_data %>%
  expand_grid(draws) %>%
  mutate(
    sigma = exp(b_sigma_Intercept + log(va.ans.x)),
    va.pred.error = rnorm(n(), mean = b_Intercept, sd = sigma),
    va.pred.y = va.pred.error + va.ans.y,
    # back-convert using this participant's pixelToMM
    phy.pred.y = phys_size(va.pred.y, distToScreen),
    px.pred.y = phy.pred.y * pixelToMM,
    num.pred.y = px.pred.y / 200
  )
```

Again, the above number of rows is obtained by participant (6) * trials/participant (48) * dots/trial (60) * draws/dot (600). First, we want to get a sense of average value per trial, that is, now the dataframe should have participant (6) * trials/participant (48) * draw/trial (600). 

To then average across participants mean that we have distinct trials (48 * 2) * draw/trial (600): 

```{r}
avg_pred <- pred %>%
  group_by(participantId, expIndex, type, flip, noise, .draw) %>%
  summarise(
    num.pred.avg.y = mean(num.pred.y),
    num.pred.median.y = median(num.pred.y)
  ) %>%
  group_by(expIndex, type, flip, noise, .draw) %>%
  summarise(
    num.pred.avg.y = mean(num.pred.avg.y),
    num.pred.median.y = mean(num.pred.median.y)
  )
```

The next step then is to get the data that is used to plot error: 

```{r}
plot_df <- avg_pred %>%
  ungroup(.) %>%
  left_join(ans_df) %>%
  mutate(
    error.median = num.pred.median.y - num.ans.y,
    error.avg = num.pred.avg.y - num.ans.y,
    noise = factor(noise, levels = c("c(0, 0)", "c(0, 0.4)"))
  )
```

Finally let's plot to see if we can get similar results: 

```{r}
#| code-fold: true
#| layout-ncol: 2

plot_df %>%
  group_by(type, noise, flip, .draw) %>%
  summarise(error = mean(error.avg)) %>%
  ggplot(aes(x = error, y = noise, fill = type)) +
  stat_halfeye(alpha = 0.8) +
  labs(
    title = "Across participants (mean)",
    x = "Error"
  ) +
  facet_wrap(~flip, ncol = 1) +
  geom_vline(xintercept = 0, linetype = "dashed")

plot_df %>%
  group_by(type, noise, flip, .draw) %>%
  summarise(error = mean(error.median)) %>%
  ggplot(aes(x = error, y = noise, fill = type)) +
  stat_halfeye(alpha = 0.8) +
  scale_fill_manual(values = c("#76B7B2", "#E15759")) +
  labs(
    title = "Across participants (median)",
    x = "Error"
  ) +
  facet_wrap(~flip, ncol = 1) +
  geom_vline(xintercept = 0, linetype = "dashed")

```

Hmm ...?? Why is this looking like this ...? Because of thinning ...?? 

---

Extend it with the weighting approach 




### Strategy 2 -- project twice 

This time we will build the posterior prediction a bit differently: 

```{r}
# 6 participant * 48 trials per participant * 60 dots per trial
# 17280 rows
new_data <- participant_df %>% 
  select(participantId, type, pixelsPerMM, distToScreen) %>% 
  left_join(stimuli_df, by = "type", relationship = "many-to-many") %>%
  # calculate pixel values from numerical values
  mutate(
    px.x = x_num_to_px(num.x),
    px.y = y_num_to_px(num.y)
  ) %>%
  # calcuate physical values (in MM) from pixel values
  mutate(
    phy.x = px_to_phys(px.x, pixelsPerMM), 
    phy.y = px_to_phys(px.y, pixelsPerMM)
  ) %>%
  # calculate visual angle
  mutate(
    va.ans.x = vis_angle(phy.x, distToScreen),
    va.ans.y = vis_angle(phy.y, distToScreen)
  )

# 17280 rows * 600 draws per row
# 1036800
pred <- new_data %>%
  expand_grid(draws) %>%
  mutate(
    sigma = exp(b_sigma_Intercept + log(va.ans.x)),
    va.pred.error = rnorm(n(), mean = b_Intercept, sd = sigma),
    # first projection to x == 60
    va.pred.y = va.pred.error + va.ans.y,
    mse = b_Intercept^2 + sigma^2,
    w = 1 / mse
  ) %>%
  group_by(participantId, distToScreen, pixelsPerMM, expIndex, type, flip, noise, .draw) %>%
  summarise(
    va.pred.y = weighted.mean(va.pred.y, w), .groups = "drop"
  )

# given that there is a horizontal line, we probably don't need to project again 

avg_pred <- pred %>% 
  mutate(
    num.pred.y = y_va_to_num(va.pred.y, distToScreen, pixelsPerMM)
  ) %>%
  group_by(expIndex, type, flip, noise, .draw) %>%
  summarise(
    # suppose participants can calculate mean
    num.pred.y = mean(num.pred.y),
    # suppose participants use the median instead
    # num.pred.median.y = median(num.pred.y),
  )



# # project from middle to left axis
# pred2 <- pred %>%
#   left_join(draws) %>%
#   mutate(
#     phy.x_ref = x_num_to_px(60) %>% px_to_phys(pixelsPerMM), 
#     # distance to y-axis is vis angle distance from x == 60
#     sigma = exp(b_sigma_Intercept + log(vis_angle(phy.x_ref, distToScreen))),
#     va.pred.error = rnorm(n(), mean = b_Intercept, sd = sigma),
#     va.pred.y = va.pred.error + va.pred.y, 
#   ) %>%
#   # turn vis angle back to numerical values
#   mutate(
#     num.pred.y = y_va_to_num(va.pred.y, distToScreen, pixelsPerMM)
#   )
```

Then get dataframe used for plotting error: 

```{r}
plot_df <- avg_pred %>% # pred2 %>%
  ungroup(.) %>%
  left_join(ans_df) %>%
  mutate(
    error = num.pred.y - num.ans.y,
  ) %>%
  group_by(type, noise, flip, .draw) %>%
  summarise(
    error.avg = mean(error),
  ) %>%
  mutate(noise = factor(noise, levels = c("c(0, 0)", "c(0, 0.4)")))
```

Then plot: 
```{r}
#| code-fold: true

plot_df %>%
  ggplot(aes(x = error.avg, y = noise, fill = type)) +
  stat_halfeye(alpha = 0.8) +
  scale_fill_manual(values = c("#76B7B2", "#E15759")) +
  labs(
    title = "Weighted Mean Strategy | Twice Projected | All pilot participants",
    x = "Error"
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  facet_wrap(~flip, ncol = 1)
```
