---
title: "Extend from current data to Moritz"
author: "Anonymous"
date: today
date-format: "iso"
format:
  html:
    reference-location: margin
---

```{r}
#| code-fold: true

# read in libraries 
library(ggplot2)
library(tidyverse)
library(tidybayes)
library(here)
library(jsonlite)
library(brms)
library(posterior)
theme_set(theme_minimal())
```

# Replicating Mortiz Exp 2

Here are the details of the original experiment: 

- **Hypothesis:** “the effect of variability will be smaller for graphs with points spaced along the x-axis than graphs with points spaced along the arc” ([Moritz et al., 2023, p. 5](zotero://select/library/items/ZPTBTRIF)) ([pdf](zotero://open-pdf/library/items/AZRSTSWZ?page=5&annotation=4IPHWX8I))
- **Experiment Design**:
  - *Within Subject*: 
    - 2 variability: `0` and `0.4`
    - 2 flipped: `upper` and `lower` (i.e., whether the plot was flipped or not, where `flip == FALSE` implies that upper-y region has more variability, and `flip == TRUE` implies that the lower-y region has more variability)
  - *Between Subject*: 
    - 3 mark type: `line`, `cartesian spaced points`, `arc spaced points`
      - Note that we will only use cartesian spaced points and arc spaced points for replication.
  - 12 different seeds for randomization -> `2 * 2 * 12 = 48` trials per participant.

## Read in Stimuli Data

First, read in the stimuli data from Mortiz experiment 2: 

```{r}
exp_json_data <- jsonlite::fromJSON(here(
  "data",
  "moritz",
  "stimuli",
  "exp2-stimuli.json"
)) %>%
  as_tibble(.)

stimuli_df <- exp_json_data %>%
  select(index, seed, noise, flip, type, data) %>%
  mutate(noise = as.character(noise))

# get the mapping between og index and the index used in exp
stimuli_df %>%
  filter(type == "point") %>%
  mutate(expIndex = 1:nrow(.)) %>%
  select(index, type, expIndex) -> point_map
stimuli_df %>%
  filter(type == "point_arc") %>%
  mutate(expIndex = 1:nrow(.)) %>%
  select(index, type, expIndex) -> pointArc_map

# calculate the answer
stimuli_df %>%
  filter(type != "line") %>%
  unnest(data) %>%
  left_join(point_map, by = join_by(index, type)) %>%
  left_join(pointArc_map, by = join_by(index, type)) %>%
  mutate(expIndex = coalesce(expIndex.x, expIndex.y)) %>%
  select(index, expIndex, type, flip, noise, seed, x, y) %>%
  ungroup(.) %>%
  rename(ogIndex = index, num.x = x, num.y = y) -> stimuli_df

stimuli_df %>% head(5)
```

**TODO**: Double check whether `data` is the correct column we should be pulling and rendering from ... I think it is correct but one can never be too careful with these things ... 

```{r}
#| eval: false
#| echo: false
#| code-fold: true

# there are three columns, data, lineData, and highNoiseData.
# when the type is `point`, data is equivalent to lineData.
# when the type if `pointArc`,  data is NOT equivalent to lineData.

check <- exp_json_data %>%
  unnest(c(data, lineData, highNoiseData), names_repair = "unique") %>%
  mutate(
    dataEqLineData = `x...6` == `x...8` & `y...7` == `y...9`,
    dataEqHighNoisedata = `x...6` == `x...10` & `y...7` == `y...11`,
    LineDataEqHighNoiseData = `x...8` == `x...10` & `y...9` == `y...11`
  ) %>%
  filter(type != "line") %>%
  mutate(noise = as.character(noise))

check %>%
  count(
    noise,
    type,
    dataEqLineData,
    dataEqHighNoisedata,
    LineDataEqHighNoiseData
  )

check %>%
  filter(dataEqLineData == TRUE, dataEqHighNoisedata == FALSE) %>%
  filter(index == 4) # %>%
ggplot(aes(x = `x...10`, y = `y...11`)) +
  geom_point() +
  coord_cartesian(ylim = c(0, 1))
```

```{r}
# function definitions
get_og_stimuli <- function(i) {
  mean <- stimuli_df %>%
    filter(ogIndex == i) %>%
    pull(num.y) %>%
    mean(.)

  stimuli_df %>%
    filter(ogIndex == i) %>%
    ggplot(aes(x = num.x, y = num.y)) +
    geom_point() +
    coord_cartesian(ylim = c(0, 1)) +
    geom_hline(yintercept = mean, color = "gray", linetype = "dashed") +
    annotate("text", x = 120, y = mean, label = mean) +
    labs(title = paste0("ogIndex", " ", i))
}
```

Let's also read in the data collected from pilot. First we get participant's numerical and pixel responses: 

```{r}
raw_df <- read_csv(here("data", "part_2", "part2_pilot.csv")) %>% as_tibble(.)

pilot_df <- raw_df %>%
  select(
    participantId,
    trialId,
    trialOrder,
    responseId,
    answer,
    duration,
    parameters_taskid,
    parameters_params
  )

# responses
pilot_df %>%
  filter(responseId %in% c("numericResponse", "pixelResponse")) %>%
  select(participantId, trialId, responseId, answer) %>%
  separate_wider_delim(
    col = trialId,
    delim = "_",
    names = c("type", NA, "index")
  ) %>%
  pivot_wider(names_from = "responseId", values_from = "answer") %>%
  mutate(
    numericResponse = as.numeric(numericResponse),
    pixelResponse = as.numeric(pixelResponse),
    index = as.integer(index)
  ) %>%
  rename(expIndex = index) %>%
  mutate(pixelResponse = 200 - pixelResponse + 5) -> resp_df

resp_df %>% head(5)
```

Then we also get a dataset that documents how far away they're sitting from the screen: 

```{r}
# visual angle related
pilot_df %>%
  filter(!responseId %in% c("numericResponse", "pixelResponse")) %>%
  group_by(participantId) %>%
  filter(
    responseId %in%
      c(
        "prolificId",
        "pixelsPerMM",
        "dist-calibration-MM",
        "dist-calibration-CM",
        "ball-positions",
        "square-position"
      )
  ) %>%
  select(participantId, responseId, answer) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  # we will not use participants who have distance calibration < 250 mm
  mutate(`dist-calibration-MM` = as.numeric(`dist-calibration-MM`)) %>%
  filter(`dist-calibration-MM` >= 250) %>%
  select(1, 3, 4) %>%
  rename(distToScreen = `dist-calibration-MM`) -> participant_df

# only keep people whose ans are somewhat useful
keep_id <- participant_df %>% pull(participantId)

participant_df %>% head(5)
```

## Translate btw spaces

The goal is to translate from the pixel space to the physical space, and then to the visual angle space. 

1. Data to Pixel: This should be given by the custom `VegaEmbed` component we built (directly fetching)
  - In this case, data is from $x \in [0, 120]$ and $y \in [0, 1]$ with the x-axis mapping to `500 px` and y-axis to `200 px`.
2. Pixel to Physical: need to know the pixel to MM ratio, which is *different per Participant*.
3. Physical to Vis Angle: need to know the distance to screen, again, which is *different per Participant*. 

```{r}
#| label: define-function
#| code-fold: true

# return visual angle in degrees and not radian
vis_angle <- function(size, distance) {
  return(2 * atan(size / (2 * distance)) * 180 / pi)
}

# define function for turning vis angle to phy size
phys_size <- function(va, distance) {
  return(2 * distance * tan(va * pi / 360))
}

# tolerance for numerical precision
tolerance <- 1e-10
```

Process the response df given the new info about participants: 

```{r}
resp_df <- resp_df %>%
  filter(participantId %in% keep_id) %>%
  left_join(participant_df) %>%
  rename(
    num.y = numericResponse,
    px.y = pixelResponse,
  ) %>%
  mutate(
    pixelsPerMM = as.numeric(pixelsPerMM),
    phy.y = px.y / pixelsPerMM,
    va.y = vis_angle(phy.y, distToScreen)
  )
resp_df %>% head(5)
```


## Read in fitted model

We assume that the baseline model is this: participants project from each of the dots to the y-axis, and then perform an averaging. We will use the **Project to Y-Axis Model**. 
Recall that we defined `error = resp - ans`. 

For an average participant, the posterior prediction is going to
- define $\log(\sigma_{i, j}) = \mu_\alpha + \log(\texttt{va.ans.x}_i) $ 
- then $\text{error}_i \sim \mathcal{N}(0, \sigma_{i, j})$ where $\sigma_{i, j}$ is already defined above.  

$$
\begin{align*}
\texttt{error}_i &\sim \mathcal{N}(\beta_0 + u_{\text{PID[i]}}, \sigma_{i, \text{PID}[i]}) \\ 
\log(\sigma_{i, j}) &= \alpha_j + \log(\texttt{va.ans.x}_i) \\ 
\text{or:} \log(\sigma_{i, j}) &\sim \mathcal{N}(\mu_\alpha + \log(\texttt{va.ans.x}_i), \sigma_\alpha)\\ 
\alpha_j &\sim N(\mu_\alpha, \sigma_\alpha) \\ 
\mu_\alpha &\sim N(0, 1) \\ 
\sigma_\alpha &\sim \mathcal{N}^+(0, 1)
\end{align*}
$$

Read in the fitted model by first reading in its formula and prior specification: 

```{r}
f <- bf(
  error_va ~ (1 | participantId),
  sigma ~ (1 | participantId) + offset(log(va.ans.x))
)

p <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 1), class = sd, group = participantId, lb = 0),
  prior(normal(0, 1), class = Intercept, dpar = "sigma"),
  prior(normal(0, 1), class = sd, group = participantId, dpar = sigma, lb = 0)
)
```

Load the previously fitted model: 

```{r}
model <- readRDS(here("analysis", "fitted_models", "task5.y.rds"))
model
```

### How do things map to each other?

```{r}
model %>% get_variables(.) %>% head(4)
```

- `b_Intercept` corresponds to $\beta_0$ 
- `b_sigma_Intercept` corresponds to $\alpha_0$
- `sd_participantId__Intercept` and `sd_participantId__sigma_Intercept` corresponds to $\tau_\mu$ and $\tau_\alpha$
- `r_participantId[<id>, Intercept]`, `r_participantId__sigma[<id>, Intercept]` correspond to $u_j$ and $v_j$ respectively. 

## Generate Predictions 

### For an "average" participant with fixed `pixelToMM` and `distToScreen`. 

We will make the predictions for an average participant, conditioning on specific values of `pixelToMM` and `distToScreen`. 

First, we calculate the visual angle values for all the stimuli. 
This is the route from `num -> px -> MM -> VA`. This will allow us to make predictions of error in the visual angle space. 

```{r}
# assume that the value of pixelToMM and distToScreen is fixed
# we get the new_data to be crossed with posterior draws

pixelToMM <- 5.12
distToScreen <- 473

# 5760 rows = 2 (plot type) * 2 (noise type) * 2 (flipped or not) * 12 repeated measures * 60 points per graph
new_data <- stimuli_df %>%
  # calculate pixel values from numerical values
  mutate(px.x = 500 / 120 * num.x, px.y = 200 * num.y) %>%
  # calcuate physical values (in MM) from pixel values
  mutate(phy.x = px.x / pixelToMM, phy.y = px.y / pixelToMM) %>%
  # calculate visual angle
  mutate(
    va.ans.x = vis_angle(phy.x, distToScreen),
    va.ans.y = vis_angle(phy.y, distToScreen)
  )

# get a smaller version of this
# consider only one kind of graph type
nd <- new_data %>%
  select(ogIndex, expIndex, type, va.ans.x, va.ans.y) %>%
  filter(type == "point")

# 3000 iterations per chain * 4 chains = 12,000 rows
# if we thin by k, then we get 12,000 / k rows
draws <- model %>%
  spread_draws(b_Intercept, b_sigma_Intercept) %>%
  as_draws_df(.) %>%
  thin_draws(thin = 2) # if too much we can also do thinning
```
  
Here we go from `VA -> mm -> px -> num`:
  
```{r}
pred <- draws %>%
  # every draw * every dot; will result in 5760 * 3000 (thinned version) row
  expand_grid(new_data) %>% 
  mutate(
    mu = b_Intercept,
    sigma = exp(b_sigma_Intercept + log(va.ans.x)),
    va.pred.error = rnorm(n(), mean = mu, sd = sigma),
    va.pred.y = va.pred.error + va.ans.y
  ) %>%
  # turn vis angle back to numerical values
  mutate(
    phy.pred.y = phys_size(va.pred.y, distToScreen),
    px.pred.y = phy.pred.y * pixelToMM,
    num.pred.y = px.pred.y / 200
  )

pred %>% head(5)
```

To calculate how many rows that `pred` has we need to look at these two other dataframes: 

- The `draws` dataframe has 3000 iterations / chain $\times$ 4 chains $\div k = 2$ = 6000 rows 
- The `new_data` dataframe has 96 trials $\times$ 60 dots per trial = 5760 rows
- Therefore `pred` is going to be 6000 $\times$ 5760 = 34560000 rows 

---

For now we will not think about things such as rejection sampling, and just calculate the average across all 60 dots for each trial^[This implies that we wil have 34560000/60 = 576000 rows for `avg_pred`]:  

```{r}
avg_pred <- pred %>%
  group_by(ogIndex, expIndex, type, flip, noise, .draw) %>%
  summarise(
    num.pred.avg.y = mean(num.pred.y), # suppose participants can calculate mean
    num.pred.median.y = median(num.pred.y), # suppor participants use the median instead
  )
```

:::{.callout-note}
**Averaging Strategy vs. Median Strategy**: 

“Equal number of points or line below and above. A strategy we did not anticipate was ensuring an equal number of points or line lengths above and below the judgment indicator. To indicate their responses, participants drug a line on the graph. By allowing participants to place a line directly on the graphs, participants could then easily count the number of points above and below the line. One participant simply wrote, “I tried to have the number of points above and below the line be approximately equal.” Unsurprisingly, this strategy was most common for participants who viewed graphs with points equally spaced along the x-axis (24%).” ([Moritz et al., 2023, p. 7](zotero://select/library/items/ZPTBTRIF)) ([pdf](zotero://open-pdf/library/items/AZRSTSWZ?page=7&annotation=4AENLNDJ))
:::

Using the same strategy that Moritz et al. used to calculate **error** and **normalized error**^[... which is, instead of calculating the actual mean for the actual dataset, for trials that share the same value for `seed` and `flip`, they share the same "default mean", i.e., the mean of the `point` type, instead of its actual mean.], we need to first calculate the answer for each trial: 

```{r}
#| code-fold: true 

temp_df <- stimuli_df %>% filter(type == "point") %>% 
  group_by(seed, flip) %>% 
  summarise(num.ans.y = mean(num.y)) %>%
  ungroup(.)

ans_df <- stimuli_df %>% select(-num.x, -num.y) %>% 
  group_by(ogIndex, expIndex, type, flip, noise, seed) %>% 
  summarise(.) %>% ungroup(.) %>% 
  left_join(temp_df)

ans_df
```


I think this is the very reasonable next step here: 

```{r}
#| code-fold: true 

avg_pred %>%
  ungroup(.) %>%
  left_join(ans_df) %>%
  mutate(
    error.median = num.pred.median.y - num.ans.y,
    norm.error.median = ifelse(flip == FALSE, 
      num.pred.median.y - num.ans.y, 
      num.ans.y - num.pred.median.y
    ),
    error.avg = num.pred.avg.y - num.ans.y,
    norm.error.avg = ifelse(flip == FALSE, 
      num.pred.avg.y - num.ans.y, 
      num.ans.y - num.pred.avg.y
    )
  ) %>%
  group_by(type, noise, .draw) %>% 
  summarise(error = mean(norm.error.avg)) %>% 
  ggplot(aes(x = error, y = noise, fill = type)) +
  stat_halfeye(alpha = 0.8) +
  labs(title = "Posterior error when an average participant use mean strategy, in data units")
```

