---
title: "Extend from current data to Moritz"
---

```{r}
#| code-fold: true
library(ggplot2)
library(tidyverse)
library(tidybayes)
library(here)
library(jsonlite)
library(brms)
```

# Replicating Mortiz Exp 2

Here are the details of the original experiment: 

- **Hypothesis:** “the effect of variability will be smaller for graphs with points spaced along the x-axis than graphs with points spaced along the arc” ([Moritz et al., 2023, p. 5](zotero://select/library/items/ZPTBTRIF)) ([pdf](zotero://open-pdf/library/items/AZRSTSWZ?page=5&annotation=4IPHWX8I))
- **Experiment Design**:
  - 2 variability: `0` and `0.4`
  - 2 flipped: `upper` and `lower` (i.e., whether the plot was flipped or not)
  - 3 mark type: `line`, `cartesian spaced points`, `arc spaced points` -> between-subjects
    - In our experiment we will only use cartesian spaced points and arc spaced points. 
  - 12 different seeds -> `2 * 2 * 12 = 48` trials per participant.


## Read in Stimuli Data

First, read in the stimuli data from Mortiz experiment 2: 

```{r}
exp_json_data <- jsonlite::fromJSON(here("data", "moritz", "stimuli", "exp2-stimuli.json" )) %>% as_tibble(.)

stimuli_df <- exp_json_data %>% select(index, seed, noise, flip, type, data) %>% 
    mutate(noise = as.character(noise))

stimuli_df %>% head(5)

# get the mapping between og index and the index used in exp
stimuli_df %>% filter(type == "point") %>% 
  mutate(expIndex = 1:nrow(.)) %>% 
  select(index, type, expIndex) -> point_map
stimuli_df %>% filter(type == "point_arc") %>% 
  mutate(expIndex = 1:nrow(.)) %>% 
  select(index, type, expIndex) -> pointArc_map

# calculate the answer
stimuli_df %>% filter(type != "line") %>% 
  unnest(data) %>% 
  group_by(index, type) %>% 
  summarise(y.avg = mean(y)) %>% 
  left_join(point_map, by = join_by(index, type)) %>% 
  left_join(pointArc_map, by = join_by(index, type)) %>% 
  mutate(expIndex = coalesce(expIndex.x, expIndex.y)) %>% 
  select(index, expIndex, type, y.avg) %>% ungroup(.) -> ans_df
```

Let's also read in the data collected from pilot: 

```{r}
raw_df <- read_csv(here("data", "part_2", "part2_pilot.csv")) %>% as_tibble(.) 

pilot_df <- raw_df %>% select(participantId, trialId, trialOrder, responseId, answer, duration, parameters_taskid, parameters_params)

# responses
pilot_df %>% filter(responseId %in% c("numericResponse", "pixelResponse")) %>%
  select(participantId, trialId, responseId, answer) %>%
  separate_wider_delim(col = trialId, delim = "_", names = c("type", NA, "index")) %>% 
  pivot_wider(names_from = "responseId", values_from = "answer") %>% 
  mutate(numericResponse = as.numeric(numericResponse), 
    pixelResponse = as.numeric(pixelResponse), 
    index = as.integer(index)
  ) %>% 
  rename(expIndex = index) %>% 
  mutate(pixelResponse = 200 - pixelResponse + 5) %>%
  left_join(ans_df) -> resp_df 

resp_df %>% head(5)
```

```{r}
# visual angle related 
pilot_df %>% filter(!responseId %in% c("numericResponse", "pixelResponse")) %>% 
  group_by(participantId) %>% 
  filter(responseId %in% c("prolificId", "pixelsPerMM", "dist-calibration-MM", "dist-calibration-CM", "ball-positions", "square-position")) %>% 
  select(participantId, responseId, answer) %>% 
  pivot_wider(names_from = responseId, values_from = answer) %>% 
  # we will not use participants who have distance calibration < 250 mm 
  mutate(`dist-calibration-MM` = as.numeric(`dist-calibration-MM`)) %>% 
  filter(`dist-calibration-MM` >= 250) %>% 
  select(1, 3, 4) %>% 
  rename(distToScreen = `dist-calibration-MM`) -> participant_df
  
# only keep people whose ans are somewhat useful
keep_id <- participant_df %>% pull(participantId)
```

## Translate btw spaces

The goal is to translate from the pixel space to the physical space, and then to the visual angle space. 

1. Data to Pixel: This should be given by the custom `VegaEmbed` component we built (directly fetching)
2. Pixel to Physical: need to know the pixel to MM ratio
3. Physical to Vis Angle: need to know the distance to screen

```{r}
# y-axis: numerical [0, 1] -> pixel [0, 200] -> phys [...?] 
# for some reason, the pixel response seems to be on the 210 scale and not the 200 scale ...? 
```

```{r}
#| label: define-function
#| code-fold: true

# origin is top left  
# data_to_pixel_y <- function(data_y) {
#   return(-395 * data_y + 410)
# }
# data_to_pixel_x <- function(data_x) {
#   return (53.5 * data_x + 317.5)
# }

# origin is bottom left 
# pixel_to_phy_x <- function(pixel, pxMM){
#   (pixel - 50) / pxMM
# }
# pixel_to_phy_y <- function(pixel, pxMM){
#   (410 - pixel) / pxMM
# }

# return visual angle in degrees and not radian
vis_angle <- function(size, distance){
  return(2 * atan(size / (2 * distance)) * 180 / pi)
}

# tolerance for numerical precision 
tolerance <- 1e-10
```

Process the response df given the new info about participants: 

```{r}
resp_df <- resp_df %>% filter(participantId %in% keep_id) %>% 
  left_join(participant_df) %>% 
  rename(num.y = numericResponse, 
    px.y = pixelResponse, 
    ans.num.y = y.avg,
    ogIndex = index
  ) %>% mutate(
    pixelsPerMM = as.numeric(pixelsPerMM),
    phy.y = px.y / pixelsPerMM,
    va.y = vis_angle(phy.y, distToScreen)
  ) 
resp_df %>% head(5)
```


## Read in fitted model

We assume that the baseline model is this: participants project from each of the dots to the y-axis, and then perform an averaging. We will use the **Project to Y-Axis Model**. 
Recall that we defined `error = resp - ans`. 

For an average participant, the posterior prediction is going to
- define $\log(\sigma_{i, j}) = \mu_\alpha + \log(\texttt{va.ans.x}_i) $ 
- then $\text{error}_i \sim \mathcal{N}(0, \sigma_{i, j})$ where $\sigma_{i, j}$ is already defined above.  

$$
\begin{align*}
\texttt{error}_i &\sim \mathcal{N}(0, \sigma_{i, \text{PID}[i]}) \\ 
\log(\sigma_{i, j}) &= \alpha_j + \log(\texttt{va.ans.x}_i) \\ 
\alpha_j &\sim N(\mu_\alpha, \sigma_\alpha) \\ 
\mu_\alpha &\sim N(0, 1) \\ 
\sigma_\alpha &\sim \mathcal{N}^+(0, 1)
\end{align*}
$$

Read in the fitted model by first reading in its formula and prior specification: 

```{r}
f <- bf(error_va ~ (1 | participantId), 
        sigma ~ (1|participantId) + offset(log(va.ans.x)))

p <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 1), class = sd, group = participantId, lb = 0), 
  prior(normal(0, 1), class = Intercept, dpar = "sigma"), 
  prior(normal(0, 1), class = sd, group = participantId, dpar = sigma, lb = 0)
)
```

```{r}
model %>% get_variables(.)
```

```{r}
# model <- brm(
#     formula = f,
#     data = .,
#     family = gaussian(),
#     prior = p,
#     chains = 4, 
#     cores = 4,
#     control = list(adapt_delta = 0.99),
#     iter = 6000, 
#     file = here("analysis", "fitted_models", "task5.y"), 
#     file_refit = "on_change", 
#     save_pars = save_pars(all = TRUE)
# )
model <- readRDS(here("analysis", "fitted_models", "task5.y.rds"))
model
```

```{r}
draws_df <- tidy_draws(model) %>% select(.chain, .iteration, .draw, b_Intercept, b_sigma_Intercept) 
draws_df %>% head(5)
```

```{r}
# assume we have fixed pixelToMM
pixelToMM <- 5.12
screenDist <- 473

stimuli_1_df <- stimuli_df %>% filter(type != "line") %>% 
  rename(ogIndex = index) %>% 
  unnest(col = data) %>% 
  rename(ans.num.x = x, ans.num.y = y) %>%
  # calculate pixel values from numerical values
  mutate(ans.px.x = 500 / 120 * ans.num.x,
         ans.px.y = 200 * ans.num.y) %>% 
  # calcuate physical values (in MM) from pixel values
  mutate(ans.phy.x = ans.px.x / pixelToMM, 
         ans.phy.y = ans.px.y / pixelToMM
  ) %>% 
  # calculate visual angle 
  mutate(va.ans.x = vis_angle(ans.phy.x, screenDist),
         va.ans.y = vis_angle(ans.phy.y, screenDist)
  ) %>% 
  select(ogIndex, va.ans.x, va.ans.y)



model %>% add_epred_draws(newdata = filter(stimuli_1_df, ogIndex == 4), dpar="sigma", re_formula = NA, 
allow_new_levels = TRUE) %>% 
  ggplot(data = ., aes(x = exp(.epred))) + 
  stat_halfeye()
```

# What does an average participant see? 

Assume the pixels to MM ratio is `3.25`. 

```{r}
# know which one it's making back to 
pixelToMM <- 3.25
screenDist <- 612

stimuli_df <- stimuli_df %>% mutate(expIndex = 1:nrow(.))
ex_df <- stimuli_df %>% filter(index == 4) %>% unnest(col = data) %>% 
  rename(numX = x, numY = y) %>%
  # calculate pixel values from numerical values
  mutate(pxX = 500 / 120 * numX,
         pxY = 200 * numY) %>% 
  # calcuate physical values (in MM) from pixel values
  mutate(phyX = pxX / pixelToMM, 
         phyY = pxY / pixelToMM
  ) %>%
  # calculate visual angle 
  mutate(va.ans.x = vis_angle(phyX, screenDist),
         va.ans.y = vis_angle(phyY, screenDist)
  )

ex_df
```

What would the posterior prediction look like? 

```{r}
# define function for turning vis angle to phy size 
phys_size <- function(va, distance) {
  return(2 * distance * tan(va * pi / 360))
}
```

```{r}
ex_df %>% select(index, va.ans.x, va.ans.y) %>% 
  mutate(participantId = "extend") %>% 
  tidybayes::add_predicted_draws(newdata = ., model, re_formula = NA) %>% 
  mutate(va.pred.y = va.ans.y + .prediction) %>% 
  group_by(.draw) %>%
  summarise(va.pred.y.avg = mean(va.pred.y)) %>%
  mutate(phy.pred.y.avg = phys_size(va.pred.y.avg, screenDist), 
    px.pred.y.avg = phy.pred.y.avg * pixelToMM, 
    num.pred.y.avg = px.pred.y.avg / 200
  ) -> pred_df

pred_df
```

```{r}
pred_df %>% median_qi(num.pred.y.avg) 
```

Let's also try to plot this uncertainty: 

```{r}
library(ggplot2)
library(tidybayes)
library(ggdist)

actual_mean <- ex_df %>% pull(numY) %>% mean(.)

pred_df %>% ggplot(data = ., aes(y = num.pred.y.avg)) +
  stat_halfeye() + 
  theme_minimal() + 
  geom_hline(yintercept = actual_mean, color = "red", linetype = "dashed") # + 
  # coord_cartesian(ylim = c(0, 1))
```

What about the actual answers? 

```{r}
pilot_df %>% filter(responseId %in% c("numericResponse", "pixelResponse")) %>%
  select(participantId, trialId, responseId, answer) %>%
  separate_wider_delim(col = trialId, delim = "_", names = c("taskType", NA, "index")) %>% 
  pivot_wider(names_from = "responseId", values_from = "answer") %>% 
  mutate(numericResponse = as.numeric(numericResponse), 
    pixelResponse = as.numeric(pixelResponse), 
    index = as.integer(index)
  ) %>% 
  mutate(pixelResponse = 200 - pixelResponse + 5) %>% 
  filter(index == 1)
```

Huh this is indeed quite close to the 