---
title: "Average Estimates in Line Graphs are Biased Towards Variability"
author: "Anonymous"
date: "June 28th, 2023"
output:
  pdf_document:
    toc: yes
  html_document:
    fig_height: 6
    fig_width: 8
    highlight: pygments
    theme: yeti
    toc: yes
    toc_float: yes
    self_contained: yes
  word_document:
    toc: yes
always_allow_html: yes
---
<br>


This is the code used for the analysis reported in 'Average Estimates in Line Graphs are Biased Towards Variability'.

# Data wrangling

Load packages used and functions:

```{r packages, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse) # Data wrangling/pipe (%>%)
require(tidybayes) # Visualizing model output
require(kableExtra) # Make nice tables
require(lme4) # Regression analysis 
require(irr) #Inter-Rater Reliability with Coehn's Kappa
source("Functions.R") # Calculating CI for mixed designs
```

Reading in data from Exp 1 and 2

```{r data}
#Exp 1
exp1 <- read.csv("data/exp1.csv", check.names=FALSE) #main study
exp1_demog <- read.csv("data/exp1-demographics.csv", check.names=FALSE) #exp 1 demographics

#Exp 2
exp2_line <- read.csv("data/exp2-line.csv", check.names=FALSE) #line
exp2_dotArc <- read.csv("data/exp2-point-arc.csv", check.names=FALSE) #point arc
exp2_dotX <- read.csv("data/exp2-point.csv", check.names=FALSE) #point
exp2_demog <- read.csv("data/exp2-demographics.csv", check.names=FALSE) #exp 2 demographics

#strategy codes
strategies <- read.csv("data/strategies.csv", check.names=FALSE)
```

Combining conditions in Exp 2

```{r combine_exp2}
exp2 <- rbind(exp2_line, exp2_dotArc, exp2_dotX)
```

Fixing the naming and coding from various conditions  

```{r naming}
#fixing variability coding for Exp 1
exp1$variability <- 
  factor(ifelse(exp1$variability=="0,0.15", "0.15",
         ifelse(exp1$variability=="0,0.4", "0.4", 999)))


#fixing variability coding for Exp 2
exp2$variability <- 
  factor(ifelse(exp2$variability=="0,0", "0",
         ifelse(exp2$variability=="0,0.4", "0.4", 999)))

#fixing the flip naming for Exp 1
exp1$flipped <- factor(ifelse(exp1$flip=="TRUE", "Flipped", "Unflipped"))
exp1$flipped <- factor(exp1$flipped, levels=c("Unflipped", "Flipped"))

#fixing the flip naming for Exp 2
exp2$flipped <- factor(ifelse(exp2$flip=="TRUE", "Flipped", "Unflipped"))
exp2$flipped <- factor(exp2$flipped, levels=c("Unflipped", "Flipped"))

#fixing the overestimation naming for Exp 1-2
exp1$estimation <- factor(ifelse(exp1$overestimated=="TRUE", "overestimated", "underestimated"))
exp2$estimation <- factor(ifelse(exp2$overestimated=="TRUE", "overestimated", "underestimated"))

#adding biased towards variability code
exp1$bias_towards_variability <- ifelse(exp1$normalized_error > "0", "True", "False")
exp2$bias_towards_variability <- ifelse(exp2$normalized_error > "0", "True", "False")
```


# Analyses

## Exp 1

### Descriptive Statistics

We now perform the analyses for Experiment 1.

Look at the demographics data first

```{r demographics exp1}
# Age
exp1_demog$age <- as.numeric(exp1_demog$age) # Make numeric
range(exp1_demog$age) # Range
round(mean(exp1_demog$age), 0) # Mean
round(sd(exp1_demog$age), 0) # SD

# Gender
(xtab <- table(exp1_demog$gender)) #Raw stats
round(prop.table(xtab), 3) * 100 # Proportions (in order)
```

We now create a plot showing the descriptive statistics for over and underestimation 

```{r Descriptive Plot}
#creating a theme
line_theme <- function() {
  theme(axis.text.y = element_blank(), 
      axis.title.y = element_blank(), 
      legend.position="top", 
      legend.justification="left",
      legend.direction="horizontal",
      legend.title = element_blank(), 
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(), 
      strip.background = element_blank(),
      strip.text.x = element_text(size = 15, face ="bold", hjust = 0))
    }

#plot 
expDesc1 <- ggplot(exp1, aes(estimation, fill = estimation))+
  geom_bar()+ #adding bar
  coord_flip()+ #flipping axis
  facet_wrap('flipped', ncol = 1)+ #breaking up by flip
  geom_text(stat='count', aes(label=..count..), hjust=-.15)+ #adding count text
  theme(legend.position="top")+ #moving legend
  scale_fill_manual(values=c("#4e79a7", "#F28e2b"))+ #color
  scale_y_continuous(breaks=seq(0, 3000, by = 500), expand = expansion(mult = c(0, 0.12)))+ #scale
  theme_classic() + #theme
  guides(y = "none")+ #removing guides
  line_theme() # adding theme
expDesc1 #print plot

#export plot
ggsave(plot=expDesc1,height=2.5,width=4.2,dpi=200, filename="plots/expDesc1.pdf", useDingbats=FALSE)
```

### Statistical Analysis

Regression Analysis for Exp 1

```{r Regression exp1}
#ordering the factors
exp1$flipped <- factor(exp1$flipped, levels=c("Unflipped", "Flipped"))

#Pre-registered regression model (updated from lm)
m.exp1.1 = lmer(error ~ variability*flipped + (1 | assignmentid), data=exp1)

#approximating coefficients
c.exp1.1 <- cbind(data.frame(coef(summary(m.exp1.1))) %>% #summary of coefficients 
  mutate(p.z = 2 * (1 - pnorm(abs(t.value)))), #p-value estimation 
  as_tibble(confint(m.exp1.1, method="Wald", level = 0.95))[-c(1:2),]) #CIs 


#creating a table printout 
knitr::kable(c.exp1.1,  
             align = "l",
             booktabs = T, 
             linesep = "", 
             digits = 3)%>%
    row_spec(0, bold = T)%>%
    kable_styling(latex_options = "striped", font_size = 12)
```

Test of H1: estimation error will be significantly different than zero. The results revealed a significant intercept of the model (b = -0.036, t(6811) = -6.6, p < .001, 95% CI [-.047, -.025]), providing evidence that the absolute estimation error (3.6%) for the referent conditions was meaningfully different than zero (supporting H1).

Test of H2: significantly more estimation error for trials with higher variability. The results also revealed a significant main effect of variability (b = -.022, t(6811) = -6.5, p < .001, 95% CI [-.029, -.016]). This finding supports H2, suggesting significantly more estimation error for trials with higher variability than lower variability

Test of H3: estimation error will occur in the direction of the increased variability. There was also a significant interaction between variability .15 vs .4 and variability upper vs. lower (b = .034, t(6811) = 7, p < .001, 95% CI [.025, .044]). To unpack the interaction, we ran the same model above but with the variability in the lower y-value graphs as the referent.


```{r}
#changing levels 
exp1$flipped <- factor(exp1$flipped, levels=c("Flipped", "Unflipped"))

#Pre-registered regression model (updated from lm)
m.exp1.2 = lmer(error ~ variability*flipped + (1|assignmentid), data=exp1)

#approximating coefficients
c.exp1.2 <- cbind(data.frame(coef(summary(m.exp1.2))) %>% #summary of coefficients 
  mutate(p.z = 2 * (1 - pnorm(abs(t.value)))), #p-value estimation 
  as_tibble(confint(m.exp1.2, method="Wald", level = 0.95))[-c(1:2),]) #CIs 


#creating a table printout 
knitr::kable(c.exp1.2,  
             align = "l",
             booktabs = T, 
             linesep = "", 
             digits = 3)%>%
    row_spec(0, bold = T)%>%
    kable_styling(latex_options = "striped", font_size = 12)

```

This model yielded a significant effect of variability but in the opposite direction (b = .012, t(6811) = 3.39, p = .001, 95% CI [.005, .018]) compared to the prior model (b = -.022).  We found positive errors when the area of highest variability was above the average y-value and negative errors when the highest variability was lower on the y-axis than the average.

### Plots
Summery statistics for the plot 
```{r}
freq_sum <- summarySEwithin(data = exp1, measurevar = "error", #data and dependent measure
                withinvars = c("variability", "flip"), #within subject variables 
                idvar = "assignmentid") #participant id
```

```{r}
#plot
plot1_Exp1_freq<- freq_sum %>%
  ggplot(aes(x = variability, y = error,  dist = "norm", arg1 = error, arg2 = se, fill = flip)) + #data and uncertainty
  stat_dist_halfeye(show_interval = TRUE, position = position_dodge(width = 0.2), .width = 0.95, size = .5)+ #halfeye
  coord_flip()+ #flip axis
  theme_tidybayes()+ #add theme
  scale_fill_manual(values=c("#EDC948", "#B07AA1"))+ #color
  scale_y_continuous(breaks=seq(-.9, .08, by = .02), expand = expansion(mult = c(.1, 0.1))) #scale
plot1_Exp1_freq #print plot

#export PDF
#ggsave(plot=plot1_Exp1_freq,height=4,width=7,dpi=200, filename="plots/Exp1Freq.pdf", useDingbats=FALSE)
```

### Replication of Xiong et al., 2020
Here we check for an overall bias toward the bottom of the screen. We will check if the absolute difference from 0 is larger for Flipped stimui compared to the Unflipped stimui.   

```{r}

xionglm = lm(absolute_error ~ flipped, data=exp1)
summary(xionglm)
```

The lack of significant difference between groups (p = .058) suggests we do not have sufficient evidence to conclude that significantly more people underestimated in the low condition (flipped) than those who overestimated in the high condition (unflipped). 

## Exp 2

We now perform the analyses for Experiment 2.

### Descriptive Statistics

First we look at the demographics data.

```{r demographics exp2}
# Age
exp2_demog$age <- as.numeric(exp2_demog$age) # Make numeric
range(exp2_demog$age) # Range
round(mean(exp2_demog$age), 0) # Mean
round(sd(exp2_demog$age), 0) # SD

# Gender
(xtab <- table(exp2_demog$gender)) #Raw stats
round(prop.table(xtab), 3) * 100 # Proportions (in order)
```

Plot

```{r}
#specifying levels 
exp2$mark <- factor(exp2$mark, levels=c("point_arc", "line", "point"))

#plot 
exp2Desc1 <- exp2%>%
  filter(bias_towards_variability == "True")%>% #only show overestimation 
  ggplot(aes(bias_towards_variability))+ #x value
  geom_bar()+ #add bar
  coord_flip()+ #flip axis
  facet_wrap('mark', ncol = 1)+ #break up by mark type
  geom_text(stat='count', aes(label=..count..), hjust=-.2)+ #add count text
  theme(legend.position="top")+ #move legend
  scale_y_continuous(breaks=seq(0, 7000, by = 1000), expand = expansion(mult = c(0, 0.12)))+ #scale
  geom_hline(yintercept = 6720)+ #add horizontal line indicating the number of judgement 
  theme_classic() + #theme
  guides(y = "none")+ #remove guides
  line_theme() #add theme
exp2Desc1 #print plot

#export plot
ggsave(plot=exp2Desc1,height=2.5,width=4.2,dpi=200, filename="plots/exp2Desc1.pdf", useDingbats=FALSE)
```

### Statistical Analysis

Regression Analysis for Exp 2

```{r Exp 2}
#converting mark to factor
exp2$mark<- as.factor(exp2$mark)

#ordering mark type
exp2$mark <- factor(exp2$mark, levels = c("line", "point", "point_arc"))

#Pre-registered regression model
m.exp2.1 = lmer(normalized_error ~ mark*variability + (1 | assignmentid), data=exp2)

#approximating coefficients
c.exp2.1 <- cbind(data.frame(coef(summary(m.exp2.1))) %>% #summary of coefficients 
  mutate(p.z = 2 * (1 - pnorm(abs(t.value)))), #p-value estimation 
  as_tibble(confint(m.exp2.1, method="Wald", level = 0.95))[-c(1:2),]) #CIs 

#creating a table printout 
knitr::kable(c.exp2.1,  
             align = "l",
             booktabs = T, 
             linesep = "", 
             digits = 3)%>%
    row_spec(0, bold = T)%>%
    kable_styling(latex_options = "striped", font_size = 12)
```


Replicating Experiment 1, the model results revealed a main effect of variability (b = 0.03, t(20153) = 10.027, p < .00, 95% CI
[0.024, 0.036]), indicating that graphs with more variability had greater estimation error.

Our results revealed the change from low to high variability was meaningfully larger for Line vs. Point (b = -0.023, t(20153) = -5.451, p < .00, 95% CI [-0.015,-0.031]).

To compare Point Arc vs. Point, we changed the referent of the model in the following analysis. 

```{r}
#ordering mark type
exp2$mark <- factor(exp2$mark, levels = c("point", "point_arc", "line"))

#Pre-registered regression model
m.exp2.2 = lmer(normalized_error ~ mark*variability + (1 | assignmentid), data=exp2)

#approximating coefficients
c.exp2.2 <- cbind(data.frame(coef(summary(m.exp2.2))) %>% #summary of coefficients 
  mutate(p.z = 2 * (1 - pnorm(abs(t.value)))), #p-value estimation 
  as_tibble(confint(m.exp2.1, method="Wald", level = 0.95))[-c(1:2),]) #CIs 

#creating a table printout 
knitr::kable(c.exp2.2,  
             align = "l",
             booktabs = T, 
             linesep = "", 
             digits = 3)%>%
    row_spec(0, bold = T)%>%
    kable_styling(latex_options = "striped", font_size = 12)
```

The change in normalized error from low to high variability was also meaningfully larger for Point Arc vs. Point (b = 0.05, t(20153) =
10.742, p < .00, 95% CI [-0.031	-0.015]).

We conducted a follow-up regression analysis to determine if the small increase in normalized estimation error was meaningful for Point.

Point only model
```{r}
#subsetting data for just point
point <- filter(exp2, mark == "point") #making a data frame with point data

#model
 m.exp2.point = lm(normalized_error ~ variability, data=point) #regression equation with point data
 summary(m.exp2.point) #print summary 
```

This analysis revealed a meaningful but small bias for Point (b = 0.007, t(6718) = 2.243, p = .025). 

### Plot 
First we visualize the main effect of variability, collapsed across mark type

```{r}
#calculating summery statistics 
freq_sum <- summarySEwithin(data = exp2, #data
                measurevar = "normalized_error", #dependent variable 
                withinvars = "variability", #within subjects variable
                idvar = "assignmentid") #participant id

#plot
plot1_Exp2_freq_sum<- freq_sum %>% #data
  ggplot(aes(x = variability, y = normalized_error,  dist = "norm", arg1 = normalized_error, arg2 = se)) + #uncertainty
  stat_dist_halfeye(show_interval = TRUE, position = position_dodge(width = 0.2), .width = 0.95, size = 4, alpha = .7)+ #adding halfeye
  coord_flip()+ #flipping axis
  theme_tidybayes()+ # adding theme
  xlab("Variability") + #adding x label
  ylab("Normalized Error")+  #adding y label
  ylim(c(0, .06)) #specifying y limits
plot1_Exp2_freq_sum #print plot

#export plot
ggsave(plot=plot1_Exp2_freq_sum,height=4,width=7,dpi=200, filename="plots/Exp2V2Freq_sum.pdf", useDingbats=FALSE)
```
Next we visualize the effect of variability for each mark type

```{r}
#calculating summery statistics 
freq_sum <- summarySEwithin(data = exp2, #data
                measurevar = "normalized_error",  #dependent variable 
                withinvars = "variability", #within subjects variable
                betweenvars = "mark", #between subject variable 
                idvar = "assignmentid") #participant id

#plot
plot1_Exp2_freq<- freq_sum %>% #data
  ggplot(aes(x = variability, y = normalized_error,  dist = "norm", arg1 = normalized_error, arg2 = se, fill = mark))+ #uncertainty
  stat_dist_halfeye(show_interval = TRUE, position = position_dodge(width = 0.2), .width = 0.95, size = .5, alpha = .8)+ #adding halfeye
  coord_flip()+ #flipping axis
  theme_tidybayes()+ #theme
  scale_fill_manual(values=c("#76B7B2", "#E15759", "#59A14F"))+ #adding color
  scale_y_continuous(breaks=seq(0, .11, by = .01), expand = expansion(mult = c(0.01, 0.01)))+ #specifying scale
  xlab("Variability") + #add x label
  ylab("Normalized Error") #add y label
plot1_Exp2_freq # print plot

#export PDF
ggsave(plot=plot1_Exp2_freq,height=6,width=7,dpi=200, filename="plots/Exp2Freq.pdf", useDingbats=FALSE)
```


# Open Responses for Exp 1 and 2 

First, we calculate an inter-rater reliability score for the strategies.

```{r IRR}
#mentioned using high or low points
kappa2(strategies[,c("r1_highlow","r2_highlow")],weight = "unweighted")
#Kappa = 0.884 

#did some type of mental averaging
kappa2(strategies[,c("r1_mental_averaging","r2_mental_averaging")],weight = "unweighted")
#Kappa = 0.83

#mentioned using the begin or end points
kappa2(strategies[,c("r1_start_end","r2_start_end")],weight = "unweighted")
#Kappa = 0.828 

#considered variability in some capacity
kappa2(strategies[,c("r1_variability","r2_variability")],weight = "unweighted")
#Kappa = 0.697 

#equal number of points or line below and above the average line
kappa2(strategies[,c("r1_number_of_points","r2_number_of_points")],weight = "unweighted")
#Kappa = 0.87 

#equal area above and below the line
kappa2(strategies[,c("r1_equal_area","r2_equal_area")],weight = "unweighted")
#Kappa = 0.889 

#didn't know their strategy 
kappa2(strategies[,c("r1_unknown","r2_unknown")],weight = "unweighted")
#Kappa = 0.844 

#other/na
kappa2(strategies[,c("r1_other_na","r2_other_na")],weight = "unweighted")
#Kappa = 0.893 
```

Table of counts of those who mentioned using high or low points
```{r} 
strategies %>%
  dplyr::mutate(highlow = (r1_highlow+r2_highlow)/2) %>%
  dplyr::group_by(Exp, Condition, highlow) %>%
  dplyr::count()
```

Table of counts those who mentioned using mental averaging
```{r} 
strategies %>%
  dplyr::mutate(mental_averaging = (r1_mental_averaging+r2_mental_averaging)/2) %>%
  dplyr::group_by(Exp, Condition, mental_averaging) %>%
  dplyr::count()
```

Table of counts those who mentioned using beginning or ending points
```{r} 
strategies %>%
  dplyr::mutate(start_end = (r1_start_end+r1_start_end)/2) %>%
  dplyr::group_by(Exp, Condition, start_end) %>%
  dplyr::count()
```

Table of counts those who mentioned considering variability
```{r} 
strategies %>%
  dplyr::mutate(variability = (r1_variability+r2_variability)/2) %>%
  dplyr::group_by(Exp, Condition, variability) %>%
  dplyr::count()
```

Table of counts those who mentioned making sure there were an equal number of points or line below and above the average line
```{r} 
strategies %>%
  dplyr::mutate(number_of_points = (r1_number_of_points+r2_number_of_points)/2) %>%
  dplyr::group_by(Exp, Condition, number_of_points) %>%
  dplyr::count()
```

Table of counts those who mentioned making sure there were an equal area below and above the average line

```{r} 
strategies %>%
  dplyr::mutate(equal_area = (r1_equal_area+r2_equal_area)/2) %>%
  dplyr::group_by(Exp, Condition, equal_area) %>%
  dplyr::count()
```


Participants who used the correct strategy 

```{r}
correct = c("32Q90QCQ1TLWMDCPHF1CL11UI7EEKK", "336YQZE83WERGIJZRQLPWKNLSMR5M0", "337RC3OW062O8UGTX52N052V5TFVLJ", "33CUSNVVNOCVIB0XVNI93MWH8ZJ88F", "33F859I567D7DKZX13WP71FHJ37BH3", "34YB12FSQZOPWF8J8M4G0HEAIYNGMH", "37Q970SNZF8VQVVLA0FZ539A62NS1C", "39DD6S19JQBRB8BBHG4M8GP7IRCEZ9", "39RP059MEITT5Y08S0C2Z6ND2BUBMA", "3CN4LGXD5YO9EG4DWOQQLN4EZT64Y0", "3EICBYG645WMEVMT33V5UY1EJP8CJ0", "3ERMJ6L4DZS63MXIFJ0MU41HHJV7M6", "3H8DHMCCWABRU7YPLKQSFJJ9EP3DK0", "3HUTX6F6VVNNHOLOMNY4QH9E1BNO21", "3IFS6Q0HJJJ6Q1RJI705X2QWSG6ISV", "3IRIK4HM3BKAQ07MJK1GLUW9WGH6CA", "3LEP4MGT3H0Q14UVF7DZGG0JRVSDBE", "3QJOXOW4XKRRC1CS2RES0ETY6UXMEK", "3ZPBJO59KQ10SHXHFZNV2R3HQUMDHV", "3ZSY5X72NYB4L8291DD5RPHMFDPROR")
```


Creating separate data frames for participants who reported correct and incorrect strategies 

```{r}
#code for not in
"%ni%" <- Negate("%in%")

#Filtering
exp2.wrong = exp2 %>%
  dplyr::filter(assignmentid %ni% correct)

exp2.correct = exp2 %>%
  dplyr::filter(assignmentid %in% correct)
```

Calculating summery statistics for each group

```{r}
freq_sum.wrong <- summarySEwithin(data = exp2.wrong, #data
                measurevar = "normalized_error",  #dependent variable 
                withinvars = "variability", #within subjects variable
                betweenvars = "mark", #between subject variable 
                idvar = "assignmentid") #participant id

freq_sum.correct <- summarySEwithin(data = exp2.correct, #data
                measurevar = "normalized_error",  #dependent variable 
                withinvars = "variability", #within subjects variable
                betweenvars = "mark", #between subject variable 
                idvar = "assignmentid") #participant id


freq_sum.correct.id <- summarySEwithin(data = exp2.correct, #data
                measurevar = "normalized_error",  #dependent variable 
                withinvars = "variability", #within subjects variable
                betweenvars = c("mark","assignmentid"), #between subject variable 
                idvar = "assignmentid") #participant id
```


Plot comparing strategies which was cleaned up in illustrator 

```{r, warning=FALSE}
good<-  ggplot()+ #uncertainty
   stat_dist_halfeye(data = freq_sum.wrong, aes(x = 1, y = normalized_error,  dist = "norm", arg1 = normalized_error, arg2 = se, fill = mark, color=mark), show_interval = TRUE, position = position_dodge(width = 1.8), .width = 0.95, size = .5, alpha = .9)+
  
   stat_dist_halfeye(data = freq_sum.correct.id, aes(x = 1, y = normalized_error,  dist = "norm", arg1 = normalized_error, arg2 = se, fill = mark), show_interval = TRUE, position = position_dodge(width = 1.5), .width = 0.95, size = .5, alpha = .35)+
    scale_y_continuous(limits = c(-.1, .15), breaks=seq(-.1, 5, by = .02), expand = expansion(mult = c(0.0, 0.01)))+ #specifying
    facet_grid(variability~.)+
    coord_flip()+ #flipping axis
    theme_tidybayes()+ #theme
    scale_fill_manual(values=c("#76B7B2", "#E15759", "#59A14F"))+ #adding color
    scale_color_manual(values=c("#76B7B2", "#E15759", "#59A14F"))+ #adding color
    xlab("Variability") + #add x label
    ylab("Normalized Error")+ #add y label
    theme(axis.text.y = element_blank(), 
      axis.title.y = element_blank())
good 

#export PDF
#ggsave(plot=good,height=8,width=8.5,dpi=200, filename="plots/Exp2_goodcompv1.pdf", useDingbats=FALSE)
```

# Sensitivity Analysis

## People who guessed the exp 

```{r}
#People who guess the study in Exp 1
Exp2_guessed = c("3LEP4MGT3H0Q14UVF7DZGG0JRVSDBE
", "35K3O9HUACDL6RKBG4AF5O9P67CEFL", "3HHRAGRYX95YH86N7WDJ04HS7DSO9G", "3P59JYT76MK3UGQWI4K3KRF10QRT2T", 
"37Q970SNZF8VQVVLA0FZ539A62NS1C", "39DD6S19JQBRB8BBHG4M8GP7IRCEZ9", "3VAR3R6G1Q1Y33ND247508MHOP78O9", "3TU5ZICBRE11OF0THHZ9PDHH4BJ8QI", "32Q90QCQ1TLWMDCPHF1CL11UI7EEKK", "3IFS6Q0HJJJ6Q1RJI705X2QWSG6ISV")

#Filtering
exp2.guess = exp2 %>%
  dplyr::filter(assignmentid %in% Exp2_guessed)

exp2.noguess = exp2 %>%
  dplyr::filter(assignmentid %ni% Exp2_guessed)
```

Rerun regression equations without those people in Exp 2 

```{r Regression exp1 guess}
#ordering the factors
exp2.noguess$flipped <- factor(exp2.noguess$flipped, levels=c("Unflipped", "Flipped"))

#Pre-registered regression model
m.exp1.1.noguess = lm(error ~ variability*flipped, data=exp2.noguess)

#approximating coefficients
c.exp1.1.noguess <- cbind(data.frame(coef(summary(m.exp1.1.noguess ))) %>% #summary of coefficients 
  mutate(p.z = 2 * (1 - pnorm(abs(t.value)))), #p-value estimation 
  as_tibble(confint(m.exp1.1.noguess , method="Wald", level = 0.95))) #CIs 

#creating a table printout 
knitr::kable(c.exp1.1.noguess,
             align = "l",
             booktabs = T, 
             linesep = "", 
             digits = 3)%>%
    row_spec(0, bold = T)%>%
    kable_styling(latex_options = "striped", font_size = 12)
```

Plot with and without those who guessed the purpose of the study. The density plots show the original data with everyone included. The quantile dotplot shows the data without those who guessed the purpose of the study. 

```{r}
#calculating summery statistics 
freq_sum.noguess <- summarySEwithin(data = exp2.noguess, #data
                measurevar = "normalized_error",  #dependent variable 
                withinvars = "variability", #within subjects variable
                betweenvars = "mark", #between subject variable 
                idvar = "assignmentid") #participant id

freq_sum <- summarySEwithin(data = exp2, #data
                measurevar = "normalized_error",  #dependent variable 
                withinvars = "variability", #within subjects variable
                betweenvars = "mark", #between subject variable 
                idvar = "assignmentid") #participant id

#ordering the factors
freq_sum.noguess$variability <- factor(freq_sum.noguess$variability, levels=c("0.4", "0"))
freq_sum$variability <- factor(freq_sum$variability, levels=c("0.4", "0"))

#plot
sensitivity<-  ggplot()+ #uncertainty
   stat_dist_dotsinterval(data = freq_sum.noguess, aes(x = 1, y = normalized_error,  dist = "norm", arg1 = normalized_error, arg2 = se, color = mark), show_interval = TRUE, position = position_dodge(width = -1.3), .width = 0.95, size = .5, alpha = .9)+
  stat_dist_halfeye(data = freq_sum, aes(x = 1, y = normalized_error,  dist = "norm", arg1 = normalized_error, arg2 = se, fill = mark), show_interval = TRUE, position = position_dodge(width = -1.5), .width = 0.95, size = .5, alpha = .7)+
   scale_y_continuous(breaks=seq(0, .11, by = .01), expand = expansion(mult = c(0.01, 0.01)))+ #specifying
    facet_grid(variability~.)+
    coord_flip()+ #flipping axis
    theme_tidybayes()+ #theme
    scale_fill_manual(values=c("#76B7B2", "#E15759", "#59A14F"))+ #adding color
    scale_color_manual(values=c("#76B7B2", "#E15759", "#59A14F"))+ #adding color
    xlab("Variability") + #add x label
    ylab("Normalized Error")+ #add y label
    theme(axis.text.y = element_blank(), 
      axis.title.y = element_blank())
sensitivity 

#export PDF
#ggsave(plot=sensitivity,height=6,width=7,dpi=200, filename="plots/Exp2sensitivity.pdf", useDingbats=FALSE)
```
