---
title: "New Pilot with combined components "
author: "Anonymous"
date: today
date-format: "iso"
format:
  html:
    reference-location: margin
---

```{r}
#| code-fold: true

library(plyr)
source(here::here("analysis", "moritz", "Functions.R"))
library(tidyverse)
library(jsonlite)
library(ggplot2)
library(tidybayes)
library(brms)
library(posterior)
library(purrr)
theme_set(theme_tidybayes())
bayesplot::bayesplot_theme_update(text = element_text(family = "sans"))
```

Read in stimuli data for experiment 2 from Moritz et al.: 

```{r}
exp_json_data <- jsonlite::fromJSON(here::here(
  "data",
  "moritz",
  "stimuli",
  "exp2-stimuli.json"
)) %>%
  as_tibble(.)

stimuli_df <- exp_json_data %>%
  select(index, seed, noise, flip, type, data) %>%
  mutate(noise = as.character(noise))

# get the mapping between og index and the index used in exp
stimuli_df %>%
  filter(type == "point") %>%
  mutate(expIndex = 1:nrow(.)) %>%
  select(index, type, expIndex) -> point_map
stimuli_df %>%
  filter(type == "point_arc") %>%
  mutate(expIndex = 1:nrow(.)) %>%
  select(index, type, expIndex) -> pointArc_map

# get the answer
part2_ans_df <- stimuli_df %>%
  filter(type != "line") %>%
  unnest(col = data) %>%
  rename(num.x = x, num.y = y, ogIndex = index) %>%
  summarise(
    num.mean.y = mean(num.y),
    num.median.y = median(num.y),
    .by = c(ogIndex, type, flip, noise, seed)
  ) |>
  mutate(
    # this is what Moritz did, and what we're not doing
    # num.ans.y = num.mean.y[type == "point"][[1]],
    num.ans.y = num.mean.y,
    .by = c(flip, noise, seed)
  )
```

Read in participant data: 

```{r}
raw_df <- read_csv(here::here("data", "part_2", "exp2-pilot.csv")) %>%
  as_tibble(.) %>%
  filter(status != "rejected")

pilot_df <- raw_df %>% filter(responseId != "windowEvents")

# participant-relevant df
# i.e., the part that is relevant for pixelsPerMM and distToScreen
participant_df <- pilot_df %>%
  filter(str_detect(trialId, "chinrest")) %>%
  select(participantId, responseId, answer) %>%
  filter(
    str_detect(responseId, "pixelsPerMM") |
      str_detect(responseId, "dist-calibration-MM")
  ) %>%
  mutate(answer = as.numeric(answer)) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  rename(distToScreen = `dist-calibration-MM`)

# part 1 relevant df
part1_df <- pilot_df %>%
  filter(str_detect(trialId, "block5_test")) %>%
  select(participantId, trialOrder, responseId, answer) %>%
  mutate(answer = as.numeric(answer)) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  rename(
    num.ans.x = `location-x`,
    num.ans.y = `location-y`,
    num.resp.x = `slider-x`,
    num.resp.y = `slider-y`,
    px.resp.x = `px-slider-x`,
    px.resp.y = `px-slider-y`,
    px.ans.x = `px-location-x`,
    px.ans.y = `px-location-y`
  )

# part 2 relevant df
part2_df <- pilot_df %>%
  filter(str_detect(trialId, "task")) %>%
  select(
    participantId,
    trialOrder,
    trialId,
    responseId,
    condition,
    answer,
    parameters_taskType,
    parameters_taskIndex
  ) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  rename(
    type = parameters_taskType,
    ogIndex = parameters_taskIndex,
    px.resp.y = pixelResponse,
    num.resp.y = numericResponse
  ) %>%
  mutate(
    px.resp.y = as.numeric(px.resp.y),
    num.resp.y = as.numeric(num.resp.y),
    type = ifelse(type == "pointArc", "point_arc", "point")
  )

# join to make the participant_df also contain condition
participant_df <- participant_df %>%
  left_join(
    part2_df %>%
      select(-condition) %>%
      separate_wider_delim(
        trialId,
        delim = "_",
        names = c("condition", "A", "B")
      ) %>%
      select(participantId, condition) %>%
      unique(.)
  )
```

Read in helper functions for transforming between spaces: 

```{r}
#| code-fold: true
#| label: helper-functions

# --- share helper functions
# return visual angle in degrees and not radian
vis_angle <- function(size, distance) {
  return(2 * atan(size / (2 * distance)) * 180 / pi)
}

# define function for turning vis angle to phy size
phys_size <- function(va, distance) {
  return(2 * distance * tan(va * pi / 360))
}

px_to_phys = \(px, px_per_mm) px / px_per_mm
phys_to_px = \(phys, px_per_mm) phys * px_per_mm

# --- Factory function
make_va_scale <- function(data_min, data_max, px_min, px_max) {
  px_extent <- abs(px_max - px_min)
  data_range <- data_max - data_min

  data_to_phys <- function(data, px_per_mm) {
    (data - data_min) / data_range * px_extent / px_per_mm
  }
  phys_to_data <- function(phys, px_per_mm) {
    phys * px_per_mm / px_extent * data_range + data_min
  }

  data_to_va <- function(data, distance, px_per_mm) {
    data |> data_to_phys(px_per_mm) |> vis_angle(distance = distance)
  }
  va_to_data <- function(va, distance, px_per_mm) {
    va |> phys_size(distance = distance) |> phys_to_data(px_per_mm)
  }

  data_to_px <- function(data) {
    (data - data_min) / data_range * (px_max - px_min) + px_min
  }
  px_to_data <- function(px) {
    (px - px_min) / (px_max - px_min) * data_range + data_min
  }

  list(
    data_to_va = data_to_va,
    va_to_data = va_to_data,
    data_to_phys = data_to_phys,
    phys_to_data = phys_to_data,
    data_to_px = data_to_px,
    px_to_data = px_to_data
  )
}

# First Part -- Block 5
x_block5 <- make_va_scale(
  data_min = -5,
  data_max = 5,
  px_min = 50,
  px_max = 585
)
y_block5 <- make_va_scale(
  data_min = 0,
  data_max = 10,
  px_min = 410,
  px_max = 15
)

# Second part -- Mortiz
x_moritz <- make_va_scale(
  data_min = 0,
  data_max = 120,
  px_min = 0,
  px_max = 500
)
y_moritz <- make_va_scale(data_min = 0, data_max = 1, px_min = 0, px_max = 200)

# # Code copied from Matt; this prevents us from having mistakes
# x_num_to_px = \(x, x_range = 120, px_range = 500) x / x_range * px_range
# y_num_to_px = \(y, y_range = 1, px_range = 200) y / y_range * px_range

# x_px_to_num = \(x, x_range = 120, px_range = 500) x * x_range / px_range
# y_px_to_num = \(y, y_range = 1, px_range = 200) y * y_range / px_range

# x_va_to_num = \(va, distance, px_per_mm, x_range = 120, px_range = 500) {
#   va |>
#     phys_size(distance) |>
#     phys_to_px(px_per_mm) |>
#     x_px_to_num(x_range, px_range)
# }

# y_va_to_num = \(va, distance, px_per_mm, y_range = 1, px_range = 200) {
#   va |>
#     phys_size(distance) |>
#     phys_to_px(px_per_mm) |>
#     y_px_to_num(y_range, px_range)
# }

# data_to_pixel_y <- function(data_y) {
#   return(-395 * data_y + 410)
# }
# data_to_pixel_x <- function(data_x) {
#   return(53.5 * data_x + 317.5)
# }

# # origin is bottom left
# pixel_to_phy_x <- function(pixel, pxMM) {
#   (pixel - 50) / pxMM
# }
# pixel_to_phy_y <- function(pixel, pxMM) {
#   (410 - pixel) / pxMM
# }
```

# See whether data is good 

```{r}
# get a sense of the distribution of part1_df
part1_df %>%
  ggplot(aes(y = num.resp.y)) +
  stat_halfeye() +
  coord_cartesian(ylim = c(0, 10))
part1_df %>%
  ggplot(aes(x = num.resp.x)) +
  stat_halfeye() +
  coord_cartesian(xlim = c(-5, 5))
part1_df %>% ggplot(aes(x = num.resp.x, y = num.resp.y)) + geom_bin2d()
```

Get a sense of whether participants answered the calibration task correctly: 

```{r}
participant_df %>% ggplot(aes(x = pixelsPerMM)) + geom_dotplot()
participant_df %>% ggplot(aes(x = distToScreen)) + geom_dotplot()
```

Calculate spearman correlation to determine whether we want to filter out participant who fail our check: 

```{r}
part1_df %>%
  summarise(
    rho = cor(num.ans.x, num.resp.x, method = "spearman"),
    .by = participantId
  )

part1_df %>%
  summarise(
    rho = cor(num.ans.y, num.resp.y, method = "spearman"),
    .by = participantId
  )
```

So far everyone seems to fit. So we are not filtering out anyone from this. 

# Fit model for part 1 

## Get the data ready: 

```{r}
# adding the visual angle related stuff

part1_df <- part1_df %>%
  left_join(participant_df) %>%
  mutate(
    va.ans.x = x_block5$data_to_va(num.ans.x, distToScreen, pixelsPerMM),
    va.ans.y = y_block5$data_to_va(num.ans.y, distToScreen, pixelsPerMM),
    va.resp.x = x_block5$data_to_va(num.resp.x, distToScreen, pixelsPerMM),
    va.resp.y = y_block5$data_to_va(num.resp.y, distToScreen, pixelsPerMM)
  ) %>%
  mutate(va.error.x = va.resp.x - va.ans.x, va.error.y = va.resp.y - va.ans.y)
```

Get a sense of error distribution 

```{r}
#| layout-ncol: 2
#| code-fold: true

part1_df %>%
  ggplot(aes(y = va.error.x, x = va.ans.y)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  labs(title = "Project to X-Axis Error (Vis Angle)")
part1_df %>%
  ggplot(aes(y = va.error.y, x = va.ans.x)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  labs(title = "Project to Y-Axis Error (Vis Angle)")
```

Note that for matching with part 2, we are interested in the second graph. 

## Model Specification 

Recall the we defined the following model for the "project from dot to axis" operator: 

$$
\begin{align*}
\texttt{error}_i &\sim \mathcal{N}(\beta_0 + u_{\text{PID[i]}}, \sigma_{i, \text{PID}[i]}) \\ 
\log(\sigma_{i, j}) &= \alpha_j + \log(\texttt{va.ans.x}_i) \\ 
\text{or:} \log(\sigma_{i, j}) &\sim \mathcal{N}(\mu_\alpha + \log(\texttt{va.ans.x}_i), \sigma_\alpha)\\ 
\alpha_j &\sim N(\mu_\alpha, \sigma_\alpha) \\ 
\mu_\alpha &\sim N(0, 1) \\ 
\sigma_\alpha &\sim \mathcal{N}^+(0, 1)
\end{align*}
$$

Here is the specification we used for fitting: 

```{r}
#| label: brms specs

# Recall that we defined error = resp - ans
f <- bf(
  va.error.y ~ (1 | participantId),
  sigma ~ (1 | participantId) + offset(log(va.ans.x))
)

p <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 1), class = sd, group = participantId, lb = 0),
  prior(normal(0, 1), class = Intercept, dpar = "sigma"),
  prior(normal(0, 1), class = sd, group = participantId, dpar = sigma, lb = 0)
)
```

Fit the model: 

```{r}
m1 <- part1_df %>%
  brm(
    formula = f,
    data = .,
    family = gaussian(),
    prior = p,
    chains = 4,
    cores = 4,
    control = list(adapt_delta = 0.99),
    iter = 6000,
    file = here::here("analysis", "fitted_models", "exp2", "m1"),
    file_refit = "on_change",
    save_pars = save_pars(all = TRUE)
  )
```

Get a sense of posterior fit: 

```{r}
summary(m1)
plot(m1)
```

Note that the Bulk_ESS and Tail_ESS all indicate good fit. 

Posterior predictive checks: 

```{r}
#| layout-ncol: 3
#| code-fold: true
pp_check(m1, ndraws = 100) +
  theme(legend.position = c(0.9, 0.5)) +
  theme(text = element_text(size = 10)) -> p1
p1

pp_check(m1, ndraws = 100, type = "pit_ecdf") +
  theme(text = element_text(size = 10)) -> p2
p2

pp_check(m1, ndraws = 100, type = "pit_ecdf", plot_diff = TRUE) +
  theme(text = element_text(size = 10)) -> p3
p3
```

Plot existing data on top of prediction, for an average participant: 

```{r}
prediction_grid <- part1_df %>%
  modelr::data_grid(
    va.ans.x = modelr::seq_range(va.ans.x, n = 100),
    participantId = unique(participantId)
  ) %>%
  add_predicted_draws(m1, ndraws = 100)

ggplot(prediction_grid, aes(x = va.ans.x, y = .prediction)) +
  stat_lineribbon(
    .width = c(.95, .80, .50),
    # alpha = 0.5,
    linewidth = 0.5
  ) +
  scale_fill_brewer(palette = "Blues") +
  geom_point(
    data = part1_df,
    aes(y = va.error.y, x = va.ans.x),
    alpha = 0.5,
    size = 0.5
  ) +
  labs(
    title = "Project to Y-axis",
    y = "Error",
    x = "Distance to Y-Axis"
  ) +
  # theme_minimal(base_size = 8) +
  # coord_cartesian(ylim = c(-0.25, 0.4)) +
  theme(legend.position = "none") -> p1

p1
```

## Prior sensitivity analysis -- Only Run if time permitting 

To test whether our model is sensitive to prior specifications, we fit them on differnt priors: 

```{r}
# Wider — less informative
p_wide <- c(
  prior(normal(0, 5), class = Intercept),
  prior(normal(0, 5), class = sd, group = participantId, lb = 0),
  prior(normal(0, 5), class = Intercept, dpar = "sigma"),
  prior(normal(0, 5), class = sd, group = participantId, dpar = sigma, lb = 0)
)

# Tighter — more informative
p_tight <- c(
  prior(normal(0, 0.1), class = Intercept),
  prior(normal(0, 0.1), class = sd, group = participantId, lb = 0),
  prior(normal(0, 0.1), class = Intercept, dpar = "sigma"),
  prior(normal(0, 0.1), class = sd, group = participantId, dpar = sigma, lb = 0)
)

# Different family — heavier tails
p_student <- c(
  prior(student_t(3, 0, 1), class = Intercept),
  prior(student_t(3, 0, 1), class = sd, group = participantId, lb = 0),
  prior(student_t(3, 0, 1), class = Intercept, dpar = "sigma"),
  prior(
    student_t(3, 0, 1),
    class = sd,
    group = participantId,
    dpar = sigma,
    lb = 0
  )
)
```

```{r}
m1_wide <- part1_df %>%
  brm(
    formula = f,
    data = .,
    family = gaussian(),
    prior = p_wide,
    chains = 4,
    cores = 4,
    control = list(adapt_delta = 0.99),
    iter = 6000,
    file = here::here("analysis", "fitted_models", "exp2", "m1_wide"),
    file_refit = "on_change",
    save_pars = save_pars(all = TRUE)
  )

m1_tight <- part1_df %>%
  brm(
    formula = f,
    data = .,
    family = gaussian(),
    prior = p_tight,
    chains = 4,
    cores = 4,
    control = list(adapt_delta = 0.99),
    iter = 6000,
    file = here::here("analysis", "fitted_models", "exp2", "m1_tight"),
    file_refit = "on_change",
    save_pars = save_pars(all = TRUE)
  )

m1_student <- part1_df %>%
  brm(
    formula = f,
    data = .,
    family = gaussian(),
    prior = p_student,
    chains = 4,
    cores = 4,
    control = list(adapt_delta = 0.99),
    iter = 6000,
    file = here::here("analysis", "fitted_models", "exp2", "m1_student"),
    file_refit = "on_change",
    save_pars = save_pars(all = TRUE)
  )
```

All models fit without convergence issues. Now we combine posterior and plot them: 

```{r}
# Combine posteriors from all models
models <- list(
  "default (0,1)" = m1,
  "wide (0,5)" = m1_wide,
  "tight (0,0.1)" = m1_tight,
  "student_t(3,0,1)" = m1_student
)

combined <- models %>%
  imap_dfr(
    ~ gather_draws(.x, b_Intercept, b_sigma_Intercept) %>%
      mutate(prior = .y)
  )
```

Interval plot: 

```{r}
combined %>%
  # uncomment to see what variance instead of log(variance) is.
  # mutate(
  #   .value = if_else(.variable == "b_sigma_Intercept", exp(.value), .value)
  # ) %>%
  ggplot(aes(x = .value, y = prior)) +
  stat_pointinterval(.width = c(.50, .95)) +
  facet_wrap(~.variable, scales = "free_x") +
  labs(x = "Estimate", y = "Prior specification")
```

Densities overlaid: 

```{r}
combined %>%
  ggplot(aes(x = .value, color = prior, fill = prior)) +
  stat_slab(alpha = 0.3) +
  facet_wrap(~.variable, scales = "free") +
  labs(x = "Estimate", y = NULL)
```

To see the actual table: 

```{r}
combined %>%
  group_by(prior, .variable) %>%
  median_qi(.value, .width = .95) %>%
  select(prior, .variable, .value, .lower, .upper) %>%
  arrange(.variable, prior)
```

We can also check random effects: 

```{r}
combined_sd <- models %>%
  imap_dfr(
    ~ gather_draws(
      .x,
      sd_participantId__Intercept,
      sd_participantId__sigma_Intercept
    ) %>%
      mutate(prior = .y)
  )
```

```{r}
combined_sd %>%
  ggplot(aes(x = .value, y = prior)) +
  stat_pointinterval(.width = c(.50, .95)) +
  facet_wrap(~.variable, scales = "free_x") +
  labs(x = "Estimate", y = "Prior specification")
```

**Takeaway**: posteriors were robust to prior specification across default, wide, and heavy-tailed priors. The tight prior $\sigma = 0.1$ constrained the residual variance parameter, confirming it is unreasonably informative for this parameter; the remaining three specifications produced essentially identical posteriors. 

# Infer model for part 2 

## Mean and Median 

First we get a cross between participant and the stimuli they saw, in order to generate **predictions**: 

```{r}
predict_df <- participant_df %>%
  rename(type = condition) %>%
  mutate(type = ifelse(type == "pointArc", "point_arc", "point")) %>%
  inner_join(
    stimuli_df %>% filter(type != "line") %>% rename(ogIndex = index),
    by = "type",
    relationship = "many-to-many"
  ) %>%
  unnest(col = data) %>%
  rename(num.ans.x = x, num.ans.y = y) %>%
  mutate(
    va.ans.x = x_moritz$data_to_va(num.ans.x, distToScreen, pixelsPerMM),
    va.ans.y = y_moritz$data_to_va(num.ans.y, distToScreen, pixelsPerMM),
  )

predict_df %>% head(5)
```

We then generate the posterior draws from out fitted model, `m1`, using `add_predicted_draws`: 

```{r}
preds <- predict_df %>%
  add_predicted_draws(m1, ndraws = 100) %>%
  mutate(va.pred.y = va.ans.y + .prediction) %>%
  group_by(
    participantId,
    ogIndex,
    type,
    flip,
    noise,
    distToScreen,
    pixelsPerMM,
    .draw
  ) %>%
  summarise(
    # calculate mean and median of va
    va.pred.mean.y = mean(va.pred.y),
    va.pred.median.y = median(va.pred.y),
  ) %>%
  # turn vis angle into numerical
  mutate(
    num.pred.mean.y = y_moritz$va_to_data(
      va.pred.mean.y,
      distToScreen,
      pixelsPerMM
    ),
    num.pred.median.y = y_moritz$va_to_data(
      va.pred.median.y,
      distToScreen,
      pixelsPerMM
    )
  )
```

### Checking Results across participants 

We can also plot them by the diagonal to see whether they predict things correcty: 

```{r}
combined <- preds %>%
  group_by(ogIndex, participantId) %>%
  median_qi(num.pred.mean.y, .width = .95) %>%
  left_join(part2_df %>% select(ogIndex, participantId, num.resp.y))

ggplot(combined, aes(x = num.pred.mean.y, y = num.resp.y)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  geom_point(alpha = 0.3) +
  labs(x = "Predicted", y = "Observed", title = "Mean Strategy")
```

```{r}
combined <- preds %>%
  group_by(ogIndex, participantId) %>%
  median_qi(num.pred.median.y, .width = .95) %>%
  left_join(part2_df %>% select(ogIndex, participantId, num.resp.y))

ggplot(combined, aes(x = num.pred.median.y, y = num.resp.y)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  geom_point(alpha = 0.3) +
  labs(x = "Predicted", y = "Observed", title = "Median Strategy")
```

Let's also do coverage check: 

```{r}
combined <- preds %>%
  group_by(ogIndex, participantId) %>%
  summarise(
    lo95 = quantile(num.pred.median.y, 0.025), # could also be num.pred.mean.y
    hi95 = quantile(num.pred.median.y, 0.975), # could also be num.pred.mean.y
    .groups = "drop"
  ) %>%
  left_join(part2_df %>% select(ogIndex, participantId, num.resp.y)) %>%
  mutate(covered = num.resp.y >= lo95 & num.resp.y <= hi95)

mean(combined$covered)
```

### Checking results for one participant 

```{r}
# Pick one participant
pid <- unique(part1_df$participantId)[1]

pred_one <- preds %>%
  filter(participantId == pid)

obs_one <- part2_df %>%
  filter(participantId == pid)
```

Let's first plot mean strategy: 

```{r}
#| code-fold: true

ggplot() +
  stat_pointinterval(
    data = pred_one,
    aes(y = ogIndex, x = num.pred.mean.y),
    .width = c(.50, .95)
  ) +
  geom_point(
    data = obs_one,
    aes(y = ogIndex, x = num.resp.y),
    color = "red",
    size = 2
  ) +
  labs(y = "Trial", x = "Response (numeric)", title = "Mean Strategy") +
  facet_wrap(~flip)
```

Then let's plot median strategy: 

```{r}
#| code-fold: true
ggplot() +
  stat_pointinterval(
    data = pred_one,
    aes(y = ogIndex, x = num.pred.median.y),
    .width = c(.50, .95)
  ) +
  geom_point(
    data = obs_one,
    aes(y = ogIndex, x = num.resp.y),
    color = "red",
    size = 2
  ) +
  labs(y = "Trial", x = "Response (numeric)", title = "Median Strategy") +
  facet_wrap(~flip)
```

## Sensor Fusion Strategy 

$$
\hat{y} = \sum w_i \times y_i, \text{ where } w_i = \frac{MSE^{-1}(y_i)}{\sum_j MSE(y_j)}
$$

```{r}
predict_df <- participant_df %>%
  rename(type = condition) %>%
  mutate(type = ifelse(type == "pointArc", "point_arc", "point")) %>%
  inner_join(
    stimuli_df %>% filter(type != "line") %>% rename(ogIndex = index),
    by = "type",
    relationship = "many-to-many"
  ) %>%
  unnest(col = data) %>%
  rename(num.ans.x = x, num.ans.y = y) %>%
  mutate(
    va.ans.x = abs(
      x_moritz$data_to_va(60, distToScreen, pixelsPerMM) -
        x_moritz$data_to_va(num.ans.x, distToScreen, pixelsPerMM)
    ),
    va.ans.y = y_moritz$data_to_va(num.ans.y, distToScreen, pixelsPerMM),
  )
```

```{r}
draws <- m1 %>%
  spread_draws(
    b_Intercept,
    r_participantId[participantId, ],
    b_sigma_Intercept,
    r_participantId__sigma[participantId, ]
  ) %>%
  as_draws_df(.) %>%
  posterior::thin_draws(thin = 60)

weights_df <- draws %>%
  mutate(
    b_Intercept = b_Intercept + r_participantId,
    log_sigma = b_sigma_Intercept + r_participantId__sigma
  )
```

```{r}

```

## pre-process data 

```{r}
# part2_df %>%
#   mutate(type = ifelse(type == "pointArc", "point_arc", "point")) %>%
#   left_join(part2_ans_df, by = c("ogIndex", "type")) %>%
#   left_join(participant_df, by = join_by(participantId)) %>%
#   mutate(
#     va.ans.y = y_moritz$data_to_va(num.ans.y, distToScreen, pixelsPerMM),
#     va.resp.y = y_moritz$data_to_va(num.resp.y, distToScreen, pixelsPerMM)
#   ) %>%
#   mutate(va.error.y = va.resp.y - va.ans.y)
```

Look at the empirical distribution: 

```{r}
part2_df %>%
  left_join(part2_ans_df) %>%
  mutate(diff = num.resp.y - num.ans.y) %>%
  ggplot(aes(x = diff, y = noise, fill = type)) +
  stat_halfeye(alpha = 0.5) +
  facet_wrap(~flip, ncol = 1) +
  geom_vline(xintercept = 0, linetype = "dashed")
```

```{r}
# plotting function for Moritz result
plot_results <- function(data, xvar) {
  data %>%
    ggplot(aes(
      y = noise,
      x = {{ xvar }},
      dist = "norm",
      arg1 = {{ xvar }},
      arg2 = se,
      fill = type
    )) + #uncertainty
    stat_dist_halfeye(
      show_interval = TRUE,
      position = position_dodge(width = 0.2),
      .width = 0.95,
      size = .5,
      alpha = .8
    ) +
    theme(legend.position = "top") +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray") +
    scale_fill_manual(values = c("#76B7B2", "#E15759")) + # take out "#59A14F"
    ylab("Variability") # + #add x label
  # xlab("Error") #add y label
}
```

```{r}
part2_df %>%
  left_join(part2_ans_df) %>%
  mutate(
    error = num.resp.y - num.ans.y,
    noise = factor(noise, levels = c("c(0, 0)", "c(0, 0.4)"))
  ) %>%
  summarySEwithin(
    data = .,
    measurevar = "error",
    withinvars = c("noise", "flip"),
    betweenvars = "type",
    idvar = "participantId"
  ) %>%
  mutate(
    flip_lab = ifelse(
      flip == FALSE,
      "Variability in top",
      "Variability in bottom"
    ),
    flip_lab = factor(
      flip_lab,
      levels = c("Variability in top", "Variability in bottom")
    )
  ) %>%
  # change error to the direction of "variability"
  # mutate(error = ifelse(flip == FALSE, error, -error)) %>%
  plot_results(., error) +
  # xlab("Error (in the direction of variability)") +
  facet_wrap(~flip_lab, ncol = 1) +
  labs(title = "N = 20")
```

Now this is not something that you would see in Moritz ... 