---
title: "New Pilot with combined components "
author: "Anonymous"
date: today
date-format: "iso"
format:
  html:
    reference-location: margin
---

```{r}
#| code-fold: true

library(plyr)
source(here::here("analysis", "moritz", "Functions.R"))
library(tidyverse)
library(jsonlite)
library(ggplot2)
library(tidybayes)
theme_set(theme_tidybayes())
```

Read in stimuli data: 

```{r}
exp_json_data <- jsonlite::fromJSON(here::here(
  "data",
  "moritz",
  "stimuli",
  "exp2-stimuli.json"
)) %>%
  as_tibble(.)

stimuli_df <- exp_json_data %>%
  select(index, seed, noise, flip, type, data) %>%
  mutate(noise = as.character(noise))

# get the mapping between og index and the index used in exp
stimuli_df %>%
  filter(type == "point") %>%
  mutate(expIndex = 1:nrow(.)) %>%
  select(index, type, expIndex) -> point_map
stimuli_df %>%
  filter(type == "point_arc") %>%
  mutate(expIndex = 1:nrow(.)) %>%
  select(index, type, expIndex) -> pointArc_map

# get the answer
part2_ans_df <- stimuli_df %>%
  filter(type != "line") %>%
  unnest(col = data) %>%
  rename(num.x = x, num.y = y, ogIndex = index) %>%
  summarise(
    num.mean.y = mean(num.y),
    num.median.y = median(num.y),
    .by = c(ogIndex, type, flip, noise, seed)
  ) |>
  mutate(
    # num.ans.y = num.mean.y[type == "point"][[1]], # this is what Moritz did, and what we're not doing
    num.ans.y = num.mean.y,
    .by = c(flip, noise, seed)
  )
```

Read in participant data: 

```{r}
raw_df <- read_csv(here::here("data", "part_2", "exp2-pilot.csv")) %>%
  as_tibble(.)

pilot_df <- raw_df %>% filter(responseId != "windowEvents")

# participant-relevant df
# i.e., the part that is relevant for pixelsPerMM and distToScreen
participant_df <- pilot_df %>%
  filter(str_detect(trialId, "chinrest")) %>%
  select(participantId, trialId, trialOrder, responseId, condition, answer) %>%
  select(participantId, responseId, answer) %>%
  filter(
    str_detect(responseId, "pixelsPerMM") |
      str_detect(responseId, "dist-calibration-MM")
  ) %>%
  mutate(answer = as.numeric(answer)) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  rename(distToScreen = `dist-calibration-MM`)

# part 1 relevant df
part1_df <- pilot_df %>%
  filter(str_detect(trialId, "block5_test")) %>%
  select(participantId, trialOrder, responseId, answer) %>%
  mutate(answer = as.numeric(answer)) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  rename(
    num.ans.x = `location-x`,
    num.ans.y = `location-y`,
    num.resp.x = `slider-x`,
    num.resp.y = `slider-y`,
    px.resp.x = `px-slider-x`,
    px.resp.y = `px-slider-y`,
    px.ans.x = `px-location-x`,
    px.ans.y = `px-location-y`
  )

# part 2 relevant df
part2_df <- pilot_df %>%
  filter(str_detect(trialId, "task")) %>%
  select(
    participantId,
    trialOrder,
    trialId,
    responseId,
    condition,
    answer,
    parameters_taskType,
    parameters_taskIndex
  ) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  rename(
    type = parameters_taskType,
    ogIndex = parameters_taskIndex,
    px.resp.y = pixelResponse,
    num.resp.y = numericResponse
  ) %>%
  mutate(px.resp.y = as.numeric(px.resp.y), num.resp.y = as.numeric(num.resp.y))
```

Read in helper functions for transforming between spaces: 

```{r}
#| code-fold: true
#| label: helper-functions

# --- share helper functions
# return visual angle in degrees and not radian
vis_angle <- function(size, distance) {
  return(2 * atan(size / (2 * distance)) * 180 / pi)
}

# define function for turning vis angle to phy size
phys_size <- function(va, distance) {
  return(2 * distance * tan(va * pi / 360))
}

px_to_phys = \(px, px_per_mm) px / px_per_mm
phys_to_px = \(phys, px_per_mm) phys * px_per_mm

# --- Factory function
make_va_scale <- function(data_min, data_max, px_min, px_max) {
  px_extent <- abs(px_max - px_min)
  data_range <- data_max - data_min

  data_to_phys <- function(data, px_per_mm) {
    (data - data_min) / data_range * px_extent / px_per_mm
  }

  data_to_va <- function(data, distance, px_per_mm) {
    data |> data_to_phys(px_per_mm) |> vis_angle(distance = distance)
  }

  # keep these around if you need them elsewhere
  data_to_px <- function(data) {
    (data - data_min) / data_range * (px_max - px_min) + px_min
  }
  px_to_data <- function(px) {
    (px - px_min) / (px_max - px_min) * data_range + data_min
  }

  list(
    data_to_va = data_to_va,
    data_to_phys = data_to_phys,
    data_to_px = data_to_px,
    px_to_data = px_to_data
  )
}

# Block 5
x_block5 <- make_va_scale(
  data_min = 0,
  data_max = 120,
  px_min = 0,
  px_max = 500
)
y_block5 <- make_va_scale(data_min = 0, data_max = 1, px_min = 0, px_max = 200)

# Moritz
x_moritz <- make_va_scale(
  data_min = -5,
  data_max = 5,
  px_min = 50,
  px_max = 585
)
y_moritz <- make_va_scale(
  data_min = 0,
  data_max = 10,
  px_min = 410,
  px_max = 15
)

# # Code copied from Matt; this prevents us from having mistakes
# x_num_to_px = \(x, x_range = 120, px_range = 500) x / x_range * px_range
# y_num_to_px = \(y, y_range = 1, px_range = 200) y / y_range * px_range

# x_px_to_num = \(x, x_range = 120, px_range = 500) x * x_range / px_range
# y_px_to_num = \(y, y_range = 1, px_range = 200) y * y_range / px_range

# x_va_to_num = \(va, distance, px_per_mm, x_range = 120, px_range = 500) {
#   va |>
#     phys_size(distance) |>
#     phys_to_px(px_per_mm) |>
#     x_px_to_num(x_range, px_range)
# }

# y_va_to_num = \(va, distance, px_per_mm, y_range = 1, px_range = 200) {
#   va |>
#     phys_size(distance) |>
#     phys_to_px(px_per_mm) |>
#     y_px_to_num(y_range, px_range)
# }

# data_to_pixel_y <- function(data_y) {
#   return(-395 * data_y + 410)
# }
# data_to_pixel_x <- function(data_x) {
#   return(53.5 * data_x + 317.5)
# }

# # origin is bottom left
# pixel_to_phy_x <- function(pixel, pxMM) {
#   (pixel - 50) / pxMM
# }
# pixel_to_phy_y <- function(pixel, pxMM) {
#   (410 - pixel) / pxMM
# }
```

# See whether data is good 

```{r}
# get a sense of the distribution of part1_df
part1_df %>%
  ggplot(aes(y = num.resp.y)) +
  stat_halfeye() +
  coord_cartesian(ylim = c(0, 10))
part1_df %>%
  ggplot(aes(x = num.resp.x)) +
  stat_halfeye() +
  coord_cartesian(xlim = c(-5, 5))
part1_df %>% ggplot(aes(x = num.resp.x, y = num.resp.y)) + geom_bin2d()
```

Get a sense of whether participants answered the calibration task correctly: 

```{r}
participant_df %>% ggplot(aes(x = pixelsPerMM)) + geom_dotplot()
participant_df %>% ggplot(aes(x = distToScreen)) + geom_dotplot()
```

# Fit model for part 1 

## Model Specification 

Recall the we defined the following model for the "project from dot to axis" operator: 

$$
\begin{align*}
\texttt{error}_i &\sim \mathcal{N}(\beta_0 + u_{\text{PID[i]}}, \sigma_{i, \text{PID}[i]}) \\ 
\log(\sigma_{i, j}) &= \alpha_j + \log(\texttt{va.ans.x}_i) \\ 
\text{or:} \log(\sigma_{i, j}) &\sim \mathcal{N}(\mu_\alpha + \log(\texttt{va.ans.x}_i), \sigma_\alpha)\\ 
\alpha_j &\sim N(\mu_\alpha, \sigma_\alpha) \\ 
\mu_\alpha &\sim N(0, 1) \\ 
\sigma_\alpha &\sim \mathcal{N}^+(0, 1)
\end{align*}
$$

Here is the specification we used for fitting: 

```{r}
#| eval: false
#| label: brms specs

# Recall that we defined error = resp - ans
f <- bf(
  error_va ~ (1 | participantId),
  sigma ~ (1 | participantId) + offset(log(va.ans.x))
)

p <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 1), class = sd, group = participantId, lb = 0),
  prior(normal(0, 1), class = Intercept, dpar = "sigma"),
  prior(normal(0, 1), class = sd, group = participantId, dpar = sigma, lb = 0)
)
```

Get the data ready: 

```{r}
# adding the visual angle related stuff

part1_df <- part1_df %>%
  left_join(participant_df) %>%
  mutate(
    va.ans.x = x_block5$data_to_va(num.ans.x, distToScreen, pixelsPerMM),
    va.ans.y = y_block5$data_to_va(num.ans.y, distToScreen, pixelsPerMM),
    va.resp.x = x_block5$data_to_va(num.resp.x, distToScreen, pixelsPerMM),
    va.resp.y = y_block5$data_to_va(num.resp.y, distToScreen, pixelsPerMM)
  ) %>%
  mutate(va.error.x = va.resp.x - va.ans.x, va.error.y = va.resp.y - va.ans.y)
```

Get a sense of error distribution 

```{r}
part1_df %>%
  ggplot(aes(x = num.resp.y - num.ans.y)) +
  geom_dotplot()
```

```{r}
part1_df %>%
  ggplot(aes(y = va.error.x, x = va.ans.y, color = participantId)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") # okay this scale is hella weird ...  +
facet_wrap(~participantId, ncol = 4)
part1_df %>%
  ggplot(aes(y = va.error.y, x = va.ans.x, color = participantId)) +
  geom_point() # +
facet_wrap(~participantId, ncol = 4)
```

Calculate spearman correlation to determine whether we want to filter out participant who fail our check: 

```{r}
part1_df %>%
  summarise(
    rho = cor(va.ans.x, va.resp.x, method = "spearman"),
    .by = participantId
  )

part1_df %>%
  summarise(
    rho = cor(va.ans.y, va.resp.y, method = "spearman"),
    .by = participantId
  )
```

So far everyone seems to fit. 

# Infer model for part 2 

```{r}
part2_df %>%
  mutate(type = ifelse(type == "pointArc", "point_arc", "point")) %>%
  left_join(part2_ans_df) %>%
  mutate(diff = num.resp.y - num.ans.y) %>%
  ggplot(aes(x = diff, y = noise, fill = type)) +
  stat_halfeye(alpha = 0.5) +
  facet_wrap(~flip, ncol = 1) +
  geom_vline(xintercept = 0, linetype = "dashed")
```

```{r}
plot_results <- function(data, xvar) {
  data %>%
    ggplot(aes(
      y = noise,
      x = {{ xvar }},
      dist = "norm",
      arg1 = {{ xvar }},
      arg2 = se,
      fill = type
    )) + #uncertainty
    stat_dist_halfeye(
      show_interval = TRUE,
      position = position_dodge(width = 0.2),
      .width = 0.95,
      size = .5,
      alpha = .8
    ) +
    theme(legend.position = "top") +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray") +
    scale_fill_manual(values = c("#76B7B2", "#E15759")) + # take out "#59A14F"
    ylab("Variability") # + #add x label
  # xlab("Error") #add y label
}
```

```{r}
part2_df %>%
  mutate(
    type = ifelse(type == "pointArc", "point_arc", "point"),
  ) %>%
  left_join(part2_ans_df) %>%
  mutate(
    error = num.resp.y - num.ans.y,
    noise = factor(noise, levels = c("c(0, 0)", "c(0, 0.4)"))
  ) %>%
  summarySEwithin(
    data = .,
    measurevar = "error",
    withinvars = c("noise", "flip"),
    betweenvars = "type",
    idvar = "participantId"
  ) %>%
  # change error to the direction of "variability"
  # mutate(error = ifelse(flip == FALSE, error, -error)) %>%
  plot_results(., error) +
  # xlab("Error (in the direction of variability)") +
  facet_wrap(~flip, ncol = 1)
```