---
title: "Reanalyze Mortiz"
---

```{r}
#| echo: false
#| label: load libraries
library(tidyverse)
library(tidybayes)
library(here)
```

# TLDR 

The goal of this file is to (1) recreate the analysis that Moritz et al. produced for their second experiment, and (2) perform a different kind of error analysis, as well as analyze particpant's whose strategies are closely mapped to the mental averaging and median strategy. 

Reading in data from Exp 2 (excluding the `line` condition): 

```{r}
#Exp 2
exp2_dotArc <- read.csv(
  here("data", "moritz", "exp", "exp2-point-arc.csv"),
  ,
  check.names = FALSE
) #point arc
exp2_dotX <- read.csv(
  here("data", "moritz", "exp", "exp2-point.csv"),
  check.names = FALSE
) #point
```

Combining conditions in Exp 2

```{r combine_exp2}
exp2 <- rbind(exp2_dotArc, exp2_dotX)
```

Fixing the naming and coding from various conditions

```{r naming}
#fixing variability coding for Exp 2
exp2$variability <-
  factor(ifelse(
    exp2$variability == "0,0",
    "0",
    ifelse(exp2$variability == "0,0.4", "0.4", 999)
  ))

#fixing the flip naming for Exp 2
exp2$flipped <- factor(ifelse(exp2$flip == "TRUE", "Flipped", "Unflipped"))
exp2$flipped <- factor(exp2$flipped, levels = c("Unflipped", "Flipped"))

#fixing the overestimation naming for Exp 2
exp2$estimation <- factor(ifelse(
  exp2$overestimated == "TRUE",
  "overestimated",
  "underestimated"
))

#adding biased towards variability code
exp2$bias_towards_variability <- ifelse(
  exp2$normalized_error > "0",
  "True",
  "False"
)

```

---

Check how the data relates to each other 

```{r}
exp2_df <- exp2 %>% as_tibble(.)

exp2_df %>% count(xdomain)

raw_data <- exp2_df %>% select(xdomain, ydomain)

# filter to only keep the ones that matter
exp2_df <- exp2_df %>%
  select(
    assignmentid,
    mark,
    reportedaverage,
    seed,
    stimuliindex,
    flip,
    trial,
    variability,
    data_average,
    average,
    normalized_average,
    error,
    normalized_error
  )
```

Play with the dataset 

```{r}
exp2_df %>% count(trial) # ok trial is just the trial order
exp2_df %>% count(seed) # TODO: confirm if this is the same as the one we did ...
exp2_df %>% count(seed, stimuliindex) # stimuliindex is the actual index used in the json I think
exp2_df %>%
  count(seed, stimuliindex, mark, flip, variability, data_average, average)
```

This tells us that for stimuli that share the same seed, same flip, they should also share the same average: 

```{r}
exp2_df %>% count(seed, flip, average)
```

---


# Analyses 

## Exp 2

Create a theme: 

```{r}
#creating a theme
line_theme <- function() {
  theme(
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal",
    legend.title = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_text(size = 15, face = "bold", hjust = 0)
  )
}
```

```{r}
#specifying levels
exp2$mark <- factor(exp2$mark, levels = c("point_arc", "line", "point"))

#plot
exp2Desc1 <- exp2 %>%
  filter(bias_towards_variability == "True") %>% #only show overestimation
  ggplot(aes(bias_towards_variability)) + #x value
  geom_bar() + #add bar
  coord_flip() + #flip axis
  facet_wrap('mark', ncol = 1) + #break up by mark type
  geom_text(stat = 'count', aes(label = after_stat(count)), hjust = -.2) + #add count text
  theme(legend.position = "top") + #move legend
  scale_y_continuous(
    breaks = seq(0, 7000, by = 1000),
    expand = expansion(mult = c(0, 0.12))
  ) + #scale
  geom_hline(yintercept = 6720) + #add horizontal line indicating the number of judgement
  theme_classic() + #theme
  guides(y = "none") + #remove guides
  line_theme() #add theme
exp2Desc1 #print plot
```

### Statistical Analysis 

```{r Exp 2}
#converting mark to factor
exp2$mark <- as.factor(exp2$mark)

#ordering mark type
exp2$mark <- factor(exp2$mark, levels = c("line", "point", "point_arc"))

#Pre-registered regression model
m.exp2.1 = lmer(
  normalized_error ~ mark * variability + (1 | assignmentid),
  data = exp2
)

#approximating coefficients
c.exp2.1 <- cbind(
  data.frame(coef(summary(m.exp2.1))) %>% #summary of coefficients
    mutate(p.z = 2 * (1 - pnorm(abs(t.value)))), #p-value estimation
  as_tibble(confint(m.exp2.1, method = "Wald", level = 0.95))[-c(1:2), ]
) #CIs

#creating a table printout
knitr::kable(c.exp2.1, align = "l", booktabs = T, linesep = "", digits = 3) %>%
  row_spec(0, bold = T) %>%
  kable_styling(latex_options = "striped", font_size = 12)
```

