---
title: "Reanalyze Mortiz"
---

```{r}
#| echo: false
#| label: load libraries
library(tidyverse)
library(tidybayes)
library(here)
library(ggplot2)
theme_set(theme_tidybayes())
```

## TLDR 

The goal of this file is to (1) recreate the analysis that Moritz et al. produced for their second experiment, and (2) perform a different kind of error analysis, as well as analyze particpant's whose strategies are closely mapped to the mental averaging and median strategy. 

Reading in data from Exp 2 (excluding the `line` condition): 

```{r}
#Exp 2
#point arc
exp2_dotArc <- read.csv(
  here("data", "moritz", "exp", "exp2-point-arc.csv"),
  ,
  check.names = FALSE
)
#point
exp2_dotX <- read.csv(
  here("data", "moritz", "exp", "exp2-point.csv"),
  check.names = FALSE
)
# Combining conditions in Exp 2
exp2 <- rbind(exp2_dotArc, exp2_dotX) %>% as_tibble(.)
```

Get a basic idea of the underlying dataset: 

```{r}
exp2 %>% count(assignmentid) # 277 participants altogether
```

## Read in data related to strategy 

```{r}
strategy_df <- read.csv(here("data", "moritz", "exp", "strategies.csv")) %>%
  as_tibble(.) %>%
  # keep only those related to experiment 2, and only related to point and point arc condition
  filter(Exp == 2, Condition %in% c("point", "point_arc")) %>%
  select(1:4, contains("mental_averaging"), contains("number_of_points"))
```

Where do the two raters (`r1` and `r2`) disagree? 

```{r}
# for mental averaging
strategy_df %>% filter(r1_mental_averaging != r2_mental_averaging)
# for equal number of points
strategy_df %>% filter(r1_number_of_points != r2_number_of_points) %>% View(.)
```

Where do they agree? 

```{r}
# for mental averaging --- 98 participants where both raters agree
strategy_df %>%
  filter(r1_mental_averaging == r2_mental_averaging) %>%
  filter(r1_mental_averaging == TRUE) %>%
  pull(ResponseId) -> mental_average_id
# for equal number of points -- 44 participants where both raters agree
strategy_df %>%
  filter(r1_number_of_points == r2_number_of_points) %>%
  filter(r1_number_of_points == TRUE) %>%
  pull(ResponseId) -> equal_points_id
```

## Calculate Error and plot empirical error distributions 

Read in the stimuli and calculate the answer for each `index` * `seed` 

```{r}
#| code-fold: true
exp_json_data <- jsonlite::fromJSON(here(
  "data",
  "moritz",
  "stimuli",
  "exp2-stimuli.json"
)) %>%
  as_tibble(.)

stimuli_df <- exp_json_data %>%
  select(index, seed, noise, flip, type, data) %>%
  mutate(noise = as.character(noise)) %>%
  filter(type != "line") %>%
  unnest(col = data)

ans_df <- stimuli_df %>%
  group_by(index, seed, noise, flip, type) %>%
  summarise(ans.mean.y = mean(y), ans.median.y = median(y))
```

```{r}
ans_df
```

Use the `data_average` to calculate error, and this is what we in terms of empirical distribution of error: 

```{r}
#| code-fold: true
#| layout-ncol: 2
exp2 %>%
  select(
    assignmentid,
    mark,
    reportedaverage,
    seed,
    stimuliindex,
    flip,
    variability,
    data_average
  ) %>%
  mutate(error = reportedaverage - data_average) %>%
  ggplot(aes(x = error, y = variability, fill = mark)) +
  stat_halfeye(alpha = 0.8)

exp2 %>%
  select(
    assignmentid,
    mark,
    reportedaverage,
    seed,
    stimuliindex,
    flip,
    variability,
    data_average
  ) %>%
  mutate(
    norm.error = ifelse(
      flip == FALSE,
      reportedaverage - data_average,
      data_average - reportedaverage
    )
  ) %>%
  ggplot(aes(x = norm.error, y = variability, fill = mark)) +
  stat_halfeye(alpha = 0.8)
```

We could also try to facet things so they're not all on the same thing: 

```{r}
exp2 %>%
  select(
    assignmentid,
    mark,
    reportedaverage,
    seed,
    stimuliindex,
    flip,
    variability,
    data_average
  ) %>%
  mutate(
    error = reportedaverage - data_average,
    variance_area = ifelse(flip == FALSE, "high", "low")
  ) %>%
  ggplot(aes(x = error, y = variability, fill = mark)) +
  stat_halfeye(alpha = 0.8) +
  facet_wrap(~variance_area)
```


# Analyses 

## Exp 2

Create a theme: 

```{r}
#creating a theme
line_theme <- function() {
  theme(
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal",
    legend.title = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_text(size = 15, face = "bold", hjust = 0)
  )
}
```

```{r}
#specifying levels
exp2$mark <- factor(exp2$mark, levels = c("point_arc", "line", "point"))

#plot
exp2Desc1 <- exp2 %>%
  filter(bias_towards_variability == "True") %>% #only show overestimation
  ggplot(aes(bias_towards_variability)) + #x value
  geom_bar() + #add bar
  coord_flip() + #flip axis
  facet_wrap('mark', ncol = 1) + #break up by mark type
  geom_text(stat = 'count', aes(label = after_stat(count)), hjust = -.2) + #add count text
  theme(legend.position = "top") + #move legend
  scale_y_continuous(
    breaks = seq(0, 7000, by = 1000),
    expand = expansion(mult = c(0, 0.12))
  ) + #scale
  geom_hline(yintercept = 6720) + #add horizontal line indicating the number of judgement
  theme_classic() + #theme
  guides(y = "none") + #remove guides
  line_theme() #add theme
exp2Desc1 #print plot
```

### Statistical Analysis 

```{r Exp 2}
#converting mark to factor
exp2$mark <- as.factor(exp2$mark)

#ordering mark type
exp2$mark <- factor(exp2$mark, levels = c("line", "point", "point_arc"))

#Pre-registered regression model
m.exp2.1 = lmer(
  normalized_error ~ mark * variability + (1 | assignmentid),
  data = exp2
)

#approximating coefficients
c.exp2.1 <- cbind(
  data.frame(coef(summary(m.exp2.1))) %>% #summary of coefficients
    mutate(p.z = 2 * (1 - pnorm(abs(t.value)))), #p-value estimation
  as_tibble(confint(m.exp2.1, method = "Wald", level = 0.95))[-c(1:2), ]
) #CIs

#creating a table printout
knitr::kable(c.exp2.1, align = "l", booktabs = T, linesep = "", digits = 3) %>%
  row_spec(0, bold = T) %>%
  kable_styling(latex_options = "striped", font_size = 12)
```

