---
title: "Reanalyze Mortiz"
author: "Anonymous" 
---

```{r}
#| echo: false
#| label: load libraries

library(plyr)
# source(paste0(getwd(), "/analysis/moritz/Functions.R"))
# library(here)
source(here::here("analysis", "moritz", "Functions.R"))
library(tidyverse)
library(tidybayes)
library(ggplot2)
library(patchwork)
theme_set(theme_tidybayes())
```

## TLDR 

The goal of this file is to (1) recreate the analysis that Moritz et al. produced for their second experiment, and (2) perform a different kind of error analysis, as well as analyze particpant's whose strategies are closely mapped to the mental averaging and median strategy. 

Reading in data from Exp 2 (excluding the `line` condition): 

# Repeating Mortiz's analysis 

```{r}
#| label: read in Mortiz data

#point arc
exp2_dotArc <- read.csv(
  here::here("data", "moritz", "exp", "exp2-point-arc.csv"),
  # paste0(getwd(), "/data/moritz/exp/exp2-point-arc.csv"),
  check.names = FALSE
)

#point
exp2_dotX <- read.csv(
  here::here("data", "moritz", "exp", "exp2-point.csv"),
  # paste0(getwd(), "/data/moritz/exp/exp2-point.csv"),
  check.names = FALSE
)

# combine
raw_exp2 <- rbind(exp2_dotArc, exp2_dotX) %>%
  as_tibble(.)

exp2 <- raw_exp2 %>%
  select(assignmentid, mark, flip, variability, error, normalized_error)

exp2 %>% head(5)

# also read in the original stimuli file
exp_json_data <- jsonlite::fromJSON(
  here::here("data", "moritz", "stimuli", "exp2-stimuli.json")
  # paste0(getwd(), "/data/moritz/stimuli/exp2-stimuli.json"),
) %>%
  as_tibble(.)

stimuli_df <- exp_json_data %>%
  select(index, seed, noise, flip, type, data) %>%
  mutate(noise = as.character(noise))
```

How many participants per task? 

```{r}
exp2 %>% group_by(mark) %>% count(assignmentid)
```

There are 277 participants altogether excluding the `line` condition. 

Define helper function for plotting: 

```{r}
#| code-fold: true

plot_results <- function(data, xvar) {
  data %>%
    ggplot(aes(
      y = variability,
      x = {{ xvar }},
      dist = "norm",
      arg1 = {{ xvar }},
      arg2 = se,
      fill = mark
    )) + #uncertainty
    stat_dist_halfeye(
      show_interval = TRUE,
      position = position_dodge(width = 0.2),
      .width = 0.95,
      size = .5,
      alpha = .8
    ) +
    theme(legend.position = "top") +
    scale_fill_manual(values = c("#76B7B2", "#E15759")) + # take out "#59A14F"
    ylab("Variability") + #add x label
    xlab("Normalized Error") #add y label
}
```

For our own sanity, recreate the plot that Moritz et al. created, using the `Functions.R` library that they used: 

```{r}
#| label: plot Moritz
#| code-fold: true

p_og <- exp2 %>%
  summarySEwithin(
    data = .,
    measurevar = "normalized_error",
    withinvars = "variability",
    betweenvars = "mark",
    idvar = "assignmentid"
  ) %>%
  plot_results(., normalized_error)

p_og
```

# Analyze signed error

We will calculate error ourselves using reported average and the og `stimuli.json`: 

```{r}
exp2 <- raw_exp2 %>%
  select(assignmentid, reportedaverage, stimuliindex)

ans_df <- stimuli_df %>%
  filter(type != "line") %>%
  rename(stimuliindex = index) %>%
  unnest(data) %>%
  group_by(stimuliindex, seed, noise, flip, type) %>%
  summarise(mean.y = mean(y), median.y = median(y))

error_df <- exp2 %>%
  left_join(ans_df) %>%
  mutate(
    error_mean = reportedaverage - mean.y, # in the original file, mean.y \equiv data_average
    error_median = reportedaverage - median.y
  ) %>%
  rename(variability = noise, mark = type) %>%
  mutate(variability = factor(variability, levels = c("c(0, 0)", "c(0, 0.4)")))

error_df %>% head(5)
```

Results with "noramlized" true error: 

```{r}
#| code-fold: true

p1 <- error_df %>%
  mutate(
    normalized_error_mean = ifelse(flip == FALSE, error_mean, -error_mean)
  ) %>%
  summarySEwithin(
    .,
    measurevar = "normalized_error_mean",
    withinvars = "variability",
    betweenvars = "mark",
    idvar = "assignmentid"
  ) %>%
  plot_results(., normalized_error_mean) +
  labs(x = "Normalized True Error") +
  geom_vline(xintercept = 0, linetype = "dashed")

p1
```

:::{.callout-note collapse="true"}
To compare the two: 

```{r}
(p_og + coord_cartesian(xlim = c(-0.01, 0.1))) /
  (p1 + coord_cartesian(xlim = c(-0.01, 0.1)) + theme(legend.position = ""))
```

One could tell that the result for `point` is preserved, while the result for `point_arc` is completely reversed ... 

:::

Results with true error: 

```{r}
#| code-fold: true

error_df %>%
  summarySEwithin(
    .,
    measurevar = "error_mean",
    withinvars = "variability",
    betweenvars = "mark",
    idvar = "assignmentid"
  ) %>%
  plot_results(., error_mean) +
  labs(x = "True Error") +
  geom_vline(xintercept = 0, linetype = "dashed")
```

Results with true error + additional withinvars factor of `flip`: 

```{r}
#| code-fold: true

error_df %>%
  summarySEwithin(
    .,
    measurevar = "error_mean",
    withinvars = c("variability", "flip"),
    betweenvars = "mark",
    idvar = "assignmentid"
  ) %>%
  plot_results(., error_mean) +
  labs(x = "True Error") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_wrap(~flip, ncol = 1)
```

It seems to me that there are several things that are unexplained in the original analysis: 

- When the original plot is flipped (when `flip == TRUE`), there seems to be a bigger bias compared to the case when it's not flipped. 
- When calculating error using the true mean, instead of the mean for the plot with the same seed, we don't see the same trend that more variability implies bigger error. 

---
  
# Do participants who use Median Strategy get answer closer to median or mean? 

## Read in data related to strategy 

Question to answer: people who claim to use mean strategy vs people who claim to use median strategy. **Is there a difference between their responses/errors?** 
  
```{r}
strategy_df <- read.csv(here::here(
  "data",
  "moritz",
  "exp",
  "strategies.csv"
)) %>%
  as_tibble(.) %>%
  # keep only those related to experiment 2, and only related to point and point arc condition
  filter(Exp == 2, Condition %in% c("point", "point_arc")) %>%
  select(1:4, contains("mental_averaging"), contains("number_of_points"))

strategy_df %>% head(.)
```

Where do the two raters (`r1` and `r2`) disagree? 

```{r}
# for mental averaging -- 22 participants
strategy_df %>% filter(r1_mental_averaging != r2_mental_averaging)
# for equal number of points -- 9 participants
strategy_df %>% filter(r1_number_of_points != r2_number_of_points)
```

Where do they agree? 

```{r}
# for mental averaging --- 98 participants where both raters agree
strategy_df %>%
  filter(r1_mental_averaging == r2_mental_averaging) %>%
  filter(r1_mental_averaging == TRUE) %>%
  pull(ResponseId) -> mental_average_id
# for equal number of points -- 44 participants where both raters agree
strategy_df %>%
  filter(r1_number_of_points == r2_number_of_points) %>%
  filter(r1_number_of_points == TRUE) %>%
  pull(ResponseId) -> equal_points_id
```

I should note that there is also some overlap between those who perform mental averaging and median: 

```{r}
intersect(mental_average_id, equal_points_id)
```

What about the distribution? Do participants who view `point_arc` and `point` use these strategies about equally or is there a preference? 

```{r}
exp2 <- raw_exp2 %>%
  select(assignmentid, mark, flip, variability, error, normalized_error)

exp2 %>%
  filter(assignmentid %in% mental_average_id) %>%
  group_by(mark) %>%
  count(.)

exp2 %>%
  filter(assignmentid %in% equal_points_id) %>%
  group_by(mark) %>%
  count(.)
```
