---
title: "Reanalyze Mortiz"
---

```{r}
#| echo: false
#| label: load libraries
library(tidyverse)
library(tidybayes)
library(here)
library(ggplot2)
theme_set(theme_tidybayes())
```

## TLDR 

The goal of this file is to (1) recreate the analysis that Moritz et al. produced for their second experiment, and (2) perform a different kind of error analysis, as well as analyze particpant's whose strategies are closely mapped to the mental averaging and median strategy. 

Reading in data from Exp 2 (excluding the `line` condition): 

```{r}
#Exp 2
#point arc
exp2_dotArc <- read.csv(
  here("data", "moritz", "exp", "exp2-point-arc.csv"),
  ,
  check.names = FALSE
)
#point
exp2_dotX <- read.csv(
  here("data", "moritz", "exp", "exp2-point.csv"),
  check.names = FALSE
)
# Combining conditions in Exp 2
exp2 <- rbind(exp2_dotArc, exp2_dotX) %>% as_tibble(.)
```

Get a basic idea of the underlying dataset: 

```{r}
exp2 %>% count(assignmentid) # 277 participants altogether
```

## Read in data related to strategy 

```{r}
strategy_df <- read.csv(here("data", "moritz", "exp", "strategies.csv")) %>%
  as_tibble(.) %>%
  # keep only those related to experiment 2, and only related to point and point arc condition
  filter(Exp == 2, Condition %in% c("point", "point_arc")) %>%
  select(1:4, contains("mental_averaging"), contains("number_of_points"))
```

Where do the two raters (`r1` and `r2`) disagree? 

```{r}
# for mental averaging
strategy_df %>% filter(r1_mental_averaging != r2_mental_averaging)
# for equal number of points
strategy_df %>% filter(r1_number_of_points != r2_number_of_points)
```

Where do they agree? 

```{r}
# for mental averaging --- 98 participants where both raters agree
strategy_df %>%
  filter(r1_mental_averaging == r2_mental_averaging) %>%
  filter(r1_mental_averaging == TRUE) %>%
  pull(ResponseId) -> mental_average_id
# for equal number of points -- 44 participants where both raters agree
strategy_df %>%
  filter(r1_number_of_points == r2_number_of_points) %>%
  filter(r1_number_of_points == TRUE) %>%
  pull(ResponseId) -> equal_points_id
```

## Calculate Error and plot empirical error distributions 

Read in the stimuli and calculate the answer for each `index` * `seed` 

```{r}
#| code-fold: true
exp_json_data <- jsonlite::fromJSON(here(
  "data",
  "moritz",
  "stimuli",
  "exp2-stimuli.json"
)) %>%
  as_tibble(.)

stimuli_df <- exp_json_data %>%
  select(index, seed, noise, flip, type, data) %>%
  mutate(noise = as.character(noise)) %>%
  filter(type != "line") %>%
  unnest(col = data)

ans_df <- stimuli_df %>%
  group_by(index, seed, noise, flip, type) %>%
  summarise(ans.mean.y = mean(y), ans.median.y = median(y))
```

```{r}
ans_df
```

Use the `data_average` to calculate error, and this is what we in terms of empirical distribution of error: 

```{r}
#| code-fold: true
#| layout-ncol: 2
exp2 %>%
  mutate(error = reportedaverage - data_average) %>%
  ggplot(aes(x = error, y = variability, fill = mark)) +
  stat_halfeye(alpha = 0.8)

exp2 %>%
  mutate(
    norm.error = ifelse(
      flip == FALSE,
      reportedaverage - data_average,
      data_average - reportedaverage
    )
  ) %>%
  ggplot(aes(x = norm.error, y = variability, fill = mark)) +
  stat_halfeye(alpha = 0.8)
```

We could also try to facet by whether y-axis is flipped, and this will allow us to separate plotting: 

```{r}
#| code-fold: true
exp2 %>%
  mutate(
    error = reportedaverage - data_average,
    variance_area = ifelse(flip == FALSE, "high", "low")
  ) %>%
  ggplot(aes(x = error, y = variability, fill = mark)) +
  stat_halfeye(alpha = 0.8) +
  facet_wrap(~variance_area)
```

If we use `average` instead of `data_average` to calculate error, this is what we will get: 

```{r}
#| code-fold: true
exp2 %>%
  mutate(
    error = reportedaverage - average,
    variance_area = ifelse(flip == FALSE, "high", "low")
  ) %>%
  ggplot(aes(x = error, y = variability, fill = mark)) +
  stat_halfeye(alpha = 0.8) +
  facet_wrap(~variance_area) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "darkgray")
```

# Analyses 

Note that the plots produced in the paper, specifically, Figure 2, it was really a summary statistics plot. So we first try to replicate the original correction: 

```{r}
#| code-fold: true
#| eval: false

# when mark == "point", you get average == data_average
# which makes sense, because for `point_arc`, the data average is NOT the actual average ...

exp2 %>%
  mutate(
    error = round(error, 5),
    normalized_error = round(normalized_error, 5)
  ) %>%
  # verified that this is how normalized error is calculated
  mutate(
    calc_error = ifelse(
      flip == FALSE,
      reportedaverage - average,
      average - reportedaverage
    ) %>%
      round(., 5),
  ) %>%
  filter(calc_error != normalized_error)
```

```{r}
#| code-fold: true
exp2 %>%
  group_by(assignmentid, mark) %>%
  mutate(subj_mean = mean(normalized_error)) %>%
  ungroup() %>%
  mutate(grand_mean = mean(normalized_error)) %>%
  mutate(corrected = normalized_error - subj_mean + grand_mean) %>%
  group_by(mark, variability) %>%
  summarise(
    mu = mean(normalized_error),
    se = sd(corrected) / sqrt(n()) * sqrt(2 / (2 - 1)), # k is 2
    .groups = "drop"
  )
```

Plot original: 

```{r}
#| label: plot Mortiz
#| code-fold: true
exp2 %>%
  group_by(assignmentid, mark) %>%
  mutate(subj_mean = mean(normalized_error)) %>%
  ungroup() %>%
  mutate(grand_mean = mean(normalized_error)) %>%
  mutate(corrected = normalized_error - subj_mean + grand_mean) %>%
  group_by(mark, variability) %>%
  summarise(
    mu = mean(normalized_error),
    se = sd(corrected) / sqrt(n()) * sqrt(2 / (2 - 1)), # k is 2
    .groups = "drop"
  ) %>%
  ggplot(aes(
    x = variability,
    y = mu,
    dist = "norm",
    arg1 = mu,
    arg2 = se,
    fill = mark
  )) + #uncertainty
  stat_dist_halfeye(
    show_interval = TRUE,
    position = position_dodge(width = 0.2), # <- WHY THIS???
    .width = 0.95,
    size = .5,
    alpha = .8
  ) + #adding halfeye
  coord_flip() + #flipping axis
  theme_tidybayes() + #theme
  theme(legend.position = "top") +
  scale_fill_manual(values = c("#76B7B2", "#E15759")) + # take out "#59A14F"
  scale_y_continuous(
    breaks = seq(0, .11, by = .01),
    expand = expansion(mult = c(0.01, 0.01))
  ) + #specifying scale
  xlab("Variability") + #add x label
  ylab("Normalized Error") #add y label
```

---

Plot the graph with `error` instead of `normalized_error` and also its facetted version: 

`#TODO:` the following needs to be checked because it does not look right ....  

```{r}
#| code-fold: true
exp2 %>%
  # Step 1: collapse to subject-level means per cell (now includes flipped)
  group_by(assignmentid, mark, variability, flip) %>%
  summarise(subj_cell_mean = mean(error), .groups = "drop") %>%

  # Step 2: subject mean across ALL within-subject conditions
  group_by(assignmentid, mark) %>%
  mutate(subj_mean = mean(subj_cell_mean)) %>%
  ungroup() %>%

  # Step 3: normalize
  mutate(grand_mean = mean(subj_cell_mean)) %>%
  mutate(corrected = subj_cell_mean - subj_mean + grand_mean) %>%

  # Step 4: SE per cell (now includes flipped)
  group_by(mark, variability, flip) %>%
  summarise(
    mu = mean(subj_cell_mean),
    n_subj = n(),
    se = sd(corrected) / sqrt(n_subj) * sqrt(4 / (4 - 1)),
    .groups = "drop"
  ) %>%
  ggplot(aes(
    x = variability,
    y = mu,
    dist = "norm",
    arg1 = mu,
    arg2 = se,
    fill = mark
  )) + #uncertainty
  stat_dist_halfeye(
    show_interval = TRUE,
    position = position_dodge(width = 0.2), # <- WHY THIS???
    .width = 0.95,
    size = .5,
    alpha = .8
  ) + #adding halfeye
  coord_flip() + #flipping axis
  theme_tidybayes() + #theme
  facet_grid(~ as.character(flip)) +
  theme(legend.position = "top") +
  scale_fill_manual(values = c("#76B7B2", "#E15759")) + # take out "#59A14F"
  # scale_y_continuous(
  #   breaks = seq(0, .11, by = .01),
  #   expand = expansion(mult = c(0.01, 0.01))
  # ) + #specifying scale
  xlab("Variability") + #add x label
  ylab("Normalized Error") + #add y label +
  geom_hline(yintercept = 0, linetype = "dashed", color = "darkgray")
```

# Do participants who use Median Strategy get answer closer to median or mean? 

```{r}
median_df <- exp2 %>%
  filter(assignmentid %in% equal_points_id) %>%
  rename(type = mark, index = stimuliindex)
```

```{r}
#| code-fold: true
#| layout-ncol: 2

median_df %>%
  left_join(ans_df) %>%
  mutate(
    error_avg = reportedaverage - ans.mean.y,
    error_median = reportedaverage - ans.median.y
  ) %>%
  ggplot(aes(x = error_avg, fill = type, y = noise)) +
  stat_halfeye() +
  geom_vline(xintercept = 0, linetype = "dashed", color = "darkgray") +
  labs(title = "error calculated using mean")

median_df %>%
  left_join(ans_df) %>%
  mutate(
    error_avg = reportedaverage - ans.mean.y,
    error_median = reportedaverage - ans.median.y
  ) %>%
  ggplot(aes(x = error_median, fill = type, y = noise)) +
  stat_halfeye() +
  geom_vline(xintercept = 0, linetype = "dashed", color = "darkgray") +
  labs(title = "error calculated using median")
```

What if we do the same analysis as did before? 

```{r}
#| code-fold: true
#| layout-ncol: 2

median_df %>%
  left_join(ans_df) %>%
  mutate(
    error = ifelse(
      flip == FALSE,
      reportedaverage - ans.mean.y,
      ans.mean.y - reportedaverage
    )
  ) %>%
  group_by(assignmentid, type) %>%
  mutate(subj_mean = mean(error)) %>%
  ungroup() %>%
  mutate(grand_mean = mean(error)) %>%
  mutate(corrected = error - subj_mean + grand_mean) %>%
  group_by(type, variability) %>%
  summarise(
    mu = mean(error),
    se = sd(corrected) / sqrt(n()) * sqrt(2 / (2 - 1)), # k is 2
    .groups = "drop"
  ) %>%
  ggplot(aes(
    x = variability,
    y = mu,
    dist = "norm",
    arg1 = mu,
    arg2 = se,
    fill = type
  )) + #uncertainty
  stat_dist_halfeye(
    show_interval = TRUE,
    position = position_dodge(width = 0.2), # <- WHY THIS???
    .width = 0.95,
    size = .5,
    alpha = .8
  ) + #adding halfeye
  coord_flip() + #flipping axis
  theme_tidybayes() + #theme
  theme(legend.position = "top") +
  scale_fill_manual(values = c("#76B7B2", "#E15759")) + # take out "#59A14F"
  scale_y_continuous(
    breaks = seq(0, .11, by = .01),
    expand = expansion(mult = c(0.01, 0.01))
  ) + #specifying scale
  xlab("Variability") + #add x label
  ylab("Normalized Error") + #add y label
  labs(title = "Error calculated using mean")

median_df %>%
  left_join(ans_df) %>%
  mutate(
    error = ifelse(
      flip == FALSE,
      reportedaverage - ans.median.y,
      ans.median.y - reportedaverage
    )
  ) %>%
  group_by(assignmentid, type) %>%
  mutate(subj_mean = mean(error)) %>%
  ungroup() %>%
  mutate(grand_mean = mean(error)) %>%
  mutate(corrected = error - subj_mean + grand_mean) %>%
  group_by(type, variability) %>%
  summarise(
    mu = mean(error),
    se = sd(corrected) / sqrt(n()) * sqrt(2 / (2 - 1)), # k is 2
    .groups = "drop"
  ) %>%
  ggplot(aes(
    x = variability,
    y = mu,
    dist = "norm",
    arg1 = mu,
    arg2 = se,
    fill = type
  )) + #uncertainty
  stat_dist_halfeye(
    show_interval = TRUE,
    position = position_dodge(width = 0.2), # <- WHY THIS???
    .width = 0.95,
    size = .5,
    alpha = .8
  ) + #adding halfeye
  coord_flip() + #flipping axis
  theme_tidybayes() + #theme
  theme(legend.position = "top") +
  scale_fill_manual(values = c("#76B7B2", "#E15759")) + # take out "#59A14F"
  scale_y_continuous(
    breaks = seq(0, .11, by = .01),
    expand = expansion(mult = c(0.01, 0.01))
  ) + #specifying scale
  xlab("Variability") + #add x label
  ylab("Normalized Error") + #add y label
  labs(title = "Error calculated using median")

```
