---
title: "Reanalyze Mortiz"
author: "Anonymous" 
date: last-modified
code-fold: true 
format: 
  html: 
    toc: true
    reference-location: margin
    embed-resources: true
---

```{r}
#| echo: false
#| label: load libraries
library(plyr)
source(here::here("analysis", "moritz", "Functions.R"))
library(tidyverse)
library(tidybayes)
library(ggplot2)
library(patchwork)
theme_set(theme_tidybayes())
```

## TLDR 

The goal is to: 

1. Recreate the analysis that Moritz et al. produced for their second experiment, and 
2. Perform a different kind of error analysis.
<!-- as well as analyze particpant's whose strategies are closely mapped to the mental averaging and median strategy.  -->

Reading in data from Exp 2 (excluding the `line` condition): 

# Repeating Mortiz's analysis

First we read in the relevant data: 

```{r}
#| label: read in Mortiz data
#point arc
exp2_dotArc <- read.csv(
  here::here("data", "moritz", "exp", "exp2-point-arc.csv"),
  check.names = FALSE
)

#point
exp2_dotX <- read.csv(
  here::here("data", "moritz", "exp", "exp2-point.csv"),
  check.names = FALSE
)

# line
exp2_line <- read.csv(
  here::here("data", "moritz", "exp", "exp2-line.csv"),
  check.names = FALSE
)

# combine
raw_exp2 <- rbind(exp2_dotArc, exp2_dotX, exp2_line) %>%
  as_tibble(.)

exp2 <- raw_exp2 %>%
  select(assignmentid, mark, flip, variability, error, normalized_error)

exp2 %>% head(5)
```

We also read in the original stimuli, with the `line` condition filtered out: 

```{r}
exp_json_data <- jsonlite::fromJSON(
  here::here("data", "moritz", "stimuli", "exp2-stimuli.json")
  # paste0(getwd(), "/data/moritz/stimuli/exp2-stimuli.json"),
) %>%
  as_tibble(.)

stimuli_df <- exp_json_data %>%
  select(index, seed, noise, flip, type, data) %>%
  mutate(noise = as.character(noise)) %>%
  filter(type != "line")

stimuli_df %>% head(5)
```

Define helper function for plotting^[Borrowed from what the og analysis file had]:

```{r}
#| label: helper function for plotting 
plot_results <- function(data, xvar) {
  data %>%
    ggplot(aes(
      y = variability,
      x = {{ xvar }},
      dist = "norm",
      arg1 = {{ xvar }},
      arg2 = se,
      fill = mark
    )) + #uncertainty
    stat_dist_halfeye(
      show_interval = TRUE,
      position = position_dodge(width = 0.2),
      .width = 0.95,
      size = .5,
      alpha = .8
    ) +
    theme(legend.position = "top") +
    scale_fill_manual(values = c("#59A14F", "#76B7B2", "#E15759")) +
    scale_y_discrete(
      labels = c("0,0" = "0", "0,0.4" = "0.4"),
      expand = expansion(add = c(0.1, 0.5))
    )
}
```

## Recreate Moritz et al. plot 

For our own sanity, recreate the plot that Moritz et al. created: 

```{r}
#| label: plot Moritz OG 
p_og <- exp2 %>%
  summarySEwithin(
    data = .,
    measurevar = "normalized_error",
    withinvars = "variability",
    betweenvars = "mark",
    idvar = "assignmentid"
  ) %>%
  # mutate(variability = factor(variability)) %>%
  plot_results(., normalized_error) +
  theme(text = element_text(size = 9), legend.position = "none") +
  scale_x_continuous(
    breaks = seq(0, 0.09, by = 0.01)
  )

p_og # + theme(aspect.ratio = 1/4)
```

```{r}
#| eval: false
#| echo: false
ggsave(
  here::here("figs/og_recreated.svg"),
  p_og,
  width = 3.5,
  height = 2,
  units = "in"
)
```

## Adding Additional Factors to OG Graph 

```{r}
p_og_facet <- exp2 %>%
  summarySEwithin(
    data = .,
    measurevar = "error",
    withinvars = c("variability", "flip"),
    betweenvars = "mark",
    idvar = "assignmentid"
  ) %>%
  plot_results(., error) +
  # facet_wrap(flip ~ ., ncol = 1) +
  theme(text = element_text(size = 9), legend.position = "none") +
  scale_x_continuous(
    breaks = seq(-1, 1, by = 0.01)
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray")

p_og_facet
```

```{r}
#| echo: false 
#| eval: false 
ggsave(
  here::here("figs/p_og_facet.svg"),
  p_og_facet,
  width = 3.5,
  height = 2,
  units = "in",
  dpi = 300
)
```

## Recalculate error 

First, let's see what's the difference between the average data average and the average that Moritz et al. used for their error analysis. Let's first look at how Moritz et al is reporting things in their dataset: 

```{r}
raw_exp2 %>%
  select(stimuliindex, seed, mark, flip, variability, average, data_average) %>%
  group_by(
    stimuliindex,
    seed,
    mark,
    flip,
    variability,
    average,
    data_average
  ) %>%
  summarise(.) %>%
  print(n = 8)
```

Observe that for stimuli that share the same `seed`, the `data_average` is the actual average of the dataset, while the `average` is the same for stimuli index that share the same `seed`, `flip`, and `variability`^[for example examine `stimuliindex == 4` and ` stimuliindex == 8`]. 

:::{.callout-note collapse="true"}
# Verify that `data_average` is true average 

To verify that it is indeed the case the `data_average` is true average, we can look at the actual stimuli that Moritz et al. provided: 

```{r}
ans_df <- stimuli_df %>%
  rename(stimuliindex = index) %>%
  unnest(data) %>%
  group_by(stimuliindex, seed, noise, flip, type) %>%
  rename(mark = type, variability = noise) %>%
  mutate(variability = ifelse(variability == "c(0, 0)", "0,0", "0,0.4")) %>%
  summarise(mean.y = mean(y))

ans_df
```

Join and check: 

```{r}
ans_df %>%
  left_join(
    raw_exp2 %>%
      select(
        stimuliindex,
        seed,
        mark,
        flip,
        variability,
        average,
        data_average
      ) %>%
      group_by(
        stimuliindex,
        seed,
        mark,
        flip,
        variability,
        average,
        data_average
      ) %>%
      summarise(.)
  ) %>%
  ungroup(.) %>%
  count(near(mean.y, data_average))
```

Thus this completes the verification. Indeed `data_average` corresponds to the actual average. 

:::

This allows us to calculate two kinds of error: (1) reference error, i.e., error that uses the reference average, and (2) actual error, i.e., error that uses the true underyling dataset. 

```{r}
exp2 <- raw_exp2 %>%
  select(
    assignmentid,
    mark,
    flip,
    variability,
    reportedaverage,
    average,
    data_average,
    error
  )

exp2 %>% head(5)
```

Check to see that the original reported `error` is obtained via difference between `reportedaverage` and `average`: 

```{r}
exp2 %>% count(near(reportedaverage - average, error))
```

Apparently there are two rows where it's not near ...? Let's check: 

```{r}
exp2 %>%
  filter(!near(reportedaverage - average, error)) %>%
  mutate(calc_error = reportedaverage - average)
```

This appears to just be due to rounding error. So this confirms that indeed `error = reportedaverage - average`. 

## Reanalyze with different error definition

Using the exact same calculations, we obtain the following: 

```{r}
p_NewError <- exp2 %>%
  rename(ref_error = error) %>%
  mutate(
    actual_error = reportedaverage - data_average,
    normalized_actual_error = ifelse(flip == TRUE, -actual_error, actual_error)
  ) %>%
  summarySEwithin(
    data = .,
    measurevar = "normalized_actual_error",
    withinvars = "variability",
    betweenvars = "mark",
    idvar = "assignmentid"
  ) %>%
  plot_results(., normalized_actual_error) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray") +
  theme(text = element_text(size = 9), legend.position = "none") +
  coord_cartesian(xlim = c(-0.01, 0.08)) +
  scale_x_continuous(
    breaks = seq(-1, 1, by = 0.01)
  )

p_NewError
```

```{r}
#| echo: false 
#| eval: false 
ggsave(
  here::here("figs/newError.svg"),
  p_NewError,
  width = 3.5,
  height = 2,
  units = "in",
  dpi = 300
)
```

If we don't use the "normalization", and ~~facet by the factor of `flip`, then we get this:~~ (following Matt's advice we are not going to facet by `flip`) 

```{r}
exp2 %>%
  rename(ref_error = error) %>%
  mutate(
    actual_error = reportedaverage - data_average,
    normalized_actual_error = ifelse(flip == TRUE, -actual_error, actual_error)
  ) %>%
  summarySEwithin(
    data = .,
    measurevar = "actual_error",
    withinvars = c("variability", "flip"),
    betweenvars = "mark",
    idvar = "assignmentid"
  ) %>%
  plot_results(., actual_error) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray") # +
# facet_grid(flip ~ .)
```

We can also make the direction look nicer and easier to read: 

```{r}
p_newError_facet <- exp2 %>%
  rename(ref_error = error) %>%
  mutate(
    actual_error = reportedaverage - data_average,
    normalized_actual_error = ifelse(flip == TRUE, -actual_error, actual_error)
  ) %>%
  summarySEwithin(
    data = .,
    measurevar = "actual_error",
    withinvars = c("variability", "flip"),
    betweenvars = "mark",
    idvar = "assignmentid"
  ) %>%
  plot_results(., actual_error) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray") +
  # facet_wrap(flip ~ ., ncol = 1) +
  theme(text = element_text(size = 9), legend.position = "none") +
  scale_x_continuous(
    breaks = seq(-1, 1, by = 0.01)
  )

p_newError_facet
```

```{r}
#| echo: false 
#| eval: false 
ggsave(
  here::here("figs/newError_facet.svg"),
  p_newError_facet,
  width = 3.5,
  height = 2,
  units = "in"
)
```

---

It seems to me that there are several things that are unexplained in the original analysis: 

- When the original plot is flipped (when `flip == TRUE`), there seems to be a bigger bias compared to the case when it's not flipped. 
- When calculating error using the true mean, instead of the mean for the plot with the same seed, we don't see the same trend that more variability implies bigger error. 

## What is the relationship between actual mean and reference mean, vs the plot type? 

As the title states, for the plots that share the same reference line, does being `pointArc` imply more variability? 

```{r}
s_df <- raw_exp2 %>%
  select(
    stimuliindex,
    seed,
    mark,
    flip,
    variability,
    average,
    data_average
  ) %>%
  group_by(
    stimuliindex,
    seed,
    mark,
    flip,
    variability,
    average,
    data_average
  ) %>%
  summarise(.)

s_df %>% head(5)
```

Now let's plot the density and mean: 

```{r}
#| layout-ncol: 2
means_df <- s_df %>%
  group_by(flip, variability, mark) %>%
  summarize(mean_val = mean(data_average), .groups = "drop")

s_df %>%
  ggplot(aes(x = data_average, color = mark)) +
  geom_density() +
  geom_vline(
    data = means_df,
    aes(xintercept = mean_val, color = mark),
    linetype = "dashed"
  ) +
  facet_wrap(flip + variability ~ .)

s_df %>%
  ggplot(aes(x = data_average, color = mark)) +
  geom_density() +
  stat_pointinterval(
    aes(y = -0.05),
    point_interval = mean_qi,
    position = position_dodge(width = 0.4)
  ) +
  facet_wrap(flip + variability ~ .)
```