---
title: "Reanalyze Mortiz"
author: "Anonymous" 
---

```{r}
#| echo: false
#| label: load libraries

# library(plyr)
# source(paste0(getwd(), "analysis/moritz/Functions.R"))
library(tidyverse)
library(tidybayes)
library(here)
library(ggplot2)
theme_set(theme_tidybayes())
```

## TLDR 

The goal of this file is to (1) recreate the analysis that Moritz et al. produced for their second experiment, and (2) perform a different kind of error analysis, as well as analyze particpant's whose strategies are closely mapped to the mental averaging and median strategy. 

Reading in data from Exp 2 (excluding the `line` condition): 

# Repeating Mortiz's analysis 

```{r}
#| label: read in Mortiz data

#point arc
exp2_dotArc <- read.csv(
  here("data", "moritz", "exp", "exp2-point-arc.csv"),
  ,
  check.names = FALSE
)
#point
exp2_dotX <- read.csv(
  here("data", "moritz", "exp", "exp2-point.csv"),
  check.names = FALSE
)
exp2 <- rbind(exp2_dotArc, exp2_dotX) %>%
  as_tibble(.) %>%
  select(assignmentid, mark, flip, variability, error, normalized_error)

exp2 %>% head(5)
```

Perform the correction: 

```{r}
#| code-fold: true
exp2 %>%
  group_by(assignmentid, mark) %>%
  mutate(subj_mean = mean(normalized_error)) %>%
  ungroup() %>%
  mutate(grand_mean = mean(normalized_error)) %>%
  mutate(corrected = normalized_error - subj_mean + grand_mean) %>%
  group_by(mark, variability) %>%
  summarise(
    mu = mean(normalized_error),
    se = sd(corrected) / sqrt(n()) * sqrt(2 / (2 - 1)), # k is 2 when the within-subj variability is this
    .groups = "drop"
  )
```

If we do it the same way by calling functions in `Functions.R` this is what we get: 

```{r}
#| eval: false
freq_sum <- summarySEwithin(
  data = exp2, #data
  measurevar = "normalized_error", #dependent variable
  withinvars = "variability", #within subjects variable
  betweenvars = "mark", #between subject variable
  idvar = "assignmentid"
) #participant id
freq_sum
```

Seems like the numbers match each other, which implies that our approach is correct. So we will use our dataframe to plot: 

Recreate the plot that Moritz et al. created: 

```{r}
#| label: plot Moritz
#| code-fold: true

# detach("package:plyr", unload = TRUE)

exp2 %>%
  group_by(assignmentid, mark) %>%
  mutate(subj_mean = mean(normalized_error)) %>%
  ungroup() %>%
  mutate(grand_mean = mean(normalized_error)) %>%
  mutate(corrected = normalized_error - subj_mean + grand_mean) %>%
  group_by(mark, variability) %>%
  summarise(
    mu = mean(normalized_error),
    se = sd(corrected) / sqrt(n()) * sqrt(2 / (2 - 1)), # k is 2
    .groups = "drop"
  ) %>%
  ggplot(aes(
    x = variability,
    y = mu,
    dist = "norm",
    arg1 = mu,
    arg2 = se,
    fill = mark
  )) + #uncertainty
  stat_dist_halfeye(
    show_interval = TRUE,
    position = position_dodge(width = 0.2), # <- WHY THIS???
    .width = 0.95,
    size = .5,
    alpha = .8
  ) + #adding halfeye
  coord_flip() + #flipping axis
  theme_tidybayes() + #theme
  theme(legend.position = "top") +
  scale_fill_manual(values = c("#76B7B2", "#E15759")) + # take out "#59A14F"
  scale_y_continuous(
    breaks = seq(0, .11, by = .01),
    expand = expansion(mult = c(0.01, 0.01))
  ) + #specifying scale
  xlab("Variability") + #add x label
  ylab("Normalized Error") #add y label
```

## Analyzed signed error 

Instead of calculating corrected error, we will first calculate error that still uses the "wrong" mean, but is not "normalized": 

```{r}
#| code-fold: true

freq_summary <- exp2 %>%
  group_by(assignmentid, mark) %>%
  mutate(subj_mean = mean(error)) %>%
  ungroup() %>%
  mutate(grand_mean = mean(error)) %>%
  mutate(corrected = error - subj_mean + grand_mean) %>%
  group_by(mark, variability, flip) %>%
  summarise(
    mu = mean(error),
    se = sd(corrected) / sqrt(n()) * sqrt(4 / (4 - 1)), # <- NEW: k is 4
    .groups = "drop"
  )

freq_summary
```

```{r}
#| code-fold: true

p1 <- freq_summary %>%
  mutate(
    flip_label = ifelse(flip == TRUE, "Variability Below", "Variability Above"),
    mu_display = ifelse(flip == FALSE, mu, -mu)
  ) %>%
  ggplot(aes(
    y = variability,
    x = mu_display,
    dist = "norm",
    arg1 = mu_display,
    arg2 = se,
    fill = mark
  )) +
  stat_dist_halfeye(
    show_interval = TRUE,
    position = position_dodge(width = 0.2),
    .width = 0.95,
    size = .5,
    alpha = .8
  ) +
  theme_tidybayes() +
  theme(legend.position = "top") +
  scale_fill_manual(values = c("#76B7B2", "#E15759")) +
  # scale_x_continuous(
  #   breaks = seq(0, .11, by = .01),
  #   expand = expansion(mult = c(0.01, 0.01))
  # ) +
  facet_wrap(~flip_label, ncol = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray") +
  xlab("Error (adjusted toward variability direction)") +
  ylab("Variability")

p1
```

Ok so the above tells us that if we look at normal, calculated by looking at the mean in which people share mean, then the above is what we get. 

## Recalculate error 

It is unclear to me how `error` is calculated in the original dataset. 

```{r}
exp2 <- rbind(exp2_dotArc, exp2_dotX) %>%
  as_tibble(.)

# normalized error is just error but flipped when flip is TRUE
# the following confirms it.
exp2 %>% count(mark, flip, error == normalized_error)

# and error is just reportedaverage - average
# confirmed by the following
exp2 %>%
  mutate(
    error.1 = reportedaverage - data_average,
    error.2 = reportedaverage - average,
    error.3 = reportedaverage - average_along_arc,
  ) %>%
  select(seed, error.1, error.2, error.3, error, normalized_error) %>%
  arrange(seed) %>%
  filter(!near(error.2, error)) # using dplyr::near becasue of floating point related stuff ....
```

So let's calculate the true error and plot things: 

```{r}
#| code-fold: true

freq_summary <- exp2 %>%
  mutate(true_error = reportedaverage - data_average) %>%
  group_by(assignmentid, mark) %>%
  mutate(subj_mean = mean(true_error)) %>%
  ungroup() %>%
  mutate(grand_mean = mean(true_error)) %>%
  mutate(corrected = true_error - subj_mean + grand_mean) %>%
  group_by(mark, variability, flip) %>%
  summarise(
    mu = mean(true_error),
    se = sd(corrected) / sqrt(n()) * sqrt(4 / (4 - 1)), # <- NEW: k is 4
    .groups = "drop"
  )

freq_summary
```

```{r}
#| code-fold: true

p2 <- freq_summary %>%
  mutate(
    flip_label = ifelse(flip == TRUE, "Variability Below", "Variability Above"),
    mu_display = ifelse(flip == FALSE, mu, -mu)
  ) %>%
  ggplot(aes(
    y = variability,
    x = mu_display,
    dist = "norm",
    arg1 = mu_display,
    arg2 = se,
    fill = mark
  )) +
  stat_dist_halfeye(
    show_interval = TRUE,
    position = position_dodge(width = 0.2),
    .width = 0.95,
    size = .5,
    alpha = .8
  ) +
  theme_tidybayes() +
  theme(legend.position = "top") +
  scale_fill_manual(values = c("#76B7B2", "#E15759")) +
  # scale_x_continuous(
  #   breaks = seq(0, .11, by = .01),
  #   expand = expansion(mult = c(0.01, 0.01))
  # ) +
  facet_wrap(~flip_label, ncol = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray") +
  xlab("Error (adjusted toward variability)") +
  ylab("Variability")

p2
```

So this new plot seems to suggest that for `point_arc`, when error is calculated toward the true mean, with more variability somehow we're more precise ...? It's kind of counter-intuitive ... But to put things into comparison: 

```{r}
library(patchwork)
(p1 +
  coord_cartesian(xlim = c(-0.01, 0.1))) /
  (p2 +
    coord_cartesian(xlim = c(-0.01, 0.1)))
```

The above shows up how different things could be if you use different "means" to calculate error ... Nothing changes for `point` because it's being calculated the right way, but for `point_arc` ... the variability trend basically reverses ... 

What is consistent, is that when the "variability" region is in the lower region, the error somehow increases ... which makes you wonder whether this could potentially be due to dragging ...? Or something else ...? 

---
  
## Read in data related to strategy 
  
```{r}
strategy_df <- read.csv(here("data", "moritz", "exp", "strategies.csv")) %>%
  as_tibble(.) %>%
  # keep only those related to experiment 2, and only related to point and point arc condition
  filter(Exp == 2, Condition %in% c("point", "point_arc")) %>%
  select(1:4, contains("mental_averaging"), contains("number_of_points"))
```

Where do the two raters (`r1` and `r2`) disagree? 

```{r}
# for mental averaging
strategy_df %>% filter(r1_mental_averaging != r2_mental_averaging)
# for equal number of points
strategy_df %>% filter(r1_number_of_points != r2_number_of_points)
```

Where do they agree? 

```{r}
# for mental averaging --- 98 participants where both raters agree
strategy_df %>%
  filter(r1_mental_averaging == r2_mental_averaging) %>%
  filter(r1_mental_averaging == TRUE) %>%
  pull(ResponseId) -> mental_average_id
# for equal number of points -- 44 participants where both raters agree
strategy_df %>%
  filter(r1_number_of_points == r2_number_of_points) %>%
  filter(r1_number_of_points == TRUE) %>%
  pull(ResponseId) -> equal_points_id
```

## Calculate Error and plot empirical error distributions 

Read in the stimuli and calculate the answer for each `index` * `seed` 

```{r}
#| code-fold: true
exp_json_data <- jsonlite::fromJSON(here(
  "data",
  "moritz",
  "stimuli",
  "exp2-stimuli.json"
)) %>%
  as_tibble(.)

stimuli_df <- exp_json_data %>%
  select(index, seed, noise, flip, type, data) %>%
  mutate(noise = as.character(noise)) %>%
  filter(type != "line") %>%
  unnest(col = data)

ans_df <- stimuli_df %>%
  group_by(index, seed, noise, flip, type) %>%
  summarise(ans.mean.y = mean(y), ans.median.y = median(y))
```





# Do participants who use Median Strategy get answer closer to median or mean? 

