---
title: "Analysis - Task 1"
format: 
  html: 
    code-fold: false
    reference-location: margin
    df-print: paged
    toc: true
    code-overflow: wrap
---

```{r}
#| echo: false
#| output: false
#| label: setup

library(tidyverse)
library(ggplot2)
library(sgt)
library(numDeriv)
library(ggdist)
# setting the theme 
theme_set(theme_minimal(base_size = 8))
# model
library(brms)
# plot results 
library(tidybayes)
library(cmdstanr)
library(posterior)
```

# Set up

Read in data 

```{r}
#| label: read-in-data
#| code-fold: true

task_df <- read_rds("task.rds")
participants <- read_rds("participants.rds")
```

```{r}
#| label: define-function
#| code-fold: true

# origin is top left  
data_to_pixel_y <- function(data_y) {
  return(-395 * data_y + 410)
}
data_to_pixel_x <- function(data_x) {
  return (53.5 * data_x + 317.5)
}

# origin is bottom left 
pixel_to_phy_x <- function(pixel, pxMM){
  (pixel - 50) / pxMM
}
pixel_to_phy_y <- function(pixel, pxMM){
  (410 - pixel) / pxMM
}

# return visual angle in degrees and not radian
vis_angle <- function(size, distance){
  return(2 * atan(size / (2 * distance)) * 180 / pi)
}

# tolerance for numerical precision 
tolerance <- 1e-10
```

# Task 1 

Load data: 

```{r}
#| code-fold: true 

task1_df <- task_df %>% filter(task == "task1") %>% # task_df
  select(-slider.x, -slider.y) %>% 
  right_join(participants, by = join_by(participantId)) %>% 
  mutate(data.select.left_area = psgt(data.select.x, param.mu, param.sigma, param.lambda, param.p, param.q, mean.cent = FALSE), 
         data.ans.x = qsgt(0.5, param.mu, param.sigma, param.lambda, param.p, param.q, mean.cent = FALSE)) %>% 
  # pixel related 
  mutate(pixel.med.x = data_to_pixel_x(data.ans.x), 
         pixel.mod.x = data_to_pixel_x(param.mu)) %>% 
  # phy related 
  mutate(phy.select.x = pixel_to_phy_x(pixel.select.x, pixelToMM), 
         phy.med.x = pixel_to_phy_x(pixel.med.x, pixelToMM), 
         phy.mod.x = pixel_to_phy_x(pixel.mod.x, pixelToMM)) %>% 
  # visual angle related 
  mutate(va.select.x = vis_angle(phy.select.x, dist_to_screen), 
         va.med.x = vis_angle(phy.med.x, dist_to_screen), 
         va.mod.x = vis_angle(phy.mod.x, dist_to_screen))
```

Prepare data for stan: 

```{r}
N <- task1_df %>% nrow(.)
J <- length(unique(task1_df$participantId))
x <- task1_df %>% pull(va.select.x)
x_med <- task1_df %>% pull(va.med.x)
x_mod <- task1_df %>% pull(va.mod.x)

# change the participantId to numeric id 
# 1. Get unique participant IDs
unique_ids <- unique(task1_df$participantId)

# 2. Create a mapping
id_mapping <- setNames(1:length(unique_ids), unique_ids)

# 3. Create the numeric ID vector
numeric_id <- as.numeric(factor(task1_df$participantId, levels = names(id_mapping)))


stan_data <- list(
  N = N, 
  J = J, 
  id = numeric_id, 
  x = x, 
  x_med = x_med, 
  x_mod = x_mod
)
```

Define custom posterior predictive check function: 

```{r}
#| label: custom pp_check 

custom_pp_check <- function(
   fitted_model, 
   fun = bayesplot::ppc_dens_overlay, 
   ndraws = 100, 
   transform = identity,
   ...
 ) {
   y_rep <- as_draws_rvars(fitted_model)$y_rep |> 
     transform() |>
     as_draws_matrix()
   if (!is.null(ndraws)) {
     y_rep = y_rep |>
       merge_chains() |>
       resample_draws(ndraws = ndraws) 
   }
   y <- fitted_model$draws("x_org", format = "matrix")[1, ] |>
     as.vector() |>
     transform()
   fun(y, y_rep, ...)
 }
```

## Centered Parameterization 

Spoiler: Has divergent transitions. 

```{r}
#| eval: false 

model <- cmdstan_model("stan_files/task1.weighted.avg.stan")
model
``` 

Fit the stan model: 

```{r}
#| output: false 
#| eval: false 

fit <- model$sample(
  data = stan_data, 
  chains = 4, 
  iter_warmup = 2000,   # Number of warmup iterations (e.g., 1000)
  iter_sampling = 4000, 
  adapt_delta = 0.95 # Number of sampling iterations (e.g., 2000)
  # parallel_chains = 4
)
```

```{r}
summary <- fit$summary()
```



```{r}
custom_pp_check(fit, bayesplot::ppc_pit_ecdf_grouped, group = with(task1_df, case_when(param.lambda < 0 ~ "neg", param.lambda == 0 ~ "zero", param.lambda > 0 ~ "pos")))

custom_pp_check(fit, bayesplot::ppc_dens_overlay_grouped, group = ifelse(task1_df$param.lambda < 0, "neg", ifelse(task1_df$param.lambda > 0, "pos", "zero")))

custom_pp_check(fit, bayesplot::ppc_pit_ecdf, transform = \(x) x - task1_df$va.med.x, ndraws = NULL)

custom_pp_check(fit, bayesplot::ppc_pit_ecdf_grouped, group = ifelse(task1_df$param.lambda < 0, "neg", ifelse(task1_df$param.lambda > 0, "pos", "zero")))
```

```{r}
task1_df %>% mutate(y_rep = as_draws_rvars(fit)$y_rep - va.med.x) %>% 
  unnest_rvars() %>% 
  ggplot(aes(x = param.lambda, y = y_rep)) + 
  ggdist::stat_lineribbon() + 
  geom_point(aes(y = va.select.x - va.med.x)) + 
  scale_fill_brewer(palette = "Blues")
```

```{r}
task1_df %>% mutate(y_rep = as_draws_rvars(fit)$y_rep - va.med.x) %>% 
  unnest_rvars() %>% 
  ggplot(aes(x = va.med.x - va.mod.x, y = y_rep)) + 
  ggdist::stat_lineribbon() + 
  geom_point(aes(y = va.select.x - va.med.x)) + 
  facet_wrap(~ participantId) + 
  scale_fill_brewer(palette = "Blues")
```

## Non-centered parameterization --- Informative prior 

```{r}
#| output: false 

model <- cmdstan_model("stan_files/task1.weighted.avg.nonc.stan")

fit <- model$sample(
  data = stan_data, 
  chains = 4, 
  parallel_chains = 4, 
  iter_warmup = 4000,   
  iter_sampling = 4000, 
  adapt_delta = 0.99,  
  max_treedepth = 15
)

fit$save_object("models/task1.informative.prior.rds")

fit <- read_rds("models/task1.informative.prior.rds")
```

Posterior Predictive Check: 

```{r}
custom_pp_check(fit, bayesplot::ppc_dens_overlay_grouped, group = ifelse(task1_df$param.lambda < 0, "neg", ifelse(task1_df$param.lambda > 0, "pos", "zero")))

custom_pp_check(fit, bayesplot::ppc_pit_ecdf, transform = \(x) x - task1_df$va.med.x, ndraws = NULL)

custom_pp_check(fit, bayesplot::ppc_pit_ecdf_grouped, group = ifelse(task1_df$param.lambda < 0, "neg", ifelse(task1_df$param.lambda > 0, "pos", "zero")))
```

```{r}
#| layout-ncol: 2 

plot_data <- task1_df %>% mutate(y_rep = as_draws_rvars(fit)$y_rep - va.med.x) %>% # signed error 
  unnest_rvars() 

# param.lambda on x-axis 
plot_data %>% ggplot(aes(x = param.lambda, y = y_rep)) + 
  ggdist::stat_lineribbon(size = 0.5) + 
  geom_point(aes(y = va.select.x - va.med.x), size = 0.5) + 
  scale_fill_brewer(palette = "Blues")

# x.med - x.mod on X axis 
plot_data %>% ggplot(aes(x = va.med.x - va.mod.x, y = y_rep)) + 
  ggdist::stat_lineribbon(size = 0.5) + 
  geom_point(aes(y = va.select.x - va.med.x), alpha = 0.5, size = 0.5) + 
  # facet_wrap(~ participantId) + 
  scale_fill_brewer(palette = "Blues") + 
  labs(x = "x^med - x^mod", y = "Signed error") + 
  theme_minimal(base_size = 8) -> p

p

ggsave("plots/task1.pdf", p, height = 5, width = 10, unit = "cm")
```

```{r}
as_draws_rvars(fit)$mu_mod_mean
as_draws_rvars(fit)$mu_med_mean
as_draws_rvars(fit)$log_sigma_mod_mean %>% exp(.)
as_draws_rvars(fit)$log_sigma_med_mean %>% exp(.)
```

## Non-centered paramterization --- Uninformative prior 

```{r}
#| output: false 

model <- cmdstan_model("stan_files/task1.weighted.avg.nonc.normal.stan")

fit <- model$sample(
  data = stan_data, 
  chains = 4, 
  parallel_chains = 4, 
  iter_warmup = 4000,   
  iter_sampling = 4000, 
  adapt_delta = 0.99,  
  max_treedepth = 15
)

fit$save_object("models/task1.non.informative.prior.rds")

# hmm I keep getting these divergent transitions ... 
# note that the saved model still has divergent transitions ... 
# maybe I should try iter_warmup = 5000 and iter_sampling = 5000 ...? 
```

Posterior Predictive Check: 

```{r}
custom_pp_check(fit, bayesplot::ppc_dens_overlay_grouped, group = ifelse(task1_df$param.lambda < 0, "neg", ifelse(task1_df$param.lambda > 0, "pos", "zero")))

custom_pp_check(fit, bayesplot::ppc_pit_ecdf, transform = \(x) x - task1_df$va.med.x, ndraws = NULL)

custom_pp_check(fit, bayesplot::ppc_pit_ecdf_grouped, group = ifelse(task1_df$param.lambda < 0, "neg", ifelse(task1_df$param.lambda > 0, "pos", "zero")))
```

```{r}
task1_df %>% mutate(y_rep = as_draws_rvars(fit)$y_rep - va.med.x) %>% 
  unnest_rvars() %>% 
  ggplot(aes(x = va.med.x - va.mod.x, y = y_rep)) + 
  ggdist::stat_lineribbon(size = 0.5) + 
  geom_point(aes(y = va.select.x - va.med.x), alpha = 0.5, size = 0.5) + 
  # facet_wrap(~ participantId) + 
  scale_fill_brewer(palette = "Blues") + 
  labs(x = "x^med - x^mod", y = "Signed error") + 
  theme_minimal(base_size = 8) -> p

p
```

```{r}
as_draws_rvars(fit)$mu_mod_mean
as_draws_rvars(fit)$mu_med_mean
as_draws_rvars(fit)$log_sigma_mod_mean %>% exp(.)
as_draws_rvars(fit)$log_sigma_med_mean %>% exp(.)
```

Most of the difference seems to be in the `sigma` related stuff ...? 
