---
title: "Data Cleaning and Descriptive Checks"
author: "Anonymous"
format:
  html:
    callout-appearance: simple
    reference-location: margin
    embed-resources: true
---

```{r}
#| echo: false
#| output: false
#| label: setup

library(tidyverse)
library(ggplot2)
library(sgt)
library(numDeriv)
library(ggdist)
# setting the theme
theme_set(theme_minimal())
# model
library(brms)
# plot results
library(tidybayes)

# source the helper functions
source(here::here("R", "helpers.R"))
```

:::{.callout-important}
# TLDR 

We perform data cleaning in this document. We exclude participants who: 

* ... have eye-to-screen distance < 20 cm
* has Pearson correlation < 0.5 to the actual answer 

:::

# Set up 

Read in data: 

```{r}
#| label: read-in-data
#| code-fold: true

# read in scale
exp1_x <- make_va_scale(data_min = -5, data_max = 5, px_min = 50, px_max = 585)
exp1_y <- make_va_scale(data_min = 0, data_max = 1, px_min = 410, px_max = 15)

# read in data from the three runs
df1 <- read.csv(here::here("data", "part1", "pilot-01.csv")) %>% as_tibble(.)
df2 <- read.csv(here::here("data", "part1", "pilot-02.csv")) %>% as_tibble(.)
df3 <- read.csv(here::here("data", "part1", "pilot-03.csv")) %>% as_tibble(.)

# combine data from the three runs
combined_df <- bind_rows(df1, df2, df3)

df <- combined_df %>%
  filter(status == "completed") %>%
  mutate(rsp_time = as.numeric(cleanedDuration) / 1000) %>% # seconds
  select(participantId, trialId, responseId, answer, rsp_time, status)

task_df <- df %>%
  filter(str_detect(trialId, "task")) %>%
  mutate(answer = as.numeric(answer)) %>%
  pivot_wider(
    names_from = responseId,
    values_from = answer,
    names_repair = "universal"
  ) %>%
  separate_wider_delim(
    trialId,
    delim = "_",
    names = c("task", "type", "id")
  ) %>%
  filter(type != "train") %>%
  select(-type) %>%
  rename(
    data.select.x = location.x,
    data.select.y = location.y,
    pixel.select.x = pixel.x,
    pixel.select.y = pixel.y
  )

# initialize empty list of participants to be excluded from analysis
exclude_participants <- c()
```

Some participants did not answer the calibration tasks truthfully. Therefore, we need to reverse engineer the pixel location of `squarePos` to verify if participants actually put in the correct effort into the calibration tasks. 

```{r}
#| label: define-special-dfs
#| code-fold: true

p <- df %>%
  filter(status == "completed") %>%
  select(participantId, trialId, responseId, answer) %>%
  filter(grepl("calibration", trialId))

ball_positions <- df %>%
  filter(status == "completed") %>%
  filter(grepl("calibration", trialId)) %>%
  select(participantId, responseId, answer) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  # filter(!is.na(`square-position`)) %>%
  mutate(`ball-positions` = gsub("\\[|\\]", "", `ball-positions`)) %>%
  mutate(split_answer = str_split(`ball-positions`, ",")) %>%
  mutate(numeric_vectors = map(split_answer, ~ as.numeric(.x))) %>%
  unnest(split_answer) %>%
  mutate(split_answer = as.numeric(split_answer)) %>%
  group_by(participantId) %>%
  summarise(
    avg.ball.pos = sum(split_answer) / 5,
    min.ball.pos = min(split_answer),
    max.ball.pos = max(split_answer),
    true_ball_pos = (avg.ball.pos * 5 - min.ball.pos - max.ball.pos) / 3
  )

calibrations <- p %>%
  filter(grepl("pixelsPerMM|dist-calibration-MM", responseId)) %>%
  select(-trialId) %>%
  mutate(answer = as.numeric(answer)) %>%
  pivot_wider(names_from = responseId, values_from = answer) %>%
  rename(pixelToMM = pixelsPerMM, dist_to_screen = `dist-calibration-MM`)

participants <- calibrations %>%
  right_join(ball_positions, by = join_by(participantId)) %>%
  mutate(
    phy = dist_to_screen * tan(13.5 * pi / 180),
    px = phy * pixelToMM,
    square_pos = round(px + avg.ball.pos)
  ) %>%
  mutate(
    ball.square.dist = (square_pos - true_ball_pos) / pixelToMM,
    d = ball.square.dist / tan(13.5 * pi / 180)
  ) %>%
  select(participantId, pixelToMM, d) %>%
  rename(dist_to_screen = d)

# plot distance to screen distribution
participants %>%
  ggplot(aes(x = dist_to_screen)) +
  geom_dotplot() +
  labs(x = "distance to screen (mm)")
```

From the above histogram, two participants have `dist_to_screen` less than 200 mm. This indicates that they did not press the space bar in the calibration tasks as instructed, thus we do not include these participants in our analysis. 

```{r}
participants %>% filter(dist_to_screen > 200) -> participants
```

# Going through each task ... 

In the following we will go through each task, and determine the participants we exclude from our analysis. Note that we index each tasks with names such as "task 1" and "task 2", etc., because that's how we named each task in the configuration file, for easier indexing and referencing. 

## Task 1 -- `BisectArea` -- split area into equal halves

Load data: 

```{r}
#| label: get task 1 data
#| code-fold: true

task1_df <- task_df %>%
  filter(task == "task1") %>%
  select(-slider.x, -slider.y) %>%
  right_join(participants, by = join_by(participantId)) %>%
  mutate(
    data.select.left_area = psgt(
      data.select.x,
      param.mu,
      param.sigma,
      param.lambda,
      param.p,
      param.q,
      mean.cent = FALSE
    ),
    data.ans.x = qsgt(
      0.5,
      param.mu,
      param.sigma,
      param.lambda,
      param.p,
      param.q,
      mean.cent = FALSE
    )
  ) %>%
  mutate(
    va.resp.x = exp1_x$data_to_va(data.select.x, dist_to_screen, pixelToMM),
    va.med.x = exp1_x$data_to_va(data.ans.x, dist_to_screen, pixelToMM),
    va.mod.x = exp1_x$data_to_va(param.mu, dist_to_screen, pixelToMM)
  )
```

Calculate correlation: 

```{r}
#| code-fold: true

task1_corr <- task1_df %>%
  group_by(participantId) %>%
  summarise(correlation.x = cor(data.select.x, data.ans.x)) %>%
  mutate(
    mean = mean(correlation.x),
    sd = sd(correlation.x),
    lb = mean - 2 * sd,
    ub = mean + 2 * sd
  ) %>%
  filter(correlation.x < 0.5)

task1_df %>%
  group_by(participantId) %>%
  summarise(correlation.x = cor(data.select.x, data.ans.x)) %>%
  mutate(
    mean = mean(correlation.x),
    sd = sd(correlation.x),
    lb = mean - 2 * sd,
    ub = mean + 2 * sd
  ) %>%
  ggplot(aes(x = correlation.x)) +
  geom_dotplot()
```

Given that most participants have correlation > 0.5, we filter out participant whose selection's correlation to the actual answer has correlation < 0.5. 

```{r}
exclude_participants <- c(
  task1_corr %>% pull(participantId),
  exclude_participants
)
```

:::{.callout-note collapse="true" appearance="simple"}
### More in-depth examination

Let's also examine the two participants who have correlation < 0.5 for task 1, by plotting the actual answer (in red) and the participants' selection (in black): 

```{r}
#| layout-ncol: 2
#| code-fold: true

task1_df %>%
  filter(participantId == "67f44a0314ad081a761de62c") %>%
  ggplot(aes(x = data.select.x, y = data.select.y)) +
  geom_point() +
  xlim(-5, 5) +
  ylim(0, 1) +
  geom_point(
    data = filter(task1_df, participantId == "67f44a0314ad081a761de62c"),
    aes(x = data.ans.x, color = "answer")
  ) +
  labs(title = "PID --- 67f44a0314ad081a761de62c")

task1_df %>%
  filter(participantId == "67ea39f872546624b6b520bf") %>%
  ggplot(aes(x = data.select.x, y = data.select.y)) +
  geom_point() +
  xlim(-5, 5) +
  ylim(0, 1) +
  geom_point(
    data = filter(task1_df, participantId == "67ea39f872546624b6b520bf"),
    aes(x = data.ans.x, color = "answer")
  ) +
  labs(title = "PID --- 67ea39f872546624b6b520bf")
```

This participant is selecting -5 or 5 all the time. Clearly demonstrated low effort. We exclude their answers from the final analysis. 
:::

```{r}
#| layout-ncol: 2
#| code-fold: true
#| eval: false
#| echo: false

# Plotting selection, answer, and `x` coordinate for highest point against lambda:

task1_df %>%
  filter(!participantId %in% exclude_participants) %>%
  ggplot(aes(y = data.select.x - data.ans.x, x = param.lambda)) +
  geom_point(alpha = 0.5) # + ylim(-1, 1)

task1_df %>%
  filter(!participantId %in% exclude_participants) %>%
  ggplot(aes(y = data.select.left_area - 0.5, x = param.lambda)) +
  geom_point(alpha = 0.5)

task1_df %>%
  filter(!participantId %in% exclude_participants) %>%
  ggplot(aes(y = va.resp.x - va.med.x, x = param.lambda)) +
  geom_point(alpha = 0.5)


task1_df %>%
  ggplot(aes(x = param.lambda, y = data.select.x - param.mu)) +
  geom_point(alpha = 0.5)

task1_df %>% ggplot(aes(x = data.select.x - param.mu)) + geom_dots()

task1_df %>%
  ggplot(aes(y = data.select.x - param.mu, x = dist_to_screen)) +
  geom_point()

task1_df %>%
  ggplot(aes(y = data.select.left_area - 0.5, x = dist_to_screen)) +
  geom_point()

task1_df %>% ggplot(aes(x = data.ans.x - data.select.x)) + geom_dots()

task1_df %>%
  ggplot(aes(y = data.ans.x - data.select.x, x = dist_to_screen)) +
  geom_point()

task1_df %>%
  ggplot(aes(y = data.select.left_area^(1 / 0.7), x = param.lambda)) +
  geom_point()
```

## Task 2 -- `HighestPoint` -- Find highest point on curve 

First we get the data: 

```{r}
#| label: get task 2 data
#| code-fold: true

task2_df <- task_df %>%
  filter(task == "task2") %>%
  select(-slider.x, -slider.y) %>%
  mutate(
    data.ans.x = param.mu,
    data.ans.y = dsgt(
      param.mu,
      param.mu,
      param.sigma,
      param.lambda,
      param.p,
      param.q,
      mean.cent = FALSE
    ),
    # this is the gradient of the left of the highest point
    data.left.grad = numDeriv::grad(
      dsgt,
      param.mu - 1,
      mu = param.mu,
      sigma = param.sigma,
      lambda = param.lambda,
      p = param.p,
      q = param.q,
      mean.cent = FALSE
    ),
    data.right.grad = numDeriv::grad(
      dsgt,
      param.mu + 1,
      mu = param.mu,
      sigma = param.sigma,
      lambda = param.lambda,
      p = param.p,
      q = param.q,
      mean.cent = FALSE
    ),
  ) %>%
  right_join(participants, by = join_by(participantId)) %>%
  # calculate things related to participants' selection
  mutate(
    va.resp.x = exp1_x$px_to_va(pixel.select.x, dist_to_screen, pixelToMM),
    va.resp.y = exp1_y$px_to_va(pixel.select.y, dist_to_screen, pixelToMM),
    va.ans.x = exp1_x$data_to_va(data.ans.x, dist_to_screen, pixelToMM),
    va.ans.y = exp1_y$data_to_va(data.ans.y, dist_to_screen, pixelToMM),
  ) %>%
  mutate(flat = (abs(data.left.grad) + abs(data.right.grad)) / 2)
```

Calculate correlation^[similar to what we did for Task 1, exclude those who have correlation to the true answer < 0.5]: 

```{r}
#| code-fold: true
task2_corr <- task2_df %>%
  group_by(participantId) %>%
  summarise(
    correlation.y = cor(data.select.y, data.ans.y),
    correlation.x = cor(data.select.x, data.ans.x)
  ) %>%
  filter(correlation.y < 0.5 | correlation.x < 0.5)

task2_corr
```

```{r}
exclude_participants <- c(
  exclude_participants,
  task2_corr %>% pull(participantId)
)
```

:::{.callout-note title="Further examination" collapse="true"}
Examine particular participants: 

```{r}
#| code-fold: true
#| layout-ncol: 2

task2_df %>%
  filter(participantId == "67ea39f872546624b6b520bf") %>%
  ggplot(aes(x = data.select.x, y = data.select.y)) +
  geom_point() +
  xlim(-5, 5) +
  ylim(0, 1) +
  geom_point(
    data = filter(task2_df, participantId == "67ea39f872546624b6b520bf"),
    aes(x = data.ans.x, color = "answer")
  )

task2_df %>%
  filter(participantId == "67eedf7be13ffb77c677d4dd") %>%
  ggplot(aes(x = data.select.x, y = data.select.y)) +
  geom_point() +
  xlim(-5, 5) +
  ylim(0, 1) +
  geom_point(
    data = filter(task2_df, participantId == "67eedf7be13ffb77c677d4dd"),
    aes(x = data.ans.x, y = data.ans.y, color = "answer")
  )
```
:::


```{r}
#| layout-ncol: 2
#| code-fold: true
#| eval: false
#| echo: false

# task2_df %>% ggplot(aes(y = phy.ans.y - phy.select.y, x = dist_to_screen)) +
#   geom_point()
#
# task2_df %>% ggplot(aes(y = va.ans.y - va.select.y, x = dist_to_screen)) +
#   geom_point()

task2_df %>%
  filter(!participantId %in% exclude_participants) %>%
  ggplot(aes(x = va.ans.y - va.resp.y)) +
  geom_dots()


task2_df %>% ggplot(aes(x = data.ans.y - data.select.y)) + geom_dots()
```

```{r}
#| layout-ncol: 2
#| code-fold: true
#| eval: false
#| echo: false

# Checking relations, focusing on the **x** axes: 

task2_df %>%
  filter(!participantId %in% exclude_participants) %>%
  ggplot(aes(x = va.resp.x - va.ans.x)) +
  geom_dots()

task2_df %>% ggplot(aes(x = data.ans.x - data.select.x)) + geom_dots()

task2_df %>%
  ggplot(aes(x = va.resp.x - va.ans.x)) +
  geom_dots() +
  facet_wrap(~participantId)

task2_df %>%
  ggplot(aes(y = va.resp.x - va.ans.x, x = dist_to_screen)) +
  geom_point(alpha = 0.5)

task2_df %>%
  ggplot(aes(y = va.resp.x - va.ans.x, x = param.lambda)) +
  geom_point(alpha = 0.5) +
  labs(title = "Signed error on x-axes against skewness")
```

## Task 3 -- `ProjectToCurve` -- find point on curve where `y == 0.5`

First we get the data:

```{r}
#| code-fold: true

task3_df <- task_df %>%
  filter(task == "task3") %>%
  select(-slider.x, -slider.y) %>%
  right_join(participants, by = join_by(participantId)) %>%
  mutate(
    data.ans.y = 0.5,
    data.ans.x = qsgt(
      0.5,
      param.mu,
      param.sigma,
      param.lambda,
      param.p,
      param.q,
      mean.cent = FALSE
    )
  ) %>%
    mutate(
    va.resp.x = exp1_x$data_to_va(data.select.x, dist_to_screen, pixelToMM),
    va.resp.y = exp1_y$data_to_va(data.select.y, dist_to_screen, pixelToMM),
    va.ans.x = exp1_x$data_to_va(data.ans.x, dist_to_screen, pixelToMM),
    va.ans.y = exp1_y$data_to_va(data.ans.y, dist_to_screen, pixelToMM)
  )
```

Calculation correlation and exclude participants: 

```{r}
#| code-fold: true
# can't calculat correlation on y because the standard deviation is 0.
# we use x as an approximation

task3_corr <- task3_df %>%
  group_by(participantId) %>%
  summarise(correlation = suppressWarnings(cor(data.select.x, data.ans.x))) %>%
  filter(is.na(correlation) | correlation < 0.9)

task3_corr
```

```{r}
exclude_participants <- c(exclude_participants, pull(task3_corr, participantId))
```

:::{.callout-note title="Further examine" collapse="true"}

Examine particular participants: 

```{r}
#| layout-ncol: 2

task3_plot <- function(pid) {
  task3_df %>%
    filter(participantId == pid) %>%
    ggplot() +
    geom_point(aes(x = data.select.x, y = data.select.y, color = "select")) +
    geom_point(aes(x = data.ans.x, y = 0.5, color = "answer")) +
    ylim(0, 1) +
    xlim(-5, 5) +
    geom_vline(xintercept = 0.5, linetype = "dashed", color = "gray") +
    labs(title = pid)
}

task3_plot("67ea39f872546624b6b520bf")
task3_plot("67ee9fb880b10fe380af1803")
task3_plot("67eedf7be13ffb77c677d4dd")
task3_plot("67fd4f955123c6d79ae66a3a")
```

:::

Verify that the scales are correct:

```{r}
#| eval: false
#| echo: false
#| code-fold: true
#| label: task 3 verify

# verify that the results are of correct scale
task3_df %>% ggplot(aes(x = data.select.x - data.ans.x)) + geom_dots()

task3_df %>% ggplot(aes(x = data.select.y - data.ans.y)) + geom_dots()

task3_df %>% ggplot(aes(x = va.resp.x - va.ans.x)) + geom_dots()

task3_df %>% ggplot(aes(x = va.resp.y - va.ans.y)) + geom_dots()
```

It is not very normal that we have a bunch of selections that have this big of a variation on y ... 

```{r}
task3_df %>% filter(abs(data.select.y - data.ans.y) > 0.2)
```

Hmm it is very clear that this participant only selected values very close to 1 ... which is NOT what we wanted ... 

Distribution of signed error:

```{r}
#| code-fold: true

task3_df %>%
  ggplot(aes(x = va.resp.y - va.ans.y)) +
  geom_dots() +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  labs(title = "Task 3 -- Find dot on curve where y == 0.5")
```

Relationship between distance from the y-axes and the signed distribution:

```{r}
#| code-fold: true
#| layout-ncol: 2

task3_df %>%
  ggplot(aes(x = va.ans.x, y = va.resp.y - va.ans.y)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black")

task3_df %>% ggplot(aes(x = va.resp.y - va.ans.y)) + geom_dots()
```

## Task 4 -- `MaxSlope` -- find maximum slope 

Define helper functions: 

```{r}
#| code-fold: true
# x: one unit in data maps to (585 - 50)/(5 - (-5)) in pixel
# y: one unit in data maps to (410 - 15)/(1 - 0) in pixel
data_to_screen_slope <- function(k) {
  return(k * 395 / 53.5)
}
data_to_va_slope <- function(data_x, data_y, dist, px_per_mm) {
  eps <- 1e-6
  s_y <- abs(
    exp1_y$data_to_va(data_y + eps, dist, px_per_mm) -
      exp1_y$data_to_va(data_y, dist, px_per_mm)
  ) /
    eps
  s_x <- (abs(
    exp1_x$data_to_va(data_x + eps, dist, px_per_mm) -
      exp1_x$data_to_va(data_x, dist, px_per_mm)
  ) /
    eps)
  return(s_y / s_x)
}
```

Get data: 

```{r}
#| label: get task 4 data 

task4_df <- task_df %>%
  filter(task == "task4") %>%
  select(-slider.x, -slider.y) %>%
  right_join(participants, by = join_by(participantId)) %>%
  mutate(
    data.resp.slope = dsgt(
      data.select.x,
      param.mu,
      param.sigma,
      param.lambda,
      param.p,
      param.q,
      mean.cent = FALSE
    ),
    data.ans.slope = dsgt(
      param.mu,
      param.mu,
      param.sigma,
      param.lambda,
      param.p,
      param.q,
      mean.cent = FALSE
    ),
    data.ans.x = param.mu,
    data.ans.y = 0.5 *
      (psgt(
        data.ans.x + tolerance,
        param.mu,
        param.sigma,
        param.lambda,
        param.p,
        param.q,
        mean.cent = FALSE
      ) +
        psgt(
          data.ans.x - tolerance,
          param.mu,
          param.sigma,
          param.lambda,
          param.p,
          param.q,
          mean.cent = FALSE
        )),
    va.resp.slope = data.resp.slope *
      data_to_va_slope(
        data.select.x,
        data.select.y,
        dist_to_screen,
        pixelToMM
      ),
    va.ans.slope = data.ans.slope *
      data_to_va_slope(
        data.ans.x,
        data.ans.y,
        dist_to_screen,
        pixelToMM
      )
  )
```

Calculate correlation: 

```{r}
#| code-fold: true
task4_corr <- task4_df %>%
  group_by(participantId) %>%
  summarise(
    correlation = cor(
      data_to_screen_slope(data.resp.slope),
      data_to_screen_slope(data.ans.slope)
    )
  ) %>%
  filter(correlation < 0.5)

task4_corr
```

```{r}
exclude_participants <- c(exclude_participants, pull(task4_corr, participantId))
```

```{r}
#| code-fold: true
#| eval: false
#| echo: false 

task4_df %>% ggplot(aes(x = va.ans.x - va.resp.x)) + geom_dots()
task4_df %>% ggplot(aes(x = va.ans.y - va.resp.y)) + geom_dots()
task4_df %>% ggplot(aes(x = atan(va.ans.y / va.ans.x) * 180 / pi)) + geom_dots()
task4_df %>%
  ggplot(aes(x = va.ans.y / va.ans.x, y = data.ans.slope)) +
  geom_point() +
  coord_equal()

task4_df %>% ggplot(aes(x = data.select.slope - data.ans.slope)) + geom_dots()

task4_df %>% filter(data.select.slope - data.ans.slope == 0)

task4_df %>%
  ggplot(aes(
    x = atan(data.select.slope) * 180 / pi - atan(data.ans.slope) * 180 / pi
  )) +
  geom_dots()

task4_df %>%
  ggplot(aes(
    x = data_to_screen_slope(data.select.slope) -
      data_to_screen_slope(data.ans.slope)
  )) +
  geom_dots()

task4_df %>%
  ggplot(aes(
    x = atan(data_to_screen_slope(data.select.slope)) *
      180 /
      pi -
      atan(data_to_screen_slope(data.ans.slope)) * 180 / pi
  )) +
  stat_ecdf()
```

If we map the mathematical slope to the actual slope, this is what we get: 

- y axis range [0, 1] correspond to 395 px -> 395px / unit 
- x axis range [-5, 5] correspond to 535 px -> 53.5 px / unit 

```{r}
task4_df %>%
  ggplot(aes(
    x = data_to_screen_slope(data.resp.slope),
    y = data_to_screen_slope(data.ans.slope)
  )) +
  geom_point(alpha = 0.5) +
  coord_equal()

task4_df %>%
  ggplot(aes(
    x = data_to_screen_slope(data.resp.slope) -
      data_to_screen_slope(data.ans.slope)
  )) +
  geom_dots()
```

```{r}
#| layout-ncol: 2
#| code-fold: true
#| output: false
#| eval: false 
#| echo: false

# Let's see if there's any obvious relation between signed error and parameters of the distribution: 

task4_df %>%
  ggplot(aes(y = data.select.slope - data.ans.slope, x = dist_to_screen)) +
  geom_point()

task4_df %>%
  ggplot(aes(y = data.select.slope - data.ans.slope, x = param.mu)) +
  geom_point()

task4_df %>%
  ggplot(aes(y = data.select.slope - data.ans.slope, x = param.sigma)) +
  geom_point()

task4_df %>%
  ggplot(aes(y = data.select.slope - data.ans.slope, x = param.lambda)) +
  geom_point()

task4_df %>%
  ggplot(aes(y = data.select.slope - data.ans.slope, x = param.p)) +
  geom_point()

task4_df %>%
  ggplot(aes(y = data.select.slope - data.ans.slope, x = param.q)) +
  geom_point()

task4_df %>% ggplot(aes(x = data.select.slope - data.ans.slope)) + geom_dots()
```

## Task 5 -- `ProjectToAxis` -- Project from dot to axes 

Get data: 

```{r}
#| label: task 5 load data
#| code-fold: true

task5_df <- task_df %>%
  filter(task == "task5") %>%
  select(
    participantId,
    task,
    id,
    data.select.x,
    data.select.y,
    slider.x,
    slider.y
  ) %>%
  rename(
    data.ans.x = data.select.x,
    data.ans.y = data.select.y,
    data.resp.x = slider.x,
    data.resp.y = slider.y
  ) %>%
  right_join(participants, by = join_by(participantId)) %>%
  # convert from data space to visual angle space
  mutate(
    va.resp.x = exp1_x$data_to_va(data.resp.x, dist_to_screen, pixelToMM),
    va.resp.y = exp1_y$data_to_va(data.resp.y, dist_to_screen, pixelToMM),
    va.ans.x = exp1_x$data_to_va(data.ans.x, dist_to_screen, pixelToMM),
    va.ans.y = exp1_y$data_to_va(data.ans.y, dist_to_screen, pixelToMM),
  )
```

Check correlation:  

```{r}
#| code-fold: true
task5_corr <- task5_df %>%
  group_by(participantId) %>%
  summarise(
    correlation.x = cor(data.resp.x, data.ans.x),
    correlation.y = cor(data.resp.y, data.ans.y)
  ) %>%
  filter(correlation.x < 0.9 | correlation.y < 0.9)

task5_corr
```

```{r}
exclude_participants <- c(
  exclude_participants,
  task5_corr %>% pull(participantId)
)
```

:::{.callout-note title="Further examination" collapse="true"}
Examine specific participants: 

```{r}
#| layout-ncol: 2

task5_plot <- function(pid) {
  task5_df %>%
    filter(participantId == pid) %>%
    ggplot() +
    geom_point(aes(x = data.ans.x, y = data.ans.y, color = "answer")) +
    geom_point(aes(x = data.resp.x, y = data.resp.y, color = "select")) +
    xlim(-5, 5) +
    ylim(0, 1)
}

task5_plot("6443ebb84fc33e703937a6f9")
task5_plot("67ea39f872546624b6b520bf")
task5_plot("67eedf7be13ffb77c677d4dd")
```

:::

```{r}
#| eval: false
#| code-fold: true
#| label: task 5 verify
#| echo: false 

# Verify that the results are in the right scale: 

task5_df %>% ggplot(aes(x = data.ans.x - data.resp.x)) + geom_dots()

task5_df %>%
  ggplot(aes(x = va.resp.x - va.ans.x)) +
  geom_dots()

task5_df %>% ggplot(aes(x = va.resp.y - va.ans.y)) + geom_dots()

```

## `ProjectToAxis_X` Project from dot to X axes

Distribution of signed error: 

```{r}
#| code-fold: true

task5_df %>%
  ggplot(aes(x = va.resp.x - va.ans.x)) +
  geom_dots() +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black")
```

Relationship between signed error (in visual angle space) and distance to X-axes: 

```{r}
#| code-fold: true

task5_df %>%
  ggplot(aes(x = va.ans.y, y = va.resp.x - va.ans.x)) +
  geom_point(alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray")
```

```{r}
#| code-fold: true

task5_df %>%
  ggplot(aes(x = va.ans.x, y = va.resp.y - va.ans.y)) +
  geom_point(alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray") 

task5_df %>% filter(va.resp.y - va.ans.y > 1)
```

```{r}
task5_df %>%
  filter(va.resp.y - va.ans.y > 1) %>%
  ggplot(aes(x = data.ans.x, y = data.ans.y, color = "answer")) +
  geom_point() +
  geom_point(aes(x = data.resp.x, y = data.resp.y, color = "select")) +
  xlim(-5, 5) +
  ylim(0, 1)
```

Have reason to suspect that this participant really isn't answering the questions nicely ... 

---

# How long on average did people spend? 

```{r}
df %>%
  filter(!participantId %in% exclude_participants) %>%
  filter(grepl("task", trialId)) %>%
  group_by(participantId, trialId, rsp_time) %>%
  summarise(.) %>%
  group_by(participantId) %>%
  summarise(sum_rsp = sum(rsp_time)) %>%
  pull(sum_rsp) %>%
  mean(.) /
  60
```

# Save the cleaned data

Count the number of times participants are excluded^[Given that we have five tasks, the maximum number of times that a participant can be excluded is 5]: 

```{r}
exclude_participants %>% as_tibble() %>% count(value)
```

Save the cleaned files: 

```{r}
#| eval: false 

task_df %>% filter(!participantId %in% exclude_participants) -> task_df

saveRDS(task_df, file = here::here("data", "part1", "task.rds"))

participants %>%
  filter(!participantId %in% exclude_participants) # %>%
  # saveRDS(., "../../data/final/participants.rds")
```


