---
title: "BisectArea Operator"
author: "Anonymous"
date: today
date-format: "iso"
html:
  reference-location: margin
  embed-resources: true
---

```{r}
#| echo: false
#| output: false
#| label: setup

library(tidyverse)
library(ggplot2)
library(sgt)
library(numDeriv)
library(ggdist)
# setting the theme
theme_set(theme_minimal(base_size = 8))
# model
library(brms)
# plot results
library(tidybayes)
library(cmdstanr)
library(posterior)

library(here)

# source the helper functions
source(here("R", "helpers.R"))

bayesplot::bayesplot_theme_update(text = element_text(family = "sans"))
```

# Set up

Read in data 

```{r}
#| label: read-in-data
#| code-fold: true

task_df <- read_rds(here("data", "part1", "task.rds"))
participants <- read_rds(here("data", "part1", "participants.rds"))
exp1_x <- make_va_scale(data_min = -5, data_max = 5, px_min = 50, px_max = 585)
exp1_y <- make_va_scale(data_min = 0, data_max = 1, px_min = 410, px_max = 15)
```

Load data for task 1 only: 

```{r}
#| code-fold: true

task1_df <- task_df %>%
  filter(task == "task1") %>% # task_df
  select(-slider.x, -slider.y) %>%
  right_join(participants, by = join_by(participantId)) %>%
  mutate(
    data.select.left_area = psgt(
      data.select.x,
      param.mu,
      param.sigma,
      param.lambda,
      param.p,
      param.q,
      mean.cent = FALSE
    ),
    data.ans.x = qsgt(
      0.5,
      param.mu,
      param.sigma,
      param.lambda,
      param.p,
      param.q,
      mean.cent = FALSE
    )
  ) %>%
  mutate(
    va.resp.x = exp1_x$data_to_va(data.select.x, dist_to_screen, pixelToMM),
    va.med.x = exp1_x$data_to_va(data.ans.x, dist_to_screen, pixelToMM),
    va.mod.x = exp1_x$data_to_va(param.mu, dist_to_screen, pixelToMM)
  )
```

```{r}
#| eval: false
#| echo: false
task1_df %>%
  as.data.frame(.) %>%
  write.csv(paste0(here(), "/data/processed/task1.csv"))
```

Prepare data for stan: 

```{r}
N <- task1_df %>% nrow(.)
J <- length(unique(task1_df$participantId))
x <- task1_df %>% pull(va.resp.x)
x_med <- task1_df %>% pull(va.med.x)
x_mod <- task1_df %>% pull(va.mod.x)

# change the participantId to numeric id
# 1. Get unique participant IDs
unique_ids <- unique(task1_df$participantId)

# 2. Create a mapping
id_mapping <- setNames(1:length(unique_ids), unique_ids)

# 3. Create the numeric ID vector
numeric_id <- as.numeric(factor(
  task1_df$participantId,
  levels = names(id_mapping)
))

# assemble to create data for stan to fit on:
stan_data <- list(
  N = N,
  J = J,
  id = numeric_id,
  x = x,
  x_med = x_med,
  x_mod = x_mod
)
```

Define custom posterior predictive check function: 

```{r}
#| code-summary: "custom pp_check"
#| code-fold: true

custom_pp_check <- function(
  fitted_model,
  fun = bayesplot::ppc_dens_overlay,
  ndraws = 100,
  transform = identity,
  ...
) {
  y_rep <- as_draws_rvars(fitted_model)$y_rep |>
    transform() |>
    as_draws_matrix()
  if (!is.null(ndraws)) {
    y_rep = y_rep |>
      merge_chains() |>
      resample_draws(ndraws = ndraws)
  }
  y <- fitted_model$draws("x_org", format = "matrix")[1, ] |>
    as.vector() |>
    transform()
  fun(y, y_rep, ...)
}
```

:::{.callout-important}
# TLDR 

* centered parameterization is the most straight-forward mapping form math model to fitted model, but it creates problems for sampling (e.g., going into a funnel)
* non-centered parameteriation makes things easier to fit; the informative prior was obtained from `HighestPoint_X` operator.  
:::

# Fit `stan` models 

## Centered Parameterization (informative priors)

**Spoiler: The following model has divergent transitions after fitting.**

We run this anyway to make out analysis complete. Note this is NOT the model we actually fit, but we list here nonetheless as the math model and the stan specifications are easier to map to each other.  

```{r}
model <- cmdstan_model(here::here(
  "analysis/stan_files/task1.weighted.avg.stan"
))
model
``` 

here is the corresponding math model: 

$$
\begin{align*}
% Likelihood
x_n &\sim \mathcal{N}\!\left(\theta_n \cdot (x^{\text{med}}_n + \mu^{\text{med}}_{j[n]}) + (1 - \theta_n) \cdot (x^{\text{mod}}_n + \mu^{\text{mod}}_{j[n]}),\ \ \tilde\sigma_n\right) \\[4pt]
\tilde\sigma_n &= \sqrt{\theta_n^2 \cdot (\sigma^{\text{med}}_{j[n]})^2 + (1 - \theta_n)^2 \cdot (\sigma^{\text{mod}}_{j[n]})^2} \\[4pt]
% Weights (deterministic, inverse-MSE pooling)
\theta_n &= \frac{\text{invMSE}^{\text{med}}_n}{\text{invMSE}^{\text{med}}_n + \text{invMSE}^{\text{mod}}_n} \\[4pt]
\text{invMSE}^{\text{med}}_n &= \left[(\mu^{\text{med}}_{j[n]})^2 + (\sigma^{\text{med}}_{j[n]})^2\right]^{-1} \\[2pt]
\text{invMSE}^{\text{mod}}_n &= \left[(x^{\text{mod}}_n - x^{\text{med}}_n + \mu^{\text{mod}}_{j[n]})^2 + (\sigma^{\text{mod}}_{j[n]})^2\right]^{-1} \\[6pt]
% Participant-level parameters
\mu^{\text{mod}}_j &\sim \mathcal{N}(\bar\mu^{\text{mod}},\ \bar\sigma^{\text{mod}}_\mu) \\
\sigma^{\text{mod}}_j &\sim \text{Lognormal}(\bar\mu^{\text{mod}}_{\log\sigma},\ \bar\sigma^{\text{mod}}_{\log\sigma}) \\
\mu^{\text{med}}_j &\sim \mathcal{N}(\bar\mu^{\text{med}},\ \bar\sigma^{\text{med}}_\mu) \\
\sigma^{\text{med}}_j &\sim \text{Lognormal}(\bar\mu^{\text{med}}_{\log\sigma},\ \bar\sigma^{\text{med}}_{\log\sigma}) \\[6pt]
% Hyperpriors — mod (informative, from brms fit)
\bar\mu^{\text{mod}} &\sim \mathcal{N}(0.01,\ 0.02) \\
\bar\sigma^{\text{mod}}_\mu &\sim \mathcal{N}^{+}(0.04,\ 0.03) \\
\bar\mu^{\text{mod}}_{\log\sigma} &\sim \mathcal{N}(-1.37,\ 0.10) \\
\bar\sigma^{\text{mod}}_{\log\sigma} &\sim \mathcal{N}^{+}(0.34,\ 0.10) \\[6pt]
% Hyperpriors — med (weakly informative)
\bar\mu^{\text{med}},\ \bar\mu^{\text{med}}_{\log\sigma} &\sim \mathcal{N}(0, 1) \\
\bar\sigma^{\text{med}}_\mu,\ \bar\sigma^{\text{med}}_{\log\sigma} &\sim \mathcal{N}^{+}(0, 1) \\[4pt]
% Indexing
n &= 1, \ldots, N, \quad j[n] \in \{1, \ldots, J\}
\end{align*}
$$

Fit the stan model: 

```{r}
#| output: false
#| eval: false

fit <- model$sample(
  data = stan_data,
  chains = 4,
  iter_warmup = 4000, # Number of warmup iterations (e.g., 1000)
  iter_sampling = 4000,
  adapt_delta = 0.99, # Number of sampling iterations (e.g., 2000)
  parallel_chains = 4
)
```

```{r}
#| echo: false
#| eval: false
fit$save_object(here::here(
  "analysis/fitted_models/part1/task1.centered.param.rds"
))
```

```{r}
#| echo: false
fit <- read_rds(here::here(
  "analysis/fitted_models/part1/task1.centered.param.rds"
))
```

```{r}
summary <- fit$summary()
summary
```

Observe that we have `rhat` values not equal to 1, suggesting that the model has not converged. 

Density overlay plot: 

```{r}
#| code-fold: true
custom_pp_check(
  fit,
  bayesplot::ppc_dens_overlay_grouped,
  group = ifelse(
    task1_df$param.lambda < 0,
    "neg",
    ifelse(task1_df$param.lambda > 0, "pos", "zero")
  )
)
```

Overall PIT-ECDF plot: 

```{r}
#| code-fold: true
custom_pp_check(
  fit,
  bayesplot::ppc_pit_ecdf,
  transform = \(x) x - task1_df$va.med.x,
  ndraws = NULL
)
```

PIT-ECDF for each "group", where group is defined by whether $\lambda$ is positive, negative, or zero: 

```{r}
#| code-fold: true
custom_pp_check(
  fit,
  bayesplot::ppc_pit_ecdf_grouped,
  group = with(
    task1_df,
    case_when(
      param.lambda < 0 ~ "neg",
      param.lambda == 0 ~ "zero",
      param.lambda > 0 ~ "pos"
    )
  )
)
```

## Non-centered parameterization --- Informative prior

Note this is also the model we use in the actual paper. 

```{r}
model <- cmdstan_model(here::here(
  "analysis/stan_files/task1.weighted.avg.nonc.stan"
))
model
```

Note: this is the part of the code that corresponds to the informative prior: 

```
mu_mod_mean ~ normal(0.01, 0.02); 
mu_mod_sd ~ normal(0.04, 0.03); 
log_sigma_mod_mean ~ normal(-1.37, 0.10); 
log_sigma_mod_sd ~ normal(0.34, 0.10);
```

```{r}
#| output: false
#| eval: true

fit <- model$sample(
  data = stan_data,
  chains = 4,
  parallel_chains = 4,
  iter_warmup = 4000,
  iter_sampling = 4000,
  adapt_delta = 0.99,
  max_treedepth = 15
)
```

```{r}
#| eval: false
#| echo: false
fit$save_object(here::here(
  "analysis/fitted_models/part1/task1.informative.prior.rds"
))
```

```{r}
#| echo: false
fit <- read_rds(here::here(
  "analysis/fitted_models/part1/task1.informative.prior.rds"
))
```

Get the summary from the fit and check the value of `rhat`: 

```{r}
summary <- fit$summary()
summary %>% filter(!grepl("\\[", variable)) %>% select(contains("rhat"))
```

All values are `1.00`, which is good. Next we check `BULK_ESS` and `TAIL_ESS`: 

```{r}
summary %>% filter(!grepl("\\[", variable)) %>% select(contains("ess"))
``` 

So far these preliminary checks indicate that the fit is good. Next we perform some posterior predictive checks: 

```{r}
#| code-fold: true

# Overall
custom_pp_check(
  fit,
  bayesplot::ppc_pit_ecdf,
  transform = \(x) x - task1_df$va.med.x,
  ndraws = NULL
)
```

```{r}
#| code-fold: true
#| layout-ncol: 2

custom_pp_check(
  fit,
  bayesplot::ppc_dens_overlay_grouped,
  group = ifelse(
    task1_df$param.lambda < 0,
    "neg",
    ifelse(task1_df$param.lambda > 0, "pos", "zero")
  )
) -> p1

custom_pp_check(
  fit,
  bayesplot::ppc_pit_ecdf_grouped,
  group = ifelse(
    task1_df$param.lambda < 0,
    "neg",
    ifelse(task1_df$param.lambda > 0, "pos", "zero")
  )
) -> p2

p1
p2
```

```{r}
#| eval: false
#| echo: false

ggsave(
  here("figs", "task1-pp-check.pdf"),
  plot = p1,
  width = 8,
  height = 6,
  units = "cm",
  dpi = 300
)
ggsave(
  here("figs", "task1-pit-ecdf-grouped.pdf"),
  plot = p2,
  width = 8,
  height = 6,
  units = "cm",
  dpi = 300
)
```

Create output for paper: 

```{r}
#| code-fold: true

plot_df <- task1_df %>%
  modelr::data_grid(
    diff = seq(
      -max(abs(va.med.x - va.mod.x)),
      max(abs(va.med.x - va.mod.x)),
      length.out = 100
    ), # pretend the mod id 0
    participantId = unique(participantId)
  ) %>%
  mutate(
    numericPID = as.numeric(factor(participantId, levels = names(id_mapping))),
    x_mod = 0,
    x_med = diff
  ) %>%
  left_join(
    spread_draws(
      fit,
      mu_med[numericPID],
      mu_mod[numericPID],
      sigma_med[numericPID],
      sigma_mod[numericPID]
    ),
    by = join_by(numericPID)
  )

plot2_df <- plot_df %>%
  mutate(
    inv_mse_med = 1 / ((mu_med)^2 + (sigma_med)^2),
    inv_mse_mod = 1 / ((x_mod - x_med + mu_mod)^2 + (sigma_mod)^2),
    theta = inv_mse_med / (inv_mse_med + inv_mse_mod)
  ) %>%
  mutate(
    y_rep = rnorm(
      n(),
      theta * (x_med + mu_med) + (1 - theta) * (x_mod + mu_mod),
      sqrt((theta)^2 * (sigma_med)^2 + (1 - theta)^2 * (sigma_mod)^2)
    )
  )

plot3_df <- plot2_df %>%
  group_by(diff) %>%
  median_qi(y_rep, .width = c(0.5, 0.8, 0.95))

plot3_df %>%
  ggplot(aes(x = diff, y = y_rep - diff)) +
  geom_lineribbon(
    aes(ymin = .lower - diff, ymax = .upper - diff),
    linewidth = 0.5
  ) +
  geom_point(
    data = task1_df,
    aes(x = va.med.x - va.mod.x, y = va.resp.x - va.med.x),
    alpha = 0.5,
    size = 0.5
  ) +
  scale_fill_brewer(palette = "Blues") +
  labs(
    x = "va(median) - va(mode)",
    y = "Signed Error in Visual Angle of Bisection Point"
  ) +
  theme(legend.position = "none") -> p # note that x^mod is fixed to be 0.

p
```

```{r}
#| echo: false
#| eval: false
ggsave(
  here("figs", "task1.pdf"),
  plot = p,
  height = 3.5 / 4 * 2,
  width = 3.5,
  units = "in",
  dpi = 300
)
```

We also check the posterior for the informative priors we used: 

:::{.column-margin}
```
mu_mod_mean ~ normal(0.01, 0.02); 
mu_mod_sd ~ normal(0.04, 0.03); 
log_sigma_mod_mean ~ normal(-1.37, 0.10); 
log_sigma_mod_sd ~ normal(0.34, 0.10);
```
:::

```{r}
as_draws_rvars(fit)$mu_mod_mean
as_draws_rvars(fit)$mu_med_mean
as_draws_rvars(fit)$log_sigma_mod_mean # %>% exp(.)
as_draws_rvars(fit)$log_sigma_med_mean # %>% exp(.)
```

## Non-centered paramterization --- Uninformative prior 

```{r}
model <- cmdstan_model(
  here::here("analysis/stan_files/task1.weighted.avg.nonc.normal.stan")
)
model
```

So this is the prior we're using now: 

```
mu_mod_mean ~ normal(0, 0.5); 
mu_mod_sd ~ normal(0.04, 0.3); 
log_sigma_mod_mean ~ normal(-1.37, 1); 
log_sigma_mod_sd ~ normal(0.34, 1); 
```

```{r}
#| eval: false
fit <- model$sample(
  data = stan_data,
  chains = 4,
  parallel_chains = 4,
  iter_warmup = 4000,
  iter_sampling = 4000,
  adapt_delta = 0.99,
  max_treedepth = 15
)
```

The above is giving us 0.4 % divergent transitions. So not bad, but also not zero. Also the entire fitting took about 10 minutes. 

```{r}
#| echo: false
#| eval: false
fit$save_object(here::here(
  "analysis/fitted_models/part1/task1.non.informative.prior.rds"
))
```

```{r}
#| echo: false

fit <- read_rds(here::here(
  "analysis/fitted_models/part1/task1.non.informative.prior.rds"
))
```

Get the summary from the fit and check the value of `rhat`: 

```{r}
summary <- fit$summary()
summary %>% filter(!grepl("\\[", variable)) %>% select(contains("rhat"))
```

Next we check `BULK_ESS` and `TAIL_ESS`: 

```{r}
summary %>% filter(!grepl("\\[", variable)) %>% select(contains("ess"))
``` 

Posterior Predictive Check: 

```{r}
#| code-fold: true
#| layout-ncol: 3

custom_pp_check(
  fit,
  bayesplot::ppc_dens_overlay_grouped,
  group = ifelse(
    task1_df$param.lambda < 0,
    "neg",
    ifelse(task1_df$param.lambda > 0, "pos", "zero")
  )
)

custom_pp_check(
  fit,
  bayesplot::ppc_pit_ecdf,
  transform = \(x) x - task1_df$va.med.x,
  ndraws = NULL
)

custom_pp_check(
  fit,
  bayesplot::ppc_pit_ecdf_grouped,
  group = ifelse(
    task1_df$param.lambda < 0,
    "neg",
    ifelse(task1_df$param.lambda > 0, "pos", "zero")
  )
)
```

Examine posterior: 

```{r}
as_draws_rvars(fit)$mu_mod_mean
as_draws_rvars(fit)$mu_med_mean
as_draws_rvars(fit)$log_sigma_mod_mean %>% exp(.)
as_draws_rvars(fit)$log_sigma_med_mean %>% exp(.)
```
