---
title: "Analysis --- Task 1 --- Bisect Area"
author: "Anonymous"
html:
  reference-location: margin
  embed-resources: true
---

```{r}
#| echo: false
#| output: false
#| label: setup

library(tidyverse)
library(ggplot2)
library(sgt)
library(numDeriv)
library(ggdist)
# setting the theme
theme_set(theme_minimal(base_size = 8))
# model
library(brms)
# plot results
library(tidybayes)
library(cmdstanr)
library(posterior)

library(here)

# source the helper functions
source(here("R", "helpers.R"))

bayesplot::bayesplot_theme_update(text = element_text(family = "sans"))
```

# Set up

Read in data 

```{r}
#| label: read-in-data
#| code-fold: true

task_df <- read_rds(paste0(here(), "/data/final/task.rds"))
participants <- read_rds(paste0(here(), "/data/final/participants.rds"))
exp1_x <- make_va_scale(data_min = -5, data_max = 5, px_min = 50, px_max = 585)
exp1_y <- make_va_scale(data_min = 0, data_max = 1, px_min = 410, px_max = 15)
```

Load data for task 1 only: 

```{r}
#| code-fold: true

task1_df <- task_df %>%
  filter(task == "task1") %>% # task_df
  select(-slider.x, -slider.y) %>%
  right_join(participants, by = join_by(participantId)) %>%
  mutate(
    data.select.left_area = psgt(
      data.select.x,
      param.mu,
      param.sigma,
      param.lambda,
      param.p,
      param.q,
      mean.cent = FALSE
    ),
    data.ans.x = qsgt(
      0.5,
      param.mu,
      param.sigma,
      param.lambda,
      param.p,
      param.q,
      mean.cent = FALSE
    )
  ) %>%
  mutate(
    va.resp.x = exp1_x$data_to_va(data.select.x, dist_to_screen, pixelToMM),
    va.med.x = exp1_x$data_to_va(data.ans.x, dist_to_screen, pixelToMM),
    va.mod.x = exp1_x$data_to_va(param.mu, dist_to_screen, pixelToMM)
  )
```

Save processed data for python fitting: 

```{r}
#| eval: false
#| echo: false
task1_df %>%
  as.data.frame(.) %>%
  write.csv(paste0(here(), "/data/processed/task1.csv"))
```

Prepare data for stan: 

```{r}
N <- task1_df %>% nrow(.)
J <- length(unique(task1_df$participantId))
x <- task1_df %>% pull(va.resp.x)
x_med <- task1_df %>% pull(va.med.x)
x_mod <- task1_df %>% pull(va.mod.x)

# change the participantId to numeric id
# 1. Get unique participant IDs
unique_ids <- unique(task1_df$participantId)

# 2. Create a mapping
id_mapping <- setNames(1:length(unique_ids), unique_ids)

# 3. Create the numeric ID vector
numeric_id <- as.numeric(factor(
  task1_df$participantId,
  levels = names(id_mapping)
))

# assemble to create data for stan to fit on:
stan_data <- list(
  N = N,
  J = J,
  id = numeric_id,
  x = x,
  x_med = x_med,
  x_mod = x_mod
)
```

Define custom posterior predictive check function: 

```{r}
#| code-summary: "custom pp_check"
#| code-fold: true

custom_pp_check <- function(
  fitted_model,
  fun = bayesplot::ppc_dens_overlay,
  ndraws = 100,
  transform = identity,
  ...
) {
  y_rep <- as_draws_rvars(fitted_model)$y_rep |>
    transform() |>
    as_draws_matrix()
  if (!is.null(ndraws)) {
    y_rep = y_rep |>
      merge_chains() |>
      resample_draws(ndraws = ndraws)
  }
  y <- fitted_model$draws("x_org", format = "matrix")[1, ] |>
    as.vector() |>
    transform()
  fun(y, y_rep, ...)
}
```

## Centered Parameterization 

Spoiler: The following model has divergent transitions after fitting.  

```{r}
#| eval: false

model <- cmdstan_model("analysis/stan_files/task1.weighted.avg.stan")
model
``` 

Fit the stan model: 

```{r}
#| output: false
#| eval: false

fit <- model$sample(
  data = stan_data,
  chains = 4,
  iter_warmup = 4000, # Number of warmup iterations (e.g., 1000)
  iter_sampling = 4000,
  adapt_delta = 0.99, # Number of sampling iterations (e.g., 2000)
  parallel_chains = 4
)
```

```{r}
summary <- fit$summary()
summary
```

Density overlay plot: 

```{r}
#| code-fold: true
custom_pp_check(
  fit,
  bayesplot::ppc_dens_overlay_grouped,
  group = ifelse(
    task1_df$param.lambda < 0,
    "neg",
    ifelse(task1_df$param.lambda > 0, "pos", "zero")
  )
)
```

Overall PIT-ECDF plot: 

```{r}
#| code-fold: true
custom_pp_check(
  fit,
  bayesplot::ppc_pit_ecdf,
  transform = \(x) x - task1_df$va.med.x,
  ndraws = NULL
)
```

PIT-ECDF for each "group", where group is defined by whether $\lambda$ is positive, negative, or zero: 

```{r}
#| code-fold: true
custom_pp_check(
  fit,
  bayesplot::ppc_pit_ecdf_grouped,
  group = with(
    task1_df,
    case_when(
      param.lambda < 0 ~ "neg",
      param.lambda == 0 ~ "zero",
      param.lambda > 0 ~ "pos"
    )
  )
)

# custom_pp_check(fit, bayesplot::ppc_pit_ecdf_grouped, group = ifelse(task1_df$param.lambda < 0, "neg", ifelse(task1_df$param.lambda > 0, "pos", "zero")))
```

```{r}
task1_df %>%
  mutate(y_rep = as_draws_rvars(fit)$y_rep - va.med.x) %>%
  unnest_rvars() %>%
  ggplot(aes(x = param.lambda, y = y_rep)) +
  ggdist::stat_lineribbon() +
  geom_point(aes(y = va.resp.x - va.med.x)) +
  scale_fill_brewer(palette = "Blues")
```

```{r}
#| echo: false
#| eval: false

# facetted plot for each participant
task1_df %>%
  mutate(y_rep = as_draws_rvars(fit)$y_rep - va.med.x) %>%
  unnest_rvars() %>%
  ggplot(aes(x = va.med.x - va.mod.x, y = y_rep)) +
  ggdist::stat_lineribbon() +
  geom_point(aes(y = va.resp.x - va.med.x)) +
  facet_wrap(~participantId) +
  scale_fill_brewer(palette = "Blues")
```

## Non-centered parameterization --- Informative prior

```{r}
#| eval: false
model <- cmdstan_model("analysis/stan_files/task1.weighted.avg.nonc.stan")
model
```

```{r}
#| output: false
#| eval: false

fit <- model$sample(
  data = stan_data,
  chains = 4,
  parallel_chains = 4,
  iter_warmup = 4000,
  iter_sampling = 4000,
  adapt_delta = 0.99,
  max_treedepth = 15
)
```

```{r}
#| eval: false
#| echo: false
fit$save_object("analysis/fitted_models/task1.informative.prior.rds")
```

```{r}
#| output: false
fit <- read_rds("analysis/fitted_models/task1.informative.prior.rds")
summary <- fit$summary()
```

Check Rhat: 

```{r}
summary %>% filter(!grepl("\\[", variable)) %>% select(contains("rhat"))
```

Check `BULK_ESS` and `TAIL_ESS`: 

```{r}
summary %>% filter(!grepl("\\[", variable)) %>% select(contains("ess"))
``` 

So far these preliminary checks indicate that the fit is good. Next we perform some posterior predictive checks: 

```{r}
#| code-fold: true

# Overall
custom_pp_check(
  fit,
  bayesplot::ppc_pit_ecdf,
  transform = \(x) x - task1_df$va.med.x,
  ndraws = NULL
)
```

```{r}
#| code-fold: true
custom_pp_check(
  fit,
  bayesplot::ppc_dens_overlay_grouped,
  group = ifelse(
    task1_df$param.lambda < 0,
    "neg",
    ifelse(task1_df$param.lambda > 0, "pos", "zero")
  )
) -> p1

custom_pp_check(
  fit,
  bayesplot::ppc_pit_ecdf_grouped,
  group = ifelse(
    task1_df$param.lambda < 0,
    "neg",
    ifelse(task1_df$param.lambda > 0, "pos", "zero")
  )
) -> p2

p1
p2
```

```{r}
#| eval: false
#| echo: false

ggsave(
  here("figs", "task1-pp-check.pdf"),
  plot = p1,
  width = 8,
  height = 6,
  units = "cm",
  dpi = 300
)
ggsave(
  here("figs", "task1-pit-ecdf-grouped.pdf"),
  plot = p2,
  width = 8,
  height = 6,
  units = "cm",
  dpi = 300
)
```

Create output for paper: 

```{r}
#| code-fold: true

plot_df <- task1_df %>%
  modelr::data_grid(
    diff = seq(
      -max(abs(va.med.x - va.mod.x)),
      max(abs(va.med.x - va.mod.x)),
      length.out = 100
    ), # pretend the mod id 0
    participantId = unique(participantId)
  ) %>%
  mutate(
    numericPID = as.numeric(factor(participantId, levels = names(id_mapping))),
    x_mod = 0,
    x_med = diff
  ) %>%
  left_join(
    spread_draws(
      fit,
      mu_med[numericPID],
      mu_mod[numericPID],
      sigma_med[numericPID],
      sigma_mod[numericPID]
    ),
    by = join_by(numericPID)
  )

plot2_df <- plot_df %>%
  mutate(
    inv_mse_med = 1 / ((mu_med)^2 + (sigma_med)^2),
    inv_mse_mod = 1 / ((x_mod - x_med + mu_mod)^2 + (sigma_mod)^2),
    theta = inv_mse_med / (inv_mse_med + inv_mse_mod)
  ) %>%
  mutate(
    y_rep = rnorm(
      n(),
      theta * (x_med + mu_med) + (1 - theta) * (x_mod + mu_mod),
      sqrt((theta)^2 * (sigma_med)^2 + (1 - theta)^2 * (sigma_mod)^2)
    )
  )

plot3_df <- plot2_df %>%
  group_by(diff) %>%
  median_qi(y_rep, .width = c(0.5, 0.8, 0.95))

plot3_df %>%
  ggplot(aes(x = diff, y = y_rep - diff)) +
  geom_lineribbon(
    aes(ymin = .lower - diff, ymax = .upper - diff),
    linewidth = 0.5
  ) +
  geom_point(
    data = task1_df,
    aes(x = va.med.x - va.mod.x, y = va.resp.x - va.med.x),
    alpha = 0.5,
    size = 0.5
  ) +
  scale_fill_brewer(palette = "Blues") +
  labs(
    x = "va(median) - va(mode)",
    y = "Signed Error in Visual Angle of Bisection Point"
  ) +
  theme(legend.position = "none") -> p # note that x^mod is fixed to be 0.

p
```

```{r}
#| echo: false
#| eval: false
ggsave(
  here("figs", "task1.pdf"),
  plot = p,
  height = 3.5 / 4 * 2,
  width = 3.5,
  units = "in",
  dpi = 300
)
```

```{r}
#| layout-ncol: 2

# plot_data <- task1_df %>%
#   mutate(y_rep = as_draws_rvars(fit)$y_rep - va.med.x) %>% # signed error
#   unnest_rvars()

# # param.lambda on x-axis
# plot_data %>%
#   ggplot(aes(x = param.lambda, y = y_rep)) +
#   ggdist::stat_lineribbon(linewidth = 0.5) +
#   geom_point(aes(y = va.resp.x - va.med.x), size = 0.5) +
#   scale_fill_brewer(palette = "Blues")

# # x.med - x.mod on X axis
# plot_data %>%
#   ggplot(aes(x = va.med.x - va.mod.x, y = y_rep)) +
#   ggdist::stat_lineribbon(size = 0.5) +
#   geom_point(aes(y = va.resp.x - va.med.x), alpha = 0.5, size = 0.5) +
#   # facet_wrap(~ participantId) +
#   scale_fill_brewer(palette = "Blues") +
#   labs(x = "x^med - x^mod", y = "Signed error") +
#   theme_minimal(base_size = 8) -> p

# p
```

```{r}
as_draws_rvars(fit)$mu_mod_mean
as_draws_rvars(fit)$mu_med_mean
as_draws_rvars(fit)$log_sigma_mod_mean %>% exp(.)
as_draws_rvars(fit)$log_sigma_med_mean %>% exp(.)
```

## Non-centered paramterization --- Uninformative prior 

```{r}
#| output: false

model <- cmdstan_model(
  "analysis/stan_files/task1.weighted.avg.nonc.normal.stan"
)

fit <- model$sample(
  data = stan_data,
  chains = 4,
  parallel_chains = 4,
  iter_warmup = 4000,
  iter_sampling = 4000,
  adapt_delta = 0.99,
  max_treedepth = 15
)

fit$save_object("analysis/fitted_models/task1.non.informative.prior.rds")

# hmm I keep getting these divergent transitions ...
# note that the saved model still has divergent transitions ...
# maybe I should try iter_warmup = 5000 and iter_sampling = 5000 ...?
```

Posterior Predictive Check: 

```{r}
custom_pp_check(
  fit,
  bayesplot::ppc_dens_overlay_grouped,
  group = ifelse(
    task1_df$param.lambda < 0,
    "neg",
    ifelse(task1_df$param.lambda > 0, "pos", "zero")
  )
)

custom_pp_check(
  fit,
  bayesplot::ppc_pit_ecdf,
  transform = \(x) x - task1_df$va.med.x,
  ndraws = NULL
)

custom_pp_check(
  fit,
  bayesplot::ppc_pit_ecdf_grouped,
  group = ifelse(
    task1_df$param.lambda < 0,
    "neg",
    ifelse(task1_df$param.lambda > 0, "pos", "zero")
  )
)
```

```{r}
task1_df %>%
  mutate(y_rep = as_draws_rvars(fit)$y_rep - va.med.x) %>%
  unnest_rvars() %>%
  ggplot(aes(x = va.med.x - va.mod.x, y = y_rep)) +
  ggdist::stat_lineribbon(size = 0.5) +
  geom_point(aes(y = va.select.x - va.med.x), alpha = 0.5, size = 0.5) +
  # facet_wrap(~ participantId) +
  scale_fill_brewer(palette = "Blues") +
  labs(x = "x^med - x^mod", y = "Signed error") +
  theme_minimal(base_size = 8) -> p

p
```

```{r}
as_draws_rvars(fit)$mu_mod_mean
as_draws_rvars(fit)$mu_med_mean
as_draws_rvars(fit)$log_sigma_mod_mean %>% exp(.)
as_draws_rvars(fit)$log_sigma_med_mean %>% exp(.)
```

<!-- Most of the difference seems to be in the `sigma` related stuff ...?  -->
