---
title: "Projection Operators"
author: "Anonymous"
date: today
date-format: "iso"
format:
  html:
    reference-location: margin
    embed-resources: true
execute: 
  freeze: auto
---

```{r}
#| echo: false
#| output: false
#| label: setup
#| include: false

library(tidyverse)
library(ggplot2)
library(sgt)
library(numDeriv)
library(ggdist)
# setting the theme
theme_set(theme_minimal(base_size = 10))
# model
library(brms)
# library(rstan)
library(cmdstanr)
options(brms.backend = "cmdstanr")
options(mc.cores = parallel::detectCores())
# rstan_options(auto_write = TRUE)
# plot results
library(tidybayes)
library(modelr)
library(patchwork)
# for making sure the path is right
library(here)

# source the helper functions
source(here::here("R", "helpers.R"))

# theme setting
bayesplot::bayesplot_theme_update(text = element_text(family = "sans"))
```

# Set up

Read in data and scales: 

```{r}
#| label: read-in-data
#| code-fold: true

task_df <- read_rds(here("data", "part1", "task.rds"))
participants <- read_rds(here("data", "part1", "participants.rds"))
exp1_x <- make_va_scale(data_min = -5, data_max = 5, px_min = 50, px_max = 585)
exp1_y <- make_va_scale(data_min = 0, data_max = 1, px_min = 410, px_max = 15)
```

# `ProjectToAxis`: Project from dot to axes 

We denote this task as `task5` in the configuration file. First, we get the data: 

```{r}
#| label: task 5 load data
#| code-fold: true

task5_df <- task_df %>%
  filter(task == "task5") %>%
  select(
    participantId,
    task,
    id,
    data.select.x,
    data.select.y,
    slider.x,
    slider.y
  ) %>%
  rename(
    data.ans.x = data.select.x,
    data.ans.y = data.select.y,
    data.resp.x = slider.x,
    data.resp.y = slider.y
  ) %>%
  right_join(participants, by = join_by(participantId)) %>%
  # convert from data space to visual angle space
  mutate(
    va.resp.x = exp1_x$data_to_va(data.resp.x, dist_to_screen, pixelToMM),
    va.resp.y = exp1_y$data_to_va(data.resp.y, dist_to_screen, pixelToMM),
    va.ans.x = exp1_x$data_to_va(data.ans.x, dist_to_screen, pixelToMM),
    va.ans.y = exp1_y$data_to_va(data.ans.y, dist_to_screen, pixelToMM),
  )
```

```{r}
#| eval: false
#| code-fold: true
#| echo: false
task5_df %>% write.csv(here("data", "processed", "task5.csv"))
```

## `ProjectToAxisX`: Project to X-Axis 

```{r}
#| code-fold: true

task5_df %>%
  ggplot(aes(y = va.ans.y, x = va.resp.x - va.ans.x, color = participantId)) +
  geom_point(alpha = 0.5)

task5_df %>%
  filter(va.resp.x - va.ans.x > 0.4) %>%
  count(participantId) %>%
  left_join(participants, by = join_by(participantId))
```

As can be seen, there are two participants, who have `va.resp.x - va.ans.x` > 0.4. Let's also plot what the actual selected answers in data space is:  

```{r}
#| code-fold: true

task5_df %>%
  filter(va.resp.x - va.ans.x > 0.4) %>%
  ggplot() +
  geom_point(aes(x = data.resp.x, y = data.resp.y, color = "resp")) +
  geom_point(aes(x = data.ans.x, y = data.ans.y, color = "ans")) +
  facet_wrap(~participantId) +
  xlim(-5, 5) +
  ylim(0, 1)
```

Recall that we are dealing with data that is already cleaned, so why are we still seeing data that is slightly shifted towards the right, for answers on the X-axis? Turns out these two participants were aligning things wrongly on the slider (i.e, they did not match the blue dot to the red dot but matched the thing on the slider to the red dot), hence the shift in their answers. 
**Thus, we exclude these two participants from our analysis for task 5**:  

```{r}
task5.x.exclude <- c("667433338e6cc9e53c828774", "67e2f5ebbe69a58c40a5d550")
```

---

Math model:

$$
\begin{align}
\texttt{error}_i &\sim \mathcal{N}(\mu_{\text{PID}[i]}, \sigma_{i, \text{PID}[i]}) \\ 
\mu_j &\sim \mathcal{N}(\overline{\mu}, \overline{\sigma}^2)\\ 
\log(\sigma_{i, j}) &= \alpha_j + \log(\texttt{va.ans.y}_i) \\ 
\alpha_j &\sim N(\mu_\alpha, \sigma_\alpha) \\ 
\mu_\alpha, \bar{\mu} &\sim N(0, 1) \\ 
\sigma_\alpha, \bar{\sigma} &\sim \mathcal{N}^+(0, 1)
\end{align}
$$

Specify formula and prior: 

```{r}
f <- bf(
  error_va ~ 1 + (1 | participantId),
  sigma ~ 1 + (1 | participantId) + offset(log(va.ans.y))
)

p <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 1), class = sd, group = participantId, lb = 0),
  prior(normal(0, 1), class = Intercept, dpar = "sigma"),
  prior(normal(0, 1), class = sd, group = participantId, dpar = "sigma", lb = 0)
)
```

```{r}
#| echo: false
#| eval: false

task5_df %>%
  mutate(error_va = va.resp.x - va.ans.x) %>%
  get_prior(formula = f, data = .)

task5_df %>%
  mutate(error_va = va.resp.x - va.ans.x) %>%
  make_stancode(f, data = ., prior = p, family = gaussian())
```

Fit model: 

```{r}
#| output: false
#| label: fit model for task 5, project to x-axis

m.5.x.va <- task5_df %>%
  mutate(error_va = va.resp.x - va.ans.x) %>%
  # exclude the two participants who answered this task wrong
  filter(!participantId %in% task5.x.exclude) %>%
  brm(
    formula = f,
    data = .,
    family = gaussian(),
    prior = p,
    chains = 4,
    cores = 4,
    control = list(adapt_delta = 0.99),
    iter = 6000,
    file = here("analysis", "fitted_models", "task5.x"),
    file_refit = "on_change",
    save_pars = save_pars(all = TRUE)
  )
```

```{r}
#| eval: false
#| echo: false

sd <- standata(m.5.x.va)
table(sd$Z_1_1)
table(sd$Z_2_sigma_1)
```

Check fit and posterior: 

```{r}
summary(m.5.x.va)
plot(m.5.x.va)
```

Posterior predictive checks:

```{r}
#| layout-ncol: 2

pp_check(m.5.x.va, ndraws = 100) +
  theme(legend.position = c(0.9, 0.5)) +
  theme(text = element_text(size = 10)) -> p1
p1

pp_check(m.5.x.va, ndraws = 100, type = "pit_ecdf") +
  theme(text = element_text(size = 10)) -> p2
p2

pp_check(m.5.x.va, ndraws = 100, type = "pit_ecdf", plot_diff = TRUE) +
  theme(text = element_text(size = 10)) -> p3
p3
```

```{r}
#| eval: false
#| echo: false
#| label: saving all plots for task 5 project to x-axis

ggsave(
  here("figs", "task5-x-pp-check.pdf"),
  p1,
  width = 10,
  height = 5,
  units = "cm",
  dpi = 300
)

ggsave(
  here("figs", "task5-x-pit-ecdf.pdf"),
  p2,
  width = 6,
  height = 6,
  units = "cm",
  dpi = 300
)

ggsave(
  here("figs", "task5-x-pit-ecdf-diff.pdf"),
  p3,
  width = 5,
  height = 5,
  units = "cm",
  dpi = 300
)
```

```{r}
#| echo: false
#| eval: false

# Posterior for population level mean and sd:
# m.5.x.va %>% get_variables(.)

m.5.x.va %>%
  spread_draws(b_Intercept, b_sigma_Intercept) %>%
  mutate(b = exp(b_Intercept), b_sigma = exp(b_sigma_Intercept)) %>%
  pivot_longer(6:7) %>%
  ggplot(aes(x = value, y = name)) +
  stat_halfeye()
```

Plot existing data on top of prediction: 

```{r}
#| label: Task 5 - X - Results
#| code-fold: true
# for an average participant

prediction_grid <- task5_df %>%
  mutate(error_va = va.resp.x - va.ans.x) %>%
  filter(!participantId %in% task5.x.exclude) %>%
  data_grid(
    va.ans.y = seq_range(va.ans.y, n = 100),
    participantId = unique(participantId)
  ) %>%
  add_predicted_draws(m.5.x.va, ndraws = 1000)

ggplot(prediction_grid, aes(x = va.ans.y, y = .prediction)) +
  stat_lineribbon(
    .width = c(.95, .80, .50),
    # alpha = 0.5,
    linewidth = 0.5
  ) +
  scale_fill_brewer(palette = "Blues") +
  geom_point(
    data = filter(task5_df, !participantId %in% task5.x.exclude),
    aes(y = va.resp.x - va.ans.x, x = va.ans.y),
    alpha = 0.5,
    size = 0.5
  ) +
  labs(
    title = "Project to X-axis",
    y = "Error",
    x = "Distance to X-Axis"
  ) +
  # theme_minimal(base_size = 8) +
  coord_cartesian(ylim = c(-0.25, 0.4)) +
  theme_minimal(base_size = 4) +
  theme(legend.position = "none") -> p1

p1
```

```{r}
#| eval: false
#| echo: false

ggsave(
  here("figs", "task5-project-x.svg"),
  p1,
  height = 330,
  width = 330,
  units = "px",
  dpi = 300
)
```

## `ProjectToAxisY` 

This time we look at error on $y$-axis: 

```{r}
task5_df %>%
  ggplot(aes(x = va.ans.x, y = va.resp.y - va.ans.y)) +
  geom_point(alpha = 0.5)
```

We use a similar model: 

$$
\begin{align*}
\texttt{error}_i &\sim \mathcal{N}(0, \sigma_{i, \text{PID}[i]}) \\ 
\log(\sigma_{i, j}) &= \alpha_j + \log(\texttt{va.ans.x}_i) \\ 
\alpha_j &\sim N(\mu_\alpha, \sigma_\alpha) \\ 
\mu_\alpha &\sim N(0, 1) \\ 
\sigma_\alpha &\sim N^+(0, 1)
\end{align*}
$$

Which corresponds to this `brms` formula and prior: 

```{r}
f <- bf(
  error_va ~ 1 + (1 | participantId),
  sigma ~ 1 + (1 | participantId) + offset(log(va.ans.x))
)

p <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 1), class = sd, group = participantId, lb = 0),
  prior(normal(0, 1), class = Intercept, dpar = "sigma"),
  prior(normal(0, 1), class = sd, group = participantId, dpar = sigma, lb = 0)
)
```

```{r}
#| echo: false
#| output: false
#| eval: false

task5_df %>%
  mutate(error_va = va.resp.y - va.ans.y) %>%
  get_prior(formula = f, data = .)

task5_df %>%
  mutate(error_va = va.resp.y - va.ans.y) %>%
  make_stancode(f, data = ., prior = p, family = gaussian())
```

Fit model: 

```{r}
#| output: false
#| label: fit task 5 y-axis model

m.5.y.va <- task5_df %>%
  mutate(error_va = va.resp.y - va.ans.y) %>%
  brm(
    formula = f,
    data = .,
    family = gaussian(),
    prior = p,
    chains = 4,
    cores = 4,
    control = list(adapt_delta = 0.99),
    iter = 6000,
    file = here("analysis", "fitted_models", "task5.y"),
    file_refit = "on_change",
    save_pars = save_pars(all = TRUE)
  )
```

Check fit and posterior: 

```{r}
summary(m.5.y.va)
plot(m.5.y.va)
```

```{r}
#| layout-ncol: 2
#| code-fold: true

pp_check(m.5.y.va, ndraws = 100) +
  theme(text = element_text(size = 10)) +
  theme(legend.position = c(0.9, 0.5)) -> p1

pp_check(m.5.y.va, ndraws = 100, type = "pit_ecdf") +
  theme(text = element_text(size = 10)) -> p2

p1
p2
```

```{r}
#| eval: false
#| echo: false

ggsave(
  here("figs", "task5-y-pp-check.pdf"),
  p1,
  height = 5,
  width = 10,
  units = "cm",
  dpi = 300
)
ggsave(
  here("figs", "task5-y-pit-ecdf.pdf"),
  p2,
  height = 5,
  width = 5,
  units = "cm",
  dpi = 300
)
```

Plot results: 

```{r}
#| code-fold: true

prediction_grid <- task5_df %>%
  data_grid(
    va.ans.x = seq_range(va.ans.x, n = 100),
    participantId = unique(participantId)
  ) %>%
  add_predicted_draws(m.5.y.va, ndraws = 1000)

ggplot(prediction_grid, aes(x = va.ans.x, y = .prediction)) +
  stat_lineribbon(
    .width = c(.95, .80, .50),
    # alpha = 0.5,
    linewidth = 0.5
  ) +
  scale_fill_brewer(palette = "Blues") +
  geom_point(
    data = task5_df,
    aes(y = va.resp.y - va.ans.y, x = va.ans.x),
    alpha = 0.5,
    size = 0.5
  ) +
  coord_cartesian(ylim = c(-0.25, 0.4)) +
  labs(
    title = "Project to Y-axis",
    y = "Error",
    x = "Distance to Y-Axis"
  ) +
  # theme_minimal(base_size = 8) +
  theme_minimal(base_size = 4) +
  theme(legend.position = "none") -> p2

p2
```

```{r}
#| echo: false
#| eval: false
#| label: Task 5 - Y - Results

ggsave(
  here("figs", "task5-project-y.svg"),
  p2,
  height = 330,
  width = 330,
  units = "px",
  dpi = 300
)

# potential to plot the combined one
# p1 + p2 -> combined
# ggsave("plots/task5-combined.pdf", combined, height = 8, width = 10, units = "cm")
# combined <- (p1 + guides(color = "none", fill = "none", size = "none")) / (p2 +
#   guides(color = "none", fill = "none", size = "none"))
#
# combined <- (p1 + theme(legend.position = "top")) + (p2 + theme(legend.position = "top"))
#
# # ggsave("plots/task5-combined.svg", combined, height = 8, width = 10, units = "cm", dpi = 300)
# ggsave("plots/task5-combined.pdf", combined, height = 5, width = 10, units = "cm", dpi = 300)
```

# `ProjectToCurve`: Project from where y == 0.5 to Curve

We denote this as `task 3` in the configuration json. First, we get task 3 specific data: 

```{r}
#| label: Get task 3 data
#| code-fold: true

task3_df <- task_df %>%
  filter(task == "task3") %>%
  select(-slider.x, -slider.y) %>%
  right_join(participants, by = join_by(participantId)) %>%
  mutate(
    data.ans.y = 0.5,
    data.ans.x = qsgt(
      0.5,
      param.mu,
      param.sigma,
      param.lambda,
      param.p,
      param.q,
      mean.cent = FALSE
    )
  ) %>%
  # do all the calculations for the actual answer
  mutate(
    va.resp.x = exp1_x$data_to_va(data.select.x, dist_to_screen, pixelToMM),
    va.resp.y = exp1_y$data_to_va(data.select.y, dist_to_screen, pixelToMM),
    va.ans.x = exp1_x$data_to_va(data.ans.x, dist_to_screen, pixelToMM),
    va.ans.y = exp1_y$data_to_va(data.ans.y, dist_to_screen, pixelToMM)
  )
```

```{r}
#| echo: false
#| eval: false
#| label: Export data of task 3 for python processing
task3_df %>% write.csv(here("data", "processed", "task3.csv"))
```

Descriptive plots: 

```{r}
#| layout-ncol: 2
#| code-fold: true
task3_df %>%
  ggplot(aes(x = va.ans.x, y = va.resp.y - va.ans.y)) +
  geom_point(alpha = 0.5)

task3_df %>% ggplot(aes(x = va.resp.y - va.ans.y)) + geom_dots()
```

Math: 

$$
\begin{align*}
\texttt{error}_i &\sim \mathcal{N}(0, \sigma_{i, \text{PID}[i]}) \\ 
\log(\sigma_{i, j}) &= \alpha_j + \log(\texttt{va.ans.x}_i) \\ 
\alpha_j &\sim N(\mu_\alpha, \sigma_\alpha) \\ 
\mu_\alpha &\sim N(0, 1) \\ 
\sigma_\alpha &\sim N(0, 1)
\end{align*}
$$

`brms` formula and prior: 

```{r}
f <- bf(
  error_va ~ (1 | participantId),
  sigma ~ (1 | participantId) + offset(log(va.ans.x))
)

p <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 1), class = sd, group = participantId, lb = 0),
  prior(normal(0, 1), class = Intercept, dpar = "sigma"),
  prior(normal(0, 1), class = sd, group = participantId, dpar = sigma, lb = 0)
)
```

```{r}
#| eval: false
#| echo: false

task3_df %>%
  mutate(error_va = va.select.y - va.ans.y, log.va.ans.x = log(va.ans.x)) %>%
  get_prior(formula = f, data = .)

task3_df %>%
  mutate(error_va = va.select.y - va.ans.y, log.va.ans.x = log(va.ans.x)) %>%
  make_stancode(formula = f, data = ., prior = p)
```

Fit model: 

```{r}
#| output: false
#| label: task 3 fit normal

m3 <- task3_df %>%
  mutate(error_va = va.resp.y - va.ans.y, log.va.ans.x = log(va.ans.x)) %>%
  brm(
    formula = f,
    data = .,
    family = gaussian(),
    prior = p,
    chains = 4,
    cores = 4,
    file = here("analysis", "fitted_models", "task3"),
    file_refit = "on_change",
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.99),
    iter = 6000
  )
```

Check fit and posterior: 

```{r}
summary(m3)
plot(m3)
```

```{r}
#| layout-ncol: 2
#| label: task 3
#| code-fold: true

pp_check(m3, ndraws = 100) +
  theme(text = element_text(size = 10)) +
  theme(legend.position = c(0.9, 0.5)) -> p1

pp_check(m3, ndraws = 100, type = "pit_ecdf") +
  theme(text = element_text(size = 10)) -> p2

p1
p2
```

```{r}
#| eval: false
#| echo: false

ggsave(
  here("figs", "task3-pp-check.pdf"),
  p1,
  height = 5,
  width = 10,
  units = "cm",
  dpi = 300
)
ggsave(
  here("figs", "task3-pit-ecdf.pdf"),
  p2,
  height = 5,
  width = 5,
  units = "cm",
  dpi = 300
)
```

Plot result: 

```{r}
#| label: Task 3 - Results
#| code-fold: true

prediction_grid <- task3_df %>%
  mutate(error_va = va.resp.y - va.ans.y) %>%
  data_grid(
    va.ans.x = seq_range(va.ans.x, n = 100),
    participantId = unique(participantId)
  ) %>%
  add_predicted_draws(m3)

prediction_grid %>%
  ggplot(aes(x = va.ans.x, y = .prediction)) +
  stat_lineribbon(
    .width = c(.95, .80, .50),
    # alpha = 0.5,
    linewidth = 0.5
  ) +
  scale_fill_brewer(palette = "Blues") +
  geom_point(
    data = task3_df,
    aes(y = va.resp.y - va.ans.y, x = va.ans.x),
    alpha = 0.5,
    size = 0.5
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none") +
  coord_cartesian(ylim = c(-0.25, 0.4)) +
  labs(title = "Project To Curve", y = "Error", x = "Distance to Y-Axis") -> p3

p3
```

```{r}
#| eval: false
#| echo: false

ggsave(
  here("figs", "task3.svg"),
  plot = p3,
  height = 330,
  width = 330,
  units = "px",
  dpi = 300
)
```
