---
title: "MaxSlope Operator"
html:
  reference-location: margin
  embed-resources: true
author: "Anonymous"
execute: 
  freeze: auto
---

```{r}
#| echo: false
#| output: false
#| label: setup

library(tidyverse)
library(ggplot2)
library(sgt)
library(numDeriv)
library(ggdist)
# setting the theme
theme_set(theme_minimal(base_size = 12))
# model
library(brms)
library(cmdstanr)
options(brms.backend = "cmdstanr")
options(mc.cores = parallel::detectCores())
# plot results
library(tidybayes)
library(here)

# source the helper functions
source(here("R", "helpers.R"))
# theme setting for bayesplot
bayesplot::bayesplot_theme_update(text = element_text(family = "sans"))
```

# Overview / TLDR 

Recall that Task 4 is the task where we ask the participant to find the highest sloped on the CDF (i.e., find mode on CDF). Aka `MaxSlope` Operator. 

We ran the following models: 

- The [Exponential model](@sec-exponential)
- The [Lomax model](@sec-lomax)
- **best-performing model**^[... by LOO model comparison] The [Weibull model](@sec-weibull)
- The [Log-Normal model](@sec-logNormal)

Finally we compared all models in [here](@sec-model-compare). 

# Set up

Read in data and relevant scales: 

```{r}
#| label: read-in-data
#| code-fold: true

task_df <- read_rds(here("data", "part1", "task.rds"))
participants <- read_rds(here("data", "part1", "participants.rds"))
exp1_x <- make_va_scale(data_min = -5, data_max = 5, px_min = 50, px_max = 585)
exp1_y <- make_va_scale(data_min = 0, data_max = 1, px_min = 410, px_max = 15)
```

Ratio related helper function: 

```{r}
# x: one unit in data maps to (585 - 50)/(5 - (-5)) in pixel
# y: one unit in data maps to (410 - 15)/(1 - 0) in pixel
data_to_screen_slope <- function(k) {
  return(k * 395 / 53.5)
}
data_to_va_slope <- function(data_x, data_y, dist, px_per_mm) {
  eps <- 1e-6
  s_y <- abs(
    exp1_y$data_to_va(data_y + eps, dist, px_per_mm) -
      exp1_y$data_to_va(data_y, dist, px_per_mm)
  ) /
    eps
  s_x <- (abs(
    exp1_x$data_to_va(data_x + eps, dist, px_per_mm) -
      exp1_x$data_to_va(data_x, dist, px_per_mm)
  ) /
    eps)
  return(s_y / s_x)
}
```

Get task-related specific data: 

```{r}
#| label: get task 4 data

task4_df <- task_df %>%
  filter(task == "task4") %>%
  select(-slider.x, -slider.y) %>%
  right_join(participants, by = join_by(participantId)) %>%
  mutate(
    data.resp.slope = dsgt(
      data.select.x,
      param.mu,
      param.sigma,
      param.lambda,
      param.p,
      param.q,
      mean.cent = FALSE
    ),
    data.ans.slope = dsgt(
      param.mu,
      param.mu,
      param.sigma,
      param.lambda,
      param.p,
      param.q,
      mean.cent = FALSE
    ),
    data.ans.x = param.mu,
    data.ans.y = 0.5 *
      (psgt(
        data.ans.x + tolerance,
        param.mu,
        param.sigma,
        param.lambda,
        param.p,
        param.q,
        mean.cent = FALSE
      ) +
        psgt(
          data.ans.x - tolerance,
          param.mu,
          param.sigma,
          param.lambda,
          param.p,
          param.q,
          mean.cent = FALSE
        )),
    va.resp.slope = data.resp.slope *
      data_to_va_slope(
        data.select.x,
        data.select.y,
        dist_to_screen,
        pixelToMM
      ),
    va.ans.slope = data.ans.slope *
      data_to_va_slope(
        data.ans.x,
        data.ans.y,
        dist_to_screen,
        pixelToMM
      )
  )
```

```{r}
#| echo: false
#| eval: false

write_csv(task4_df, here("data", "processed", "task4.csv"))
```

```{r}
#| code-fold: true
#| layout-ncol: 2
#| eval: false

## Plot relations

task4_df %>% ggplot(aes(x = data.resp.slope - data.ans.slope)) + geom_dots()
task4_df %>% ggplot(aes(x = va.resp.slope - va.ans.slope)) + geom_dots()
task4_df %>% ggplot(aes(x = abs(va.resp.slope - va.ans.slope))) + geom_dots()
task4_df %>%
  ggplot(aes(x = va.resp.slope - va.ans.slope, y = dist_to_screen)) +
  geom_point(alpha = 0.5)
```

```{r}
#| layout-ncol: 2
#| code-fold: true
#| eval: false

# Checking to see if there's any obvious relation between signed error and parameters of the distribution.
# nothing apparent.

task4_df %>%
  ggplot(aes(y = data.resp.slope - data.ans.slope, x = dist_to_screen)) +
  geom_point()

task4_df %>%
  ggplot(aes(y = data.resp.slope - data.ans.slope, x = param.mu)) +
  geom_point()

task4_df %>%
  ggplot(aes(y = data.resp.slope - data.ans.slope, x = param.sigma)) +
  geom_point()

task4_df %>%
  ggplot(aes(y = data.resp.slope - data.ans.slope, x = param.lambda)) +
  geom_point()

task4_df %>%
  ggplot(aes(y = data.resp.slope - data.ans.slope, x = param.p)) +
  geom_point()

task4_df %>%
  ggplot(aes(y = data.resp.slope - data.ans.slope, x = param.q)) +
  geom_point()

task4_df %>% ggplot(aes(x = data.resp.slope - data.ans.slope)) + geom_dots()
```

# Exponential Model {#sec-exponential}

Math model: 

$$
\begin{align*}
\text{error_i} &\sim \text{Exponential}(\lambda_{\text{PID}[i]}) \\ 
\log(\lambda_j) &\sim \mathcal{N}(\mu_\lambda, \sigma_\lambda) \\ 
\mu_\lambda &\sim \mathcal{N}(0, 1) \\ 
\sigma_\lambda &\sim \mathcal{N}^+(0, 1) 
\end{align*}
$$

Formula and prior: 

```{r}
f <- bf(error ~ 0, mu ~ (1 | participantId), family = exponential())

# mu has log-link, so we don't pose any bounds on it
p <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 1), class = sd, group = participantId, lb = 0)
)
```

```{r}
#| eval: false
#| echo: false
#| label: check stan code

task4_df %>%
  mutate(error = abs(va.resp.slope - va.ans.slope)) %>%
  get_prior(f, data = .)

task4_df %>%
  mutate(error = abs(va.resp.slope - va.ans.slope)) %>%
  make_stancode(f, data = ., prior = p)
```

Fit the model: 

```{r}
#| output: false
#| label: fit exponential

model.exponential <- task4_df %>%
  mutate(error = abs(va.resp.slope - va.ans.slope)) %>%
  # brms need rsp > 0 for exp
  mutate(error = ifelse(error < tolerance, tolerance, error)) %>%
  brm(
    formula = f,
    data = .,
    prior = p,
    chains = 4,
    cores = 4,
    file = here("analysis", "fitted_models", "task4.exponential"),
    file_refit = "on_change",
    save_pars = save_pars(all = TRUE),
    iter = 6000
  )
```

Check fit and posterior:

```{r}
summary(model.exponential)
plot(model.exponential)
```

```{r}
#| layout-ncol: 2
#| code-fold: true
pp_check(model.exponential, ndraws = 100)
pp_check(model.exponential, ndraws = 100, type = "pit_ecdf")
```

```{r}
#| eval: false
#| echo: false

# model.exponential %>% get_variables()

model.exponential %>%
  spread_draws(b_Intercept, sd_participantId__Intercept) %>%
  mutate(lambda = exp(b_Intercept)) %>%
  pivot_longer(5:6) %>%
  ggplot(aes(x = value, y = name)) +
  stat_halfeye()
```

Overlay plot: 

```{r}
#| code-fold: true
prediction <- task4_df %>%
  mutate(error = abs(va.resp.slope - va.ans.slope)) %>%
  mutate(error = ifelse(error < tolerance, tolerance, error)) %>%
  add_predicted_draws(model.exponential, re_formula = NA)

prediction %>%
  ggplot() +
  stat_slab(aes(x = .prediction), fill = "skyblue", alpha = 0.8) +
  geom_dots(
    data = task4_df,
    aes(x = abs(va.resp.slope - va.ans.slope))
  )
```

```{r}
#| code-fold: true
task4_df %>%
  modelr::data_grid(error = 0) %>%
  add_predicted_draws(model.exponential, re_formula = NA) %>%
  ggplot() +
  stat_slab(aes(x = .prediction), fill = "lightblue", alpha = 1) +
  geom_dots(
    data = task4_df,
    aes(x = abs(va.resp.slope - va.ans.slope))
  )
```

# Lomax Model {#sec-lomax}

Lomax is Pareto Type II distribution, which is implemented in `cmdstanr`. We had similar models for task 2. 

```{r}
lomax_family <- custom_family(
  "lomax",
  dpars = c("mu", "lambda", "alpha"), # includes mu as required
  links = c("identity", "log", "log"), # appropriate links
  lb = c(-Inf, 0, 0), # appropriate bounds
  type = "real"
)

stan_funs <- "
  real lomax_lpdf(real y, real mu, real lambda, real alpha) {
    return pareto_type_2_lpdf(y | 0, lambda, alpha);  
  }

  real lomax_rng(real mu, real lambda, real alpha) {
    return pareto_type_2_rng(0, lambda, alpha);  
  }
"
```

Specify formula and prior^[Note that the Half-Cauchy prior for hyperpriors led to divergent transitions, so we switched to truncated normal for priors]: 

```{r}
# Define the formula for the model
f <- bf(
  error ~ 0,
  mu ~ 0,
  alpha ~ (1 | participantId),
  lambda ~ (1 | participantId)
)

p <- c(
  prior(normal(0, 1), class = Intercept, dpar = alpha),
  prior(normal(0, 1), class = sd, group = participantId, dpar = alpha, lb = 0),
  prior(normal(0, 1), class = Intercept, dpar = lambda),
  prior(normal(0, 1), class = sd, group = participantId, dpar = lambda, lb = 0)
)
```

```{r}
#| eval: false
#| echo: false

task4_df %>%
  mutate(error = abs(va.resp.slope - va.ans.slope)) %>%
  mutate(error = ifelse(error < tolerance, tolerance, error)) %>%
  get_prior(formula = f, family = lomax_family)

task4_df %>%
  mutate(error = abs(va.resp.slope - va.ans.slope) + tolerance) %>%
  make_stancode(formula = f, family = lomax_family, prior = p)
```

```{r}
#| output: false
#| label: fit lomax

model.lomax <- task4_df %>%
  mutate(error = abs(va.resp.slope - va.ans.slope)) %>%
  mutate(error = ifelse(error < tolerance, tolerance, error)) %>%
  brm(
    formula = f,
    data = .,
    family = lomax_family,
    prior = p,
    stanvars = stanvar(scode = stan_funs, block = "functions"),
    chains = 4,
    cores = 4,
    file = here("analysis", "fitted_models", "task4.lomax"),
    file_refit = "always",
    control = list(adapt_delta = 0.99),
    iter = 6000,
    save_pars = save_pars(all = TRUE)
  )
```

Check fit:

```{r}
# pairs(model.lomax)
summary(model.lomax)
plot(model.lomax)
```

```{r}
#| eval: false
#| echo: false

# Check the posterior distribution of parameters, alpha and lambda:

model.lomax %>%
  spread_draws(`^b_.*`, regex = TRUE) %>%
  mutate(
    b_alpha = exp(b_alpha_Intercept),
    b_lambda = exp(b_lambda_Intercept)
  ) %>%
  pivot_longer(6:7) %>%
  ggplot(aes(x = value, y = name)) +
  stat_halfeye() +
  facet_wrap(~name, scales = "free") +
  geom_vline(xintercept = 1e-10, linetype = "dashed", color = "gray")
```

Define custom functions^[Check out this [link](https://cran.r-project.org/web/packages/brms/vignettes/brms_customfamilies.html?utm_source=chatgpt.com) for more details.] for posterior predictive check (`pp_check`) and model comparison (`loo`):

```{r}
#| eval: false
# update(model.lomax, recompile = TRUE)
# TODO: use this https://rok-cesnovar.github.io/misc/exposing_cmdstanr_udf.html to debug
expose_functions(model.lomax, vectorize = TRUE)

# Define the posterior_predict function in R as per vignette
posterior_predict_lomax <- function(i, draws, ...) {
  alpha <- draws$dpars$alpha[i]
  lambda <- draws$dpars$lambda[i]
  # Use Stan's rng function indirectly via brms
  lomax_rng(0, lambda, alpha)
}

log_lik_lomax <- function(i, prep) {
  alpha <- brms::get_dpar(prep, "alpha", i = i)
  lambda <- brms::get_dpar(prep, "lambda", i = i)
  y <- prep$data$Y[i]
  lomax_lpdf(y, 0, lambda, alpha)
}
```

```{r}
#| layout-ncol: 2
#| eval: false
pp_check(model.lomax, ndraws = 100)
pp_check(model.lomax, ndraws = 100, type = "pit_ecdf")
```

Overlay plot:

```{r}
#| eval: false
prediction <- task4_df %>%
  mutate(error = abs(va.resp.slope - va.ans.slope)) %>%
  add_predicted_draws(model.lomax, ndraws = 100)

prediction %>%
  ggplot() +
  stat_slab(aes(x = .prediction), fill = "lightblue", alpha = 1) +
  geom_dots(
    data = task4_df,
    aes(x = abs(va.resp.slope - va.ans.slope))
  )
```

# Weibull distribution {#sec-weibull}

- [wikipedia](https://en.wikipedia.org/wiki/Weibull_distribution)
- [stan](https://mc-stan.org/docs/functions-reference/positive_continuous_distributions.html#weibull-distribution)
- it's also implemented in brms! So unlike the [Lomax model](@sec-lomax), we don't need to write custom functions. 

Specify formula and prior: 

```{r}
f <- bf(
  error ~ 0,
  mu ~ (1 | participantId), # log link
  shape ~ (1 | participantId), # log link
  family = weibull
)

p <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 1), class = sd, group = participantId, lb = 0),
  prior(normal(0, 1), class = Intercept, dpar = shape),
  prior(normal(0, 1), class = sd, group = participantId, dpar = shape, lb = 0)
)
```

```{r}
#| eval: false
#| echo: false

task4_df %>%
  mutate(error = abs(va.resp.slope - va.ans.slope)) %>%
  mutate(error = ifelse(error < tolerance, tolerance, error)) %>%
  get_prior(formula = f)

task4_df %>%
  mutate(error = abs(va.resp.slope - va.ans.slope) + tolerance) %>%
  make_stancode(formula = f, prior = p)
```

Fit the model:

```{r}
#| output: false
#| label: fit weibull

model.weibull <- task4_df %>%
  mutate(error = abs(va.resp.slope - va.ans.slope)) %>%
  mutate(error = ifelse(error < tolerance, tolerance, error)) %>% # brms needs value to be > 0
  brm(
    formula = f,
    data = .,
    prior = p,
    chains = 4,
    cores = 4,
    file = here("analysis", "fitted_models", "task4.weibull"),
    file_refit = "on_change",
    save_pars = save_pars(all = TRUE),
    iter = 6000
  )
```

Check fit and posterior:

```{r}
summary(model.weibull)
plot(model.weibull)
```

```{r}
#| layout-ncol: 2
#| code-fold: true

pp_check(model.weibull, ndraws = 100) +
  coord_cartesian(xlim = c(0, 10)) +
  theme(text = element_text(size = 12)) +
  theme(legend.position = c(0.9, 0.5)) -> p1

pp_check(model.weibull, ndraws = 100, type = "pit_ecdf") +
  theme(text = element_text(size = 4)) -> p2

p1
p2
```

```{r}
#| echo: false
#| eval: false

ggsave(
  here("figs", "task4-pp-check.pdf"),
  p1,
  height = 5,
  width = 10,
  units = "cm",
  dpi = 300
)

ggsave(
  here("figs", "task4-pit-ecdf.pdf"),
  p2,
  height = 2,
  width = 2,
  units = "in",
  dpi = 300
)
```

```{r}
#| eval: false
#| echo: false

# model.weibull %>% get_variables()

model.weibull %>%
  spread_draws(b_Intercept, b_shape_Intercept) %>%
  mutate(mu = exp(b_Intercept), shape = exp(b_shape_Intercept)) %>%
  pivot_longer(6:7) %>%
  ggplot(aes(x = value, y = name)) +
  stat_halfeye()
```

Overlay plot:

```{r}
#| code-fold: true
#| layout-ncol: 2

# prediction <- task4_df %>%
#   mutate(error = abs(data_to_screen_slope(data.ans.slope) - data_to_screen_slope(data.select.slope))) %>%
#   add_predicted_draws(model.weibull, ndraws = 100)

task4_df %>%
  modelr::data_grid(error = 0) %>%
  add_predicted_draws(model.weibull, re_formula = NA) %>%
  ggplot() +
  stat_slab(aes(x = .prediction), fill = "lightblue", alpha = 1) +
  geom_dots(
    data = task4_df,
    aes(x = abs(va.resp.slope - va.ans.slope))
  )

task4_df %>%
  modelr::data_grid(error = 0) %>%
  add_predicted_draws(model.weibull, re_formula = NA) %>%
  ggplot() +
  stat_slab(aes(x = .prediction), fill = "lightblue", alpha = 1) +
  geom_dots(
    data = task4_df,
    aes(x = abs(va.resp.slope - va.ans.slope))
  ) +
  labs(y = "Probability", x = "Signed error") +
  coord_cartesian(xlim = c(0, 5))
```

Posterior draws + strip plot of actual answers: 

```{r}
#| code-fold: true
#| layout-ncol: 3

task4_df %>%
  modelr::data_grid(error = 0, participantId = unique(participantId)) %>%
  add_predicted_draws(model.weibull, ndraws = 100) %>% # ndraws = 100
  ggplot() +
  stat_slab(
    aes(x = .prediction, group = .draw),
    color = "lightblue",
    alpha = 0.5,
    fill = NA,
    n = 3000,
    linewidth = 0.5
  ) +
  scale_thickness_shared() +
  geom_rug(
    data = task4_df,
    aes(x = abs(va.resp.slope - va.ans.slope)),
    linewidth = 0.5
  ) +
  stat_slab(
    data = task4_df,
    aes(x = abs(va.resp.slope - va.ans.slope)),
    color = "darkblue",
    fill = NA,
    n = 3000,
    linewidth = 0.5
  ) +
  labs(y = "density", x = "signed error") +
  coord_cartesian(xlim = c(0, 4)) +
  scale_y_continuous(breaks = 0) -> p

p
```

```{r}
#| echo: false
#| eval: false
ggsave(
  here("figs", "task4.pdf"),
  plot = p,
  height = 6,
  width = 12,
  units = "cm",
  dpi = 300
)
```

# Log-Normal distribution {#sec-logNormal}

Formula and prior: 

```{r}
f <- bf(
  error ~ 0,
  mu ~ (1 | participantId),
  sigma ~ (1 | participantId),
  family = lognormal()
)

p <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 1), class = sd, group = participantId, lb = 0),
  prior(normal(0, 1), class = Intercept, dpar = sigma),
  prior(normal(0, 1), class = sd, group = participantId, dpar = sigma, lb = 0)
)
```

```{r}
#| echo: false
#| eval: false

task4_df %>%
  mutate(error = abs(va.resp.slope - va.ans.slope)) %>%
  get_prior(data = ., formula = f)
```

Fit the model: 

```{r}
#| output: false
#| label: fit lognormal

model.lognormal <- task4_df %>%
  mutate(error = abs(va.resp.slope - va.ans.slope)) %>%
  mutate(error = ifelse(error < tolerance, tolerance, error)) %>% # brms needs value to be > 0
  brm(
    formula = f,
    data = .,
    prior = p,
    chains = 4,
    cores = 4,
    file = here("analysis", "fitted_models", "task4.lognormal"),
    file_refit = "always",
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.95),
    iter = 6000
  )
```

Check fit and posterior:

```{r}
summary(model.lognormal)
plot(model.lognormal)
```

```{r}
#| layout-ncol: 2
#| code-fold: true
pp_check(model.lognormal, ndraws = 100)
pp_check(model.lognormal, ndraws = 100, type = "pit_ecdf")
```

Overlay plot:

```{r}
#| code-fold: true
prediction <- task4_df %>%
  mutate(
    error = abs(va.resp.slope - va.ans.slope)
  ) %>%
  add_predicted_draws(model.lognormal, ndraws = 100)

prediction %>%
  ggplot() +
  stat_slab(aes(x = .prediction), fill = "lightblue", alpha = 1) +
  geom_dots(
    data = task4_df,
    aes(
      x = abs(va.resp.slope - va.ans.slope)
    )
  )
```

```{r}
#| code-fold: true
task4_df %>%
  modelr::data_grid(error = 0) %>%
  add_predicted_draws(model.lognormal, re_formula = NA, ndraws = 100) %>%
  ggplot() +
  stat_slab(aes(x = .prediction), fill = "lightblue", alpha = 1) +
  geom_dots(
    data = task4_df,
    aes(
      x = abs(va.resp.slope - va.ans.slope)
    )
  )
```

# Model Comparison {#sec-model-compare}

```{r}
#| eval: false
loo(
  model.exponential,
  model.laplace,
  # model.lomax,
  model.weibull,
  model.lognormal #,
  # moment_match = TRUE
)
```

The Weibull model performs the best. 

```{r}
#| layout-ncol: 2

pp_check(model.exponential, ndraws = 100, type = "pit_ecdf") +
  labs(title = "Exponential")
# pp_check(model.lomax, ndraws = 100, type = "pit_ecdf") +
#   labs(title = "Lomax")
# pp_check(model.weibull, ndraws = 100, type = "pit_ecdf") +
#   labs(title = "Weibull")
pp_check(model.lognormal, ndraws = 100, type = "pit_ecdf") +
  labs(title = "Log-Normal")
```

`pit_ecdf` is the empirical CDF of PIT (Probability Integral Transform) values. This helps us assess whether the model is well-calibrated; if it is, then the PIT ECDF plot should follow the 45 degree diagonal line, which is what we see for Weibull. 

