---
title: "Generate ReVISit configuration"
---

# Overview 

For easier presentation, I will use `.qmd` to section off things. 

```{r}
#| echo: false 

library(jsonlite)
library(purrr)
library(jsonvalidate)
library(httr)
library(here)
library(tidyverse)
```

## Initialize JSON validator 
... for checking whether the generated json file is schematically correct with ReVISit. 

```{r}
# --- initialize json validator 
schema_url <- "https://raw.githubusercontent.com/revisit-studies/study/v2.3.1/src/parser/StudyConfigSchema.json"
schema <- content(GET(schema_url), as = "text")
validator <- json_validator(schema, engine = "ajv") 
# need to use ajv to handle the newer ones
```

## Obtain data 

```{r}
# --- Get the list of numbers to pull from
# --- since the experiment design is mixed, with the between-subject factor being the visual form, whether its point or point arc. 

exp_json_data <- jsonlite::fromJSON(here("data", "moritz", "stimuli", "exp2-stimuli.json" )) %>% as_tibble(.)

stimuli_df <- exp_json_data %>% select(index, seed, noise, flip, type, data) %>% 
    mutate(noise = as.character(noise))

point_ids <- stimuli_df %>% filter(type == "point") %>% 
  pull(index)

pointArc_ids <- stimuli_df %>% filter(type == "point_arc") %>% 
  pull(index)
```

## Define and generate One-off components, such as UI and metadata 

```{r}
metadata <- list(
  title = "Visual Decoding Operators (Retrieve Value)",
  version = "pilot",
  authors = list("Sheng Long"),
  date = "2025-12-10",
  description = "A study designed to elicit resposnes from participants", 
  organizations = list("Northwestern University")
)

ui_config = list(
  contactEmail = "shenglong@u.northwestern.edu", 
  helpTextPath = "vis-decode-retrieve-value/assets/help.md",
  logoPath = "revisitAssets/revisitLogoSquare.svg",
  withProgressBar = TRUE, 
  autoDownloadStudy = FALSE,
  withSidebar = FALSE,
  minHeightSize = 400,
  minWidthSize = 400, 
  urlParticipantIdParam = "PROLIFIC_PID",
  studyEndMsg = "**Thank you for completing the study.** \n\n You may click this link and return to Prolific: [https://app.prolific.com/submissions/complete?cc=C1O3ZUHV](https://app.prolific.com/submissions/complete?cc=C1O3ZUHV)"
)
```

## Define functions for generating baseComponents 

```{r}
#| code-fold: true 

StudyResponse <- function(id, type, prompt, required = FALSE, hidden = FALSE) {
  list(
    id = id,
    prompt = prompt,
    required = required,
    type = type,
    hidden = hidden
  )
}

BaseComponent <- function(
  type, 
  path, 
  instruction, 
  instructionLocation = c("sidebar", "aboveStimulus", "belowStimulus"), 
  nextButtonLocation = c("sidebar", "aboveStimulus", "belowStimulus"), 
  parameters, 
  response_list = NULL) {
  list(
    type = type,
    path = path,
    instruction = instruction, 
    instructionLocation = match.arg(instructionLocation),
    nextButtonLocation = match.arg(nextButtonLocation),
    parameters = parameters,
    response = if (is.null(response_list)) list() else response_list
  )
}

Component <- function(baseComponent, parameters){
  list(
    baseComponent = baseComponent,
    parameters = parameters
  )
}

convert_to_json <- function(c){
  toJSON(c, auto_unbox = TRUE, pretty = TRUE)
}
```

## Construct specific instance and response 

```{r}
moritz_response = StudyResponse(
  id="weightedAverage-value",
  prompt="The value of response",
  required = FALSE,
  type = "numerical",
  hidden = TRUE
)
```

```{r}
moritz_component <- BaseComponent(
  type="react-component",
  path="vis-decode-retrieve-value/assets/Moritz.tsx",
  instruction= "Drag the slider to where the values are on average; imagine that this is some stock prices",
  instructionLocation = "aboveStimulus",
  nextButtonLocation = "belowStimulus",
  parameters = list(
    taskid = "Moritz",
    taskType = "moritz"
  ),
  response_list = list(moritz_response)
)

intro_component <- list(
    type="markdown", 
    path="TODO",
    response = list(list(
        id = "prolificId",
        prompt = "Please enter your Prolific ID (without any spaces):",
        required = TRUE, 
        location = "belowStimulus", 
        type = "shortText",
        placeholder = "Prolific ID",
        paramCapture = "PROLIFIC_ID"
    ))
)

consent_comp <- list(
    type = "markdown", 
    path = "TODO", 
    nextButtonText = "I agree",
    response = list(list(
        id = "consentApproval",
        prompt = "Do you consent to the study and wish to continue?",
        required = TRUE,
        location = "belowStimulus",
        type = "radio",
        options = list("Decline", "Accept")
    ))
)

post_survey_comp <- list(
    type = "questionnaire",
    response = list(
        list(
            id = "gender",
            prompt = "TODO",
            type = "radio",
            options = list("Male", "Female", "Prefer to self describe", "Prefer not to ...")
        ),
        list(
            id = "self-gender",
            prompt = "TODO",
            location = "belowStimulus",
            type = "shortText",
            required = FALSE
        ),
        list(
            id = "age"
            # TODO 
        ),
        list(
            id = "feedback" 
            # TODO 
        ),
        list(
            id = "feedback-text",
            # TODO 
        ),
        list(
            id = "strategy",
            # TODO 
        )
    )
)
```

```{r}
#| eval: false 
#| echo: false 

convert_to_json(moritz_component)
```

## Construct sequence of components 

```{r}
point_components <- point_ids %>% imap(~ Component(
  baseComponent = "Moritz", 
  parameters = list(params = list(index = .x))
)) %>% 
    set_names(paste0("point_task_", seq_along(point_ids)))

point_arc_components <- pointArc_ids %>% imap(~ Component(
  baseComponent = "Moritz", 
  parameters = list(params = list(index = .x))
)) %>% 
  set_names(paste0("pointArc_task_", seq_along(pointArc_ids)))
```

Build the sequence: 

```{r}
sequence <- list(
  order = "fixed", 
  components = list(
    # "introduction", 
    "$virtual-chinrest.se.full",
    list(
      order = "latinSquare",
      numSamples = 1, # randomly select 1 
      components = list(
        list(
          order = "random", 
          components = c(point_task_names)
        ), 
        list(
          order = "random", 
          components = c(pointArc_task_names)
        )
      )
    )
  )
)
```

```{r}
#| eval: false 
#| code-fold: false 

sequence %>% convert_to_json(.)
```

**TODO: Generate attention check code, as well as trial and introduction, etc.**

## Generate final output 

The final output has the following components: 
1. schema 
2. studyMetadata 
3. uiConfig
4. importedLibraries 
5. baseComponents 
6. components 
7. sequence 


```{r}
final_output = list(
  `$schema` = "https://raw.githubusercontent.com/revisit-studies/study/v2.3.1/src/parser/StudyConfigSchema.json",
  studyMetadata = metadata, 
  uiConfig = ui_config,
  importedLibraries = list("virtual-chinrest"), 
  baseComponents = list(
    Moritz = moritz_component
  ), 
  components = c(point_components, point_arc_components),
  sequence=sequence
)
```

Generate json and validate: 

```{r}
config_json <- final_output %>% convert_to_json(.)
# validate 
validator(config_json, verbose = FALSE)
```

Save output: 

```{r}
config_json %>% write(here("R", "config-r.json"))
```